---
title: "ABCD ADHD menarche all longitudinal model sensitivity analyses"
format: html
editor: visual
---

## Load libraries

```{r load libraries}
library(tidyverse)
library(knitr)
library(tidyr)
library(readxl)
library(readr)
library(brms)
library(pscl)
library(posterior)
library(rstan)
library(bayesplot)
library(tidybayes)
library(emmeans)
```

## Import imputed and orginal datasets

```{r read csv}
imp_df <- read_csv("imputed_dataset.csv", show_col_types = FALSE)
orig_df <- read_csv("2_adhd_menarche_clean.csv", show_col_types = FALSE)
```

### Check dataframe heads

```{r check head}
head(imp_df)
head(orig_df)
```

### Grand mean centre continuous predictors in original dataframe

```{r grand mean centre}
# Variables to grand mean center
vars_to_center <- c("age_years", "inr", "cbcl_scr_syn_attention_r",
                    "cbcl_scr_syn_internal_r", "cbcl_scr_syn_external_r",  
                    "pds_f4_p", "bmi_baseline")

# Compute grand means
grand_means <- orig_df %>%
  summarise(across(all_of(vars_to_center), \(x) mean(x, na.rm = TRUE)))

# Center the variables
orig_df <- orig_df %>%
  mutate(across(all_of(vars_to_center), 
                ~ . - grand_means[[cur_column()]], 
                .names = "{.col}_c"))
```

### Pivot to wide format based on wave

```{r pivot wide}
# Pivot to wide format for longitudinal analyses
imp_df_wide <- imp_df %>%
  pivot_wider(
    id_cols = src_subject_id,
    names_from = wave,
    names_glue = "{.value}_{wave}",
    values_from = -c(src_subject_id, wave)
  )

orig_df_wide <- orig_df %>%
  pivot_wider(
    id_cols = src_subject_id,
    names_from = wave,
    names_glue = "{.value}_{wave}",
    values_from = -c(src_subject_id, wave)
  )
```

### Convert to factors: site ID, family ID, ethnicity, and menarche status factors

```{r factors}
imp_df_wide <- imp_df_wide %>%
  mutate(
    family_id = as.factor(family_id_wave_0),
    site_id = as.factor(site_id_wave_0),
    ethnicity = as.factor(ethnicity_wave_0),
    adhd_diagnosis = as.factor(adhd_diagnosis_wave_0),
    ethnicity_wave_0 <- as.factor(ethnicity_wave_0),
    menarche_status_p_wave_2 <- as.factor(imp_df_wide$menarche_status_p_wave_2),
  )

orig_df_wide <- orig_df_wide %>%
  mutate(
    family_id = as.factor(family_id_wave_0),
    site_id = as.factor(site_id_wave_0),
    ethnicity = as.factor(ethnicity_wave_0),
    adhd_diagnosis = as.factor(adhd_diagnosis_wave_0),
    ethnicity_wave_0 <- as.factor(ethnicity_wave_0),
    menarche_status_p_wave_2 <- as.factor(imp_df_wide$menarche_status_p_wave_2)
  )
```

### Convert to integers: all outcomes

```{r outcomes as integers}
imp_df_wide <- imp_df_wide %>%
  mutate(
    cbcl_scr_syn_attention_r_wave_3 = as.integer(cbcl_scr_syn_attention_r_wave_3)
  )

orig_df_wide <- orig_df_wide %>%
  mutate(
    cbcl_scr_syn_attention_r_wave_3 = as.integer(cbcl_scr_syn_attention_r_wave_3)
  )
```

### Check structure of wide dataframes

```{r check structure wide df}
str(imp_df_wide)
str(orig_df_wide)
```

### Identify all participants with stimulant medications recorded at any wave in imputed dataframe

```{r identitfy stimulant medication participants}
# Identify and count participants who ever used stimulant medications
stimulant_users <- imp_df %>%
  filter(stimulant_use == 1) %>%
  distinct(src_subject_id)

# Count how many unique participants used stimulants
num_stimulant_users <- nrow(stimulant_users)

# Print the count
cat("Number of participants with stimulant use:", num_stimulant_users, "\n")
```

### Remove participants using stimulant medications for nomeds dataset

```{r remove participants using stimulants}
# Remove those participants from the dataframe
imp_df_nomeds <- imp_df %>%
  filter(!src_subject_id %in% stimulant_users$src_subject_id)

# Identify and count participants who ever used stimulant medications
stimulant_users_check <- imp_df_nomeds %>%
  filter(stimulant_use == 1) %>%
  distinct(src_subject_id)

# Count how many unique participants used stimulants
num_stimulant_users <- nrow(stimulant_users_check)

# Print the count
cat("Number of participants with stimulant use:", num_stimulant_users, "\n")
```

### Check nomeds df structure

```{r str nomeds df}
str(imp_df_nomeds)
```

### Nomeds: pivot to wide

```{r pivot wide}
imp_df_nomeds_wide <- imp_df_nomeds %>%
  pivot_wider(
    id_cols = src_subject_id,
    names_from = wave,
    names_glue = "{.value}_{wave}",
    values_from = -c(src_subject_id, wave)
  )
```

### Convert to factors: site ID, family ID, ethnicity, menarche status and ethnicity factors

```{r factors}
imp_df_nomeds_wide <- imp_df_nomeds_wide %>%
  mutate(
    family_id = as.factor(family_id_wave_0),
    site_id = as.factor(site_id_wave_0),
    ethnicity = as.factor(ethnicity_wave_0),
    adhd_diagnosis = as.factor(adhd_diagnosis_wave_0),
    menarche_status_p_wave_2 <- as.factor(menarche_status_p_wave_2),
    ethnicity_wave_0 <- as.factor(ethnicity_wave_0)
  )
```

### Convert to integers: all outcomes

```{r outcomes as integers}
imp_df_nomeds_wide <- imp_df_nomeds_wide %>%
  mutate(
    cbcl_scr_syn_attention_r_wave_3 = as.integer(cbcl_scr_syn_attention_r_wave_3)
  )
```

### Check structure of nomeds wide df

```{r structure nomeds wide}
str(imp_df_nomeds_wide)
```

## Model 1 longitudinal original cleaned dataframe

```{r model 1 long orig}
model_1_long_orig <- brm(
  cbcl_scr_syn_attention_r_wave_3 ~ 
    menarche_status_p_wave_2 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_attention_r_c_wave_0 +
    (1| family_id) + (1 | site_id),
  data = orig_df_wide,  
  family = poisson(),  
  prior = set_prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 1 long output original

```{r model 1 long orig long output}
# Model R2
bayes_R2(model_1_long_orig)

# Model estimates
print(model_1_long_orig, digits = 4)
```

## Model 1b longitudinal original cleaned dataframe

```{r model 1b long orig}
model_1b_long_orig <- brm(
  cbcl_scr_syn_attention_r_wave_3 ~ 
    pds_f4_p_c_wave_2 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_attention_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = orig_df_wide,  
  family = poisson(),  
  prior = set_prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 1b long output original

```{r model 1b long orig long output}
# Model R2
bayes_R2(model_1b_long_orig)

# Model estimates
print(model_1b_long_orig, digits = 4)
```

## Model 1 longitudinal covarying for baseline BMI

```{r model 1 long bmi}
model_1_long_bmi <- brm(
  cbcl_scr_syn_attention_r_wave_3 ~ 
    menarche_status_p_wave_2 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_attention_r_c_wave_0 + bmi_baseline_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = set_prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 1 long output bmi

```{r model 1 long bmi long output}
# Model R2
bayes_R2(model_1_long_bmi)

# Model estimates
print(model_1_long_bmi, digits = 4)
```

## Model 1b longitudinal covarying for baseling BMI

```{r model 1b long bmi}
model_1b_long_bmi <- brm(
  cbcl_scr_syn_attention_r_wave_3 ~ 
    pds_f4_p_c_wave_2 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_attention_r_c_wave_0 + bmi_baseline_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 1b long output bmi

```{r model 1b long bmi long output}
# Model R2
bayes_R2(model_1b_long_bmi)

# Model estimates
print(model_1b_long_bmi, digits = 4)
```

## Model 1 longitudinal prior 1 (normal 0, 0.25)

```{r model 1 long prior 1}
model_1_long_prior1 <- brm(
  cbcl_scr_syn_attention_r_wave_3 ~ 
    menarche_status_p_wave_2 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_attention_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = prior("normal(0, 0.25)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 1 long output prior 1

```{r model 1 long prior 1 long output}
# Model R2
bayes_R2(model_1_long_prior1)

# Model estimates
print(model_1_long_prior1, digits = 4)
```

## Model 1b longitudinal prior 1 (normal 0, 0.25)

```{r model 1b long prior 1}
model_1b_long_prior1 <- brm(
  cbcl_scr_syn_attention_r_wave_3 ~ 
    pds_f4_p_c_wave_2 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_attention_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = prior("normal(0, 0.25)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 1b long output prior 1

```{r model 1b long prior 1 long output}
# Model R2
bayes_R2(model_1b_long_prior1)

# Model estimates
print(model_1b_long_prior1, digits = 4)
```

## Model 1 longitudinal prior 2 (normal 0, 10)

```{r model 1 long prior 2}
model_1_long_prior2 <- brm(
  cbcl_scr_syn_attention_r_wave_3 ~ 
    menarche_status_p_wave_2 +
    age_years_c_wave_2 + ethnicity_wave_0 + inr_c_wave_0 + 
    cbcl_scr_syn_attention_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = prior("normal(0, 10)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 1 long output prior 2

```{r model 1 long prior 2 long output}
# Model R2
bayes_R2(model_1_long_prior2)

# Model estimates
print(model_1_long_prior2, digits = 4)
```

## Model 1b longitudinal prior 2 (normal 0, 10)

```{r model 1b long prior 2}
model_1b_long_prior2 <- brm(
  cbcl_scr_syn_attention_r_wave_3 ~ 
    pds_f4_p_c_wave_2 +
    age_years_c_wave_2 + ethnicity_wave_0 + inr_c_wave_0 + 
    cbcl_scr_syn_attention_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = prior("normal(0, 10)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 1b long output prior 2

```{r model 1b long prior 2 long output}
# Model R2
bayes_R2(model_1b_long_prior2)

# Model estimates
print(model_1b_long_prior2, digits = 4)
```

## Model 1 longitudinal excluding participants taking stimulant medications

```{r model 1 long nomeds}
model_1_long_nomeds <- brm(
  cbcl_scr_syn_attention_r_wave_3 ~ 
    menarche_status_p_wave_2 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_attention_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_nomeds_wide,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 1 long output no meds

```{r model 1 long nomeds long output}
# Model R2
bayes_R2(model_1_long_nomeds)

# Model estimates
print(model_1_long_nomeds, digits = 4)
```

## Model 1b longitudinal excluding participants taking stimulant medication

```{r model 1b long nomeds}
model_1b_long_nomeds <- brm(
  cbcl_scr_syn_attention_r_wave_3 ~ 
    pds_f4_p_c_wave_2 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_attention_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_nomeds_wide,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 1b long output no meds

```{r model 1b long nomeds long output}
# Model R2
bayes_R2(model_1b_long_nomeds)

# Model estimates
print(model_1b_long_nomeds, digits = 4)
```
