---
title: "ABCD ADHD menarche analysis: Removing participants with recorded stimulant medications sensitivity analysis Models 1 - 5 cross-sectional"
format: html
editor: visual
---

## Load libraries

```{r load libraries}
# load libraries
library(tidyverse)
library(knitr)
library(tidyr)
library(readxl)
library(readr)
library(brms)
library(pscl)
library(posterior)
library(rstan)
library(bayesplot)
library(tidybayes)
library(emmeans)
```

## Import imputed dataframe

```{r import imputed dataframe}
imp_df <- read_csv("imputed_dataset.csv", show_col_types = FALSE)
```

### Check imputed dataframe

```{r check head}
str(imp_df)
```

### Identify all participants with stimulant medications recorded at any wave

```{r identify stimulant records}
# Identify and count participants who ever used stimulant medications
stimulant_users <- imp_df %>%
  filter(stimulant_use == 1) %>%
  distinct(src_subject_id)

# Count how many unique participants used stimulants
num_stimulant_users <- nrow(stimulant_users)

# Print the count
cat("Number of participants with stimulant medications recorded:", num_stimulant_users, "\n")
```

### Remove participants with stimulant medications recorded

```{r remove stimulant use}
# Remove those participants from the dataframe
imp_df_nomeds <- imp_df %>%
  filter(!src_subject_id %in% stimulant_users$src_subject_id)

# Identify and count participants who ever used stimulant medications
stimulant_users_check <- imp_df_nomeds %>%
  filter(stimulant_use == 1) %>%
  distinct(src_subject_id)

# Count how many unique participants used stimulants
num_stimulant_users <- nrow(stimulant_users_check)

# Print the count
cat("Number of participants with stimulant use:", num_stimulant_users, "\n")
```

### Convert to factors: subject ID, site ID, family ID, ethnicity, and menarche status

```{r factors}
imp_df_nomeds <- imp_df_nomeds %>%
  mutate(across(c(src_subject_id, family_id, site_id, ethnicity, menarche_status_p, adhd_diagnosis), as.factor))
```

### Convert to integers: all outcomes

```{r outcomes as integers}
imp_df_nomeds <- imp_df_nomeds %>%
  mutate(across(c(cbcl_scr_syn_attention_r, cbcl_scr_syn_internal_r, cbcl_scr_syn_external_r), as.integer))
```

### Create observations ID variable

```{r obs id}
imp_df_nomeds$obs_id <- seq_len(nrow(imp_df_nomeds))
```

### Structure of nomeds dataframe

```{r df structure}
str(imp_df_nomeds)
```

## Model 1b

```{r model 1b}
model_1b_nomeds <- brm(
  cbcl_scr_syn_attention_r ~ pds_f4_p_c
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id),
  data = imp_df_nomeds,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 1b output

```{r model 1b output}
# Model R2
bayes_R2(model_1b_nomeds)

# Model estimates
print(model_1b_nomeds, digits = 4)
```

## Model 2

```{r model 2}
model_2_nomeds <- brm(
  cbcl_scr_syn_internal_r ~  menarche_status_p * adhd_diagnosis
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = imp_df_nomeds,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 2 output

```{r model 2 output}
# Model R2
bayes_R2(model_2_nomeds)

# Model estimates
print(model_2_nomeds, digits = 4)
```

## Model 2b

```{r model 2b}
model_2b_nomeds <- brm(
  cbcl_scr_syn_internal_r ~  pds_f4_p_c * adhd_diagnosis 
    + age_years_c + ethnicity + inr_c + 
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id), 
  data = imp_df_nomeds,
  family = poisson(),
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2,
  seed = 1234
)
```

### Model 2b output

```{r model 2b output}
# Model R2
bayes_R2(model_2b_nomeds)

# Model estimates
print(model_2b_nomeds, digits = 4)
```

## Model 3

```{r model 3}
model_3_nomeds <- brm(
  cbcl_scr_syn_external_r ~ menarche_status_p * adhd_diagnosis 
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = imp_df_nomeds,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 3 output

```{r model 3 output}
# Model R2
bayes_R2(model_3_nomeds)

# Model estimates
print(model_3_nomeds, digits = 4)
```

## Model 3b

```{r model 3b}
model_3b_nomeds <- brm(
  cbcl_scr_syn_external_r ~  pds_f4_p_c * adhd_diagnosis 
    + age_years_c + ethnicity + inr_c + 
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id), 
  data = imp_df_nomeds,
  family = poisson(), 
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2,
  seed = 1234
)
```

### Model 3b output

```{r model 3b output}
# Model R2
bayes_R2(model_3b_nomeds)

# Model estimates
print(model_3b_nomeds, digits = 4)
```

## Model 4

```{r model 4}
model_4_nomeds <- brm(
  cbcl_scr_syn_internal_r ~  menarche_status_p * cbcl_scr_syn_attention_r_c
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = imp_df_nomeds,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 4 output

```{r model 4 output}
# Model R2
bayes_R2(model_4_nomeds)

# Model estimates
print(model_4_nomeds, digits = 4)
```

## Model 5

```{r model 5}
model_5_nomeds <- brm(
  cbcl_scr_syn_external_r ~  menarche_status_p * cbcl_scr_syn_attention_r_c
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = imp_df_nomeds,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 5 output

```{r model 5 output}
# Model R2
bayes_R2(model_5_nomeds)

# Model estimates
print(model_5_nomeds, digits = 4)
```
