---
title: "ABCD ADHD menarche analysis: Original cleaned dataframe sensitivity analysis Models 1 - 5 cross-sectional"
format: html
editor: visual
---

## Load libraries

```{r load libraries}
library(tidyverse)
library(knitr)
library(tidyr)
library(readxl)
library(readr)
library(brms)
library(pscl)
library(posterior)
library(rstan)
library(bayesplot)
library(tidybayes)
library(emmeans)
```

## Import original dataframe

```{r import imputed dataframe}
orig_df <- read_csv("2_adhd_menarche_clean.csv", show_col_types = FALSE)
```

### Check original dataframs

```{r check head}
head(orig_df)
```

### Remove wave 4 data since only half collected

```{r remove wave 4}
orig_df <- orig_df %>%
  filter(wave != "wave_4")
```

### Grand mean centre continuous predictors

```{r grand mean centre}
# Variables to grand mean center
vars_to_center <- c("age_years", "inr", "cbcl_scr_syn_attention_r", "pds_f4_p")

# Compute grand means
grand_means <- orig_df %>%
  summarise(across(all_of(vars_to_center), mean, na.rm = TRUE))

# Center the variables
orig_df <- orig_df %>%
  mutate(across(all_of(vars_to_center), 
                ~ . - grand_means[[cur_column()]], 
                .names = "{.col}_c"))
```

### Family and site id variables and propagate across all waves

```{r create family and side id}
orig_df <- orig_df %>%
  group_by(src_subject_id) %>%
  mutate(family_id = first(na.omit(rel_family_id[wave == "wave_0"]))) %>%
  ungroup()

orig_df <- orig_df %>%
  group_by(src_subject_id) %>%
  mutate(site_id = first(na.omit(site_id_l[wave == "wave_0"]))) %>%
  ungroup()
```

### Convert to factors: subject ID, site ID, family ID, ethnicity, and menarche status

```{r factors}
orig_df <- orig_df %>%
  mutate(across(c(src_subject_id, family_id, site_id, ethnicity, menarche_status_p, adhd_diagnosis), as.factor))
```

### Convert to integers: all outcomes

```{r outcomes as integers}
orig_df <- orig_df %>%
  mutate(across(c(cbcl_scr_syn_attention_r, cbcl_scr_syn_internal_r, cbcl_scr_syn_external_r), as.integer))
```

### Create observations ID variable

```{r obs id}
orig_df$obs_id <- seq_len(nrow(orig_df))
```

### Structure of dataframe

```{r df structure}
str(orig_df)
```

## Model 1b

```{r Model 1b}
model_1b_orig <- brm(
  cbcl_scr_syn_attention_r ~ pds_f4_p_c
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id),
  data = orig_df,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b",) ,
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234 
)
```

### Model 1b output

```{r model 1b output}
# Model R2
bayes_R2(model_1b_orig)

# Model estimates
print(model_1b_orig, digits = 4)
```

## Model 2

```{r model 2}
model_2_orig <- brm(
  cbcl_scr_syn_internal_r ~ adhd_diagnosis * menarche_status_p
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = orig_df,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b",) ,
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234 
)
```

### Model 2 output

```{r model 2 output}
# Model R2
bayes_R2(model_2_orig)

# Model estimates
print(model_2_orig, digits = 4)
```

## Model 2b

```{r model 2b}
model_2b_orig <- brm(
  cbcl_scr_syn_internal_r ~ adhd_diagnosis * pds_f4_p_c 
    + age_years_c + ethnicity + inr_c + 
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id), 
  data = orig_df,
  family = poisson(),
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 2b output

```{r model 2b output}
# Model R2
bayes_R2(model_2b_orig)

# Model estimates
print(model_2b_orig, digits = 4)
```

## Model 3

```{r model 3}
model_3_orig <- brm(
  cbcl_scr_syn_external_r ~ adhd_diagnosis * menarche_status_p
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = orig_df,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b",) ,
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234 
)
```

### Model 3 output

```{r model 3 output}
# Model R2
bayes_R2(model_3_orig)

# Model estimates
print(model_3_orig, digits = 4)
```

## Model 3b

```{r model 3b}
model_3b_orig <- brm(
  cbcl_scr_syn_external_r ~ adhd_diagnosis * pds_f4_p_c 
    + age_years_c + ethnicity + inr_c + 
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id), 
  data = orig_df,
  family = poisson(),
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 3b output

```{r model 3b output}
# Model R2
bayes_R2(model_3b_orig)

# Model estimates
print(model_3b_orig, digits = 4)
```

## Model 4

```{r model 4}
model_4_orig <- brm(
  cbcl_scr_syn_internal_r ~ cbcl_scr_syn_attention_r_c * menarche_status_p
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = orig_df,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b",) ,
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234 
)
```

### Model 4 output

```{r model 4 output}
# Model R2
bayes_R2(model_4_orig)

# Model estimates
print(model_4_orig, digits = 4)
```

## Model 5

```{r model 5}
model_5_orig <- brm(
  cbcl_scr_syn_external_r ~ cbcl_scr_syn_attention_r_c * menarche_status_p  
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = orig_df,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b",) ,
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234 
)
```

### Model 5 output

```{r model 5 output}
# Model R2
bayes_R2(model_5_orig)

# Model estimates
print(model_5_orig, digits = 4)
```
