---
title: "ABCD ADHD menarche analysis: Prior 1 sensitivity analysis Models 1 - 5 cross-sectional"
format: html
editor: visual
---

## Load libraries

```{r load libraries}
library(tidyverse)
library(knitr)
library(tidyr)
library(readxl)
library(readr)
library(brms)
library(pscl)
library(posterior)
library(rstan)
library(bayesplot)
library(tidybayes)
library(emmeans)
```

## Import imputed dataframe

```{r import imputed dataframe}
imp_df <- read_csv("imputed_dataset.csv", show_col_types = FALSE)
```

### Check imputed dataframe

```{r check head}
head(imp_df)
```

### Convert to factors: subject ID, site ID, family ID, ethnicity, and menarche status

```{r factors}
imp_df <- imp_df %>%
  mutate(across(c(src_subject_id, family_id, site_id, ethnicity, menarche_status_p, adhd_diagnosis), as.factor))
```

### Convert to integers: all outcomes

```{r outcomes as integers}
imp_df <- imp_df %>%
  mutate(across(c(cbcl_scr_syn_attention_r, cbcl_scr_syn_internal_r, cbcl_scr_syn_external_r), as.integer))
```

### Create observations ID variable

```{r obs id}
imp_df$obs_id <- seq_len(nrow(imp_df))
```

### Structure of dataframe

```{r df structure}
str(imp_df)
```

## Model 1b

```{r model 1b}
model_1b_prior1 <- brm(
  cbcl_scr_syn_attention_r ~ pds_f4_p_c
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id),
  data = imp_df,  
  family = poisson(),  
  prior = set_prior("normal(0, 0.25)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 1b output

```{r model 1b output}
# Model R2
bayes_R2(model_1b_prior1)

# Model estimates
print(model_1b_prior1, digits = 4)
```

## Model 2

```{r model 2}
model_2_prior1 <- brm(
  cbcl_scr_syn_internal_r ~  menarche_status_p * adhd_diagnosis
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = imp_df,  
  family = poisson(),  
  prior = prior("normal(0, 0.25)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 2 output

```{r model 2 output}
# Model R2
bayes_R2(model_2_prior1)

# Model estimates
print(model_2_prior1, digits = 4)
```

## Model 2b

```{r model 2b}
model_2b_prior1 <- brm(
  cbcl_scr_syn_internal_r ~  pds_f4_p_c * adhd_diagnosis 
    + age_years_c + ethnicity + inr_c + 
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id), 
  data = imp_df,
  family = poisson(),
  prior = prior("normal(0, 0.25)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2,
  seed = 1234
)
```

### Model 2b output

```{r model 2b output}
# Model R2
bayes_R2(model_2b_prior1)

# Model estimates
print(model_2b_prior1, digits = 4)
```

## Model 3

```{r model 3}
model_3_prior1 <- brm(
  cbcl_scr_syn_external_r ~ menarche_status_p * adhd_diagnosis 
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = imp_df,  
  family = poisson(),  
  prior = prior("normal(0, 0.25)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 3 output

```{r model 3 output}
# Model R2
bayes_R2(model_3_prior1)

# Model estimates
print(model_3_prior1, digits = 4)
```

## Model 3b

```{r model 3b}
model_3b_prior1 <- brm(
  cbcl_scr_syn_external_r ~  pds_f4_p_c * adhd_diagnosis 
    + age_years_c + ethnicity + inr_c + 
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id), 
  data = imp_df,
  family = poisson(), 
  prior = prior("normal(0, 0.25)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2,
  seed = 1234
)
```

### Model 3b output

```{r model 3b output}
# Model R2
bayes_R2(model_3b_prior1)

# Model estimates
print(model_3b_prior1, digits = 4)
```

## Model 4

```{r model 4}
model_4_prior1 <- brm(
  cbcl_scr_syn_internal_r ~  menarche_status_p * cbcl_scr_syn_attention_r_c
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = imp_df,  
  family = poisson(),  
  prior = prior("normal(0, 0.25)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 4 output

```{r model 4 output}
# Model R2
bayes_R2(model_4_prior1)

# Model estimates
print(model_4_prior1, digits = 4)
```

## Model 5

```{r model 5}
model_5_prior1 <- brm(
  cbcl_scr_syn_external_r ~  menarche_status_p * cbcl_scr_syn_attention_r_c
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = imp_df,  
  family = poisson(),  
  prior = prior("normal(0, 0.25)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Model 5 output

```{r model 5 output}
# Model R2
bayes_R2(model_5_prior1)

# Model estimates
print(model_5_prior1, digits = 4)
```
