---
title: "ABCD ADHD menarche cross-sectional analyses"
author: "Emily Friedel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

### Load libraries

```{r libraries}
library(tidyverse)
library(ggplot2)
library(mice)
library(brms)
library(posterior)
library(bayestestR)
library(emmeans)
library(purrr)
```

### Import saved imputed dfs

```{r import-data}
analytic_df <- read_csv("analytic_df.csv")
```

### Initial checks

```{r initial-checks}
summary(unique(analytic_df$participant_id))
summary(is.na(analytic_df$breast_development_p_sc))
length(analytic_df)
colnames(analytic_df)
```

## 1. Pre model variable checks

### 1.1 Convert to factors: subject ID, site ID, family ID, race/ethnicity, and menarche status

```{r factors}
analytic_df <- analytic_df %>%
  mutate(across(c(participant_id, family_id, site_id, ethnrace, menarche_status_p, adhd_status_life), as.factor))

# set white_nh as reference for enthrace
analytic_df$ethnrace <- relevel(analytic_df$ethnrace, ref = "white_nh")
levels(analytic_df$ethnrace)
```

### 1.2 Convert outcomes integers

```{r outcomes-as-integers}
analytic_df <- analytic_df %>%
  mutate(across(c(adhd_traits, internalising, externalising), as.integer))
```

### 1.3 Structure of dataframe

```{r df-structure}
str(analytic_df)
```

## 2. Set up hypothesis formulas and prior

```{r hypothesis-formulas}
formula_1 <- bf(adhd_traits ~ menarche_status_p
                + age_sc + ethnrace + inr_sc
                + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                )

formula_1b <- bf(adhd_traits ~ breast_development_p_sc
                 + age_sc + ethnrace + inr_sc
                 + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                 )


formula_2 <- bf(internalising ~ menarche_status_p * adhd_status_life
                + age_sc + ethnrace + inr_sc
                + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                + (1 | obs_id)
                )

formula_2b <- bf(internalising ~ breast_development_p_sc * adhd_status_life
                 + age_sc + ethnrace + inr_sc
                 + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                 + (1 | obs_id)
                 )

formula_3 <- bf(externalising ~ menarche_status_p * adhd_status_life
                + age_sc + ethnrace + inr_sc
                + (1 | participant_id) + (1 | family_id) + (1 | site_id) 
                + (1| obs_id)
                )

formula_3b <- bf(externalising ~ breast_development_p_sc * adhd_status_life
                 + age_sc + ethnrace + inr_sc 
                 + (1 | participant_id) + (1 | family_id) + (1 | site_id) 
                 + (1| obs_id)
                 )

formula_4 <- bf(internalising ~ menarche_status_p * adhd_traits_sc
                + age_sc + ethnrace + inr_sc 
                + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                + (1| obs_id)
                )

formula_4b <- bf(internalising ~ breast_development_p_sc * adhd_traits_sc
                 + age_sc + ethnrace + inr_sc 
                 + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                 + (1| obs_id)
                 )

formula_5 <- bf(externalising ~ menarche_status_p * adhd_traits_sc
                + age_sc + ethnrace + inr_sc 
                + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                + (1| obs_id)
                )

formula_5b <- bf(externalising ~ breast_development_p_sc * adhd_traits_sc
                 + age_sc + ethnrace + inr_sc 
                 + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                 + (1| obs_id)
                 )

# Set prior
prior <- prior(normal(0, 1), class = "b")
```

### 2.1 Check default priors

```{r default-prior}
get_prior(
  formula = formula_1,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1b,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_2,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_2b,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_3,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_3b,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_4,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_4b,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5b,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)
```

## 3. Run models

### Model 1

```{r model-1}
fit_1 <- brm(
  formula = formula_1,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1, "fit_1.rds")
```

### Model 1 Output

```{r model-1-output}
# Model R2
bayes_R2(fit_1)

# Model estimates
print(fit_1, digits = 3)
```

### Model 1 Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-1-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_1)) |>
  rownames_to_column("term")

# Pull coefficient and CIs
beta  <- fe$Estimate[fe$term == "menarche_status_pY"]
lower <- fe$Q2.5[fe$term == "menarche_status_pY"]
upper <- fe$Q97.5[fe$term == "menarche_status_pY"]

# Convert to % difference
result_1 <- tibble(
  Variable = "Menarche (yes vs no)",
  Estimate = round((exp(beta)  - 1) * 100, 2),
  `Lower 95% CI` = round((exp(lower) - 1) * 100, 2),
  `Upper 95% CI` = round((exp(upper) - 1) * 100, 2)
)

# View
result_1
```

### Model 1 Diagnostics

```{r model-1-diagnostics}
# Prior Summary
prior_summary(fit_1)

# Posterior Predictive Checks
plot(fit_1, ask = FALSE)

# PPC plots
pp_check(fit_1, type = "bars", ndraws = 100)
pp_check(fit_1, type = "hist", ndraws = 10)
pp_check(fit_1, type = "error_hist")
pp_check(fit_1, type = "stat", stat = "mean")
pp_check(fit_1, type = "stat", stat = "sd")
pp_check(fit_1, type = "scatter_avg")
```

### Model 1B

```{r model-1b}
fit_1b <- brm(
  formula = formula_1b,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1b, "fit_1b.rds")
```

### Model 1B Output

```{r model-1b-output}
# Model R2
bayes_R2(fit_1b)

# Model estimates
print(fit_1b, digits = 3)
```

### Model 1B Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-1b-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_1b)) |>
  tibble::rownames_to_column("term")

# Pull the coefficient and CIs for the variable of interest
beta  <- fe$Estimate[fe$term == "breast_development_p_sc"]
lower <- fe$Q2.5[fe$term == "breast_development_p_sc"]
upper <- fe$Q97.5[fe$term == "breast_development_p_sc"]

# Get SD of the original (unscaled) variable
sd_x <- sd(analytic_df$breast_development_p, na.rm = TRUE)

# Convert to % and round
result_1b <- tibble(
  Variable = "Breast development",
  Estimate = round((exp(beta  / (2 * sd_x)) - 1) * 100, 2),
  `Lower 95% CI` = round((exp(lower / (2 * sd_x)) - 1) * 100, 2),
  `Upper 95% CI` = round((exp(upper / (2 * sd_x)) - 1) * 100, 2)
)

# View
result_1b
```

### Model 1B Diagnostics

```{r model-1b-diagnostics}
# Prior Summary
prior_summary(fit_1b)

# Posterior Predictive Checks
plot(fit_1b, ask = FALSE)

# PPC plots
pp_check(fit_1b, type = "bars", ndraws = 100)
pp_check(fit_1b, type = "hist", ndraws = 10)
pp_check(fit_1b, type = "error_hist")
pp_check(fit_1b, type = "stat", stat = "mean")
pp_check(fit_1b, type = "stat", stat = "sd")
pp_check(fit_1b, type = "scatter_avg")
```
  
### Model 2

```{r model-2}
fit_2 <- brm(
  formula = formula_2,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_2, "fit_2.rds")
```

### Model 2 Output

```{r model-2-output}
# Model R2
bayes_R2(fit_2)

# Model estimates
print(fit_2, digits = 3)
```

### Model 2 Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-2-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_2)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# Pull terms and convert
result_2 <- tibble(
  Variable = c(
    "Menarche (yes vs no) [ADHD = 0]",
    "ADHD (yes vs no) [Menarche = 0]",
    "Interaction (Menarche × ADHD)"
  ),
  
  Estimate = round(c(
    to_pct(fe$Estimate[fe$term == "menarche_status_pY"]),
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"]),
    to_pct(fe$Estimate[fe$term == "menarche_status_pY:adhd_status_lifeY"])
  ), 2),
  
  `Lower 95% CI` = round(c(
    to_pct(fe$Q2.5[fe$term == "menarche_status_pY"]),
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"]),
    to_pct(fe$Q2.5[fe$term == "menarche_status_pY:adhd_status_lifeY"])
  ), 2),
  
  `Upper 95% CI` = round(c(
    to_pct(fe$Q97.5[fe$term == "menarche_status_pY"]),
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"]),
    to_pct(fe$Q97.5[fe$term == "menarche_status_pY:adhd_status_lifeY"])
  ), 2)
)

# View
result_2
```

### Model 2 Diagnostics

```{r model-2-diagnostics}
# Prior Summary
prior_summary(fit_2)

# MCMC sampling diagnostics
plot(fit_2)

# Posterior predictive checks
pp_check(fit_2, type = "bars", ndraws = 100)
pp_check(fit_2, type = "hist", ndraws = 10)
pp_check(fit_2, type = "error_hist")
pp_check(fit_2, type = "stat", stat = "mean")
pp_check(fit_2, type = "stat", stat = "sd")
pp_check(fit_2, type = "scatter_avg")
```

### Model 2B

```{r model-2b}
fit_2b <- brm(
  formula = formula_2b,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_2b, "fit_2b.rds")
```

### Model 2B Output

```{r model-2b-output}
# Model R2
bayes_R2(fit_2b)

# Model estimates
print(fit_2b, digits = 3)
```

### Model 2B Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-2b-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_2b)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# SD of original breast development variable
sd_bd <- sd(analytic_df$breast_development_p, na.rm = TRUE)

# Build table
result_2b <- tibble(
  Variable = c(
    "Breast development (per 1-unit) [ADHD = 0]",
    "ADHD (yes vs no) [Breast development = mean]",
    "Interaction (Breast development × ADHD)"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"]),
    (exp(fe$Estimate[fe$term == "breast_development_p_sc:adhd_status_lifeY"] / (2 * sd_bd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"]),
    (exp(fe$Q2.5[fe$term == "breast_development_p_sc:adhd_status_lifeY"] / (2 * sd_bd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"]),
    (exp(fe$Q97.5[fe$term == "breast_development_p_sc:adhd_status_lifeY"] / (2 * sd_bd)) - 1) * 100
  ), 2)
)

# View
result_2b
```

### Model 2B Diagnostics

```{r model-2b-diagnostics}
# Prior Summary
prior_summary(fit_2b)

# Posterior Predictive Checks
plot(fit_2b, ask = FALSE)

# PPC plots
pp_check(fit_2b, type = "bars", ndraws = 100)
pp_check(fit_2b, type = "hist", ndraws = 10)
pp_check(fit_2b, type = "error_hist")
pp_check(fit_2b, type = "stat", stat = "mean")
pp_check(fit_2b, type = "stat", stat = "sd")
pp_check(fit_2b, type = "scatter_avg")
```

### Model 3

```{r model-3}
fit_3 <- brm(
  formula = formula_3,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_3, "fit_3.rds") 
```

### Model 3 Output

```{r model-3-output}
# Model R2
bayes_R2(fit_3)

# Model estimates
print(fit_3, digits = 3)
```

### Model 3 Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-3-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_3)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# Pull terms and convert
result_3 <- tibble(
  Variable = c(
    "Menarche (yes vs no) [ADHD = 0]",
    "ADHD (yes vs no) [Menarche = 0]",
    "Interaction (Menarche × ADHD)"
  ),
  
  Estimate = round(c(
    to_pct(fe$Estimate[fe$term == "menarche_status_pY"]),
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"]),
    to_pct(fe$Estimate[fe$term == "menarche_status_pY:adhd_status_lifeY"])
  ), 2),
  
  `Lower 95% CI` = round(c(
    to_pct(fe$Q2.5[fe$term == "menarche_status_pY"]),
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"]),
    to_pct(fe$Q2.5[fe$term == "menarche_status_pY:adhd_status_lifeY"])
  ), 2),
  
  `Upper 95% CI` = round(c(
    to_pct(fe$Q97.5[fe$term == "menarche_status_pY"]),
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"]),
    to_pct(fe$Q97.5[fe$term == "menarche_status_pY:adhd_status_lifeY"])
  ), 2)
)

# View
result_3
```

### Model 3 Diagnostics

```{r model-3-diagnostics}
# Prior Summary
prior_summary(fit_3)

# Posterior Predictive Checks
plot(fit_3, ask = FALSE)

# PPC plots
pp_check(fit_3, type = "bars", ndraws = 100)
pp_check(fit_3, type = "hist", ndraws = 10)
pp_check(fit_3, type = "error_hist")
pp_check(fit_3, type = "stat", stat = "mean")
pp_check(fit_3, type = "stat", stat = "sd")
pp_check(fit_3, type = "scatter_avg")
```

### Model 3B

```{r model-3b}
fit_3b <- brm(
  formula = formula_3b,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_3b, "fit_3b.rds")
```

### Model 3B Output

```{r model-3b-output}
# Model R2
bayes_R2(fit_3b)

# Model estimates
print(fit_3b, digits = 3)
```

### Model 3B Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-3b-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_3b)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# SD of original breast development variable
sd_bd <- sd(analytic_df$breast_development_p, na.rm = TRUE)

# Build table
result_3b <- tibble(
  Variable = c(
    "Breast development (per 1-unit) [ADHD = 0]",
    "ADHD (yes vs no) [Breast development = mean]",
    "Interaction (Breast development × ADHD)"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"]),
    (exp(fe$Estimate[fe$term == "breast_development_p_sc:adhd_status_lifeY"] / (2 * sd_bd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"]),
    (exp(fe$Q2.5[fe$term == "breast_development_p_sc:adhd_status_lifeY"] / (2 * sd_bd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"]),
    (exp(fe$Q97.5[fe$term == "breast_development_p_sc:adhd_status_lifeY"] / (2 * sd_bd)) - 1) * 100
  ), 2)
)

# View
result_3b
```

### Model 3B Diagnostics

```{r model-3b-diagnostics}
# Prior Summary
prior_summary(fit_3b)

# Posterior Predictive Checks
plot(fit_3b, ask = FALSE)

# PPC plots
pp_check(fit_3b, type = "bars", ndraws = 100)
pp_check(fit_3b, type = "hist", ndraws = 10)
pp_check(fit_3b, type = "error_hist")
pp_check(fit_3b, type = "stat", stat = "mean")
pp_check(fit_3b, type = "stat", stat = "sd")
pp_check(fit_3b, type = "scatter_avg")
```

### Model 4

```{r model-4}
fit_4 <- brm(
  formula = formula_4,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_4, "fit_4.rds")
```

### Model 4 Output

```{r model-4-output}
# Model R2
bayes_R2(fit_4)

# Model estimates
print(fit_4, digits = 3)
```

### Model 4 Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-4-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_4)) |>
  rownames_to_column("term")

# Convert log scale to %
to_pct <- function(x) (exp(x) - 1) * 100

# SD of original ADHD traits 
sd_adhd <- sd(analytic_df$adhd_traits, na.rm = TRUE)

# Build table
result_4 <- tibble(
  Variable = c(
    "Menarche (yes vs no) [ADHD traits = mean]",
    "ADHD traits (per 1-unit increase) [Menarche = 0]",
    "Interaction (Menarche × ADHD traits)"
  ),
  
  Estimate = round(c(
    to_pct(fe$Estimate[fe$term == "menarche_status_pY"]),
    (exp(fe$Estimate[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "menarche_status_pY:adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    to_pct(fe$Q2.5[fe$term == "menarche_status_pY"]),
    (exp(fe$Q2.5[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "menarche_status_pY:adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    to_pct(fe$Q97.5[fe$term == "menarche_status_pY"]),
    (exp(fe$Q97.5[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "menarche_status_pY:adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2)
)

# View
result_4
```

### Model 4 Interaction Analysis

```{r model-4-interaction}
# Get emtrends
emtrends_4 <- emtrends(
  fit_4, 
  pairwise ~ menarche_status_p, 
  var = "adhd_traits_sc"
)

# Extract summary
emtrends_df <- as.data.frame(summary(emtrends_4$emtrends))

# SD of original ADHD traits
sd_adhd <- sd(analytic_df$adhd_traits, na.rm = TRUE)

# Function to convert to % change (same as before)
to_pct <- function(x) (exp(x) - 1) * 100

# Convert slopes back to original units and %
result_emtrends_4 <- emtrends_df |>
  mutate(
    Estimate = to_pct(adhd_traits_sc.trend / (2 * sd_adhd)),
    `Lower 95% CI` = to_pct(lower.HPD / (2 * sd_adhd)),
    `Upper 95% CI` = to_pct(upper.HPD / (2 * sd_adhd))
  ) |>
  select(
    menarche_status_p,
    Estimate,
    `Lower 95% CI`,
    `Upper 95% CI`
  ) |>
mutate(across(where(is.numeric), ~ sprintf("%.2f", .x)))

result_emtrends_4
```

### Model 4 Diagnostics

```{r model-4-diagnostics}
# Prior Summary
prior_summary(fit_4)

# Posterior Predictive Checks
plot(fit_4, ask = FALSE)

# PPC plots
pp_check(fit_4, type = "bars", ndraws = 100)
pp_check(fit_4, type = "hist", ndraws = 10)
pp_check(fit_4, type = "error_hist")
pp_check(fit_4, type = "stat", stat = "mean")
pp_check(fit_4, type = "stat", stat = "sd")
pp_check(fit_4, type = "scatter_avg")
```

### Model 4B

```{r model-4b}
fit_4b <- brm(
  formula = formula_4b,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_4b, "fit_4b.rds")
```

### Model 4B Output

```{r model-4b-output}
# Model R2
bayes_R2(fit_4b)

# Model estimates
print(fit_4b, digits = 3)
```

### Model 4B Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-4b-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_4b)) |>
  rownames_to_column("term")

# Convert log scale to %
to_pct <- function(x) (exp(x) - 1) * 100

# SDs of original variables
sd_adhd <- sd(analytic_df$adhd_traits, na.rm = TRUE)
sd_bd   <- sd(analytic_df$breast_development_p, na.rm = TRUE)

# Build table
result_4b <- tibble(
  Variable = c(
    "Breast development (per 1-unit) [ADHD traits = mean]",
    "ADHD traits (per 1-unit) [Breast development = mean]",
    "Interaction (Breast development × ADHD traits)"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "breast_development_p_sc:adhd_traits_sc"] / (4 * sd_bd * sd_adhd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "breast_development_p_sc:adhd_traits_sc"] / (4 * sd_bd * sd_adhd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "breast_development_p_sc:adhd_traits_sc"] / (4 * sd_bd * sd_adhd)) - 1) * 100
  ), 2)
)

# View
result_4b
```

### Model 4B Diagnostics

```{r model-4b-diagnostics}
# Prior Summary
prior_summary(fit_4b)

# Posterior Predictive Checks
plot(fit_4b, ask = FALSE)

# PPC plots
pp_check(fit_4b, type = "bars", ndraws = 100)
pp_check(fit_4b, type = "hist", ndraws = 10)
pp_check(fit_4b, type = "error_hist")
pp_check(fit_4b, type = "stat", stat = "mean")
pp_check(fit_4b, type = "stat", stat = "sd")
pp_check(fit_4b, type = "scatter_avg")
```

### Model 5

```{r model-5}
fit_5 <- brm(
  formula = formula_5,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_5, "fit_5.rds")
```

### Model 5 Output

```{r model-5-output}
# Model R2
bayes_R2(fit_5)

# Model estimates
print(fit_5, digits = 3)
```

### Model 5 Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-5-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_5)) |>
  rownames_to_column("term")

# Convert log scale to %
to_pct <- function(x) (exp(x) - 1) * 100

# SD of original ADHD traits 
sd_adhd <- sd(analytic_df$adhd_traits, na.rm = TRUE)

# Build table
result_5 <- tibble(
  Variable = c(
    "Menarche (yes vs no) [ADHD traits = mean]",
    "ADHD traits (per 1-unit increase) [Menarche = 0]",
    "Interaction (Menarche × ADHD traits)"
  ),
  
  Estimate = round(c(
    to_pct(fe$Estimate[fe$term == "menarche_status_pY"]),
    (exp(fe$Estimate[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "menarche_status_pY:adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    to_pct(fe$Q2.5[fe$term == "menarche_status_pY"]),
    (exp(fe$Q2.5[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "menarche_status_pY:adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    to_pct(fe$Q97.5[fe$term == "menarche_status_pY"]),
    (exp(fe$Q97.5[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "menarche_status_pY:adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2)
)

# View
result_5
```

### Model 5 Interaction Analysis

```{r model-5-interaction}
# Get emtrends
emtrends_5 <- emtrends(
  fit_5, 
  pairwise ~ menarche_status_p, 
  var = "adhd_traits_sc"
)

# Extract summary
emtrends_df <- as.data.frame(summary(emtrends_5$emtrends))

# SD of original ADHD traits
sd_adhd <- sd(analytic_df$adhd_traits, na.rm = TRUE)

# Function to convert to % change (same as before)
to_pct <- function(x) (exp(x) - 1) * 100

# Convert slopes back to original units and %
result_emtrends_5 <- emtrends_df |>
  mutate(
    Estimate = to_pct(adhd_traits_sc.trend / (2 * sd_adhd)),
    `Lower 95% CI` = to_pct(lower.HPD / (2 * sd_adhd)),
    `Upper 95% CI` = to_pct(upper.HPD / (2 * sd_adhd))
  ) |>
  select(
    menarche_status_p,
    Estimate,
    `Lower 95% CI`,
    `Upper 95% CI`
  ) |>
mutate(across(where(is.numeric), ~ sprintf("%.2f", .x)))

result_emtrends_5
```

### Model 5 Diagnostics

```{r model-5-diagnostics}
# Prior Summary
prior_summary(fit_5)

# Posterior Predictive Checks
plot(fit_5, ask = FALSE)

# PPC plots
pp_check(fit_5, type = "bars", ndraws = 100)
pp_check(fit_5, type = "hist", ndraws = 10)
pp_check(fit_5, type = "error_hist")
pp_check(fit_5, type = "stat", stat = "mean")
pp_check(fit_5, type = "stat", stat = "sd")
pp_check(fit_5, type = "scatter_avg")
```

### Model 5B

```{r model-5b}
fit_5b <- brm(
  formula = formula_5b,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_5b, "fit_5b.rds")
```

### Model 5B output

```{r model-5b-output}
# Model R2
bayes_R2(fit_5b)

# Model estimates
print(fit_5b, digits = 3)
```

### Model 5B Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-5b-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_5b)) |>
  rownames_to_column("term")

# Convert log scale to %
to_pct <- function(x) (exp(x) - 1) * 100

# SDs of original variables
sd_adhd <- sd(analytic_df$adhd_traits, na.rm = TRUE)
sd_bd   <- sd(analytic_df$breast_development_p, na.rm = TRUE)

# Build table
result_5b <- tibble(
  Variable = c(
    "Breast development (per 1-unit) [ADHD traits = mean]",
    "ADHD traits (per 1-unit) [Breast development = mean]",
    "Interaction (Breast development × ADHD traits)"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "breast_development_p_sc:adhd_traits_sc"] / (4 * sd_bd * sd_adhd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "breast_development_p_sc:adhd_traits_sc"] / (4 * sd_bd * sd_adhd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "breast_development_p_sc:adhd_traits_sc"] / (4 * sd_bd * sd_adhd)) - 1) * 100
  ), 2)
)

# View
result_5b
```

### Model 5B Diagnostics

```{r model-5b-diagnostics}
# Prior Summary
prior_summary(fit_5b)

# Posterior Predictive Checks
plot(fit_5b, ask = FALSE)

# PPC plots
pp_check(fit_5b, type = "bars", ndraws = 100)
pp_check(fit_5b, type = "hist", ndraws = 10)
pp_check(fit_5b, type = "error_hist")
pp_check(fit_5b, type = "stat", stat = "mean")
pp_check(fit_5b, type = "stat", stat = "sd")
pp_check(fit_5b, type = "scatter_avg")
```
