---
title: "ABCD ADHD menarche cross-sectional analyses (no interaction term)"
author: "Emily Friedel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

### Load libraries

```{r libraries}
library(tidyverse)
library(ggplot2)
library(mice)
library(brms)
library(posterior)
library(bayestestR)
library(emmeans)
library(purrr)
```

### Import saved imputed df

```{r import-data}
analytic_df <- read_csv("analytic_df.csv")
```

### Initial checks

```{r initial-checks}
summary(unique(analytic_df$participant_id))
summary(is.na(analytic_df$breast_development_p_sc))
length(analytic_df)
colnames(analytic_df)
```

## 1. Pre model variable checks

### 1.1 Convert to factors: subject ID, site ID, family ID, ethnicity, and menarche status

```{r factors}
analytic_df <- analytic_df %>%
  mutate(across(c(participant_id, family_id, site_id, ethnrace, menarche_status_p, adhd_status_life), as.factor))

# set white_nh as reference for enthrace
analytic_df$ethnrace <- relevel(analytic_df$ethnrace, ref = "white_nh")
levels(analytic_df$ethnrace)
```

### 1.2 Convert to integers: all outcomes

```{r outcomes-as-integers}
analytic_df <- analytic_df %>%
  mutate(across(c(adhd_traits, internalising, externalising), as.integer))
```

### 1.3 Structure of dataframe

```{r df-structure}
str(analytic_df)
```

## 2. Set up hypothesis formulas and prior

```{r hypothesis-formulas}
# Formulas
formula_2_nointeract <- bf(internalising ~ menarche_status_p + adhd_status_life
                + age_sc + ethnrace + inr_sc
                + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                + (1 | obs_id)
                )

formula_2b_nointeract <- bf(internalising ~ breast_development_p_sc + adhd_status_life
                 + age_sc + ethnrace + inr_sc
                 + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                 + (1 | obs_id)
                 )

formula_3_nointeract <- bf(externalising ~ menarche_status_p + adhd_status_life
                + age_sc + ethnrace + inr_sc
                + (1 | participant_id) + (1 | family_id) + (1 | site_id) 
                + (1| obs_id)
                )

formula_3b_nointeract <- bf(externalising ~ breast_development_p_sc + adhd_status_life
                 + age_sc + ethnrace + inr_sc 
                 + (1 | participant_id) + (1 | family_id) + (1 | site_id) 
                 + (1| obs_id)
                 )

formula_4b_nointeract <- bf(internalising ~ breast_development_p_sc + adhd_traits_sc
                 + age_sc + ethnrace + inr_sc 
                 + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                 + (1| obs_id)
                 )

formula_5b_nointeract <- bf(externalising ~ breast_development_p_sc + adhd_traits_sc
                 + age_sc + ethnrace + inr_sc 
                 + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                 + (1| obs_id)
                 )


# Set prior
prior <- prior(normal(0, 1), class = "b")
```

### 2.1 Check default prior

```{r default-prior}
get_prior(
  formula = formula_2_nointeract,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_2b_nointeract,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_3_nointeract,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_3b_nointeract,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_4b_nointeract,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5b_nointeract,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)
```

## 3. Run models
  
### Model 2 No Interaction

```{r model-2}
fit_2_nointeract <- brm(
  formula = formula_2_nointeract,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_2_nointeract, "fit_2_nointeract.rds")
```

### Model 2 Output

```{r model-2-output}
# Model R2
bayes_R2(fit_2_nointeract)

# Model estimates
print(fit_2_nointeract, digits = 3)
```

### Model 2 Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-2-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_2_nointeract)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# Pull terms and convert
result_2_nointeract <- tibble(
  Variable = c(
    "Menarche (yes vs no)",
    "ADHD (yes vs no)"
  ),
  
  Estimate = round(c(
    to_pct(fe$Estimate[fe$term == "menarche_status_pY"]),
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Lower 95% CI` = round(c(
    to_pct(fe$Q2.5[fe$term == "menarche_status_pY"]),
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Upper 95% CI` = round(c(
    to_pct(fe$Q97.5[fe$term == "menarche_status_pY"]),
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"])
  ), 2)
)

# View
result_2_nointeract
```

### Model 2 Diagnostics

```{r model-2-diagnostics}
# Prior Summary
prior_summary(fit_2_nointeract)

# MCMC sampling diagnostics
plot(fit_2_nointeract)

# Posterior predictive checks
pp_check(fit_2_nointeract, type = "bars", ndraws = 100)
pp_check(fit_2_nointeract, type = "hist", ndraws = 10)
pp_check(fit_2_nointeract, type = "error_hist")
pp_check(fit_2_nointeract, type = "stat", stat = "mean")
pp_check(fit_2_nointeract, type = "stat", stat = "sd")
pp_check(fit_2_nointeract, type = "scatter_avg")
```

### Model 2B

```{r model-2b}
fit_2b_nointeract <- brm(
  formula = formula_2b_nointeract,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_2b_nointeract, "fit_2b_nointeract.rds")
```

### Model 2B Output

```{r model-2b-output}
# Model R2
bayes_R2(fit_2b_nointeract)

# Model estimates
print(fit_2b_nointeract, digits = 3)
```

### Model 2B Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-2b-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_2b_nointeract)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# SD of original breast development variable
sd_bd <- sd(analytic_df$breast_development_p, na.rm = TRUE)

# Build table
result_2b_nointeract <- tibble(
  Variable = c(
    "Breast development (per 1-unit)",
    "ADHD (yes vs no)"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"])
  ), 2)
)

# View
result_2b_nointeract
```

### Model 2B Diagnostics

```{r model-2b-diagnostics}
# Prior Summary
prior_summary(fit_2b_nointeract)

# Posterior Predictive Checks
plot(fit_2b_nointeract, ask = FALSE)

# PPC plots
pp_check(fit_2b_nointeract, type = "bars", ndraws = 100)
pp_check(fit_2b_nointeract, type = "hist", ndraws = 10)
pp_check(fit_2b_nointeract, type = "error_hist")
pp_check(fit_2b_nointeract, type = "stat", stat = "mean")
pp_check(fit_2b_nointeract, type = "stat", stat = "sd")
pp_check(fit_2b_nointeract, type = "scatter_avg")
```

### Model 3

```{r model-3}
fit_3_nointeract <- brm(
  formula = formula_3_nointeract,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_3_nointeract, "fit_3_nointeract.rds") 
```

### Model 3 Output

```{r model-3-output}
# Model R2
bayes_R2(fit_3_nointeract)

# Model estimates
print(fit_3_nointeract, digits = 3)
```

### Model 3 Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-3-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_3_nointeract)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# Pull terms and convert
result_3_nointeract <- tibble(
  Variable = c(
    "Menarche (yes vs no)",
    "ADHD (yes vs no)"
  ),
  
  Estimate = round(c(
    to_pct(fe$Estimate[fe$term == "menarche_status_pY"]),
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Lower 95% CI` = round(c(
    to_pct(fe$Q2.5[fe$term == "menarche_status_pY"]),
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Upper 95% CI` = round(c(
    to_pct(fe$Q97.5[fe$term == "menarche_status_pY"]),
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"])
  ), 2)
)

# View
result_3_nointeract
```

### Model 3 Diagnostics

```{r model-3-diagnostics}
# Prior Summary
prior_summary(fit_3_nointeract)

# Posterior Predictive Checks
plot(fit_3_nointeract, ask = FALSE)

# PPC plots
pp_check(fit_3_nointeract, type = "bars", ndraws = 100)
pp_check(fit_3_nointeract, type = "hist", ndraws = 10)
pp_check(fit_3_nointeract, type = "error_hist")
pp_check(fit_3_nointeract, type = "stat", stat = "mean")
pp_check(fit_3_nointeract, type = "stat", stat = "sd")
pp_check(fit_3_nointeract, type = "scatter_avg")
```

### Model 3B

```{r model-3b}
fit_3b_nointeract <- brm(
  formula = formula_3b_nointeract,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_3b_nointeract, "fit_3b_nointeract.rds")
```

### Model 3B Output

```{r model-3b-output}
# Model R2
bayes_R2(fit_3b_nointeract)

# Model estimates
print(fit_3b_nointeract, digits = 3)
```

### Model 3B Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-3b-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_3b_nointeract)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# SD of original breast development variable
sd_bd <- sd(analytic_df$breast_development_p, na.rm = TRUE)

# Build table
result_3b_nointeract <- tibble(
  Variable = c(
    "Breast development (per 1-unit)",
    "ADHD (yes vs no)"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"])
  ), 2)
)

# View
result_3b_nointeract
```

### Model 3B Diagnostics

```{r model-3b-diagnostics}
# Prior Summary
prior_summary(fit_3b_nointeract)

# Posterior Predictive Checks
plot(fit_3b_nointeract, ask = FALSE)

# PPC plots
pp_check(fit_3b_nointeract, type = "bars", ndraws = 100)
pp_check(fit_3b_nointeract, type = "hist", ndraws = 10)
pp_check(fit_3b_nointeract, type = "error_hist")
pp_check(fit_3b_nointeract, type = "stat", stat = "mean")
pp_check(fit_3b_nointeract, type = "stat", stat = "sd")
pp_check(fit_3b_nointeract, type = "scatter_avg")
```

### Model 4B

```{r model-4b}
fit_4b_nointeract <- brm(
  formula = formula_4b_nointeract,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_4b_nointeract, "fit_4b_nointeract.rds")
```

### Model 4B Output

```{r model-4b-output}
# Model R2
bayes_R2(fit_4b_nointeract)

# Model estimates
print(fit_4b_nointeract, digits = 3)
```

### Model 4B Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-4b-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_4b_nointeract)) |>
  rownames_to_column("term")

# Convert log scale to %
to_pct <- function(x) (exp(x) - 1) * 100

# SDs of original variables
sd_adhd <- sd(analytic_df$adhd_traits, na.rm = TRUE)
sd_bd   <- sd(analytic_df$breast_development_p, na.rm = TRUE)

# Build table
result_4b_nointeract <- tibble(
  Variable = c(
    "Breast development (per 1-unit)",
    "ADHD traits (per 1-unit)"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2)
)

# View
result_4b_nointeract
```

### Model 4B Diagnostics

```{r model-4b-diagnostics}
# Prior Summary
prior_summary(fit_4b_nointeract)

# Posterior Predictive Checks
plot(fit_4b_nointeract, ask = FALSE)

# PPC plots
pp_check(fit_4b_nointeract, type = "bars", ndraws = 100)
pp_check(fit_4b_nointeract, type = "hist", ndraws = 10)
pp_check(fit_4b_nointeract, type = "error_hist")
pp_check(fit_4b_nointeract, type = "stat", stat = "mean")
pp_check(fit_4b_nointeract, type = "stat", stat = "sd")
pp_check(fit_4b_nointeract, type = "scatter_avg")
```

### Model 5B

```{r model-5b}
fit_5b_nointeract <- brm(
  formula = formula_5b_nointeract,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_5b_nointeract, "fit_5b_nointeract.rds")
```

### Model 5B output

```{r model-5b-output}
# Model R2
bayes_R2(fit_5b_nointeract)

# Model estimates
print(fit_5b_nointeract, digits = 3)
```

### Model 5B Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-5b-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_5b_nointeract)) |>
  rownames_to_column("term")

# Convert log scale to %
to_pct <- function(x) (exp(x) - 1) * 100

# SDs of original variables
sd_adhd <- sd(analytic_df$adhd_traits, na.rm = TRUE)
sd_bd   <- sd(analytic_df$breast_development_p, na.rm = TRUE)

# Build table
result_5b_nointeract <- tibble(
  Variable = c(
    "Breast development (per 1-unit)",
    "ADHD traits (per 1-unit)"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "adhd_traits_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2)
)

# View
result_5b_nointeract
```

### Model 5B Diagnostics

```{r model-5b-diagnostics}
# Prior Summary
prior_summary(fit_5b_nointeract)

# Posterior Predictive Checks
plot(fit_5b_nointeract, ask = FALSE)

# PPC plots
pp_check(fit_5b_nointeract, type = "bars", ndraws = 100)
pp_check(fit_5b_nointeract, type = "hist", ndraws = 10)
pp_check(fit_5b_nointeract, type = "error_hist")
pp_check(fit_5b_nointeract, type = "stat", stat = "mean")
pp_check(fit_5b_nointeract, type = "stat", stat = "sd")
pp_check(fit_5b_nointeract, type = "scatter_avg")
```
