---
title: "ABCD ADHD menarche longitudinal analyses"
author: "Emily Friedel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

### Load libraries

```{r libraries}
library(tidyverse)
library(ggplot2)
library(mice)
library(brms)
library(posterior)
library(bayestestR)
library(emmeans)
library(patchwork)
```

### Import saved imputed df

```{r import-data}
analytic_df_wide <- read_csv("analytic_df_wide.csv")
```

### Initial checks

```{r initial-checks}
summary(unique(analytic_df_wide$participant_id))
length(analytic_df_wide)
colnames(analytic_df_wide)
```

## 1. Pre model processing

### Convert to factors: site ID, family ID, ethnicity, and menarche status factors

```{r factors}
analytic_df_wide <- analytic_df_wide %>%
  mutate(
    family_id = as.factor(family_id),
    site_id = as.factor(site_id),
    ethnrace = as.factor(ethnrace),
    adhd_status_life = as.factor(adhd_status_life),
    menarche_status_p_wave_2 = as.factor(menarche_status_p_wave_2)
  )

# set white_nh as reference for enthorace
analytic_df_wide$ethnrace <- relevel(analytic_df_wide$ethnrace, ref = "white_nh")
levels(analytic_df_wide$ethnrace)
```

### Convert to integers: all outcomes

```{r outcomes-as-integers}
analytic_df_wide <- analytic_df_wide %>%
  mutate(
    adhd_traits_wave_3 = as.integer(adhd_traits_wave_3),
    internalising_wave_3 = as.integer(internalising_wave_3),
    externalising_wave_3 = as.integer(externalising_wave_3)
  )
```

### Check head of wide dataframe

```{r check-df}
head(analytic_df_wide)
str(analytic_df_wide)
```

## 2. Set up hypothesis formulas and prior

```{r hypothesis-formulas}
formula_1_long <- bf(adhd_traits_wave_3 ~ menarche_status_p_wave_2
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc 
                  + adhd_traits_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_1b_long <- bf(adhd_traits_wave_3 ~ breast_development_p_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc 
                  + adhd_traits_wave_0_sc 
                  + (1 | family_id) + (1 | site_id)
                  )

formula_2_long <- bf(internalising_wave_3 ~ menarche_status_p_wave_2 * adhd_status_life
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + internalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_2b_long <- bf(internalising_wave_3 ~ breast_development_p_wave_2_sc * adhd_status_life
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + internalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_3_long <- bf(externalising_wave_3 ~ menarche_status_p_wave_2 * adhd_status_life
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc 
                  + externalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_3b_long <- bf(externalising_wave_3 ~ breast_development_p_wave_2_sc * adhd_status_life
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + externalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_4_long <- bf(internalising_wave_3 ~ menarche_status_p_wave_2 * adhd_traits_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + internalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_4b_long <- bf(internalising_wave_3 ~ breast_development_p_wave_2_sc * adhd_traits_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + internalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_5_long <- bf(externalising_wave_3 ~ menarche_status_p_wave_2 * adhd_traits_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + externalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_5b_long <- bf(externalising_wave_3 ~ breast_development_p_wave_2_sc * adhd_traits_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + externalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

prior <- prior(normal(0, 1), class = "b")
```

### 2.1 Check default priors

```{r default-prior}
get_prior(
  formula = formula_1_long,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1b_long,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_2_long,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_2b_long,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_3_long,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_3b_long,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_4_long,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_4b_long,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5_long,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5b_long,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)
```

## 3. Run long models

### Model 1 Long

```{r model-1-long}
fit_1_long <- brm(
  formula = formula_1_long,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1_long, "fit_1_long.rds")
```

### Model 1 Long Output

```{r model-1-long-output}
# Posterior Summary Table
bayes_R2(fit_1_long)

# Model estimates
print(fit_1_long, digits = 3)
```

### Model 1 Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-1_long-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_1_long)) |>
  rownames_to_column("term")

# Pull coefficient and CIs
beta  <- fe$Estimate[fe$term == "menarche_status_p_wave_2Y"]
lower <- fe$Q2.5[fe$term == "menarche_status_p_wave_2Y"]
upper <- fe$Q97.5[fe$term == "menarche_status_p_wave_2Y"]

# Convert to % difference (NO SD adjustment)
result_1_long <- tibble(
  Variable = "Menarche Y (Wave 2)",
  Estimate = round((exp(beta)  - 1) * 100, 2),
  `Lower 95% CI` = round((exp(lower) - 1) * 100, 2),
  `Upper 95% CI` = round((exp(upper) - 1) * 100, 2)
)

# View
result_1_long
```

### Model 1 Long Diagnostics

```{r model-1-diagnostics}
# Prior Summary
prior_summary(fit_1_long)

# Posterior Predictive Checks

plot(fit_1_long, ask = FALSE)

# PPC plots
pp_check(fit_1_long, type = "bars", ndraws = 100)
pp_check(fit_1_long, type = "hist", ndraws = 10)
pp_check(fit_1_long, type = "error_hist")
pp_check(fit_1_long, type = "stat", stat = "mean")
pp_check(fit_1_long, type = "stat", stat = "sd")
pp_check(fit_1_long, type = "scatter_avg")
```

### Model 1B Long

```{r model-1b-long}
fit_1b_long <- brm(
  formula = formula_1b_long,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1b_long, "fit_1b_long.rds")
```

### Model 1B Long Output

```{r model-1b-long-output}
# Posterior Summary Table
bayes_R2(fit_1b_long)

# Model estimates
print(fit_1b_long, digits = 3)
```

### Model 1B Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-1b-long-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_1b_long)) |>
  tibble::rownames_to_column("term")

# Pull the coefficient and CIs for the variable of interest
beta  <- fe$Estimate[fe$term == "breast_development_p_wave_2_sc"]
lower <- fe$Q2.5[fe$term == "breast_development_p_wave_2_sc"]
upper <- fe$Q97.5[fe$term == "breast_development_p_wave_2_sc"]

# Get SD of the original (unscaled) variable
sd_x <- sd(analytic_df_wide$breast_development_p_wave_2, na.rm = TRUE)

# Convert to % and round
result_1b_long <- tibble(
  Variable = "Breast development (Wave 2)",
  Estimate = round((exp(beta  / (2 * sd_x)) - 1) * 100, 2),
  `Lower 95% CI` = round((exp(lower / (2 * sd_x)) - 1) * 100, 2),
  `Upper 95% CI` = round((exp(upper / (2 * sd_x)) - 1) * 100, 2)
)

# View
result_1b_long
```

### Model 1B Long Diagnostics

```{r model-1b-long-diagnostics}
# Prior Summary
prior_summary(fit_1b_long)

# Posterior Predictive Checks

plot(fit_1b_long, ask = FALSE)

# PPC plots
pp_check(fit_1b_long, type = "bars", ndraws = 100)
pp_check(fit_1b_long, type = "hist", ndraws = 10)
pp_check(fit_1b_long, type = "error_hist")
pp_check(fit_1b_long, type = "stat", stat = "mean")
pp_check(fit_1b_long, type = "stat", stat = "sd")
pp_check(fit_1b_long, type = "scatter_avg")
```

### Model 1 and 1B Plots

```{r model-1-1b-plots}
# 1a) Menarche model
emm_menarche <- emmeans(
  fit_1_long,
  ~ menarche_status_p_wave_2,
  type = "response"
)

plot_menarche <- as.data.frame(emm_menarche) %>%
  mutate(
    menarche_status_p_wave_2 = factor(
      menarche_status_p_wave_2,
      levels = c("N", "Y"),
      labels = c("Pre-menarche", "Post-menarche")
    )
  )

p1 <- ggplot(plot_menarche, aes(x = menarche_status_p_wave_2, y = rate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower.HPD, ymax = upper.HPD), width = 0.1) +
  geom_line(aes(group = 1), linewidth = 0.8) +
  coord_cartesian(ylim = c(0, 3)) +
  labs(
    x = NULL,
    y = "Predicted ADHD traits (wave 3)"
  ) +
  theme_classic(base_size = 14) +
  annotate("text", x = 1, y = Inf, label = "1a)", vjust = 1.5, hjust = 1, size = 5)

# 1b) Breast development model

mean_bd <- mean(analytic_df_wide$breast_development_p_wave_2, na.rm = TRUE)
sd_bd   <- sd(analytic_df_wide$breast_development_p_wave_2, na.rm = TRUE)

bd_grid_raw <- seq(1, 4, by = 0.1)
bd_grid_sc  <- (bd_grid_raw - mean_bd) / (2 * sd_bd)

emm_breast <- emmeans(
  fit_1b_long,
  ~ breast_development_p_wave_2_sc,
  at = list(
    breast_development_p_wave_2_sc = bd_grid_sc
  ),
  type = "response"
)

plot_breast <- as.data.frame(emm_breast) %>%
  mutate(
    breast_dev_raw = bd_grid_raw
  )

p2 <- ggplot(plot_breast, aes(x = breast_dev_raw, y = rate)) +
  geom_ribbon(
    aes(ymin = lower.HPD, ymax = upper.HPD),
    alpha = 0.2
  ) +
  geom_line(linewidth = 1) +
  scale_x_continuous(
    breaks = 1:4,
    limits = c(1, 4)
  ) +
  scale_y_continuous(
    breaks = 0:3,
    limits = c(0, 3)
  ) +
  coord_cartesian(ylim = c(0, 3)) +
  labs(
    x = "Breast development",
    y = ""
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.title.x = element_text(margin = margin(t = 8))
  ) +
  theme(
  axis.title.x = element_text(size = 11)
  )+
  annotate(
    "text",
    x = -Inf, y = Inf,
    label = "1b)",
    vjust = 1.5, hjust = -1.5,
    size = 5
  )

# Combine side-by-side plots

(p1 + p2 + plot_layout(ncol = 2)) +
  plot_annotation(caption = "Wave 2 predictors") &
  theme(plot.caption = element_text(hjust = 0.5, size = 12))
```

### Model 2 Long

```{r model-2-long}
fit_2_long <- brm(
  formula = formula_2_long,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_2_long, "fit_2_long.rds")
```

### Model 2 Output

```{r model-2-long-output}
# Posterior Summary Table
bayes_R2(fit_2_long)

# Model estimates
print(fit_2_long, digits = 3)
```

### Model 2 Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-2-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_2_long)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# Pull terms and convert
result_2_long <- tibble(
  Variable = c(
    "Menarche Y (Wave 2) [ADHD = 0]",
    "ADHD Y [Menarche = 0]",
    "Menarche × ADHD"
  ),
  
  Estimate = round(c(
    to_pct(fe$Estimate[fe$term == "menarche_status_p_wave_2Y"]),
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"]),
    to_pct(fe$Estimate[fe$term == "menarche_status_p_wave_2Y:adhd_status_lifeY"])
  ), 2),
  
  `Lower 95% CI` = round(c(
    to_pct(fe$Q2.5[fe$term == "menarche_status_p_wave_2Y"]),
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"]),
    to_pct(fe$Q2.5[fe$term == "menarche_status_p_wave_2Y:adhd_status_lifeY"])
  ), 2),
  
  `Upper 95% CI` = round(c(
    to_pct(fe$Q97.5[fe$term == "menarche_status_p_wave_2Y"]),
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"]),
    to_pct(fe$Q97.5[fe$term == "menarche_status_p_wave_2Y:adhd_status_lifeY"])
  ), 2)
)

# View
result_2_long
```

### Model 2 Long Diagnostics

```{r model-2-diagnostics}
# Prior Summary
prior_summary(fit_2_long)

# Posterior Predictive Checks

plot(fit_2_long, ask = FALSE)

# PPC plots
pp_check(fit_2_long, type = "bars", ndraws = 100)
pp_check(fit_2_long, type = "hist", ndraws = 10)
pp_check(fit_2_long, type = "error_hist")
pp_check(fit_2_long, type = "stat", stat = "mean")
pp_check(fit_2_long, type = "stat", stat = "sd")
pp_check(fit_2_long, type = "scatter_avg")
```

### Model 2B Long

```{r model-2b-long}
fit_2b_long <- brm(
  formula = formula_2b_long,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_2b_long, "fit_2b_long.rds")
```

### Model 2B Long Output

```{r model-2b-long-output}
# Posterior Summary Table
bayes_R2(fit_2b_long)

# Model estimates
print(fit_2b_long, digits = 3)
```

### Model 2B Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-2b-long-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_2b_long)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# SD of original wave 2 breast development variable
sd_bd <- sd(analytic_df_wide$breast_development_p_wave_2, na.rm = TRUE)

# Build table
result_2b_long <- tibble(
  Variable = c(
    "Breast development (Wave 2) [ADHD = 0]",
    "ADHD Y [Breast development = 0]",
    "Breast development × ADHD"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"]),
    (exp(fe$Estimate[fe$term == "breast_development_p_wave_2_sc:adhd_status_lifeY"] / (2 * sd_bd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"]),
    (exp(fe$Q2.5[fe$term == "breast_development_p_wave_2_sc:adhd_status_lifeY"] / (2 * sd_bd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"]),
    (exp(fe$Q97.5[fe$term == "breast_development_p_wave_2_sc:adhd_status_lifeY"] / (2 * sd_bd)) - 1) * 100
  ), 2)
)

# View
result_2b_long
```

### Model 2B Long Diagnostics

```{r model-2b-diagnostics}
# Prior Summary
prior_summary(fit_2b_long)

# Posterior Predictive Checks
plot(fit_2b_long, ask = FALSE)

# PPC plots
pp_check(fit_2b_long, type = "bars", ndraws = 100)
pp_check(fit_2b_long, type = "hist", ndraws = 10)
pp_check(fit_2b_long, type = "error_hist")
pp_check(fit_2b_long, type = "stat", stat = "mean")
pp_check(fit_2b_long, type = "stat", stat = "sd")
pp_check(fit_2b_long, type = "scatter_avg")
```

### Model 3 Long 

```{r model-3-long}
fit_3_long <- brm(
  formula = formula_3_long,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  control = list(adapt_delta = 0.95),
  cores = 2,
  seed = 1234
)

saveRDS(fit_3_long, "fit_3_long.rds")
```

### Model 3 Long Output

```{r model-3-long-output}
# Posterior Summary Table
bayes_R2(fit_3_long)

# Model estimates
print(fit_3_long, digits = 3)
```

### Model 3 Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-3-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_3_long)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# Pull terms and convert
result_3_long <- tibble(
  Variable = c(
    "Menarche Y (Wave 2) [ADHD = 0]",
    "ADHD Y [Menarche = 0]",
    "Menarche × ADHD"
  ),
  
  Estimate = round(c(
    to_pct(fe$Estimate[fe$term == "menarche_status_p_wave_2Y"]),
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"]),
    to_pct(fe$Estimate[fe$term == "menarche_status_p_wave_2Y:adhd_status_lifeY"])
  ), 2),
  
  `Lower 95% CI` = round(c(
    to_pct(fe$Q2.5[fe$term == "menarche_status_p_wave_2Y"]),
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"]),
    to_pct(fe$Q2.5[fe$term == "menarche_status_p_wave_2Y:adhd_status_lifeY"])
  ), 2),
  
  `Upper 95% CI` = round(c(
    to_pct(fe$Q97.5[fe$term == "menarche_status_p_wave_2Y"]),
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"]),
    to_pct(fe$Q97.5[fe$term == "menarche_status_p_wave_2Y:adhd_status_lifeY"])
  ), 2)
)

# View
result_3_long
```

### Model 3 Long Diagnostics

```{r model-3-diagnostics}
# Prior Summary
prior_summary(fit_3_long)

# Posterior Predictive Checks
plot(fit_3_long, ask = FALSE)

# PPC plots
pp_check(fit_3_long, type = "bars", ndraws = 100)
pp_check(fit_3_long, type = "hist", ndraws = 10)
pp_check(fit_3_long, type = "error_hist")
pp_check(fit_3_long, type = "stat", stat = "mean")
pp_check(fit_3_long, type = "stat", stat = "sd")
pp_check(fit_3_long, type = "scatter_avg")
```

### Model 3B Long 

```{r model-3b-long}
fit_3b_long <- brm(
  formula = formula_3b_long,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  control = list(adapt_delta = 0.95),
  cores = 2,
  seed = 1234
)

saveRDS(fit_3b_long, "fit_3b_long.rds")
```

### Model 3B Long Output

```{r model-3b-long-output}
# Posterior Summary Table
bayes_R2(fit_3b_long)

# Model estimates
print(fit_3b_long, digits = 3)
```

### Model 3B Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-3b-long-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_3b_long)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# SD of original wave 2 breast development
sd_bd <- sd(analytic_df_wide$breast_development_p_wave_2, na.rm = TRUE)

# Build table
result_3b_long <- tibble(
  Variable = c(
    "Breast development (Wave 2) [ADHD = 0]",
    "ADHD Y [Breast development = 0]",
    "Breast development × ADHD"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"]),
    (exp(fe$Estimate[fe$term == "breast_development_p_wave_2_sc:adhd_status_lifeY"] / (2 * sd_bd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"]),
    (exp(fe$Q2.5[fe$term == "breast_development_p_wave_2_sc:adhd_status_lifeY"] / (2 * sd_bd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"]),
    (exp(fe$Q97.5[fe$term == "breast_development_p_wave_2_sc:adhd_status_lifeY"] / (2 * sd_bd)) - 1) * 100
  ), 2)
)

# View
result_3b_long
```

### Model 3B Long Diagnostics

```{r model-3b-diagnostics}
# Prior Summary
prior_summary(fit_3b_long)

# Posterior Predictive Checks
plot(fit_3b_long, ask = FALSE)

# PPC plots
pp_check(fit_3b_long, type = "bars", ndraws = 100)
pp_check(fit_3b_long, type = "hist", ndraws = 10)
pp_check(fit_3b_long, type = "error_hist")
pp_check(fit_3b_long, type = "stat", stat = "mean")
pp_check(fit_3b_long, type = "stat", stat = "sd")
pp_check(fit_3b_long, type = "scatter_avg")
```

### Model 4 Long 

```{r model-4-long}
fit_4_long <- brm(
  formula = formula_4_long,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_4_long, "fit_4_long.rds")
```

### Model 4 Long Output

```{r model-4-long-output}
# Posterior Summary Table
bayes_R2(fit_4_long)

# Model estimates
print(fit_4_long, digits = 3)
```

### Model 4 Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-4-long-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_4_long)) |>
  rownames_to_column("term")

# Convert log scale to %
to_pct <- function(x) (exp(x) - 1) * 100

# SD of original wave 2 ADHD traits
sd_adhd <- sd(analytic_df_wide$adhd_traits_wave_2, na.rm = TRUE)

# Build table
result_4_long <- tibble(
  Variable = c(
    "Menarche (Wave 2) [ADHD traits = mean]",
    "ADHD traits (Wave 2) [Menarche = 0]",
    "Menarche × ADHD traits"
  ),
  
  Estimate = round(c(
    to_pct(fe$Estimate[fe$term == "menarche_status_p_wave_2Y"]),
    (exp(fe$Estimate[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "menarche_status_p_wave_2Y:adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    to_pct(fe$Q2.5[fe$term == "menarche_status_p_wave_2Y"]),
    (exp(fe$Q2.5[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "menarche_status_p_wave_2Y:adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    to_pct(fe$Q97.5[fe$term == "menarche_status_p_wave_2Y"]),
    (exp(fe$Q97.5[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "menarche_status_p_wave_2Y:adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2)
)

# View
result_4_long
```

### Model 4 Long Diagnostics

```{r model-4-diagnostics}
# Prior Summary
prior_summary(fit_4_long)

# Posterior Predictive Checks
plot(fit_4_long, ask = FALSE)

# PPC plots
pp_check(fit_4_long, type = "bars", ndraws = 100)
pp_check(fit_4_long, type = "hist", ndraws = 10)
pp_check(fit_4_long, type = "error_hist")
pp_check(fit_4_long, type = "stat", stat = "mean")
pp_check(fit_4_long, type = "stat", stat = "sd")
pp_check(fit_4_long, type = "scatter_avg")
```

### Model 4B Long 

```{r model-4b-long}
fit_4b_long <- brm(
  formula = formula_4b_long,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_4b_long, "fit_4b_long.rds")
```

### Model 4B Long Output

```{r model-4B-long-output}
# Posterior Summary Table
bayes_R2(fit_4b_long)

# Model estimates
print(fit_4b_long, digits = 3)
```

### Model 4B Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-4b-long-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_4b_long)) |>
  rownames_to_column("term")

# Convert log scale to %
to_pct <- function(x) (exp(x) - 1) * 100

# SDs of original wave 2 variables
sd_adhd <- sd(analytic_df_wide$adhd_traits_wave_2, na.rm = TRUE)
sd_bd   <- sd(analytic_df_wide$breast_development_p_wave_2, na.rm = TRUE)

# Build table
result_4b_long <- tibble(
  Variable = c(
    "Breast development (Wave 2) [ADHD traits = mean]",
    "ADHD traits (Wave 2) [Breast development = 0]",
    "Breast development × ADHD traits"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "breast_development_p_wave_2_sc:adhd_traits_wave_2_sc"] / (4 * sd_bd * sd_adhd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "breast_development_p_wave_2_sc:adhd_traits_wave_2_sc"] / (4 * sd_bd * sd_adhd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "breast_development_p_wave_2_sc:adhd_traits_wave_2_sc"] / (4 * sd_bd * sd_adhd)) - 1) * 100
  ), 2)
)

# View
result_4b_long
```

### Model 4B Long Diagnostics

```{r model-4b-diagnostics}
# Prior Summary
prior_summary(fit_4b_long)

# Posterior Predictive Checks
plot(fit_4b_long, ask = FALSE)

# PPC plots
pp_check(fit_4b_long, type = "bars", ndraws = 100)
pp_check(fit_4b_long, type = "hist", ndraws = 10)
pp_check(fit_4b_long, type = "error_hist")
pp_check(fit_4b_long, type = "stat", stat = "mean")
pp_check(fit_4b_long, type = "stat", stat = "sd")
pp_check(fit_4b_long, type = "scatter_avg")
```

### Model 5 Long 

```{r model-5-long}
fit_5_long <- brm(
  formula = formula_5_long,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  control = list(adapt_delta = 0.95),
  cores = 2,
  seed = 1234
)

saveRDS(fit_5_long, "fit_5_long.rds")
```

### Model 5 Long Output

```{r model-5-long-output}
# Posterior Summary Table
bayes_R2(fit_5_long)

# Model estimates
print(fit_5_long, digits = 3)
```

### Model 5 Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-5-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_5_long)) |>
  rownames_to_column("term")

# Convert log scale to %
to_pct <- function(x) (exp(x) - 1) * 100

# SD of original (unscaled) ADHD traits 
sd_adhd <- sd(analytic_df_wide$adhd_traits_wave_2, na.rm = TRUE)

# Build table
result_5_long <- tibble(
  Variable = c(
    "Menarche (Wave 2) [ADHD traits = mean]",
    "ADHD traits (Wave 2) [Menarche = 0]",
    "Menarche × ADHD traits"
  ),
  
  Estimate = round(c(
    to_pct(fe$Estimate[fe$term == "menarche_status_p_wave_2Y"]),
    (exp(fe$Estimate[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "menarche_status_p_wave_2Y:adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    to_pct(fe$Q2.5[fe$term == "menarche_status_p_wave_2Y"]),
    (exp(fe$Q2.5[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "menarche_status_p_wave_2Y:adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    to_pct(fe$Q97.5[fe$term == "menarche_status_p_wave_2Y"]),
    (exp(fe$Q97.5[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "menarche_status_p_wave_2Y:adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2)
)

# View
result_5_long
```

### Model 5 Long Diagnostics

```{r model-5-diagnostics}
# Prior Summary
prior_summary(fit_5_long)

# Posterior Predictive Checks
plot(fit_5_long, ask = FALSE)

# PPC plots
pp_check(fit_5_long, type = "bars", ndraws = 100)
pp_check(fit_5_long, type = "hist", ndraws = 10)
pp_check(fit_5_long, type = "error_hist")
pp_check(fit_5_long, type = "stat", stat = "mean")
pp_check(fit_5_long, type = "stat", stat = "sd")
pp_check(fit_5_long, type = "scatter_avg")
```

### Model 5B Long 

```{r model-5b-long}
fit_5b_long <- brm(
  formula = formula_5b_long,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  control = list(adapt_delta = 0.95),
  cores = 2,
  seed = 1234
)

saveRDS(fit_5b_long, "fit_5b_long.rds")
```

### Model 5B Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-5b-long-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_5b_long)) |>
  rownames_to_column("term")

# Convert log scale to %
to_pct <- function(x) (exp(x) - 1) * 100

# SDs of original wave 2 variables
sd_adhd <- sd(analytic_df_wide$adhd_traits_wave_2, na.rm = TRUE)
sd_bd   <- sd(analytic_df_wide$breast_development_p_wave_2, na.rm = TRUE)

# Build table
result_5b_long <- tibble(
  Variable = c(
    "Breast development (Wave 2) [ADHD traits = mean]",
    "ADHD traits (Wave 2) [Breast development = mean]",
    "Breast development × ADHD traits"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "breast_development_p_wave_2_sc:adhd_traits_wave_2_sc"] / (4 * sd_bd * sd_adhd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "breast_development_p_wave_2_sc:adhd_traits_wave_2_sc"] / (4 * sd_bd * sd_adhd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "breast_development_p_wave_2_sc:adhd_traits_wave_2_sc"] / (4 * sd_bd * sd_adhd)) - 1) * 100
  ), 2)
)

# View
result_5b_long
```

### Model 5B Long Output

```{r model-5B-long-output}
# Posterior Summary Table
bayes_R2(fit_5b_long)

# Model estimates
print(fit_5b_long, digits = 3)
```

### Model 5B Long Interaction Analysis

```{r model-5b-interaction}
# Get SDs and means of original variables
sd_adhd <- sd(analytic_df_wide$adhd_traits_wave_2, na.rm = TRUE)
sd_bd   <- sd(analytic_df_wide$breast_development_p_wave_2, na.rm = TRUE)
mean_bd <- mean(analytic_df_wide$breast_development_p_wave_2, na.rm = TRUE)

# Convert raw breast development values (1 and 4) to Gelman-scaled units
bd_1_sc <- (1 - mean_bd) / (2 * sd_bd)
bd_4_sc <- (4 - mean_bd) / (2 * sd_bd)

# Get estimated trends at stages 1 and 4
emtrends_5b <- emtrends(
  fit_5b_long,
  specs = ~ breast_development_p_wave_2_sc,
  var = "adhd_traits_wave_2_sc",
  at = list(
    breast_development_p_wave_2_sc = c(bd_1_sc, bd_4_sc),
    age_wave_2_sc = 0,
    ethnrace = "white_nh",
    inr_wave_0_sc = 0,
    externalising_wave_0_sc = 0
  )
)

# Convert to dataframe
df <- as.data.frame(summary(emtrends_5b))

# Convert slopes to % change per 1-unit RAW ADHD
df <- df %>%
  mutate(
    Estimate = (exp(adhd_traits_wave_2_sc.trend / (2 * sd_adhd)) - 1) * 100,
    `Lower 95% CI` = (exp(lower.HPD / (2 * sd_adhd)) - 1) * 100,
    `Upper 95% CI` = (exp(upper.HPD / (2 * sd_adhd)) - 1) * 100,
    breast_dev = c(1, 4)
  )

# Build table (slopes at each stage)
result_slopes <- df %>%
  mutate(
    Variable = paste0(
      "ADHD traits (Wave 2) → Outcome [Breast development = ",
      breast_dev,
      "]"
    )
  ) %>%
  select(
    Variable,
    Estimate,
    `Lower 95% CI`,
    `Upper 95% CI`
  ) %>%
  mutate(across(where(is.numeric), ~ sprintf("%.2f", .x)))

# Formal comparison of slopes (stage 4 vs stage 1)
contrast_df <- as.data.frame(pairs(emtrends_5b))

# Convert contrast to % scale (same transformation)
contrast_df <- contrast_df %>%
  mutate(
    Estimate = (exp(estimate / (2 * sd_adhd)) - 1) * 100,
    `Lower 95% CI` = (exp(lower.HPD / (2 * sd_adhd)) - 1) * 100,
    `Upper 95% CI` = (exp(upper.HPD / (2 * sd_adhd)) - 1) * 100,
    Variable = "Difference in slope (Stage 4 vs Stage 1)"
  ) %>%
  select(
    Variable,
    Estimate,
    `Lower 95% CI`,
    `Upper 95% CI`
  ) %>%
  mutate(across(where(is.numeric), ~ sprintf("%.2f", .x)))

# Combine into one table
result_5b_long_interaction <- bind_rows(result_slopes, contrast_df)

# View final table
result_5b_long_interaction
```

### Plot 5B Long Interaction Plot

```{r plot-model-5b-long-interaction}
sd_adhd  <- sd(analytic_df_wide$adhd_traits_wave_2, na.rm = TRUE)
mean_adhd <- mean(analytic_df_wide$adhd_traits_wave_2, na.rm = TRUE)

cat("ADHD mean:", mean_adhd, "\n")
cat("ADHD SD:", sd_adhd, "\n")
cat("ADHD range:", range(analytic_df_wide$adhd_traits_wave_2, na.rm = TRUE), "\n")

# Generate prediction grid
adhd_grid_raw <- seq(
  min(analytic_df_wide$adhd_traits_wave_2, na.rm = TRUE),
  max(analytic_df_wide$adhd_traits_wave_2, na.rm = TRUE),
  by = 0.1
)
adhd_grid_sc <- (adhd_grid_raw - mean_adhd) / (2 * sd_adhd)

emm_5b <- emmeans(
  fit_5b_long,
  ~ adhd_traits_wave_2_sc * breast_development_p_wave_2_sc,
  at = list(
    adhd_traits_wave_2_sc = adhd_grid_sc,
    breast_development_p_wave_2_sc = c(bd_1_sc, bd_4_sc),
    age_wave_2_sc = 0,
    ethnrace = "white_nh",
    inr_wave_0_sc = 0,
    externalising_wave_0_sc = 0
  ),
  type = "response"
)

plot_5b_line <- as.data.frame(emm_5b) %>%
  mutate(
    adhd_raw = rep(adhd_grid_raw, 2),
    breast_dev = factor(
      rep(c("Score 1", "Score 4"), each = length(adhd_grid_raw))
    )
  )

head(plot_5b_line)

p_5b <- ggplot(plot_5b_line, aes(
  x = adhd_raw,
  y = rate,
  colour = breast_dev,
  fill = breast_dev,
  linetype = breast_dev
)) +
  geom_ribbon(
    aes(ymin = lower.HPD, ymax = upper.HPD),
    alpha = 0.15,
    colour = NA
  ) +
  geom_line(linewidth = 1) +
  scale_colour_manual(
    values = c("Score 1" = "black", "Score 4" = "black"),
    name = "Breast development\n(wave 2)"
  ) +
  scale_fill_manual(
    values = c("Score 1" = "black", "Score 4" = "black"),
    name = "Breast development\n(wave 2)"
  ) +
  scale_linetype_manual(
    values = c("Score 1" = "solid", "Score 4" = "dashed"),
    name = "Breast development\n(wave 2)"
  ) +
  labs(
    x = "ADHD traits (wave 2)",
    y = "Predicted externalising (wave 3)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    legend.position = c(0.25, 0.85),
    legend.background = element_blank(),
    legend.key.width = unit(1.5, "cm")  # try smaller, e.g. unit(1, "cm")
  )

p_5b
```

### Model 5B Long Diagnostics

```{r model-5b-diagnostics}
# Prior Summary
prior_summary(fit_5b_long)

# Posterior Predictive Checks
plot(fit_5b_long, ask = FALSE)

# PPC plots
pp_check(fit_5b_long, type = "bars", ndraws = 100)
pp_check(fit_5b_long, type = "hist", ndraws = 10)
pp_check(fit_5b_long, type = "error_hist")
pp_check(fit_5b_long, type = "stat", stat = "mean")
pp_check(fit_5b_long, type = "stat", stat = "sd")
pp_check(fit_5b_long, type = "scatter_avg")
```




