---
title: "ABCD ADHD menarche longitudinal analyses (no interaction term)"
author: "Emily Friedel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

### Load libraries

```{r libraries}
library(tidyverse)
library(ggplot2)
library(mice)
library(brms)
library(posterior)
library(bayestestR)
library(emmeans)
library(patchwork)
```

### Import saved imputed df

```{r import-data}
analytic_df_wide <- read_csv("analytic_df_wide.csv")
```

### Initial checks

```{r initial-checks}
summary(unique(analytic_df_wide$participant_id))
length(analytic_df_wide)
colnames(analytic_df_wide)
```

## 1. Pre model processing

### Convert to factors: site ID, family ID, ethnicity, and menarche status factors

```{r factors}
analytic_df_wide <- analytic_df_wide %>%
  mutate(
    family_id = as.factor(family_id),
    site_id = as.factor(site_id),
    ethnrace = as.factor(ethnrace),
    adhd_status_life = as.factor(adhd_status_life),
    menarche_status_p_wave_2 = as.factor(menarche_status_p_wave_2)
  )

# set white_nh as reference for enthorace
analytic_df_wide$ethnrace <- relevel(analytic_df_wide$ethnrace, ref = "white_nh")
levels(analytic_df_wide$ethnrace)
```

### Convert to integers: all outcomes

```{r outcomes-as-integers}
analytic_df_wide <- analytic_df_wide %>%
  mutate(
    adhd_traits_wave_3 = as.integer(adhd_traits_wave_3),
    internalising_wave_3 = as.integer(internalising_wave_3),
    externalising_wave_3 = as.integer(externalising_wave_3)
  )
```

### Check head of wide dataframe

```{r check-df}
head(analytic_df_wide)
str(analytic_df_wide)
```

## 2. Set up hypothesis formulas and prior

```{r hypothesis-formulas}

formula_2_long_nointeract <- bf(internalising_wave_3 ~ menarche_status_p_wave_2 + adhd_status_life
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + internalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_2b_long_nointeract <- bf(internalising_wave_3 ~ breast_development_p_wave_2_sc + adhd_status_life
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + internalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_3_long_nointeract <- bf(externalising_wave_3 ~ menarche_status_p_wave_2 + adhd_status_life
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc 
                  + externalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_3b_long_nointeract <- bf(externalising_wave_3 ~ breast_development_p_wave_2_sc + adhd_status_life
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + externalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_4_long_nointeract <- bf(internalising_wave_3 ~ menarche_status_p_wave_2 + adhd_traits_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + internalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_4b_long_nointeract <- bf(internalising_wave_3 ~ breast_development_p_wave_2_sc + adhd_traits_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + internalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_5_long_nointeract <- bf(externalising_wave_3 ~ menarche_status_p_wave_2 + adhd_traits_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + externalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

prior <- prior(normal(0, 1), class = "b")
```

### 2.1 Check default priors

```{r default-prior}
get_prior(
  formula = formula_2_long_nointeract,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_2b_long_nointeract,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_3_long_nointeract,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_3b_long_nointeract,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_4_long_nointeract,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_4b_long_nointeract,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5_long_nointeract,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)
```

## 3. Run long models

### Model 2 Long

```{r model-2-long}
fit_2_long_nointeract <- brm(
  formula = formula_2_long_nointeract,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_2_long_nointeract, "fit_2_long_nointeract.rds")
```

### Model 2 Output

```{r model-2-long-output}
# Posterior Summary Table
bayes_R2(fit_2_long_nointeract)

# Model estimates
print(fit_2_long_nointeract, digits = 3)
```

### Model 2 Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-2-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_2_long_nointeract)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# Pull terms and convert
result_2_long_nointeract <- tibble(
  Variable = c(
    "Menarche Y (Wave 2)",
    "ADHD Y"
  ),
  
  Estimate = round(c(
    to_pct(fe$Estimate[fe$term == "menarche_status_p_wave_2Y"]),
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Lower 95% CI` = round(c(
    to_pct(fe$Q2.5[fe$term == "menarche_status_p_wave_2Y"]),
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Upper 95% CI` = round(c(
    to_pct(fe$Q97.5[fe$term == "menarche_status_p_wave_2Y"]),
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"])
  ), 2)
)

# View
result_2_long_nointeract
```

### Model 2 Long Diagnostics

```{r model-2-diagnostics}
# Prior Summary
prior_summary(fit_2_long_nointeract)

# Posterior Predictive Checks

plot(fit_2_long_nointeract, ask = FALSE)

# PPC plots
pp_check(fit_2_long_nointeract, type = "bars", ndraws = 100)
pp_check(fit_2_long_nointeract, type = "hist", ndraws = 10)
pp_check(fit_2_long_nointeract, type = "error_hist")
pp_check(fit_2_long_nointeract, type = "stat", stat = "mean")
pp_check(fit_2_long_nointeract, type = "stat", stat = "sd")
pp_check(fit_2_long_nointeract, type = "scatter_avg")
```

### Model 2B Long

```{r model-2b-long}
fit_2b_long_nointeract <- brm(
  formula = formula_2b_long_nointeract,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_2b_long_nointeract, "fit_2b_long_nointeract.rds")
```

### Model 2B Long Output

```{r model-2b-long-output}
# Posterior Summary Table
bayes_R2(fit_2b_long_nointeract)

# Model estimates
print(fit_2b_long_nointeract, digits = 3)
```

### Model 2B Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-2b-long-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_2b_long_nointeract)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# SD of original wave 2 breast development variable
sd_bd <- sd(analytic_df_wide$breast_development_p_wave_2, na.rm = TRUE)

# Build table
result_2b_long_nointeract <- tibble(
  Variable = c(
    "Breast development (Wave 2)",
    "ADHD Y"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"])
  ), 2)
)

# View
result_2b_long_nointeract
```

### Model 2B Long Diagnostics

```{r model-2b-diagnostics}
# Prior Summary
prior_summary(fit_2b_long_nointeract)

# Posterior Predictive Checks
plot(fit_2b_long_nointeract, ask = FALSE)

# PPC plots
pp_check(fit_2b_long_nointeract, type = "bars", ndraws = 100)
pp_check(fit_2b_long_nointeract, type = "hist", ndraws = 10)
pp_check(fit_2b_long_nointeract, type = "error_hist")
pp_check(fit_2b_long_nointeract, type = "stat", stat = "mean")
pp_check(fit_2b_long_nointeract, type = "stat", stat = "sd")
pp_check(fit_2b_long_nointeract, type = "scatter_avg")
```

### Model 3 Long 

```{r model-3-long}
fit_3_long_nointeract <- brm(
  formula = formula_3_long_nointeract,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  control = list(adapt_delta = 0.95),
  cores = 2,
  seed = 1234
)

saveRDS(fit_3_long_nointeract, "fit_3_long_nointeract.rds")
```

### Model 3 Long Output

```{r model-3-long-output}
# Posterior Summary Table
bayes_R2(fit_3_long_nointeract)

# Model estimates
print(fit_3_long_nointeract, digits = 3)
```

### Model 3 Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-3-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_3_long_nointeract)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# Pull terms and convert
result_3_long_nointeract <- tibble(
  Variable = c(
    "Menarche Y (Wave 2)",
    "ADHD Y"
  ),
  
  Estimate = round(c(
    to_pct(fe$Estimate[fe$term == "menarche_status_p_wave_2Y"]),
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Lower 95% CI` = round(c(
    to_pct(fe$Q2.5[fe$term == "menarche_status_p_wave_2Y"]),
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Upper 95% CI` = round(c(
    to_pct(fe$Q97.5[fe$term == "menarche_status_p_wave_2Y"]),
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"])
  ), 2)
)

# View
result_3_long_nointeract
```

### Model 3 Long Diagnostics

```{r model-3-diagnostics}
# Prior Summary
prior_summary(fit_3_long_nointeract)

# Posterior Predictive Checks
plot(fit_3_long_nointeract, ask = FALSE)

# PPC plots
pp_check(fit_3_long_nointeract, type = "bars", ndraws = 100)
pp_check(fit_3_long_nointeract, type = "hist", ndraws = 10)
pp_check(fit_3_long_nointeract, type = "error_hist")
pp_check(fit_3_long_nointeract, type = "stat", stat = "mean")
pp_check(fit_3_long_nointeract, type = "stat", stat = "sd")
pp_check(fit_3_long_nointeract, type = "scatter_avg")
```

### Model 3B Long 

```{r model-3b-long}
fit_3b_long_nointeract <- brm(
  formula = formula_3b_long_nointeract,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  control = list(adapt_delta = 0.95),
  cores = 2,
  seed = 1234
)

saveRDS(fit_3b_long_nointeract, "fit_3b_long_nointeract.rds")
```

### Model 3B Long Output

```{r model-3b-long-output}
# Posterior Summary Table
bayes_R2(fit_3b_long_nointeract)

# Model estimates
print(fit_3b_long_nointeract, digits = 3)
```

### Model 3B Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-3b-long-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_3b_long_nointeract)) |>
  rownames_to_column("term")

# Function to convert log estimates to %
to_pct <- function(x) (exp(x) - 1) * 100

# SD of original wave 2 breast development
sd_bd <- sd(analytic_df_wide$breast_development_p_wave_2, na.rm = TRUE)

# Build table
result_3b_long_nointeract <- tibble(
  Variable = c(
    "Breast development (Wave 2)",
    "ADHD Y"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Estimate[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q2.5[fe$term == "adhd_status_lifeY"])
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    to_pct(fe$Q97.5[fe$term == "adhd_status_lifeY"])
  ), 2)
)

# View
result_3b_long_nointeract
```

### Model 3B Long Diagnostics

```{r model-3b-diagnostics}
# Prior Summary
prior_summary(fit_3b_long_nointeract)

# Posterior Predictive Checks
plot(fit_3b_long_nointeract, ask = FALSE)

# PPC plots
pp_check(fit_3b_long_nointeract, type = "bars", ndraws = 100)
pp_check(fit_3b_long_nointeract, type = "hist", ndraws = 10)
pp_check(fit_3b_long_nointeract, type = "error_hist")
pp_check(fit_3b_long_nointeract, type = "stat", stat = "mean")
pp_check(fit_3b_long_nointeract, type = "stat", stat = "sd")
pp_check(fit_3b_long_nointeract, type = "scatter_avg")
```

### Model 4 Long 

```{r model-4-long}
fit_4_long_nointeract <- brm(
  formula = formula_4_long_nointeract,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_4_long_nointeract, "fit_4_long_nointeract.rds")
```

### Model 4 Long Output

```{r model-4-long-output}
# Posterior Summary Table
bayes_R2(fit_4_long_nointeract)

# Model estimates
print(fit_4_long_nointeract, digits = 3)
```

### Model 4 Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-4-long-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_4_long_nointeract)) |>
  rownames_to_column("term")

# Convert log scale to %
to_pct <- function(x) (exp(x) - 1) * 100

# SD of original wave 2 ADHD traits
sd_adhd <- sd(analytic_df_wide$adhd_traits_wave_2, na.rm = TRUE)

# Build table
result_4_long_nointeract <- tibble(
  Variable = c(
    "Menarche (Wave 2)",
    "ADHD traits (Wave 2)"
  ),
  
  Estimate = round(c(
    to_pct(fe$Estimate[fe$term == "menarche_status_p_wave_2Y"]),
    (exp(fe$Estimate[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    to_pct(fe$Q2.5[fe$term == "menarche_status_p_wave_2Y"]),
    (exp(fe$Q2.5[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    to_pct(fe$Q97.5[fe$term == "menarche_status_p_wave_2Y"]),
    (exp(fe$Q97.5[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2)
)

# View
result_4_long_nointeract
```

### Model 4 Long Diagnostics

```{r model-4-diagnostics}
# Prior Summary
prior_summary(fit_4_long_nointeract)

# Posterior Predictive Checks
plot(fit_4_long_nointeract, ask = FALSE)

# PPC plots
pp_check(fit_4_long_nointeract, type = "bars", ndraws = 100)
pp_check(fit_4_long_nointeract, type = "hist", ndraws = 10)
pp_check(fit_4_long_nointeract, type = "error_hist")
pp_check(fit_4_long_nointeract, type = "stat", stat = "mean")
pp_check(fit_4_long_nointeract, type = "stat", stat = "sd")
pp_check(fit_4_long_nointeract, type = "scatter_avg")
```

### Model 4B Long 

```{r model-4b-long}
fit_4b_long_nointeract <- brm(
  formula = formula_4b_long_nointeract,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_4b_long_nointeract, "fit_4b_long_nointeract.rds")
```

### Model 4B Long Output

```{r model-4b-long-output}
# Posterior Summary Table
bayes_R2(fit_4b_long_nointeract)

# Model estimates
print(fit_4b_long_nointeract, digits = 3)
```

### Model 4B Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-4b-long-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_4b_long_nointeract)) |>
  rownames_to_column("term")

# Convert log scale to %
to_pct <- function(x) (exp(x) - 1) * 100

# SDs of original wave 2 variables
sd_adhd <- sd(analytic_df_wide$adhd_traits_wave_2, na.rm = TRUE)
sd_bd   <- sd(analytic_df_wide$breast_development_p_wave_2, na.rm = TRUE)

# Build table
result_4b_long_nointeract <- tibble(
  Variable = c(
    "Breast development (Wave 2)",
    "ADHD traits (Wave 2)"
  ),
  
  Estimate = round(c(
    (exp(fe$Estimate[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Estimate[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    (exp(fe$Q2.5[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Q2.5[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    (exp(fe$Q97.5[fe$term == "breast_development_p_wave_2_sc"] / (2 * sd_bd)) - 1) * 100,
    (exp(fe$Q97.5[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2)
)

# View
result_4b_long_nointeract
```

### Model 4B Long Diagnostics

```{r model-4b-diagnostics}
# Prior Summary
prior_summary(fit_4b_long_nointeract)

# Posterior Predictive Checks
plot(fit_4b_long_nointeract, ask = FALSE)

# PPC plots
pp_check(fit_4b_long_nointeract, type = "bars", ndraws = 100)
pp_check(fit_4b_long_nointeract, type = "hist", ndraws = 10)
pp_check(fit_4b_long_nointeract, type = "error_hist")
pp_check(fit_4b_long_nointeract, type = "stat", stat = "mean")
pp_check(fit_4b_long_nointeract, type = "stat", stat = "sd")
pp_check(fit_4b_long_nointeract, type = "scatter_avg")
```

### Model 5 Long 

```{r model-5-long}
fit_5_long_nointeract <- brm(
  formula = formula_5_long_nointeract,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  control = list(adapt_delta = 0.95),
  cores = 2,
  seed = 1234
)

saveRDS(fit_5_long_nointeract, "fit_5_long_nointeract.rds")
```

### Model 5 Long Output

```{r model-5-long-output}
# Posterior Summary Table
bayes_R2(fit_5_long_nointeract)

# Model estimates
print(fit_5_long_nointeract, digits = 3)
```

### Model 5 Long Effects expressed as % change in expected outcome per 1-unit increase in original predictor scale

```{r model-5-results}
# Extract fixed effects
fe <- as.data.frame(fixef(fit_5_long_nointeract)) |>
  rownames_to_column("term")

# Convert log scale to %
to_pct <- function(x) (exp(x) - 1) * 100

# SD of original (unscaled) ADHD traits 
sd_adhd <- sd(analytic_df_wide$adhd_traits_wave_2, na.rm = TRUE)

# Build table
result_5_long_nointeract <- tibble(
  Variable = c(
    "Menarche (Wave 2)",
    "ADHD traits (Wave 2)"
  ),
  
  Estimate = round(c(
    to_pct(fe$Estimate[fe$term == "menarche_status_p_wave_2Y"]),
    (exp(fe$Estimate[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Lower 95% CI` = round(c(
    to_pct(fe$Q2.5[fe$term == "menarche_status_p_wave_2Y"]),
    (exp(fe$Q2.5[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2),
  
  `Upper 95% CI` = round(c(
    to_pct(fe$Q97.5[fe$term == "menarche_status_p_wave_2Y"]),
    (exp(fe$Q97.5[fe$term == "adhd_traits_wave_2_sc"] / (2 * sd_adhd)) - 1) * 100
  ), 2)
)

# View
result_5_long_nointeract
```

### Model 5 Long Diagnostics

```{r model-5-diagnostics}
# Prior Summary
prior_summary(fit_5_long_nointeract)

# Posterior Predictive Checks
plot(fit_5_long_nointeract, ask = FALSE)

# PPC plots
pp_check(fit_5_long_nointeract, type = "bars", ndraws = 100)
pp_check(fit_5_long_nointeract, type = "hist", ndraws = 10)
pp_check(fit_5_long_nointeract, type = "error_hist")
pp_check(fit_5_long_nointeract, type = "stat", stat = "mean")
pp_check(fit_5_long_nointeract, type = "stat", stat = "sd")
pp_check(fit_5_long_nointeract, type = "scatter_avg")
```
