---
title: "ABCD ADHD menarche analysis: Models 1 - 5 cross-sectional"
format: html
editor: visual
---

## Load libraries

```{r load libraries}
library(tidyverse)
library(knitr)
library(tidyr)
library(readxl)
library(readr)
library(brms)
library(pscl)
library(posterior)
library(rstan)
library(bayesplot)
library(tidybayes)
library(emmeans)
```

## Import imputed dataframe

```{r import imputed dataframe}
imp_df <- read_csv("imputed_dataset.csv", show_col_types = FALSE)
```

### Check imputed dataframe

```{r check head}
head(imp_df)
```

### Convert to factors: subject ID, site ID, family ID, ethnicity, and menarche status

```{r factors}
imp_df <- imp_df %>%
  mutate(across(c(src_subject_id, family_id, site_id, ethnicity, menarche_status_p, adhd_diagnosis), as.factor))
```

### Convert to integers: all outcomes

```{r outcomes as integers}
imp_df <- imp_df %>%
  mutate(across(c(cbcl_scr_syn_attention_r, cbcl_scr_syn_internal_r, cbcl_scr_syn_external_r), as.integer))
```

### Create observation ID variable

```{r obs id}
imp_df$obs_id <- seq_len(nrow(imp_df))
```

### Structure of dataframe

```{r df structure}
str(imp_df)
```

## Model 1

```{r model 1}
model_1 <- brm(
  cbcl_scr_syn_attention_r ~ menarche_status_p
    + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id),
  data = imp_df,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2,
  seed = 1234
)
```

### Save model 1

```{r save model 1}
saveRDS(model_1, file = "model_1.rds")
```

### Model 1 output

```{r model 1 output}
# Model R2
bayes_R2(model_1)

# Model estimates
print(model_1, digits = 4)
```

### Model 1 posterior predictive diagnostics

```{r model 1 ppcs}
# Posterior predictive checks
pp_check(model_1, type = "bars", ndraws = 100)
pp_check(model_1, type = "hist", ndraws = 10)
pp_check(model_1, type = "error_hist")
pp_check(model_1, type = "stat", stat = "mean")
pp_check(model_1, type = "stat", stat = "sd")
pp_check(model_1, type = "scatter_avg")
```

### Model 1 MCMC sampling diagnostics

```{r plot model 1}
plot(model_1)
```

## Model 1b

```{r model 1b}
model_1b <- brm(
  cbcl_scr_syn_attention_r ~ pds_f4_p_c
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id),
  data = imp_df,  
  family = poisson(),
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234 
)
```

### Save Model 1b

```{r save model 1b}
saveRDS(model_1b, file = "model_1b.rds")
```

### Model 1b output

```{r model 1b output}
# Model R2
bayes_R2(model_1b)

# Model estimates
print(model_1b, digits = 4)
```

### Model 1b posterior predictive diagnostics

```{r model 1b ppcs}
# Posterior predictive checks
pp_check(model_1b, type = "bars", ndraws = 100)
pp_check(model_1b, type = "hist", ndraws = 10)
pp_check(model_1b, type = "error_hist")
pp_check(model_1b, type = "stat", stat = "mean")
pp_check(model_1b, type = "stat", stat = "sd")
pp_check(model_1b, type = "scatter_avg")
```

### Model 1b MCMC sampling diagnostics

```{r plot model 1b}
plot(model_1b)
```

## Model 2

```{r model 2}
model_2 <- brm(
  cbcl_scr_syn_internal_r ~  menarche_status_p * adhd_diagnosis
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = imp_df,  
  family = poisson(),
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234 
)
```

### Save Model 2

```{r save model 2}
saveRDS(model_2, file = "model_2.rds")
```

### Model 2 output

```{r model 2 output}
# Model R2
bayes_R2(model_2)

# Model estimates
print(model_2, digits = 4)
```

### Model 2 posterior predictive diagnostics

```{r model 2 ppcs}
# Posterior predictive checks
pp_check(model_2, type = "bars", ndraws = 100)
pp_check(model_2, type = "hist", ndraws = 10)
pp_check(model_2, type = "error_hist")
pp_check(model_2, type = "stat", stat = "mean")
pp_check(model_2, type = "stat", stat = "sd")
pp_check(model_2, type = "scatter_avg")
```

### Model 2 MCMC sampling diagnostics

```{r plot model 2}
plot(model_2)
```

### Model 2 interaction analysis

```{r model 2 interaction}
# Get marginal means for each combination of adhd_diagnosis and menarche_status_p
emm_2 <- emmeans(model_2, ~ adhd_diagnosis * menarche_status_p)

# View estimated means for each group
summary(emm_2)

# Pairwise comparisons of menarche effect within each ADHD status
contrast(emm_2, method = "revpairwise", by = "adhd_diagnosis")
```

## Model 2b

```{r model 2b}
model_2b <- brm(
  cbcl_scr_syn_internal_r ~  pds_f4_p_c * adhd_diagnosis 
    + age_years_c + ethnicity + inr_c + 
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id), 
  data = imp_df,
  family = poisson(),
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Save Model 2b

```{r save model 2b}
saveRDS(model_2b, file = "model_2b.rds")
```

### Model 2b output

```{r model 2 output}
# Model R2
bayes_R2(model_2b)

# Model estimates
print(model_2b, digits = 4)
```

### Model 2b posterior predictive diagnostics

```{r model 2b ppcs}
# Posterior predictive checks
pp_check(model_2b, type = "bars", ndraws = 100)
pp_check(model_2b, type = "hist", ndraws = 10)
pp_check(model_2b, type = "error_hist")
pp_check(model_2b, type = "stat", stat = "mean")
pp_check(model_2b, type = "stat", stat = "sd")
pp_check(model_2b, type = "scatter_avg")
```

### Model 2b MCMC sampling diagnostics

```{r plot model 2b}
plot(model_2b)
```

### Model 2b interaction analysis

```{r model 2b interaction}
# Get estimated marginal trends (simple slopes)
emtrends_2b <- emtrends(model_2b, 
                       pairwise ~ adhd_diagnosis, 
                       var = "pds_f4_p_c")

# View the simple slopes
summary(emtrends_2b)
```

## Model 3

```{r model 3}
model_3 <- brm(
  cbcl_scr_syn_external_r ~ menarche_status_p * adhd_diagnosis 
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = imp_df,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234 
)
```

### Save Model 3

```{r save model 3}
saveRDS(model_3, file = "model_3.rds")
```

### Model 3 output

```{r model 3 output}
# Model R2
bayes_R2(model_3)

# Model estimates
print(model_3, digits = 4)
```

### Model 3 posterior predictive diagnostics

```{r model 3 ppcs}
# Posterior predictive checks
pp_check(model_3, type = "bars", ndraws = 100)
pp_check(model_3, type = "hist", ndraws = 10)
pp_check(model_3, type = "error_hist")
pp_check(model_3, type = "stat", stat = "mean")
pp_check(model_3, type = "stat", stat = "sd")
pp_check(model_3, type = "scatter_avg")
```

### Model 3 MCMC sampling diagnostics

```{r plot model 3}
plot(model_3)
```

### Model 3 interaction analysis

```{r model 3 interaction}
# Get marginal means for each combination of adhd_diagnosis and menarche_status_p
emm_3 <- emmeans(model_3, ~ adhd_diagnosis * menarche_status_p)

# View estimated means for each group
summary(emm_3)

# Pairwise comparisons of menarche effect within each ADHD status
contrast(emm_3, method = "revpairwise", by = "adhd_diagnosis")
```

## Model 3b

```{r model 3b}
model_3b <- brm(
  cbcl_scr_syn_external_r ~  pds_f4_p_c * adhd_diagnosis 
    + age_years_c + ethnicity + inr_c + 
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id), 
  data = imp_df,
  family = poisson(),
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Save Model 3b

```{r save model 3b}
saveRDS(model_3b, file = "model_3b.rds")
```

### Model 3b output

```{r model 3b output}
# Model R2
bayes_R2(model_3b)

# Model estimates
print(model_3b, digits = 4)
```

### Model 3b posterior predictive diagnostics

```{r model 3b ppcs}
# Posterior predictive checks
pp_check(model_3b, type = "bars", ndraws = 100)
pp_check(model_3b, type = "hist", ndraws = 10)
pp_check(model_3b, type = "error_hist")
pp_check(model_3b, type = "stat", stat = "mean")
pp_check(model_3b, type = "stat", stat = "sd")
pp_check(model_3b, type = "scatter_avg")
```

### Model 3b MCMC sampling diagnostics

```{r plot model 3b}
plot(model_3b)
```

### Model 3b interaction analysis

```{r model 3b interaction}
# Get estimated marginal trends (simple slopes)
emtrends_3b <- emtrends(model_3b, 
                       pairwise ~ adhd_diagnosis, 
                       var = "pds_f4_p_c")

# View the simple slopes
summary(emtrends_3b)
```

## Model 4

```{r model 4}
model_4 <- brm(
  cbcl_scr_syn_internal_r ~  menarche_status_p * cbcl_scr_syn_attention_r_c
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = imp_df,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234 
)
```

### Save Model 4

```{r save model 4}
saveRDS(model_4, file = "model_4.rds")
```

### Model 4 output

```{r model 4 output}
# Model R2
bayes_R2(model_4)

# Model estimates
print(model_4, digits = 4)
```

### Model 4 posterior predictive diagnostics

```{r model 4 ppcs}
# Posterior predictive checks
pp_check(model_4, type = "bars", ndraws = 100)
pp_check(model_4, type = "hist", ndraws = 10)
pp_check(model_4, type = "error_hist")
pp_check(model_4, type = "stat", stat = "mean")
pp_check(model_4, type = "stat", stat = "sd")
pp_check(model_4, type = "scatter_avg")
```

### Model 4 MCMC sampling diagnostics

```{r plot model 4}
plot(model_4)
```

### Model 4 interaction analysis

```{r model 4 interaction}
# Get estimated marginal trends (simple slopes)
emtrends_4 <- emtrends(model_4, 
                       pairwise ~ menarche_status_p, 
                       var = "cbcl_scr_syn_attention_r_c")

# View the simple slopes
summary(emtrends_4)
```

## Model 4b

```{r model 4b}
model_4b <- brm(
  cbcl_scr_syn_internal_r ~  pds_f4_p_c * cbcl_scr_syn_attention_r_c
    + age_years_c + ethnicity + inr_c + 
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = imp_df,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Save Model 4b

```{r save model 4b}
saveRDS(model_4b, file = "model_4b.rds")
```

### Model 4b output

```{r model 4b output}
# Model R2
bayes_R2(model_4b)

# Model estimates
print(model_4b, digits = 4)
```

### Model 4b diagnostics

```{r model 4b ppcs}
# Posterior predictive checks
pp_check(model_4b, type = "bars", ndraws = 100)
pp_check(model_4b, type = "hist", ndraws = 10)
pp_check(model_4b, type = "error_hist")
pp_check(model_4b, type = "stat", stat = "mean")
pp_check(model_4b, type = "stat", stat = "sd")
pp_check(model_4b, type = "scatter_avg")
```

### Model 4b MCMC sampling diagnostics

```{r plot model 4b}
plot(model_4b)
```

## Model 5

```{r model 5}
model_5 <- brm(
  cbcl_scr_syn_external_r ~  menarche_status_p * cbcl_scr_syn_attention_r_c
  + age_years_c + ethnicity + inr_c +
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id),
  data = imp_df,  
  family = poisson(),  
  prior = set_prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234 
)
```

### Save Model 5

```{r save model 5}
saveRDS(model_5, file = "model_5.rds")
```

### Model 5 output

```{r model 5 output}
# Model R2
bayes_R2(model_5)

# Model estimates
print(model_5, digits = 4)
```

### Model 5 posterior predictive diagnostics

```{r model 5 ppcs}
# Posterior predictive checks
pp_check(model_5, type = "bars", ndraws = 100)
pp_check(model_5, type = "hist", ndraws = 10)
pp_check(model_5, type = "error_hist")
pp_check(model_5, type = "stat", stat = "mean")
pp_check(model_5, type = "stat", stat = "sd")
pp_check(model_5, type = "scatter_avg")
```

### Model 5 MCMC sampling diagnostics

```{r plot model 5}
plot(model_5)
```

### Model 5 interaction analysis

```{r model 5 interaction}
# Get estimated marginal trends (simple slopes)
emtrends_5 <- emtrends(model_5, 
                       pairwise ~ menarche_status_p, 
                       var = "cbcl_scr_syn_attention_r_c")

# View the simple slopes
summary(emtrends_5)
```

## Model 5b

```{r model 5b}
model_5b <- brm(
  cbcl_scr_syn_external_r ~  pds_f4_p_c * cbcl_scr_syn_attention_r_c 
    + age_years_c + ethnicity + inr_c + 
    (1 | src_subject_id) + (1 | family_id) + (1 | site_id) + (1 | obs_id), 
  data = imp_df,
  family = poisson(),
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Save Model 5b

```{r save model 5b}
saveRDS(model_5b, file = "model_5b.rds")
```

### Model 5b output

```{r model 5b output}
# Model R2
bayes_R2(model_5b)

# Model estimates
print(model_5b, digits = 4)
```

### Model 5b posterior predictive diagnostics

```{r model 5b ppcs}
# Posterior predictive checks
pp_check(model_5b, type = "bars", ndraws = 100)
pp_check(model_5b, type = "hist", ndraws = 10)
pp_check(model_5b, type = "error_hist")
pp_check(model_5b, type = "stat", stat = "mean")
pp_check(model_5b, type = "stat", stat = "sd")
pp_check(model_5b, type = "scatter_avg")
```

### Model 5b MCMC sampling diagnostics

```{r plot model 5b}
plot(model_5b)
```
