---
title: "ABCD ADHD menarche analysis: Models 1 - 5 longitudinal"
format: html
editor: visual
---

## Load libraries

```{r load libraries, results='hide'}
library(tidyverse)
library(knitr)
library(tidyr)
library(readxl)
library(readr)
library(brms)
library(pscl)
library(posterior)
library(rstan)
library(bayesplot)
library(tidybayes)
library(emmeans)
```

## Import imputed dataset

```{r read csv}
# Read csv
imp_df <- read_csv("imputed_dataset.csv", show_col_types = FALSE)
```

### Check imputed dataframe

```{r check head}
head(imp_df)
```

### Pivot to wide format based on wave

```{r pivot wide}
# Pivot to wide format for longitudinal analyses
imp_df_wide <- imp_df %>%
  pivot_wider(
    id_cols = src_subject_id,
    names_from = wave,
    names_glue = "{.value}_{wave}",
    values_from = -c(src_subject_id, wave)
  )
```

### Convert to factors: site ID, family ID, ethnicity, and menarche status factors

```{r factors}
imp_df_wide <- imp_df_wide %>%
  mutate(
    family_id = as.factor(family_id_wave_0),
    site_id = as.factor(site_id_wave_0),
    ethnicity = as.factor(ethnicity_wave_0),
    adhd_diagnosis = as.factor(adhd_diagnosis_wave_0),
    menarche_status_p_wave_2 <- as.factor(imp_df_wide$menarche_status_p_wave_2)
  )
```

### Convert to integers: all outcomes

```{r outcomes as integers}
imp_df_wide <- imp_df_wide %>%
  mutate(
    cbcl_scr_syn_attention_r_wave_3 = as.integer(cbcl_scr_syn_attention_r_wave_3),
    cbcl_scr_syn_internal_r_wave_3 = as.integer(cbcl_scr_syn_internal_r_wave_3),
    cbcl_scr_syn_external_r_wave_3 = as.integer(cbcl_scr_syn_external_r_wave_3)
  )
```

### Check head of wide dataframe

```{r check head wide df}
head(imp_df_wide)
```

### Check wide imputed df structure

```{r check structure wide df}
str(imp_df_wide)
```

## Model 1 longitudinal

```{r model 1 long}
model_1_long <- brm(
  cbcl_scr_syn_attention_r_wave_3 ~ 
    menarche_status_p_wave_2 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_attention_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = set_prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Save model 1 long

```{r save model 1 long}
saveRDS(model_1_long, file = "model_1_long.rds")
```

### Model 1 long output

```{r model 1 long output}
# Model R2
bayes_R2(model_1_long)

# Model estimates
print(model_1_long, digits = 4)
```

### Model 1 long posterior predictive diagnostics

```{r model 1 long ppcs}
# Posterior predictive checks
pp_check(model_1_long, type = "bars", ndraws = 100)
pp_check(model_1_long, type = "hist", ndraws = 10)
pp_check(model_1_long, type = "error_hist")
pp_check(model_1_long, type = "stat", stat = "mean")
pp_check(model_1_long, type = "stat", stat = "sd")
pp_check(model_1_long, type = "scatter_avg")
```

### Model 1 long MCMC sampling diagnostics

```{r plot model long 1}
plot(model_1_long)
```

## Model 1b longitudinal

```{r model 1b long}
model_1b_long <- brm(
  cbcl_scr_syn_attention_r_wave_3 ~ 
    pds_f4_p_c_wave_2 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_attention_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Save model 1b long

```{r save model 1b long}
saveRDS(model_1b_long, file = "model_1b_long.rds")
```

### Model 1b long output

```{r model 1b long output}
# Model R2
bayes_R2(model_1b_long)

# Model estimates
print(model_1b_long, digits = 4)
```

### Model 1b long posterior predictive diagnostics

```{r model 1b long ppcs}
# Posterior predictive checks
pp_check(model_1b_long, type = "bars", ndraws = 100)
pp_check(model_1b_long, type = "hist", ndraws = 10)
pp_check(model_1b_long, type = "error_hist")
pp_check(model_1b_long, type = "stat", stat = "mean")
pp_check(model_1b_long, type = "stat", stat = "sd")
pp_check(model_1b_long, type = "scatter_avg")
```

### Model 1b long MCMC sampling diagnostics

```{r plot model long 1b}
plot(model_1b_long)
```

## Model 2 longitudinal

```{r model 2 long}
model_2_long <- brm(
  cbcl_scr_syn_internal_r_wave_3 ~ 
    menarche_status_p_wave_2 * adhd_diagnosis_wave_0 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 +
    cbcl_scr_syn_internal_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Save model 2 long

```{r save model 2 long}
saveRDS(model_2_long, file = "model_2_long.rds")
```

### Model 2 long output

```{r model 2 long output}
# Model R2
bayes_R2(model_2_long)

# Model estimates
print(model_2_long, digits = 4)
```

### Model 2 long posterior predictive diagnostics

```{r model 2 long ppcs}
# Posterior predictive checks
pp_check(model_2_long, type = "bars", ndraws = 100)
pp_check(model_2_long, type = "hist", ndraws = 10)
pp_check(model_2_long, type = "error_hist")
pp_check(model_2_long, type = "stat", stat = "mean")
pp_check(model_2_long, type = "stat", stat = "sd")
pp_check(model_2_long, type = "scatter_avg")
```

### Model 2 long MCMC sampling diagnostics

```{r plot model long 2}
plot(model_2_long)
```

## Model 2b longitudinal

```{r model 2b long}
model_2b_long <- brm(
  cbcl_scr_syn_internal_r_wave_3 ~ 
    pds_f4_p_c_wave_2 * adhd_diagnosis +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_internal_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
  )
```

### Save model 2b long

```{r save model 2b long}
saveRDS(model_2b_long, file = "model_2b_long.rds")
```

### Model 2b long output

```{r model 2b long output}
# Model R2
bayes_R2(model_2b_long)

# Model estimates
print(model_2b_long, digits = 4)
```

### Model 2b long posterior predictive diagnostics

```{r model 2b long ppcs}
# Posterior predictive checks
pp_check(model_2b_long, type = "bars", ndraws = 100)
pp_check(model_2b_long, type = "hist", ndraws = 10)
pp_check(model_2b_long, type = "error_hist")
pp_check(model_2b_long, type = "stat", stat = "mean")
pp_check(model_2b_long, type = "stat", stat = "sd")
pp_check(model_2b_long, type = "scatter_avg")
```

### Model 2b long MCMC sampling diagnostics

```{r plot model long 2b}
plot(model_2b_long)
```

## Model 3 longitudinal

```{r model 3 long}
model_3_long <- brm(
  cbcl_scr_syn_external_r_wave_3 ~ 
    menarche_status_p_wave_2 * adhd_diagnosis +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_external_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Save model 3 long

```{r save model 3 long}
saveRDS(model_3_long, file = "model_3_long.rds")
```

### Model 3 long output

```{r model 3 long output}
# Model R2
bayes_R2(model_3_long)

# Model estimates
print(model_3_long, digits = 4)
```

### Model 3 long posterior predictive diagnostics

```{r model 3 long ppcs}
# Posterior predictive checks
pp_check(model_3_long, type = "bars", ndraws = 100)
pp_check(model_3_long, type = "hist", ndraws = 10)
pp_check(model_3_long, type = "error_hist")
pp_check(model_3_long, type = "stat", stat = "mean")
pp_check(model_3_long, type = "stat", stat = "sd")
pp_check(model_3_long, type = "scatter_avg")
```

### Model 3 long MCMC sampling diagnostics

```{r plot model long 3}
plot(model_3_long)
```

## Model 3b longitudinal

```{r model 3b long}
model_3b_long <- brm(
  cbcl_scr_syn_external_r_wave_3 ~ 
    pds_f4_p_c_wave_2 * adhd_diagnosis +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_external_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Save model 3b long

```{r save model 3b long}
saveRDS(model_3b_long, file = "model_3b_long.rds")
```

### Model 3b long output

```{r model 3b long output}
# Model R2
bayes_R2(model_3b_long)

# Model estimates
print(model_3b_long, digits = 4)
```

### Model 3b long posterior predictive diagnostics

```{r model 3b long ppcs}
# Posterior predictive checks
pp_check(model_3b_long, type = "bars", ndraws = 100)
pp_check(model_3b_long, type = "hist", ndraws = 10)
pp_check(model_3b_long, type = "error_hist")
pp_check(model_3b_long, type = "stat", stat = "mean")
pp_check(model_3b_long, type = "stat", stat = "sd")
pp_check(model_3b_long, type = "scatter_avg")
```

### Model 3b long MCMC sampling diagnostics

```{r plot model long 3b}
plot(model_3b_long)
```

## Model 4 longitudinal

```{r model 4 long}
model_4_long <- brm(
  cbcl_scr_syn_internal_r_wave_3 ~ 
    menarche_status_p_wave_2 * cbcl_scr_syn_attention_r_c_wave_2 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_internal_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Save model 4 long

```{r save model 4 long}
saveRDS(model_4_long, file = "model_4_long.rds")
```

### Model 4 long output

```{r model 4 long output}
# Model R2
bayes_R2(model_4_long)

# Model estimates
print(model_4_long, digits = 4)
```

### Model 4 long posterior predictive diagnostics

```{r model 4 long ppcs}
# Posterior predictive checks
pp_check(model_4_long, type = "bars", ndraws = 100)
pp_check(model_4_long, type = "hist", ndraws = 10)
pp_check(model_4_long, type = "error_hist")
pp_check(model_4_long, type = "stat", stat = "mean")
pp_check(model_4_long, type = "stat", stat = "sd")
pp_check(model_4_long, type = "scatter_avg")
```

### Model 4 long MCMC sampling diagnostics

```{r plot model long 4}
plot(model_4_long)
```

## Model 4b longitudinal

```{r model 4 long}
model_4b_long <- brm(
  cbcl_scr_syn_internal_r_wave_3 ~ 
    pds_f4_p_c_wave_2 * cbcl_scr_syn_attention_r_c_wave_2 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_internal_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  cores = 2, 
  seed = 1234
)
```

### Save model 4b long

```{r save model 4b long}
saveRDS(model_4b_long, file = "model_4b_long.rds")
```

### Model 4b long output

```{r model 4b long output}
# Model R2
bayes_R2(model_4b_long)

# Model estimates
print(model_4b_long, digits = 4)
```

### Model 4b long posterior predictive diagnostics

```{r model 4b long ppcs}
# Posterior predictive checks
pp_check(model_4b_long, type = "bars", ndraws = 100)
pp_check(model_4b_long, type = "hist", ndraws = 10)
pp_check(model_4b_long, type = "error_hist")
pp_check(model_4b_long, type = "stat", stat = "mean")
pp_check(model_4b_long, type = "stat", stat = "sd")
pp_check(model_4b_long, type = "scatter_avg")
```

### Model 4b long MCMC sampling diagnostics

```{r plot model long 4b}
plot(model_4b_long)
```

## Model 5 longitudinal

```{r model 5 long}
model_5_long <- brm(
  cbcl_scr_syn_external_r_wave_3 ~ 
    menarche_status_p_wave_2 * cbcl_scr_syn_attention_r_c_wave_2 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_external_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  control = list(adapt_delta = 0.85),
  cores = 2, 
  seed = 1234
)
```

### Save model 5 long

```{r save model 5 long}
saveRDS(model_5_long, file = "model_5_long.rds")
```

### Model 5 long output

```{r model 5 long output}
# Model R2
bayes_R2(model_5_long)

# Model estimates
print(model_5_long, digits = 4)
```

### Model 5 long posterior predictive diagnostics

```{r model 5 long ppcs}
# Posterior predictive checks
pp_check(model_5_long, type = "bars", ndraws = 100)
pp_check(model_5_long, type = "hist", ndraws = 10)
pp_check(model_5_long, type = "error_hist")
pp_check(model_5_long, type = "stat", stat = "mean")
pp_check(model_5_long, type = "stat", stat = "sd")
pp_check(model_5_long, type = "scatter_avg")
```

### Model 5 long MCMC sampling diagnostics

```{r plot model long 5}
plot(model_5_long)
```

## Model 5b longitudinal

```{r model 5b long}
model_5b_long <- brm(
  cbcl_scr_syn_external_r_wave_3 ~ 
    pds_f4_p_c_wave_2 * cbcl_scr_syn_attention_r_c_wave_2 +
    age_years_c_wave_2 + ethnicity + inr_c_wave_0 + 
    cbcl_scr_syn_external_r_c_wave_0 +
    (1 | family_id) + (1 | site_id),
  data = imp_df_wide,  
  family = poisson(),  
  prior = prior("normal(0, 1)", class = "b"),
  chains = 2, iter = 8000, warmup = 4000, thin = 1,
  control = list(adapt_delta = 0.85),
  cores = 2, 
  seed = 1234
)
```

### Save model 5b long

```{r save model 5b long}
saveRDS(model_5b_long, file = "model_5b_long.rds")
```

### Model 5b long output

```{r model 5b long output}
# Model R2
bayes_R2(model_5b_long)

# Model estimates
print(model_5b_long, digits = 4)
```

### Model 5b long posterior predictive diagnostics

```{r model 5b long ppcs}
# Posterior predictive checks
pp_check(model_5b_long, type = "bars", ndraws = 100)
pp_check(model_5b_long, type = "hist", ndraws = 10)
pp_check(model_5b_long, type = "error_hist")
pp_check(model_5b_long, type = "stat", stat = "mean")
pp_check(model_5b_long, type = "stat", stat = "sd")
pp_check(model_5b_long, type = "scatter_avg")
```

### Model 5b long MCMC sampling diagnostics

```{r plot model long 5b}
plot(model_5b_long)
```

### Model 5b interaction analysis

```{r model 5b interaction}
# Get estimated marginal trends (simple slopes)
emtrends_5b_clean <- emtrends(
  model_5b_long,
  var = "cbcl_scr_syn_attention_r_c_wave_2",
  at = list(
    pds_f4_p_c_wave_2 = c(-1, 0, 1),
    age_years_c_wave_2 = 0,
    ethnicity = 1,
    inr_c_wave_0 = 0,
    cbcl_scr_syn_external_r_c_wave_0 = 0
  )
)

# View simple slopes
summary(emtrends_5b_clean, infer = c(TRUE, TRUE))
```
