---
title: "ABCD ADHD menarche raw dataset preparation (step 1)"
author: "Emily Friedel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up

### Load libraries

```{r load-libraries}
library(tidyverse)
library(dplyr)
```

## 1. Make raw dataframe with all variables relevent to analyses

### 1.1. Read all relevant ABCD files

```{r read-abcd-files}
# Site ID
ab_g_dyn <- read_csv("ab_g_dyn.csv", show_col_types = FALSE)

# Sex, Race/Ethnicity, Family ID
ab_g_stc <- read_csv("ab_g_stc.csv", show_col_types = FALSE)

# Demographic
ab_p_demo <- read_csv("ab_p_demo.csv", show_col_types = FALSE)

# CBCL
mh_p_cbcl <- read_csv("mh_p_cbcl.csv", show_col_types = FALSE)

# Puberty
ph_p_pds <- read_csv("ph_p_pds.csv", show_col_types = FALSE)
ph_y_pds <- read_csv("ph_y_pds.csv", show_col_types = FALSE)

# ADHD
mh_p_ksads__adhd <- read_csv("mh_p_ksads__adhd.csv", show_col_types = FALSE)

# Co-occurring mental health conditions
mh_p_ksads__dep <- read_csv("mh_p_ksads__dep.csv", show_col_types = FALSE)
mh_p_ksads__gad <- read_csv("mh_p_ksads__gad.csv", show_col_types = FALSE)
mh_p_ksads__cond <- read_csv("mh_p_ksads__cond.csv", show_col_types = FALSE)
mh_p_ksads__odd <- read_csv("mh_p_ksads__odd.csv", show_col_types = FALSE)

# Exclusion conditions
mh_p_ksads__psych <- read_csv("mh_p_ksads__psych.csv", show_col_types = FALSE)
mh_p_ksads__bpd <- read_csv("mh_p_ksads__bpd.csv", show_col_types = FALSE)
ab_p_screen <- read_csv("ab_p_screen.csv", show_col_types = FALSE)
ph_p_otbi <- read_csv("ph_p_otbi.csv", show_col_types = FALSE)
mh_p_ksads__tic <- read_csv("mh_p_ksads__tic.csv", show_col_types = FALSE)
nc_y_nihtb <- read_csv("nc_y_nihtb.csv", show_col_types = FALSE)

# Anthropometric
ph_y_anthr <- read_csv("ph_y_anthr.csv", show_col_types = FALSE)

# Medications
ph_p_meds <- read_csv("ph_p_meds.csv", show_col_types = FALSE)

# Area Deprivation Index
le_l_adi <- read_csv("le_l_adi.csv", show_col_types = FALSE)
```

### 1.2 Check for parsing problems

```{r check-parsing-problems}
# Check for parsing problems
datasets <- list(
  ab_g_dyn = ab_g_dyn,
  ab_g_stc = ab_g_stc,
  ab_p_demo = ab_p_demo,
  mh_p_cbcl = mh_p_cbcl,
  nc_y_nihtb = nc_y_nihtb,
  ph_p_pds = ph_p_pds,
  mh_p_ksads__adhd = mh_p_ksads__adhd,
  mh_p_ksads__psych = mh_p_ksads__psych,
  mh_p_ksads__bpd = mh_p_ksads__bpd,
  mh_p_ksads__cond = mh_p_ksads__cond,
  mh_p_ksads__odd = mh_p_ksads__odd,
  mh_p_ksads__dep = mh_p_ksads__dep,
  mh_p_ksads__gad = mh_p_ksads__gad,
  ab_p_screen = ab_p_screen,
  ph_p_otbi = ph_p_otbi,
  ph_y_anthr = ph_y_anthr,
  ph_p_meds = ph_p_meds,
  le_l_adi = le_l_adi
)

# Run problems() on each and print only if issues exist
for (name in names(datasets)) {
  dat <- datasets[[name]]
  prob <- problems(dat)
  
  if (nrow(prob) > 0) {
    cat("\nParsing issues in", name, ":\n")
    
    # Add column names to the problems dataframe
    prob$col_name <- names(dat)[prob$col]
    
    # Print full problem details including column names
    print(as.data.frame(prob))
    
    # Print unique columns affected
    cat("\nColumns with parsing problems in", name, ":\n")
    print(unique(prob$col_name))
    
  } else {
    cat("\nNo parsing issues in", name, "\n")
  }
}
```

### 1.3 Meds dataframe parsing issues 

Meds df has issues that affect variables needed in analysis. All other parsing issues do not affect columsn that we require for analysis.

```{r meds-parsing-issues}
# Get all parsing problems from the original tibble
problems(ph_p_meds)
```

### 1.4 Convert meds columns to characters. 

Other variables with parsing issues are not needed for the analysis.

```{r meds-columns-to-character}
ph_p_meds <- ph_p_meds %>%
  mutate(across(matches("^ph_p_meds__rx__(id|label)_"), as.character))
```

### 1.5 Check parsing problems fixed

```{r meds-parsing-issues-check}
# Get all meds parsing problems from the original tibble
meds_problems <- problems(ph_p_meds)

meds_problems

if (nrow(meds_problems) == 0) {
  message("No parsing problems.")
} else {
  message("Parsing problems remain.")
}
```

### 1.6 Select columns from ABCD files

```{r select-abcd-cols}
# Site ID
dynamic_select <- ab_g_dyn %>%
  dplyr::select(
    participant_id, session_id,
    ab_g_dyn__visit_age,
    ab_g_dyn__design_site
  )

# Sex, Race/Ethnicity, Family ID
static_select <- ab_g_stc %>%
  dplyr::select(
    participant_id,
    ab_g_stc__cohort_sex,
    ab_g_stc__design_id__fam,
    ab_g_stc__cohort_ethnrace__leg
  )

# Demographic
demo_select <- ab_p_demo %>%
  dplyr::select(
    participant_id, session_id,
    ab_p_demo_age,
    ab_p_demo__roster_001,
    ab_p_demo__income__hhold_001,
    ab_p_demo__income__hhold_001__v01
  )

# CBCL
cbcl_select = mh_p_cbcl %>%
  dplyr::select(
    participant_id, session_id,
    mh_p_cbcl_age,
    mh_p_cbcl__synd__attn_sum,
    mh_p_cbcl__synd__ext_sum,
    mh_p_cbcl__synd__int_sum
  )

# Puberty
ph_p_pds_select <- ph_p_pds %>%
  dplyr::select(
    participant_id, session_id,
    ph_p_pds_age,
    ph_p_pds_001,
    ph_p_pds_002,
    ph_p_pds_003,
    ph_p_pds__f_001,
    ph_p_pds__f_002,
    ph_p_pds__f_002__01
  )

ph_y_pds_select <- ph_y_pds %>%
  dplyr::select(
    participant_id, session_id,
    ph_y_pds_age,
    ph_y_pds_001,
    ph_y_pds_002,
    ph_y_pds_003,
    ph_y_pds__f_001,
    ph_y_pds__f_002,
    ph_y_pds__f_002__01
  )

# ADHD
adhd_select <- mh_p_ksads__adhd %>%
  dplyr::select(
    participant_id, session_id,
    mh_p_ksads__adhd__past_dx,
    mh_p_ksads__adhd__pres_dx,
    mh_p_ksads__adhd__fullrem_dx,
    mh_p_ksads__adhd__partrem_dx,
    mh_p_ksads__adhd__unspec_dx,
    mh_p_ksads__adhd__oth_dx
  )

# Co-occurring mental health
dep_select <- mh_p_ksads__dep %>%
  dplyr::select(
    participant_id, session_id,
    mh_p_ksads__dep__mdd__past_dx,
    mh_p_ksads__dep__mdd__pres_dx,
    mh_p_ksads__dep__mdd__partrem_dx,
    mh_p_ksads__dep__pdd__past_dx,
    mh_p_ksads__dep__pdd__pres_dx,
    mh_p_ksads__dep__pdd__partrem_dx
  )

gad_select <- mh_p_ksads__gad %>%
  dplyr::select(
    participant_id, session_id,
    mh_p_ksads__gad__past_dx,
    mh_p_ksads__gad__pres_dx,
    mh_p_ksads__gad__oth__mindur__past_dx,
    mh_p_ksads__gad__oth__mindur__pres_dx
  )

cond_select <- mh_p_ksads__cond %>%
  dplyr::select(
    participant_id, session_id,
    mh_p_ksads__cond__onset__adoles__past_dx,
    mh_p_ksads__cond__onset__adoles__pres_dx,
    mh_p_ksads__cond__onset__chld__past_dx,
    mh_p_ksads__cond__onset__chld__pres_dx
  )

odd_select <- mh_p_ksads__odd %>%
  dplyr::select(
    participant_id, session_id,
    mh_p_ksads__odd__past_dx,
    mh_p_ksads__odd__pres_dx
  )

# Exclusion conditions
schiz_select <- mh_p_ksads__psych %>%
  dplyr::select(
    participant_id, session_id,
    mh_p_ksads__psych__schiz__pres_dx,
    mh_p_ksads__psych__schizaffct__pres_dx,
    mh_p_ksads__psych__schizform__pres_dx,
    mh_p_ksads__psych__schiz__oth__past_dx,
    mh_p_ksads__psych__schiz__oth__pres_dx
  )

bpd_select <- mh_p_ksads__bpd %>%
  dplyr::select(
    participant_id, session_id,
    mh_p_ksads__bpd__bpd1__rcnt__manic__pres_dx,
    mh_p_ksads__bpd__bpd1__rcnt__manicmix__pres_dx,
    mh_p_ksads__bpd__bpd2__curdep__pres_dx,
    mh_p_ksads__bpd__bpd2__curdep__partrem_dx,
    mh_p_ksads__bpd__bpd2__curhypo__pres_dx,
    mh_p_ksads__bpd__bpd2__rcnt__hypo__pres_dx
  )

screen_select <- ab_p_screen %>%
  dplyr::select(
    ab_p_screen_age,
    participant_id, session_id,
    ab_p_screen__med_009
  )

otbi_select <- ph_p_otbi %>%
  dplyr::select(
    participant_id, session_id,
    ph_p_otbi_tbiworst
  )

tic_select <- mh_p_ksads__tic %>%
  dplyr::select(
    participant_id, session_id,
    mh_p_ksads__tic__ts__pres_dx,
    mh_p_ksads__tic__ts__nosymp__rem_dx
  )

nc_y_nihtb_select <- nc_y_nihtb %>%
  dplyr::select(
    participant_id, session_id,
    nc_y_nihtb__comp__tot__agecor_score
  )

# Anthropometric
anthro_select <- ph_y_anthr %>%
  dplyr::select(
    participant_id, session_id,
    ph_y_anthr_age,
    ph_y_anthr__height_mean,
    ph_y_anthr__weight_mean
  )

# Medications
meds_select <- ph_p_meds %>%
  dplyr::select(
    participant_id, session_id,
    ph_p_meds_age,
    matches("^ph_p_meds__rx__(label)_")
  )

# Area Deprivation Index
adi_select <- le_l_adi %>%
  dplyr::select(
    participant_id,
    le_l_adi__addr1_wsum
  )

# Join columns based on participant ID and wave
adhd_menarche_raw <- dynamic_select %>%
  full_join(static_select, by = "participant_id") %>%   # no session_id
  full_join(demo_select, by = c("participant_id", "session_id")) %>%
  full_join(cbcl_select, by = c("participant_id", "session_id")) %>%
  full_join(nc_y_nihtb_select, by = c("participant_id", "session_id")) %>%
  full_join(ph_p_pds_select, by = c("participant_id", "session_id")) %>%
  full_join(ph_y_pds_select, by = c("participant_id", "session_id")) %>%
  full_join(adhd_select, by = c("participant_id", "session_id")) %>%
  full_join(schiz_select, by = c("participant_id", "session_id")) %>%
  full_join(bpd_select, by = c("participant_id", "session_id")) %>%
  full_join(screen_select, by = c("participant_id", "session_id")) %>%
  full_join(otbi_select, by = c("participant_id", "session_id")) %>%
  full_join(tic_select, by = c("participant_id", "session_id")) %>%
  full_join(gad_select, by = c("participant_id", "session_id")) %>%
  full_join(dep_select, by = c("participant_id", "session_id")) %>%
  full_join(cond_select, by = c("participant_id", "session_id")) %>%
  full_join(odd_select, by = c("participant_id", "session_id")) %>%
  full_join(anthro_select, by = c("participant_id", "session_id")) %>%
  full_join(meds_select, by = c("participant_id", "session_id")) %>%
  full_join(adi_select, by = "participant_id")
```

## 2. Refine raw dataframe

### 2.1 Remove waves 5 and 6

```{r remove-waves-5-6}
adhd_menarche_raw <- adhd_menarche_raw %>%
  filter(!session_id %in% c("ses-05A", "ses-06A"))
```

### 2.2 Check for duplicates

```{r check-duplicates}
sum(duplicated(adhd_menarche_raw))

# Identify participants with any duplicated session_id
duplicates <- adhd_menarche_raw %>%
  group_by(participant_id) %>%
  filter(duplicated(session_id) | duplicated(session_id, fromLast = TRUE))

# View the duplicated rows
print(duplicates)
```

### 2.3 Check raw dataframe

```{r check-raw-df}
head(adhd_menarche_raw)
str(adhd_menarche_raw)
```

## 3. Sample size

```{r sample-size}
summary(unique(adhd_menarche_raw$participant_id))

female_count <- adhd_menarche_raw %>%
  filter(ab_g_stc__cohort_sex == 2) %>%
  summarise(n = n_distinct(participant_id))

male_count <- adhd_menarche_raw %>%
  filter(ab_g_stc__cohort_sex == 1) %>%
  summarise(n = n_distinct(participant_id))

female_count
male_count
```

## 4. Save raw dataframe

```{r export-raw-df}
write_csv(adhd_menarche_raw, "adhd_menarche_raw.csv")
```
