---
title: "ABCD ADHD menarche processing menarche data (step 2)"
author: "Emily Friedel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up

### Load libraries

```{r load-libraries}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

### Import raw df, check, rename

```{r import}
adhd_menarche_raw <- read.csv("adhd_menarche_raw.csv")

head(adhd_menarche_raw)
colnames(adhd_menarche_raw)

adhd_menarche_clean_menstat <- adhd_menarche_raw
```

### Check for duplicates

```{r check-duplicates}
sum(duplicated(adhd_menarche_clean_menstat))

# Identify participants with any duplicated eventname
duplicates <- adhd_menarche_clean_menstat %>%
  group_by(participant_id) %>%
  filter(duplicated(session_id) | duplicated(session_id, fromLast = TRUE))

# View the duplicated rows
print(duplicates)
```

## 1. Prepare df for menarche processing

### 1.1 Recode waves

```{r recode-waves}
adhd_menarche_clean_menstat <- adhd_menarche_clean_menstat %>%
  filter(session_id %in% c("ses-00A",
                           "ses-01A",
                           "ses-02A",
                           "ses-03A",
                           "ses-04A")) %>%
  mutate(wave = recode_factor(session_id,
                              "ses-00A" = "wave_0",
                              "ses-01A" = "wave_1",
                              "ses-02A" = "wave_2",
                              "ses-03A" = "wave_3",
                              "ses-04A" = "wave_4"),
         wave = factor(wave, levels = c("wave_0", "wave_1", "wave_2", "wave_3", "wave_4")))

unique(adhd_menarche_clean_menstat$wave)
```

### 1.2 Select female participants

```{r select-females}
adhd_menarche_clean_menstat <- adhd_menarche_clean_menstat %>% filter(ab_g_stc__cohort_sex == 2)
unique(adhd_menarche_clean_menstat$ab_g_stc__cohort_sex)

summary(unique(adhd_menarche_clean_menstat$participant_id))
```

### 1.3 Recode menarche status

```{r recode-menarche-status}
# Count 777 and 999 by wave
adhd_menarche_clean_menstat %>%
  group_by(wave) %>%
  summarise(
    y_777 = sum(ph_y_pds__f_002 == 777, na.rm = TRUE),
    y_999 = sum(ph_y_pds__f_002 == 999, na.rm = TRUE),
    p_777 = sum(ph_p_pds__f_002 == 777, na.rm = TRUE),
    p_999 = sum(ph_p_pds__f_002 == 999, na.rm = TRUE),
    .groups = "drop"
  )

# Create menarche status (self report) variable column 0 = N, 1 = Y, other values = NA
adhd_menarche_clean_menstat <- adhd_menarche_clean_menstat %>%
  mutate(menarche_status_y = dplyr::recode(as.character(ph_y_pds__f_002), 
                                    `0` = "N", `1` = "Y",
                                    .default = NA_character_))

# Create menarche status (parent report) variable column 0 = N, 1 = Y, other values = NA
adhd_menarche_clean_menstat <- adhd_menarche_clean_menstat %>%
  mutate(menarche_status_p = dplyr::recode(as.character(ph_p_pds__f_002), 
                                    `0` = "N", `1` = "Y",
                                    .default = NA_character_))

# Note that menarche status needs to remain as a character for cleaning code to work properly
```

### 1.4 Check female dataframe

```{r check-df}
str(adhd_menarche_clean_menstat)
```

### 1.5 Missing menarche status (parent) by wave

```{r missing-menarche-status}
# Count missing menarche status data by wave
missing_menarche_by_wave <- adhd_menarche_clean_menstat %>%
  group_by(wave) %>%
  summarise(missing_menarche = sum(is.na(menarche_status_p)),
            total = n(),
            missing_percent = (missing_menarche / total) * 100)

# Print the results
print(missing_menarche_by_wave)

# Create a bar graph for menarche status by wave
ggplot(adhd_menarche_clean_menstat, aes(x = wave, fill = menarche_status_p)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Menarche Status by Wave",
    x = "Wave",
    y = "Count",
    fill = "Menarche Status"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("Y" = "lightblue", "N" = "salmon"))
```

## 2. Begin menarche status cleaning

### 2.1 Identify participants with correctable menarche status

This includes those with impossible changes in menarche status (going from post-menarche back to pre-menarche) and NAs that can be reasonably inferred.

```{r impossible-menarche-change}
# Identify participants with "Y" at one wave followed by "N" at any later wave
inconsistent_menarche_status_p <- adhd_menarche_clean_menstat %>%
  filter(wave != "wave_4") %>%
  arrange(participant_id, wave) %>%
  filter(!is.na(menarche_status_p)) %>%
  group_by(participant_id) %>%
  summarise(
    has_Y_before_N = any(menarche_status_p == "Y" & lead(menarche_status_p) == "N"),
    .groups = "drop"
  ) %>%
  filter(has_Y_before_N) %>%
  dplyr::select(participant_id) %>%
  distinct()

# Print the result
print(inconsistent_menarche_status_p)

# Create vector with participant IDs
inconsistent_menarche_participants <- inconsistent_menarche_status_p$participant_id
```

### 2.2 Visual inspection of inconsistent menarche status

```{r visual-inspection-inconsistencies}
# Create dataframe of of participants with inconsistent menarche status including subject ID, wave, age, parent & youth menarche status reports, and parent and youth age at menarche report  
inconsistent_menarche_status_p_df <- adhd_menarche_clean_menstat %>%
  filter(participant_id %in% inconsistent_menarche_status_p$participant_id) %>%
  dplyr::select(participant_id, wave, ab_p_demo_age,
                menarche_status_p,menarche_status_y, 
                ph_p_pds__f_002__01, ph_y_pds__f_002__01) 

# Export as csv to inspect
write_csv(inconsistent_menarche_status_p_df, "inconsistent_menarche_status_p_df.csv")
```

### 2.3 Check for patterns

Three patterns identified for inconsistent menarche status group:

1)  Participants with N at baseline and YN elsewhere across waves. In all these cases, the N in YN appears to be incorrect.
2)  Participants with Y at baseline followed by N at another wave. Comparison of youth reports and parent reports of status and age at menarche suggests these vary in whether baseline, wave 1, 2 or 3 is most likely incorrect

### 2.4 Identify those with NA at baseline and N at wave 1

```{r baseline-na-wave-1-n}
# Identify participants where baseline = NA and wave 1 = N
baseline_na_wave1_n <- adhd_menarche_clean_menstat %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  reframe(
    baseline_status = menarche_status_p[wave == "wave_0"], 
    wave1_status = menarche_status_p[wave == "wave_1"],
    .groups = "drop"
  ) %>%
  filter(is.na(baseline_status) & wave1_status == "N")

# View results
print(baseline_na_wave1_n)

# Create vector with participant ids
baseline_na_participants <- baseline_na_wave1_n$participant_id
```

### 2.5 Identify N NA N and Y NA Y (sandwich NA status)

```{r sandwich-na}
# Collapse the sequences for each participant into a single string
pattern_data <- adhd_menarche_clean_menstat %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  summarise(pattern = paste(menarche_status_p, collapse = "_"), .groups = 'drop')

# Identify participants who have the exact "N_NA_N" or "Y_NA_Y" patterns
sandwich_NA <- pattern_data %>%
  filter(grepl("(?<=^|_)(N_NA_N|Y_NA_Y)(?=$|_)", pattern, perl = TRUE)) %>%
  dplyr::select(participant_id, pattern)

# Print the participants with the identified pattern
print(sandwich_NA)

# Create vector with only participant ids
sandwich_NA_participants <- sandwich_NA$participant_id
```

### 2.6 Check unique patterns of sandwich NAs identified

```{r sandwich-na-patterns}
# Get unique sandwich patterns
unique_sandwich_patterns <- sandwich_NA %>%
  distinct(pattern)

# Print unique sandwich patterns
print(unique_sandwich_patterns)
```

### 2.7 Put all participants who can have menarche status corrected into a vector

```{r list-participants-to-correct}
# Combine all participant IDs into one vector
participants_to_correct <- c(
  sandwich_NA_participants, 
  inconsistent_menarche_participants, 
  baseline_na_participants
  )

# Check for duplicates
sum(duplicated(participants_to_correct))
```

### 2.7 Remove duplicates from list

```{r remove-duplicates-participants-to-correct}
participants_to_correct <- unique(participants_to_correct)

str(participants_to_correct)
```

### 2.8 Extract rows for participants with menarche status data to correct from main df

```{r extract-participants-to-correct}
# Extract data for participants needing corrections
participants_to_correct_data <- adhd_menarche_clean_menstat %>%
  filter(participant_id %in% participants_to_correct)

# View extracted data
str(participants_to_correct_data)

# Check no. of participants to correct
length(unique(participants_to_correct_data$participant_id))
```

## 3. Correction type allocation

### 3.1 Create correction type variable

```{r create-correction-type}
# Create correction type variable so that each participant can be assigned this to avoid correcting more than once
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(correction_type = NA_character_)
```

### 3.2 Correction 1: Participants with N at baseline, YN elsewhere

```{r correction-1}
# Identify participants with "N" at baseline
participants_with_N_baseline <- participants_to_correct_data %>%
  filter(wave != "wave_4") %>%
  filter(wave == "wave_0" & menarche_status_p == "N") %>%
  pull(participant_id)

# Check for a "Y" followed by an "N" in later waves
correction_1_ids <- participants_to_correct_data %>%
  filter(participant_id %in% participants_with_N_baseline) %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  summarise(
    has_Y_before_N = any(menarche_status_p == "Y" & lead(menarche_status_p) == "N"),
    .groups = "drop"
  ) %>%
  filter(has_Y_before_N) %>%
  pull(participant_id)

# Assign "correction_1" only to those participants
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(correction_type = ifelse(participant_id %in% correction_1_ids, "correction_1", correction_type))

# Count unique participants who were assigned correction_1
num_correction_1 <- participants_to_correct_data %>%
  filter(correction_type == "correction_1") %>%
  summarise(unique_subjects = n_distinct(participant_id)) %>%
  pull(unique_subjects)

# Print the count
print(num_correction_1)

View(
  participants_to_correct_data %>%
    filter(wave != "wave_4") %>%
    filter(correction_type == "correction_1") %>%
    mutate(wave_num = as.numeric(gsub("wave_", "", wave))) %>%
    dplyr::select(participant_id, wave_num, menarche_status_p, correction_type) %>%
    pivot_wider(
      names_from = wave_num,
      values_from = menarche_status_p,
      names_sort = TRUE
    ) %>%
    dplyr::select(participant_id, correction_type, order(colnames(.)))
)
```

### 3.3 Correction 2: Y at baseline, N at other waves

```{r correction-2}
# Identify participants with "Y" at baseline (wave_0)
participants_with_Y_baseline <- participants_to_correct_data %>%
  filter(wave == "wave_0" & menarche_status_p == "Y") %>%
  pull(participant_id)

# Check if they have "N" in any later wave
correction_2_ids <- participants_to_correct_data %>%
  filter(participant_id %in% participants_with_Y_baseline) %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  summarise(has_N_later = any(wave != "wave_0" & menarche_status_p == "N"), .groups = "drop") %>%
  filter(has_N_later) %>%
  pull(participant_id)

# Assign "correction_2" to these participants
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(correction_type = ifelse(is.na(correction_type) & participant_id %in% correction_2_ids, "correction_2", correction_type))

# Count unique participants who were assigned correction_2
num_correction_2 <- participants_to_correct_data %>%
  filter(correction_type == "correction_2") %>%
  summarise(unique_subjects = n_distinct(participant_id)) %>%
  pull(unique_subjects)

# Print the count
print(num_correction_2)

View(participants_to_correct_data %>%
    filter(correction_type == "correction_2") %>%
    dplyr::select(participant_id, wave, menarche_status_p, correction_type) %>%
    pivot_wider(
      names_from = wave,
      values_from = menarche_status_p
    )
)
```

### 3.3 Visual check of correction 2 participants

```{r view-correction-2}
correction_2_view <- participants_to_correct_data %>%
  filter(participant_id %in% correction_2_ids) %>%
  dplyr::select(participant_id, wave, menarche_status_p, menarche_status_y,
        ph_p_pds__f_002__01, ph_y_pds__f_002__01, ab_p_demo_age, 
        correction_type)

write_csv(correction_2_view, "correction_2_view.csv")
```

### 3.4 Correction 3: NA at baseline and N at wave 1

```{r correction-3}
# Identify participants with NA at baseline (wave_0)
participants_with_NA_baseline <- participants_to_correct_data %>%
  filter(wave == "wave_0" & is.na(menarche_status_p)) %>%
  pull(participant_id)

# Check if they have "N" at wave 1
correction_3_ids <- participants_to_correct_data %>%
  filter(participant_id %in% participants_with_NA_baseline) %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  summarise(has_N_at_wave_1 = any(wave == "wave_1" & menarche_status_p == "N"), .groups = "drop") %>%
  filter(has_N_at_wave_1) %>%
  pull(participant_id)

# Assign "correction_3" to these participants
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(correction_type = ifelse(is.na(correction_type) & participant_id %in% correction_3_ids, "correction_3", correction_type))

# Count unique participants who were assigned correction_4
num_correction_3 <- participants_to_correct_data %>%
  filter(correction_type == "correction_3") %>%
  summarise(unique_subjects = n_distinct(participant_id)) %>%
  pull(unique_subjects)

# Print the count
print(num_correction_3)

View(participants_to_correct_data %>%
    filter(correction_type == "correction_3") %>%
    dplyr::select(participant_id, wave, menarche_status_p, correction_type) %>%
    pivot_wider(
      names_from = wave,
      values_from = menarche_status_p
    )
)
```

### 3.5 Correction 4: N NA N pattern

```{r correction-4}
# Collapse the menarche_status_p for each participant across waves
pattern_data <- participants_to_correct_data %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  summarise(pattern = paste(menarche_status_p, collapse = "_"), .groups = 'drop')

# Identify participants with the pattern "N_NA_N"
correction_4_ids <- pattern_data %>%
  filter(grepl("(?<=^|_)N_NA_N(?=$|_)", pattern, perl = TRUE)) %>%
  pull(participant_id)

# Assign "correction_4" only to these participants (if no correction already exists)
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(correction_type = ifelse(is.na(correction_type) & participant_id %in% correction_4_ids, "correction_4", correction_type))

# Count unique participants who were assigned correction_4
num_correction_4 <- participants_to_correct_data %>%
  filter(correction_type == "correction_4") %>%
  summarise(unique_subjects = n_distinct(participant_id)) %>%
  pull(unique_subjects)

# Print the count
print(num_correction_4)
```

### 3.6 Correction 5: Y NA Y

```{r correction-5}
# Collapse the menarche_status_p for each participant across waves
pattern_data <- participants_to_correct_data %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  summarise(pattern = paste(menarche_status_p, collapse = "_"), .groups = 'drop')

# Identify participants with the pattern "Y_NA_Y"
correction_5_ids <- pattern_data %>%
  filter(grepl("(?<=^|_)Y_NA_Y(?=$|_)", pattern, perl = TRUE)) %>%
  pull(participant_id)

# Assign "correction_5" only to these participants (if no correction already exists)
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(correction_type = ifelse(is.na(correction_type) & participant_id %in% correction_5_ids, "correction_5", correction_type))

# Count unique participants who were assigned correction_5
num_correction_5 <- participants_to_correct_data %>%
  filter(correction_type == "correction_5") %>%
  summarise(unique_subjects = n_distinct(participant_id)) %>%
  pull(unique_subjects)

# Print the count
print(num_correction_5)
```

### 3.7 Check number of participants with each correction type

```{r correction-type-totals}
participants_to_correct_data %>%
  group_by(correction_type) %>%
  summarise(unique_subjects = n_distinct(participant_id))
```

## 4. Apply corrections

### 4.1 Apply Correction 1

```{r apply-correction-1}
# Where there is Y --> N, change N to Y
participants_to_correct_data <- participants_to_correct_data %>%
  filter(correction_type == "correction_1") %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  mutate(
    menarche_status_p = as.character(menarche_status_p),
    menarche_status_p = {
      status <- menarche_status_p
      seen_y <- FALSE
      
      for (i in seq_along(status)) {
        if (!is.na(status[i]) && status[i] == "Y") {
          seen_y <- TRUE
        }
        if (seen_y && !is.na(status[i]) && status[i] == "N") {
          status[i] <- "Y"
        }
      }
      status
    }
  ) %>%
  ungroup()
```

### 4.2 Check correction 1 worked

```{r check-correction-1}
# Check for a "Y" followed by an "N" in later waves, only for correction_1 participants
correction_1_check <- participants_to_correct_data %>%
  filter(correction_type == "correction_1") %>%
  arrange(participant_id, wave) %>% 
  group_by(participant_id) %>%
  summarise(
    has_Y_before_N = any(menarche_status_p == "Y" & lead(menarche_status_p) == "N"),
    .groups = "drop"
  ) %>%
  filter(has_Y_before_N) %>%
  pull(participant_id)

# Count how many participants still have the "Y N" pattern after correction
print(length(correction_1_check))
```

### 4.3 Apply Correction 2

```{r apply-correction-2}
# Apply individual corrections to menarche_status_p based on participant ID and wave
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(menarche_status_p = case_when(
    participant_id %in% c("sub-BU31791C", 
                          "sub-DETP7ENE", 
                          "sub-GE3NT7ZK") & wave == "wave_0" ~ "N",
    participant_id %in% c("sub-0HL5D3DF", 
                          "sub-L48MK0DM") & wave == "wave_1" ~ "Y",
    participant_id %in% c("sub-1CYTVBWT", 
                          "sub-Z1WNHT63") & wave == "wave_2" ~ "Y",
    participant_id == "sub-B8G9KFMC" & wave %in% c("wave_0", "wave_1") ~ "Y",
    participant_id %in% c("sub-ECM7TBEL", 
                          "sub-LG5EUWL6", 
                          "sub-WMM5T2PL") & wave == "wave_3" ~ "Y",
    
    TRUE ~ menarche_status_p
  ))

# Print the modified rows to check the changes
View(participants_to_correct_data %>%
  filter(participant_id %in% c(
    "sub-0HL5D3DF", "sub-L48MK0DM",
    "sub-1CYTVBWT", "sub-Z1WNHT63",
    "sub-B8G9KFMC",
    "sub-BU31791C", "sub-DETP7ENE", "sub-GE3NT7ZK",
    "sub-ECM7TBEL", "sub-LG5EUWL6", "sub-WMM5T2PL",
    "sub-7VG73XLG", "sub-VVDFELL0"
  )) %>%
    dplyr::select(participant_id, wave, menarche_status_p) %>%
    pivot_wider(
      names_from = wave,
      values_from = menarche_status_p
    )
)
```

### 4.4 Check Correction 2 worked

```{r check-correction-2}
# Identify participants with "Y" at baseline (wave_0)
participants_with_Y_baseline_check <- participants_to_correct_data %>%
  filter(wave == "wave_0" & menarche_status_p == "Y") %>%
  pull(participant_id)

# Check if they have "N" in any later wave
correction_2_check <- participants_to_correct_data %>%
  filter(participant_id %in% participants_with_Y_baseline_check) %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  summarise(has_N_later = any(wave != "wave_0" & menarche_status_p == "N"), .groups = "drop") %>%
  filter(has_N_later) %>%
  pull(participant_id)

# Count unique participants with correction_2
print(length(correction_2_check))
```

### 4.5 Apply Correction 3

```{r apply-correction-3}
# Correct baseline when baseline is NA and wave_1 is N, only for correction_3 participants
participants_to_correct_data <- participants_to_correct_data %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  mutate(
    menarche_status_p = ifelse(
      correction_type == "correction_3" & 
      wave == "wave_0" & is.na(menarche_status_p) &
      any(wave == "wave_1" & menarche_status_p == "N"),
      "N",
      menarche_status_p
    )
  ) %>%
  ungroup()
```

### 4.6 Check Correction 3 worked

```{r check-correction-3}
# Identify participants who still have NA at baseline when wave_1 is N
correction_3_check <- participants_to_correct_data %>%
  filter(correction_type == "correction_3") %>% 
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  summarise(
    still_na_at_baseline = any(wave == "wave_0" & is.na(menarche_status_p) &  
                               any(wave == "wave_1" & menarche_status_p == "N")),
    .groups = "drop"
  ) %>%
  filter(still_na_at_baseline) %>%
  pull(participant_id)

# Count the number of participants who still need correction
print(length(correction_3_check))
```

### 4.7 Apply Correction 4

```{r apply-correction-4}
# Apply correction for "N_NA_N" pattern across waves for participants needing correction 4
participants_to_correct_data <- participants_to_correct_data %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  mutate(
    menarche_status_p = case_when(
      correction_type == "correction_4" &
      is.na(menarche_status_p) &
      lag(menarche_status_p, default = "X") == "N" &
      lead(menarche_status_p, default = "X") == "N" ~ "N",
      TRUE ~ menarche_status_p
    )
  ) %>%
  ungroup()
```

### 4.8 Check Correction 4 worked

```{r check-correction-4}
# Collapse the sequences for each participant into a single string
correction_4_check <- participants_to_correct_data %>%
  filter(correction_type == "correction_4") %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  summarise(pattern = paste(menarche_status_p, collapse = "_"), .groups = 'drop')

# Identify participants who have the exact "N_NA_N"
correction_4_check <- correction_4_check %>%
  filter(grepl("(?<=^|_)N_NA_N(?=$|_)", pattern, perl = TRUE)) %>%
  pull(participant_id)

# Print the number of participants who still have "N_NA_N"
print(length(correction_4_check))
```

### 4.9 Apply Correction 5

```{r apply-correction-5}
# Apply correction for "Y_NA_Y" pattern across waves for participants needing correction 5
participants_to_correct_data <- participants_to_correct_data %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  mutate(
    menarche_status_p = case_when(
      correction_type == "correction_5" &
      is.na(menarche_status_p) &
      lag(menarche_status_p, default = "X") == "Y" &
      lead(menarche_status_p, default = "X") == "Y" ~ "Y",
      TRUE ~ menarche_status_p
    )
  ) %>%
  ungroup()
```

### 4.10 Check Correction 5 worked

```{r check-correction-5}
# Collapse the sequences for each participant into a single string
correction_5_check <- participants_to_correct_data %>%
  filter(correction_type == "correction_5") %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  summarise(pattern = paste(menarche_status_p, collapse = "_"), .groups = 'drop')

# Identify participants who have the exact subsequence "N_NA_N"
correction_5_check <- correction_5_check %>%
  filter(grepl("(?<=^|_)Y_NA_Y(?=$|_)", pattern, perl = TRUE)) %>%
  pull(participant_id)

# Print the number of participants who still have "N_NA_N"
print(length(correction_5_check))
```

## 5. Check all original corrections fixed

### 5.1 Check inconsistencies in menarche status

```{r check-remaining-menarche-inconsistencies}
# Check participants with inconsistent menarche status
inconsistent_menarche_status_p_check <- participants_to_correct_data %>%
  arrange(participant_id, wave) %>%
  filter(!is.na(menarche_status_p)) %>%
  group_by(participant_id) %>%
  mutate(
    prev_status = lag(menarche_status_p),
    inconsistent = ifelse(
      menarche_status_p == "N" & prev_status == "Y",
      TRUE,
      FALSE
    )
  ) %>%
  ungroup() %>% 
  filter(inconsistent == TRUE) %>%
  dplyr::select(participant_id) %>%
  distinct()

# Display the result
print(inconsistent_menarche_status_p_check)
```

### 5.2 Check those with NA at baseline and N at wave 1

```{r baseline-na-wave-1-n-check}
# Check participants where baseline = NA and wave 1 = N
baseline_na_wave1_n_check <- participants_to_correct_data %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  reframe(
    baseline_status = menarche_status_p[wave == "wave_0"], 
    wave1_status = menarche_status_p[wave == "wave_1"]
  ) %>%
  filter(is.na(baseline_status) & wave1_status == "N")

# View results
print(baseline_na_wave1_n_check)
```

### 5.3 Check N NA N and Y NA Y (sandwich NA status)

```{r check-sandwich-na}
# Collapse the sequences for each participant into a single string
pattern_data_check <- participants_to_correct_data %>%
  arrange(participant_id, wave) %>% 
  group_by(participant_id) %>%
  summarise(pattern = paste(menarche_status_p, collapse = "_"), .groups = 'drop')

# Identify participants who have the exact "N_NA_N" or "Y_NA_Y" pattern
sandwich_NA_check <- pattern_data_check %>%
  filter(grepl("(?<=^|_)(N_NA_N|Y_NA_Y)(?=$|_)", pattern, perl = TRUE)) %>%
  dplyr::select(participant_id, pattern)

# Print the participants with the identified pattern
print(sandwich_NA_check)
```

## 6. Return corrected participant to main df

```{r return-corrected-participants}
adhd_menarche_clean_menstat <- adhd_menarche_clean_menstat %>%
  left_join(
    participants_to_correct_data %>%
      dplyr::select(participant_id, wave, menarche_status_p),
    by = c("participant_id", "wave"),
    suffix = c(".original", ".corrected")
  ) %>%
  mutate(
    menarche_status_p = coalesce(menarche_status_p.corrected, as.character(menarche_status_p.original)),
  ) %>%
  dplyr::select(-menarche_status_p.original, -menarche_status_p.corrected)

# Check the updated dataframe
head(adhd_menarche_clean_menstat)
```

### 6.1 Check menarche status variable

```{r check-menarche-status}
head(adhd_menarche_clean_menstat$menarche_status_p)
```

### 6.2 Check inconsistencies in menarche status in main df

```{r check-menarche-inconsistencies-main-df}
# Check participants with inconsistent menarche status
inconsistent_menarche_status_p_check_2 <- adhd_menarche_clean_menstat %>%
  arrange(participant_id, wave) %>%
  filter(!is.na(menarche_status_p)) %>%
  group_by(participant_id) %>%
  mutate(
    prev_status = lag(menarche_status_p),
    inconsistent = ifelse(
      menarche_status_p == "N" & prev_status == "Y",
      TRUE,
      FALSE
    )
  ) %>%
  ungroup() %>% 
  filter(inconsistent == TRUE) %>%
  dplyr::select(participant_id) %>%
  distinct()

# Display the result
print(inconsistent_menarche_status_p_check_2)
```

### 6.4 Check those with NA at baseline and N at wave 1

```{r baseline-na-wave-1-n-main-df}
# Check participants where baseline = NA and wave 1 = N
baseline_na_wave1_n_check_2 <- adhd_menarche_clean_menstat %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  reframe(
    baseline_status = menarche_status_p[wave == "wave_0"], 
    wave1_status = menarche_status_p[wave == "wave_1"],
    .groups = "drop"
  ) %>%
  filter(is.na(baseline_status) & wave1_status == "N")

# View results
print(baseline_na_wave1_n_check)
```

### 6.5 Check N NA N and Y NA Y (sandwich NA status)

```{r sandwich-na-main-df}
# Collapse the sequences for each participant into a single string
pattern_data_check_2 <- adhd_menarche_clean_menstat %>%
  arrange(participant_id, wave) %>%
  group_by(participant_id) %>%
  summarise(pattern = paste(menarche_status_p, collapse = "_"), .groups = 'drop')

# Identify participants who have the exact subsequence "N_NA_N" or "Y_NA_Y"
sandwich_NA_check_2 <- pattern_data_check_2 %>%
  filter(grepl("(?<=^|_)(N_NA_N|Y_NA_Y)(?=$|_)", pattern, perl = TRUE)) %>%
  dplyr::select(participant_id, pattern)

# Print the participants with the identified pattern
print(sandwich_NA_check_2)
```

### 6.6 Duplicates checks of corrected data

```{r check-duplicates}
sum(duplicated(adhd_menarche_clean_menstat))

# Identify participants with duplicated eventname
duplicates_2 <- adhd_menarche_clean_menstat %>%
  group_by(participant_id) %>%
  filter(duplicated(wave) | duplicated(wave, fromLast = TRUE))

# View the duplicated rows
print(duplicates_2)
```

### 6.7 Check reduction in missing menarche status

```{r final-check-menarche-missing}
# Count missing menarche data by wave
missing_menarche_by_wave_2 <- adhd_menarche_clean_menstat %>%
  group_by(wave) %>%
  summarise(missing_menarche = sum(is.na(menarche_status_p)),
            total = n(),
            missing_percent = (missing_menarche / total) * 100)

# Print the results
print(missing_menarche_by_wave_2)
```

### 6.8 Remove wave 4

```{r remove-wave-4}
adhd_menarche_clean_menstat <- adhd_menarche_clean_menstat %>%
  filter(wave != "wave_4")
```

### 6.9 Final check of cleaned df

```{r final-check-clean-df}
head(adhd_menarche_clean_menstat)
summary(unique(adhd_menarche_clean_menstat$participant_id))
```

## 7. Export copy of csv with clean menarche status (parent) data

```{r export}
write_csv(adhd_menarche_clean_menstat, "adhd_menarche_clean_menstat.csv")
```


