---
title: "ABCD ADHD menarche processing all other variables (step 3)"
author: "Emily Friedel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up

### Load libraries

```{r load-libraries}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(scales) 
library(gt)
```

### Read saved raw dataframe and check

```{r read-cleaned-df}
adhd_menarche_clean <- read.csv("adhd_menarche_clean_menstat.csv")
summary(unique(adhd_menarche_clean$participant_id))
head(adhd_menarche_clean)
colnames(adhd_menarche_clean)
```

## 1. Sample integrity checks

### 1.1 Convert 777 and 999 to NA

```{r convert-777-999}
convert_vars <- c(
  "ph_p_pds_001",
  "ph_p_pds_002",
  "ph_p_pds_003",
  "ph_p_pds__f_001",
  "ph_p_pds__f_002",
  "ab_p_demo__income__hhold_001",
  "ab_p_demo__income__hhold_001__v01",
  "ab_p_demo_age",
  "ph_y_anthr__weight_mean",
  "ph_y_anthr__height_mean",
  "ab_p_demo__roster_001",
  "le_l_adi__addr1_wsum",
  "ab_g_stc__cohort_ethnrace__leg",
  "ab_g_dyn__design_site",
  "ab_g_stc__design_id__fam"
)

# Convert 777 and 999 to NA
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(
    across(all_of(convert_vars), ~ na_if(., 777))
  ) %>%
  mutate(
    across(all_of(convert_vars), ~ na_if(., 999))
  )
```

### 1.2 Check sample size by wave

```{r sample-by-wave}
adhd_menarche_clean %>%
  group_by(wave) %>%
  mutate(N_total = n_distinct(participant_id)) %>%
  ungroup() %>%
  group_by(wave) %>%
  summarize(n_participants = n_distinct(participant_id),
            .groups = "drop")
```

### 1.3 Create household income variable

```{r household-income}
income_vars <- c(
  "ab_p_demo__income__hhold_001",
  "ab_p_demo__income__hhold_001__v01"
)

# Ensure numeric, then convert 777 and 999 to NA
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(
    across(all_of(income_vars), ~ na_if(., 777))
  ) %>%
  mutate(
    across(all_of(income_vars), ~ na_if(., 999))
  )

# Merge dfs with baseline income and all other waves
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(
    household_income = coalesce(ab_p_demo__income__hhold_001, 
                              ab_p_demo__income__hhold_001__v01))
```

### 1.4 Remove implausible roster values

```{r remove-implausible-roster}
unique_roster <- adhd_menarche_clean %>%
  pull(ab_p_demo__roster_001) %>%
  unique()

print(unique_roster)

adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(
    ab_p_demo__roster_001 = if_else(
      ab_p_demo__roster_001 > 50,
      NA_real_,
      ab_p_demo__roster_001
    )
  )

unique_roster_check <- adhd_menarche_clean %>%
  pull(ab_p_demo__roster_001) %>%
  unique()

print(unique_roster_check)
```

### 1.5 Check missingness/implausible values in key grouping & demographic variables

```{r key-variables-missing}
key_demo_vars <- c(
  "ab_p_demo_age",
  "ph_y_anthr__weight_mean",
  "ph_y_anthr__height_mean",
  "household_income",
  "ab_p_demo__roster_001",
  "le_l_adi__addr1_wsum",
  "ab_g_stc__cohort_ethnrace__leg",
  "ab_g_dyn__design_site",
  "ab_g_stc__design_id__fam"
)

adhd_menarche_clean %>%
  group_by(wave) %>%
  summarize(
    n_total = n(),
    across(
      all_of(key_demo_vars),
      list(
        n_missing = ~sum(is.na(.)),
        pct_missing = ~mean(is.na(.)) * 100
      ),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  )

sapply(
  adhd_menarche_clean[, c(
    "ab_p_demo_age",
    "ph_y_anthr__weight_mean",
    "ph_y_anthr__height_mean"
  )],
  function(x) range(x, na.rm = TRUE)
)

adhd_menarche_clean <- adhd_menarche_clean %>%
    mutate(bmi = (ph_y_anthr__weight_mean / (ph_y_anthr__height_mean^2)) * 703)

# Check where BMI <10
adhd_menarche_clean %>%
  filter(wave == "wave_0", bmi < 10) %>%
  dplyr::select(participant_id,
         ab_p_demo_age,
         ph_y_anthr__weight_mean,
         ph_y_anthr__height_mean,
         bmi)
```

### 1.6 Remove implausible weight/height

```{r remove-implausible-weight-height}
# Implausible BMI appears to be due to weight being recorded as kg instead of lbs, so weight converted to NA for these
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(
    ph_y_anthr__weight_mean = if_else(
      bmi < 10,
      NA_real_,
      ph_y_anthr__weight_mean
    )
  )
```

### 1.7 Coalesce age to address missingness

```{r coalesce-age}
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(
    age = coalesce(ab_p_demo_age, ab_g_dyn__visit_age, ph_p_pds_age, 
                   ph_y_pds_age, mh_p_cbcl_age, ph_y_anthr_age,
                   ph_p_meds_age)
  )

adhd_menarche_clean %>%
  group_by(wave) %>%
  summarize(
    n_missing   = sum(is.na(age)),
    pct_missing = mean(is.na(age)) * 100,
    .groups = "drop"
  )
```

### 1.8 Demographic summaries

```{r demographic-summaries}
continuous_demo_vars <- c(
  "ab_p_demo_age",
  "ph_y_anthr__weight_mean",
  "ph_y_anthr__height_mean",
  "le_l_adi__addr1_wsum",
  "household_income",
  "ab_p_demo__roster_001"
)

# Summary statistics by wave
summary_stats_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    across(
      all_of(continuous_demo_vars),
      list(
        mean   = ~ mean(.x, na.rm = TRUE),
        sd     = ~ sd(.x, na.rm = TRUE),
        min    = ~ min(.x, na.rm = TRUE),
        max    = ~ max(.x, na.rm = TRUE)
      ),
      .names = "{col}_{fn}"
    ),
    .groups = "drop"
  ) %>%
  pivot_longer(-wave,
               names_to = c("variable", "statistic"),
               names_sep = "_(?=[^_]+$)") %>%
  pivot_wider(names_from = statistic, values_from = value) %>%
  arrange(variable, wave)

print(summary_stats_by_wave)

# Histograms for each variable across waves
for (var in continuous_demo_vars) {
  p <- ggplot(adhd_menarche_clean, aes(x = .data[[var]])) +
    geom_histogram(fill = "skyblue", color = "black", bins = 30) +
    facet_wrap(~ wave, scales = "free_y") +
    labs(
      x = var,
      y = "Count",
      title = paste("Distribution of", var, "by Wave")
    )
  
  print(p)
}
```

## 2. Measure-level data quality checks

### 2.1 CBCL and NIHTB summaries

```{r cbcl-nihtb-summaries}
cbcl_vars <- c(
  "mh_p_cbcl__synd__attn_sum",
  "mh_p_cbcl__synd__int_sum",
  "mh_p_cbcl__synd__ext_sum")

summary_stats_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    across(
      all_of(cbcl_vars),
      list(
        mean            = ~ mean(.x, na.rm = TRUE),
        sd              = ~ sd(.x, na.rm = TRUE),
        min             = ~ min(.x, na.rm = TRUE),
        max             = ~ max(.x, na.rm = TRUE),
        n_missing       = ~ sum(is.na(.x)),
        percent_missing = ~ mean(is.na(.x)) * 100
      ),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  ) %>%
  pivot_longer(
    -wave,
    names_to = c("variable", "statistic"),
    names_pattern = "(.+)_(mean|sd|min|max|n_missing|percent_missing)"
  ) %>%
  pivot_wider(
    names_from = statistic,
    values_from = value
  ) %>%
  arrange(variable, wave)

print(summary_stats_by_wave)

# Histograms for each variable across waves
for (var in continuous_demo_vars) {
  p <- ggplot(adhd_menarche_clean, aes(x = .data[[var]])) +
    geom_histogram(fill = "skyblue", color = "black", bins = 30) +
    facet_wrap(~ wave, scales = "free_y") +
    labs(
      x = var,
      y = "Count",
      title = paste("Distribution of", var, "by Wave")
    )
  
  print(p)
}

adhd_menarche_clean %>%
  filter(wave == "wave_0") %>%
  summarise(
    mean_score = mean(nc_y_nihtb__comp__tot__agecor_score, na.rm = TRUE),
    sd_score = sd(nc_y_nihtb__comp__tot__agecor_score, na.rm = TRUE),
    min_score = min(nc_y_nihtb__comp__tot__agecor_score, na.rm = TRUE),
    max_score = max(nc_y_nihtb__comp__tot__agecor_score, na.rm = TRUE),
    n_valid = sum(!is.na(nc_y_nihtb__comp__tot__agecor_score)),
    n_missing = sum(is.na(nc_y_nihtb__comp__tot__agecor_score)),
    pct_missing = mean(is.na(nc_y_nihtb__comp__tot__agecor_score)) * 100
  )
```

### 2.2 PDS items summary

```{r pds-summary}
pds_items <- c(  
  "ph_p_pds_001",
  "ph_p_pds_002",
  "ph_p_pds_003",
  "ph_p_pds__f_001",
  "ph_p_pds__f_002"
  )

# PDS items
summary_stats <- adhd_menarche_clean %>%
  pivot_longer(
    cols = all_of(pds_items),
    names_to = "item",
    values_to = "value"
  ) %>%
  group_by(wave, item) %>%
  summarise(
    n_valid = sum(!is.na(value)),
    n_missing = sum(is.na(value)),
    pct_missing = 100 * mean(is.na(value)),

    mean   = if (all(is.na(value))) NA_real_ else mean(value, na.rm = TRUE),
    sd     = if (all(is.na(value))) NA_real_ else sd(value, na.rm = TRUE),
    min    = if (all(is.na(value))) NA_real_ else min(value, na.rm = TRUE),
    max    = if (all(is.na(value))) NA_real_ else max(value, na.rm = TRUE),

    .groups = "drop"
  )

summary_stats
```

### 2.3 Create a breast development variable

```{r breast-development}
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(breast_development_p =
           as.numeric(ph_p_pds__f_001)
  )

head(adhd_menarche_clean$breast_development_p)
```

## 3. ADHD status variables

```{r adhd-variables}
# List of ADHD variables
adhd_vars <- c(
  "mh_p_ksads__adhd__past_dx",
  "mh_p_ksads__adhd__pres_dx",
  "mh_p_ksads__adhd__fullrem_dx",
  "mh_p_ksads__adhd__partrem_dx",
  "mh_p_ksads__adhd__unspec_dx",
  "mh_p_ksads__adhd__oth_dx"
)

adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    across(
      all_of(adhd_vars),
      ~ sum(. == 1, na.rm = TRUE),
      .names = "{.col}_n1"
    ),
    .groups = "drop"
  )

adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    across(
      all_of(adhd_vars),
      list(
        n_555 = ~ sum(. == 555, na.rm = TRUE),
        n_888 = ~ sum(. == 888, na.rm = TRUE)
      ),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  )

# Convert 888 and 555 to NA 
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(
    across(
      all_of(adhd_vars),
      ~ na_if(., 555)
    )
  ) %>%
  mutate(
    across(
      all_of(adhd_vars),
      ~ na_if(., 888)
    )
  )
```

### 3.1 ADHD status lifetime

```{r adhd-lifetime}
# Wave-level counts
adhd_status_life_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  reframe(
    n_flagged = sum(if_any(all_of(adhd_vars), ~ . == 1), na.rm = TRUE),
    n_missing = sum(if_all(all_of(adhd_vars), ~ is.na(.))),
    n_nonADHD = sum(
      apply(across(all_of(adhd_vars)), 1, function(x) {
        non_missing <- x[!is.na(x)]
        length(non_missing) > 0 && all(non_missing == 0)
      })
    ),
    n_participants = n_distinct(participant_id)
  ) %>%
  mutate(
    pct_flagged = round(100 * n_flagged / n_participants, 1),
    pct_nonADHD = round(100 * n_nonADHD / n_participants, 1),
    pct_missing = round(100 * n_missing / n_participants, 1)
  )

# Create participant-level ADHD status (lifetime)
adhd_status_life <- adhd_menarche_clean %>%
  dplyr::select(participant_id, all_of(adhd_vars)) %>%
  group_by(participant_id) %>%
  summarise(
    adhd_status_life = case_when(
      any(c_across(all_of(adhd_vars)) == 1, na.rm = TRUE) ~ 1L,
      all(is.na(c_across(all_of(adhd_vars)))) ~ NA_integer_,
      TRUE ~ 0L
    ),
    .groups = "drop"
  )

# Join variable back to main dataset
adhd_menarche_clean <- adhd_menarche_clean %>%
  left_join(adhd_status_life, by = "participant_id")

# View wave-level summary
adhd_status_life_wave

total_adhd_life_status <- adhd_menarche_clean %>%
  dplyr::select(participant_id, adhd_status_life) %>% 
  distinct() %>%                          
  summarize(
    n_adhd = sum(adhd_status_life, na.rm = TRUE),
    n_total = n(),
    pct_adhd = round(100 * n_adhd / n_total, 1)
  )

total_adhd_life_status
```

### 3.2 Insufficient ADHD status lifetime data flag

```{r ahdd-insufficient-flag}
# Check participants never assigned to either group
unassigned <- adhd_menarche_clean %>%
  group_by(participant_id) %>%
  summarise(
    all_na = all(is.na(adhd_status_life)),
    any_na = any(is.na(adhd_status_life)),
    n_na_waves = sum(is.na(adhd_status_life)),
    .groups = "drop"
  ) %>%
  filter(any_na)

nrow(unassigned)  # how many participants have at least one NA wave
unassigned %>% count(n_na_waves)  # how many waves are NA per participant
```
### 3.3 ADHD status present

```{r adhd-present}
adhd_status_pres_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    n_flagged = sum(mh_p_ksads__adhd__pres_dx == 1, na.rm = TRUE),
    n_missing = sum(is.na(mh_p_ksads__adhd__pres_dx)),
    n_nonADHD = sum(mh_p_ksads__adhd__pres_dx == 0, na.rm = TRUE),
    n_participants = n_distinct(participant_id),
    .groups = "drop"
  ) %>%
  mutate(
    pct_flagged = round(100 * n_flagged / n_participants, 1),
    pct_nonADHD = round(100 * n_nonADHD / n_participants, 1),
    pct_missing = round(100 * n_missing / n_participants, 1)
  )

# View wave-level summary
adhd_status_pres_wave

adhd_status_pres <- adhd_menarche_clean %>%
  dplyr::select(participant_id, mh_p_ksads__adhd__pres_dx) %>%
  group_by(participant_id) %>%
  summarise(
    adhd_status_pres = case_when(
      any(mh_p_ksads__adhd__pres_dx == 1, na.rm = TRUE) ~ 1L,  
      all(is.na(mh_p_ksads__adhd__pres_dx)) ~ NA_integer_,       
      TRUE ~ 0L                                                  
    ),
    .groups = "drop"
  )

adhd_menarche_clean <- adhd_menarche_clean %>%
  left_join(adhd_status_pres, by = "participant_id")

total_adhd_pres_status <- adhd_menarche_clean %>%
  dplyr::select(participant_id, adhd_status_pres) %>% 
  distinct() %>%                          
  summarize(
    n_adhd = sum(adhd_status_pres, na.rm = TRUE),
    n_total = n(),
    pct_adhd = round(100 * n_adhd / n_total, 1)
  )

total_adhd_pres_status
```

## 4. Co-occurring conditions

### 4.1 Co-occurring internalising conditions

```{r cooccurring-internalising}
# List of depression variables
dep_vars <- c(
  "mh_p_ksads__dep__mdd__past_dx",
  "mh_p_ksads__dep__mdd__pres_dx",
  "mh_p_ksads__dep__mdd__partrem_dx",
  "mh_p_ksads__dep__pdd__past_dx",
  "mh_p_ksads__dep__pdd__pres_dx",
  "mh_p_ksads__dep__pdd__partrem_dx"
)

# Create participant-level flag
depression <- adhd_menarche_clean %>%
  dplyr::select(participant_id, all_of(dep_vars)) %>%
  group_by(participant_id) %>%
  reframe(
    depression = as.integer(any(c_across(all_of(dep_vars)) == 1, na.rm = TRUE))
  )

# Join back
adhd_menarche_clean <- adhd_menarche_clean %>%
  left_join(depression, by = "participant_id")

# Summary
total_depression <- adhd_menarche_clean %>%
  dplyr::select(participant_id, depression) %>% 
  distinct() %>%                          
  summarize(
    n_depression = sum(depression, na.rm = TRUE),
    n_total = n(),
    pct_depression = round(100 * n_depression / n_total, 1)
  )

total_depression

# List of GAD variables
gad_vars <- c(
  "mh_p_ksads__gad__past_dx",
  "mh_p_ksads__gad__pres_dx",
  "mh_p_ksads__gad__oth__mindur__past_dx", 
  "mh_p_ksads__gad__oth__mindur__pres_dx"
)

# Create participant-level flag
gad <- adhd_menarche_clean %>%
  dplyr::select(participant_id, all_of(gad_vars)) %>%
  group_by(participant_id) %>%
  reframe(
    gad = as.integer(any(c_across(all_of(gad_vars)) == 1, na.rm = TRUE))
  )

# Join back
adhd_menarche_clean <- adhd_menarche_clean %>%
  left_join(gad, by = "participant_id")

# Summary
total_gad <- adhd_menarche_clean %>%
  dplyr::select(participant_id, gad) %>% 
  distinct() %>%                          
  summarize(
    n_gad = sum(gad, na.rm = TRUE),
    n_total = n(),
    pct_gad = round(100 * n_gad / n_total, 1)
  )

total_gad
```

### 4.2 Co-occurring externalising conditions

```{r cooccurring-externalising}
# List of CD variables
cd_vars <- c(
  "mh_p_ksads__cond__onset__adoles__past_dx",
  "mh_p_ksads__cond__onset__adoles__pres_dx",
  "mh_p_ksads__cond__onset__chld__past_dx",
  "mh_p_ksads__cond__onset__chld__pres_dx"
)

# Create participant-level flag
cd <- adhd_menarche_clean %>%
  dplyr::select(participant_id, all_of(cd_vars)) %>%
  group_by(participant_id) %>%
  reframe(
    cd = as.integer(any(c_across(all_of(cd_vars)) == 1, na.rm = TRUE))
  )

# Join back
adhd_menarche_clean <- adhd_menarche_clean %>%
  left_join(cd, by = "participant_id")

# Summary
total_cd <- adhd_menarche_clean %>%
  dplyr::select(participant_id, cd) %>% 
  distinct() %>%                          
  summarize(
    n_cd = sum(cd, na.rm = TRUE),
    n_total = n(),
    pct_cd = round(100 * n_cd / n_total, 1)
  )

total_cd

# List of ODD variables
odd_vars <- c(
  "mh_p_ksads__odd__past_dx",
  "mh_p_ksads__odd__pres_dx"
)

# Create participant-level flag
odd <- adhd_menarche_clean %>%
  dplyr::select(participant_id, all_of(odd_vars)) %>%
  group_by(participant_id) %>%
  reframe(
    odd = as.integer(any(c_across(all_of(odd_vars)) == 1, na.rm = TRUE))
  )

# Join back
adhd_menarche_clean <- adhd_menarche_clean %>%
  left_join(odd, by = "participant_id")

# Summary
total_odd <- adhd_menarche_clean %>%
  dplyr::select(participant_id, odd) %>% 
  distinct() %>%                          
  summarize(
    n_odd = sum(odd, na.rm = TRUE),
    n_total = n(),
    pct_odd= round(100 * n_odd / n_total, 1)
  )

total_odd
```

## 5. Exclusion flags

### 5.1 Schizophrenia flag

```{r schizophrenia-flag}
# List of schizophrenia variables
schizophrenia_vars <- c(
  "mh_p_ksads__psych__schiz__pres_dx",
  "mh_p_ksads__psych__schizaffct__pres_dx",
  "mh_p_ksads__psych__schizform__pres_dx",
  "mh_p_ksads__psych__schiz__oth__past_dx",
  "mh_p_ksads__psych__schiz__oth__pres_dx"
)

# Create participant-level flag
schizophrenia_flag <- adhd_menarche_clean %>%
  dplyr::select(participant_id, all_of(schizophrenia_vars)) %>%
  group_by(participant_id) %>% 
  reframe( schizophrenia_flag = as.integer(any(across(all_of(schizophrenia_vars)) == 1)) )

# Join participant flag back to main dataset if needed
adhd_menarche_clean <- adhd_menarche_clean %>%
  left_join(schizophrenia_flag, by = "participant_id")

# Wave-level counts
schizophrenia_flag_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  reframe(
    n_flagged = sum(if_any(all_of(schizophrenia_vars), ~ . == 1), na.rm = TRUE),
    n_participants = n_distinct(participant_id)
  ) %>%
  mutate(
    pct_flagged = round(100 * n_flagged / n_participants, 1)
  )

# View wave-level summary
schizophrenia_flag_wave

total_schizophrenia_flag <- schizophrenia_flag %>%
  left_join(
    adhd_menarche_clean %>% dplyr::select(participant_id) %>% dplyr::distinct(),
    by = "participant_id"
  ) %>%
  summarize(
    n_schizophrenia = sum(schizophrenia_flag, na.rm = TRUE),
    n_total = n(),
    pct_schizophrenia = round(100 * n_schizophrenia / n_total, 1)
  )

total_schizophrenia_flag
```

### 5.2 Bipolar flag

```{r bipolar-flag}
# List of bipolar variables
bipolar_vars <- c(
  "mh_p_ksads__bpd__bpd1__rcnt__manic__pres_dx",
  "mh_p_ksads__bpd__bpd1__rcnt__manicmix__pres_dx",
  "mh_p_ksads__bpd__bpd2__curdep__pres_dx",
  "mh_p_ksads__bpd__bpd2__curdep__partrem_dx",
  "mh_p_ksads__bpd__bpd2__curhypo__pres_dx",
  "mh_p_ksads__bpd__bpd2__rcnt__hypo__pres_dx"
)
# Create participant-level flag
bipolar_flag <- adhd_menarche_clean %>%
  dplyr::select(participant_id, all_of(bipolar_vars)) %>%
  group_by(participant_id) %>%
  reframe(
    bipolar_flag = as.integer(any(across(all_of(bipolar_vars)) == 1))
  )

# Join participant flag back to main dataset if needed
adhd_menarche_clean <- adhd_menarche_clean %>%
  left_join(bipolar_flag, by = "participant_id")

# Wave-level counts
bipolar_flag_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  reframe(
    n_flagged = sum(if_any(all_of(bipolar_vars), ~ . == 1), na.rm = TRUE),
    n_participants = n_distinct(participant_id)
  ) %>%
  mutate(
    pct_flagged = round(100 * n_flagged / n_participants, 1)
  )

# View wave-level summary
bipolar_flag_wave

total_bipolar_flag <- adhd_menarche_clean %>%
  dplyr::select(participant_id, bipolar_flag) %>% 
  distinct() %>%                          
  summarize(
    n_bipolar = sum(bipolar_flag, na.rm = TRUE),
    n_total = n(),
    pct_bipolar = round(100 * n_bipolar / n_total, 1)
  )

total_bipolar_flag
```

### 5.3 Tic flag

```{r tic-flag}
# List of tic variables
tic_vars <- c(
  "mh_p_ksads__tic__ts__pres_dx",
  "mh_p_ksads__tic__ts__nosymp__rem_dx"
)

# Create participant-level flag
tic_flag <- adhd_menarche_clean %>%
  dplyr::select(participant_id, all_of(tic_vars)) %>%
  group_by(participant_id) %>%
  reframe(
    tic_flag = as.integer(any(across(all_of(tic_vars)) == 1))
  )

# Join participant flag back to main dataset
adhd_menarche_clean <- adhd_menarche_clean %>%
  left_join(tic_flag, by = "participant_id")

# Wave-level counts
tic_flag_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  reframe(
    n_flagged = sum(if_any(all_of(tic_vars), ~ . == 1), na.rm = TRUE),
    n_participants = n_distinct(participant_id)
  ) %>%
  mutate(
    pct_flagged = round(100 * n_flagged / n_participants, 1)
  )

# View wave-level summary
tic_flag_wave

total_tic_flag <- adhd_menarche_clean %>%
  dplyr::select(participant_id, tic_flag) %>% 
  distinct() %>%                          
  summarize(
    n_tic = sum(tic_flag, na.rm = TRUE),
    n_total = n(),
    pct_tic = round(100 * n_tic / n_total, 1)
  )

total_tic_flag
```

### 5.4 TBI flag

```{r tbi-flag}
# Create participant-level TBI flag (worst injury = 4 or 5 at any wave)
tbi_flag <- adhd_menarche_clean %>%
  dplyr::select(participant_id, ph_p_otbi_tbiworst) %>%
  group_by(participant_id) %>%
  reframe(
    tbi_flag = as.integer(any(ph_p_otbi_tbiworst %in% c(4, 5), na.rm = TRUE))
  )

# Join back to main dataset
adhd_menarche_clean <- adhd_menarche_clean %>%
  left_join(tbi_flag, by = "participant_id")

tbi_flag_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  reframe(
    n_flagged = sum(ph_p_otbi_tbiworst %in% c(4, 5), na.rm = TRUE),
    n_participants = n_distinct(participant_id)
  ) %>%
  mutate(
    pct_flagged = round(100 * n_flagged / n_participants, 2)
  )

tbi_flag_wave

total_tbi_flag <- adhd_menarche_clean %>%
  dplyr::select(participant_id, tbi_flag) %>%
  distinct() %>%
  summarize(
    n_tbi = sum(tbi_flag, na.rm = TRUE),
    n_total = n(),
    pct_tbi = round(100 * n_tbi / n_total, 2)
  )

total_tbi_flag
```

### 5.5 Low NIHTB flag (>2D below mean baseline NITHB total)

```{r nihtb-flag}
### 5.5 Low NIHTB flag (>2SD below mean baseline NIHTB total)

# Calculate mean and SD using baseline (wave_0) NIHTB scores only
nihtb_baseline_stats <- adhd_menarche_clean %>%
  filter(wave == "wave_0") %>%
  summarise(
    nihtb_mean = mean(nc_y_nihtb__comp__tot__agecor_score, na.rm = TRUE),
    nihtb_sd = sd(nc_y_nihtb__comp__tot__agecor_score, na.rm = TRUE)
  )

nihtb_mean <- nihtb_baseline_stats$nihtb_mean
nihtb_sd <- nihtb_baseline_stats$nihtb_sd

# Create low-score flag: 1 = >2 SD below mean at baseline, 0 = otherwise
# NA for non-baseline waves, since NIHTB total is only assessed at baseline
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(
    nihtb_low_flag = if_else(
      wave == "wave_0",
      as.integer(nc_y_nihtb__comp__tot__agecor_score < (nihtb_mean - 2 * nihtb_sd)),
      NA_integer_
    )
  )

# Participant-level summary (baseline only, since the flag is only defined there)
total_nihtb_low_flag <- adhd_menarche_clean %>%
  filter(wave == "wave_0") %>%
  dplyr::select(participant_id, nihtb_low_flag) %>%
  distinct() %>%
  summarise(
    n_nihtb_low = sum(nihtb_low_flag, na.rm = TRUE),
    n_total = n(),
    pct_nihtb_low = round(100 * n_nihtb_low / n_total, 2)
  )

total_nihtb_low_flag
```

## 6. ADHD stimulant medication use variable

### 6.1 Find stimulant medications recorded 

```{r search-stimulants}
# list of stimulant medications
stimulants <- c(
  # Methylphenidate-based
  "methylphenidate",
  "adhansia",
  "aptensio",
  "concerta",
  "cotempla",
  "daytrana",
  "jornay",
  "metadate",
  "methylin",
  "quillichew",
  "quillivant",
  "relexxii",
  "ritalin",
  "focalin",
  "azstarys",
  
  # Amphetamine-based
  "amphetamine",
  "dextroamphetamine",
  "adzenys",
  "dyanavel",
  "dexedrine",
  "procentra",
  "xelstrym",
  "zenzedi",
  "adderall",
  "mydayis",
  "evekeo",
  "vyvanse",
  "lisdexamfetamine",
  
  # Methamphetamine
  "desoxyn"
)

# Get all medication columns
med_cols <- grep("^ph_p_meds__rx__(label)_", names(adhd_menarche_clean), value = TRUE)

# Identify unique stimulant medications in the dataset (case-insensitive)
detected_stimulants <- unique(
  unlist(
    apply(
      adhd_menarche_clean[, med_cols],
      1,
      function(row) {
        row_lower <- tolower(row)  # Lowercase for case-insensitive matching
        meds_found <- sapply(stimulants, function(med) {
          grep(med, row_lower, fixed = TRUE, value = TRUE)
        })
        unlist(meds_found)
      }
    )
  )
)

# Remove any empty strings from detected list
detected_stimulants <- detected_stimulants[detected_stimulants != ""]

# Print results
print(detected_stimulants)
```

### 6.2 Create stimulant status variable

```{r stimulant-flag}
# Medication label columns
med_label_cols <- grep("label", med_cols, value = TRUE)

# Only use medication columns
med_long <- adhd_menarche_clean %>%
  dplyr::select(participant_id, wave, all_of(med_label_cols)) %>%
  pivot_longer(
    cols = all_of(med_label_cols),
    names_to = "med_col",
    values_to = "med_value"
  ) %>%
  filter(!is.na(med_value)) %>% 
  mutate(
    med_value_clean = tolower(trimws(gsub("[[:punct:]]", "", med_value)))
  )

med_long <- med_long %>%
  mutate(
    stimulant_status = as.integer(
      sapply(med_value_clean, function(x) any(grepl(paste(stimulants, collapse = "|"), x)))
    )
  )

# Create wave-level stimulant status (Y / N)
stimulant_status_wave <- med_long %>%
  group_by(participant_id, wave) %>%
  summarise(
    stimulant_status = if_else(any(stimulant_status == 1, na.rm = TRUE), "Y", "N"),
    .groups = "drop"
  )


# Merge into main dataset
adhd_menarche_clean %>%
  left_join(stimulant_status_wave, by = c("participant_id", "wave")) %>%
  glimpse()

adhd_menarche_clean <- adhd_menarche_clean %>%
  left_join(stimulant_status_wave, by = c("participant_id", "wave")) %>%
  mutate(
    stimulant_status = if_else(is.na(stimulant_status), "N", stimulant_status)
  )

# Wave-level counts
stimulant_status_wave_summary <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    n_Y = sum(stimulant_status == "Y", na.rm = TRUE),
    n_N = sum(stimulant_status == "N", na.rm = TRUE),
    n_participants = n_distinct(participant_id),
    .groups = "drop"
  ) %>%
  mutate(
    pct_Y = round(100 * n_Y / n_participants, 1)
  )

stimulant_status_wave_summary


# Overall (person-level) summary
total_stimulant_status <- adhd_menarche_clean %>%
  select(participant_id, stimulant_status) %>%
  distinct() %>%
  summarize(
    n_Y = sum(stimulant_status == "Y", na.rm = TRUE),
    n_total = n(),
    pct_Y = round(100 * n_Y / n_total, 2)
  )

total_stimulant_status
```

### 6.3 Cross tab ADHD flag with stimulant flag

```{r crosstab-adhd-stimulants}
# Cross-tabulate ADHD flag by stimulant medication flag
adhd_life_stimulant_table <- adhd_menarche_clean %>%
  group_by(adhd_status_life, stimulant_status) %>%
  summarise(
    n_participants = n_distinct(participant_id),
    .groups = "drop"
  ) %>%
  tidyr::pivot_wider(
    names_from = stimulant_status,
    values_from = n_participants,
    names_prefix = "stimulant_",
    values_fill = 0
  ) %>%
  mutate(
    total = rowSums(across(starts_with("stimulant_")))
  )

adhd_life_stimulant_table

# Cross-tabulate ADHD flag by stimulant medication flag
adhd_pres_stimulant_table <- adhd_menarche_clean %>%
  group_by(adhd_status_pres, stimulant_status) %>%
  summarise(
    n_participants = n_distinct(participant_id),
    .groups = "drop"
  ) %>%
  tidyr::pivot_wider(
    names_from = stimulant_status,
    values_from = n_participants,
    names_prefix = "stimulant_",
    values_fill = 0
  ) %>%
  mutate(
    total = rowSums(across(starts_with("stimulant_")))
  )

adhd_pres_stimulant_table
```

## 7. Observed dataframe for descriptives and sensitivity analyses

### 7.1 Apply exclusions

```{r apply-exclusions}
n_original <- length(unique(adhd_menarche_clean$participant_id))

# Participant-level exclusions
excluded_ids <- adhd_menarche_clean %>%
  group_by(participant_id) %>%
  summarise(
    tic = any(tic_flag == 1, na.rm = TRUE),
    tbi = any(tbi_flag == 1, na.rm = TRUE),
    schizophrenia = any(schizophrenia_flag == 1, na.rm = TRUE),
    bipolar = any(bipolar_flag == 1, na.rm = TRUE),
    nihtb = any(nihtb_low_flag == 1, na.rm = TRUE),
    .groups = "drop"
  )

# Initialise remaining IDs (all participants at start)
remaining_ids <- excluded_ids$participant_id

# Initialise dataframe to track sequential exclusions
exclusion_counts <- data.frame(
  flag = character(),
  n_excluded = integer(),
  stringsAsFactors = FALSE
)

# Function to apply exclusion and count
exclude_flag <- function(flag_name, current_ids) {
  to_exclude <- excluded_ids %>%
    filter(participant_id %in% current_ids) %>%
    filter(.data[[flag_name]] == TRUE) %>%
    pull(participant_id)
  
  n_excl <- length(to_exclude)
  
  # Update exclusion counts
  exclusion_counts <<- rbind(exclusion_counts,
                             data.frame(flag = flag_name,
                                        n_excluded = n_excl))
  
  # Return IDs that remain after exclusion
  setdiff(current_ids, to_exclude)
}

# Exclude participant with missing ethnrace first
missing_ethnrace <- adhd_menarche_clean %>%
  filter(participant_id %in% remaining_ids) %>%
  filter(is.na(ab_g_stc__cohort_ethnrace__leg)) %>%
  distinct(participant_id) %>%
  pull(participant_id)

exclusion_counts <- rbind(exclusion_counts,
                          data.frame(flag = "missing_ethnrace",
                                     n_excluded = length(missing_ethnrace)))

remaining_ids <- setdiff(remaining_ids, missing_ethnrace)

# Apply exclusions sequentially
remaining_ids <- exclude_flag("tic", remaining_ids)
remaining_ids <- exclude_flag("tbi", remaining_ids)
remaining_ids <- exclude_flag("schizophrenia", remaining_ids)
remaining_ids <- exclude_flag("bipolar", remaining_ids)
remaining_ids <- exclude_flag("nihtb", remaining_ids)

# Final observed dataframe
obs_df <- adhd_menarche_clean %>%
  filter(participant_id %in% remaining_ids)

n_obs <- length(unique(obs_df$participant_id))

# Total excluded
total_excluded <- n_original - n_obs

# Print results
print(exclusion_counts)
cat("Total participants excluded:", total_excluded, "\n")
cat("Total participants remaining:", length(unique(obs_df$participant_id)), "\n")
```

### 7.2 Create obs_df

```{r obs-df}
obs_vars <- c(
  # IDs / basic info
  "participant_id",
  "wave",
  
  # Demographics
  "ab_p_demo_age",
  "ph_y_anthr__weight_mean",
  "ph_y_anthr__height_mean",
  "household_income",
  "ab_p_demo__roster_001",
  "le_l_adi__addr1_wsum",
  "ab_g_stc__cohort_ethnrace__leg",
  "ab_g_dyn__design_site",
  "ab_g_stc__design_id__fam",
  
  # CBCL
  "mh_p_cbcl__synd__attn_sum",
  "mh_p_cbcl__synd__int_sum",
  "mh_p_cbcl__synd__ext_sum",
  
  # Neurocognitive measures
  "nc_y_nihtb__comp__tot__agecor_score",
  
  # PDS items
  "ph_p_pds_001", "ph_p_pds_002", "ph_p_pds_003",
  "ph_p_pds__f_001", "ph_p_pds__f_002", "ph_y_pds__f_002",
  "breast_development_p", "menarche_status_p",
  
  # Clinical flags
  "adhd_status_life",
  "adhd_status_pres",
  "depression",
  "gad",
  "cd",
  "odd",
  
  # Medication / stimulant
  "stimulant_status"
)

# Create imputation/analysis dataframe
obs_df <- obs_df %>%
  dplyr::select(all_of(obs_vars))
```

### 7.3 Recoded variables for obs df

```{r obs-df-recoding}
obs_df <- obs_df %>%
  mutate(
    adhd_status_life = factor(
      adhd_status_life, 
      levels = c(0, 1), 
      labels = c("N", "Y")
    ),
    adhd_status_pres = factor(
      adhd_status_pres, 
      levels = c(0, 1), 
      labels = c("N", "Y")
    ),
    ethnrace = factor(
      ab_g_stc__cohort_ethnrace__leg,
      levels = c(2, 1, 3, 4, 13),
      labels = c(
        "white_nh",
        "hispanic",
        "black_nh",
        "asian",
        "other"
      )
    )
  )

obs_df$ph_y_anthr__weight_mean <- as.numeric(obs_df$ph_y_anthr__weight_mean)
obs_df$ph_y_anthr__height_mean <- as.numeric(obs_df$ph_y_anthr__height_mean)

# Spot check
head(obs_df$adhd_status_life)
head(obs_df$adhd_status_pres)
head(obs_df$ethnrace)
head(obs_df$menarche_status_p)

str(obs_df[, c(
  "adhd_status_life",
  "adhd_status_pres",
  "menarche_status_p",
  "ph_p_pds__f_001",
  "ph_y_anthr__weight_mean"
)])

```

### 7.4 Obs BMI

```{r obs-bmi}
obs_df <- obs_df %>%
    mutate(bmi = (ph_y_anthr__weight_mean / (ph_y_anthr__height_mean^2)) * 703)

head(obs_df$bmi)
min(obs_df$bmi, na.rm = TRUE)
max(obs_df$bmi, na.rm = TRUE)
```

### 7.5 Obs INR variable

```{r obs-inr}
US_poverty_thresholds <- read.csv("US_poverty_thresholds.csv")

US_poverty_thresholds <- US_poverty_thresholds %>%
  mutate(wave_people = paste0(wave, "_", people))

income_mapping <- c(
  2500,   # <$5,000
  8500,   # $5,000–$11,999
  14000,  # $12,000–$15,999
  20500,  # $16,000–$24,999
  30000,  # $25,000–$34,999
  42500,  # $35,000–$49,999
  62500,  # $50,000–$74,999
  87500,  # $75,000–$99,999
  150000, # $100,000–$199,999
  300000  # $200,000+
)

obs_df <- obs_df %>%
    mutate(
      # Map ordinal income to midpoing
      mean_income = income_mapping[household_income],

      # Create merge key
      wave_people = paste0(wave, "_", ab_p_demo__roster_001)
    ) %>%

    # Join poverty thresholds
    left_join(US_poverty_thresholds, by = c("wave_people", "wave")) %>%

    # Calculate Income-to-Needs Ratio
    mutate(
      inr = mean_income / poverty_threshold
    )

names(obs_df)
head(obs_df$inr)
```

### 7.6  Rename analytic variables

```{r rename-analytic-vars}
obs_df <- obs_df %>%
    rename(
      age = ab_p_demo_age,
      family_id = ab_g_stc__design_id__fam,
      site_id = ab_g_dyn__design_site,
      adhd_traits = mh_p_cbcl__synd__attn_sum,
      internalising = mh_p_cbcl__synd__int_sum,
      externalising = mh_p_cbcl__synd__ext_sum
    )

names(obs_df)
```

### 7.7 Check variable types

```{r check-var-types}
check_vars <- c(
  "age",
  "bmi",
  "adhd_traits",
  "internalising",
  "externalising",
  "menarche_status_p",
  "breast_development_p",
  "inr",
  "ethnrace"
)

# Check class of these variables
sapply(obs_df[check_vars], class)

unique(obs_df$ethnrace)
```

### 7.8 Scaling

```{r scaling-centring}
continuous_vars <- c(
    "age",
    "inr",
    "adhd_traits",
    "breast_development_p"
)

obs_df <- obs_df %>%
  mutate(
    across(
      all_of(continuous_vars),
      ~ (.x - mean(.x, na.rm = TRUE)) / (2 * sd(.x, na.rm = TRUE)),
      .names = "{.col}_sc"
    )
  )

# Spot checks
mean(obs_df$inr_sc, na.rm = TRUE)
sd(obs_df$inr_sc, na.rm = TRUE)
mean(obs_df$age_sc, na.rm = TRUE)
sd(obs_df$age_sc, na.rm = TRUE)
mean(obs_df$breast_development_p_sc, na.rm = TRUE)
sd(obs_df$breast_development_p_sc, na.rm = TRUE)
```

### 7.9 Create wide version with scaling

```{r wide-obs-df}
obs_df_wide <- obs_df %>%
  pivot_wider(
    id_cols = participant_id,
    names_from = wave,
    names_glue = "{.value}_{wave}",
    values_from = -c(participant_id, wave)
  )

obs_df_wide <- obs_df_wide %>%
    rename(
      family_id = family_id_wave_0,
      site_id = site_id_wave_0,
      ethnrace = ethnrace_wave_0,
      adhd_status_life = adhd_status_life_wave_0
    )
  
continuous_vars <- c(
    "age_wave_2",
    "adhd_traits_wave_0",
    "adhd_traits_wave_2",
    "internalising_wave_0",
    "externalising_wave_0",
    "inr_wave_0",
    "breast_development_p_wave_2"
)

obs_df_wide <- obs_df_wide %>%
  mutate(
    across(
      all_of(continuous_vars),
      ~ (.x - mean(.x, na.rm = TRUE)) / (2 * sd(.x, na.rm = TRUE)),
      .names = "{.col}_sc"
    )
  )

# Spot checks
mean(obs_df_wide$inr_wave_0_sc, na.rm = TRUE)
sd(obs_df_wide$inr_wave_0_sc, na.rm = TRUE)
mean(obs_df_wide$age_wave_2_sc, na.rm = TRUE)
sd(obs_df_wide$age_wave_2_sc, na.rm = TRUE)
```

### 7.10 Export observed df for descriptives and sensitivity

```{r export-obs-df}
write_csv(obs_df, "obs_df.csv")

write_csv(obs_df_wide, "obs_df_wide.csv")
```

## 8. Prepare df for imputation

```{r imputation-df}
# Vector of variables to include in imputation/analysis dataframe
imputation_vars <- c(
  # IDs / basic info
  "participant_id",
  "wave",
  
  # Demographics
  "age",
  "ph_y_anthr__weight_mean",
  "ph_y_anthr__height_mean",
  "household_income",
  "ab_p_demo__roster_001",
  "le_l_adi__addr1_wsum",
  "ab_g_stc__cohort_ethnrace__leg",
  "ab_g_dyn__design_site",
  "ab_g_stc__design_id__fam",
  
  # CBCL
  "mh_p_cbcl__synd__attn_sum",
  "mh_p_cbcl__synd__ext_sum",
  "mh_p_cbcl__synd__int_sum",
  
  # Neurocognitive measures
  "nc_y_nihtb__comp__tot__agecor_score",
  
  # PDS items
  "ph_p_pds_001", "ph_p_pds_002", "ph_p_pds_003",
  "ph_p_pds__f_001", "ph_p_pds__f_002", 
  "breast_development_p", "menarche_status_p",
  "ph_y_pds_001", "ph_y_pds_002", "ph_y_pds_003",
  "ph_y_pds__f_001", "ph_y_pds__f_002",
  
  # Clinical flags
  "adhd_status_life",
  "adhd_status_pres",
  "schizophrenia_flag",
  "bipolar_flag",
  "tic_flag",
  "tbi_flag",
  "nihtb_low_flag",
  
  # Medication / stimulant
  "stimulant_status"
)

# Create imputation/analysis dataframe
imputation_df <- adhd_menarche_clean %>%
  dplyr::select(all_of(imputation_vars))

# Export to CSV
write_csv(imputation_df, "imputation_df.csv")
```


