---
title: "ABCD ADHD menarche imputations (step 4)"
author: "Emily Friedel"
date: "`r Sys.Date()`"
editor: source
---

## Set up

### Load libraries

```{r load-libraries}
library(tidyverse)
library(knitr)
library(tidyr)
library(readxl)
library(readr)
library(mice)
library(future)
library(future.apply)
library(miceadds)
library(furrr)
library(MASS)
```

### Import cleaned dataframe

```{r read-cleaned-df}
df <- read.csv("imputation_df.csv")
str(df)
```

## 1. Imputation preparation

```{r preparation}
# Recode PDS menarche status yes/no (0/1) to "N"/"Y"
df$ph_p_pds__f_002 <- case_when(
  df$ph_p_pds__f_002 == 0 ~ "N",  
  df$ph_p_pds__f_002 == 1 ~ "Y",  
  TRUE ~ NA_character_
)

# Convert categorical variables to factors
df$ph_p_pds__f_002 <- factor(df$ph_p_pds__f_002, levels = c("N","Y"))
df$wave <- factor(df$wave, levels  = c("wave_0", "wave_1", "wave_2", "wave_3"))
df$stimulant_status <- factor(df$stimulant_status, levels = c("N","Y"))
```

## 2. Imputation

### 2.1 Imputation variables

```{r impute-vars}
impute_vars <- c(
  # Biological/demographics
  "age",
  "ph_y_anthr__weight_mean",
  "ph_y_anthr__height_mean",
  "household_income",
  "ab_p_demo__roster_001",
  "le_l_adi__addr1_wsum",
  "ab_g_stc__cohort_ethnrace__leg",
  
  # PDS
  "ph_p_pds_001", 
  "ph_p_pds_002", 
  "ph_p_pds_003",
  "ph_p_pds__f_001",

  # Menarche
  "ph_p_pds__f_002",

  # CBCL
  "mh_p_cbcl__synd__attn_sum",
  "mh_p_cbcl__synd__ext_sum",
  "mh_p_cbcl__synd__int_sum",
  
  # Auxiliaries
  "ab_g_dyn__design_site",
  "ab_g_stc__design_id__fam",
  "participant_id",
  
  # Not included
  "wave",
  "schizophrenia_flag",
  "bipolar_flag",
  "tic_flag",
  "tbi_flag",
  "nihtb_low_flag",
  "stimulant_status",
  "adhd_status_life",
  "adhd_status_pres",
  "menarche_status_p",
  "breast_development_p"
)
```

### 2.2 Define methods

```{r methods}
method <- make.method(df)

# Define method vector
method <- rep("", length(impute_vars))
names(method) <- impute_vars

# Continuous / quasi-continuous
pmm_vars <- c(
  "ph_y_anthr__weight_mean",
  "ph_y_anthr__height_mean",
  "household_income",
  "ab_p_demo__roster_001",
  "le_l_adi__addr1_wsum",
  "mh_p_cbcl__synd__attn_sum",
  "mh_p_cbcl__synd__ext_sum",
  "mh_p_cbcl__synd__int_sum"
)
method[pmm_vars] <- "pmm"

# Define ordinal PDS items
pds_ord_vars <- c(
  "ph_p_pds_001",
  "ph_p_pds_002",
  "ph_p_pds_003",
  "ph_p_pds__f_001"
)

# Force them to ordered factors 1–4
df[pds_ord_vars] <- lapply(df[pds_ord_vars], function(x) {
  x <- as.numeric(x)  # ensure numeric first
  factor(x, levels = 1:4, ordered = TRUE)
})

method[pds_ord_vars] <- "polr"

# Menarche (binary)
method["ph_p_pds__f_002"] <- "logreg"

# Excluded vars (never imputed)
excluded_vars <- c(
  "schizophrenia_flag", 
  "bipolar_flag",
  "tic_flag", 
  "tbi_flag", 
  "nihtb_low_flag",
  "menarche_status_p",
  "breast_development_p",
  "adhd_status_pres",
  "participant_id",
  "ab_g_stc__design_id__fam",
  "ab_g_dyn__design_site"
)

method[excluded_vars] <- ""

not_imputed <- c(
  "age",
  "ab_g_stc__cohort_ethnrace__leg",
  "adhd_status_life",
  "stimulant_status",
  "wave"
)

method[not_imputed] <- ""
```

### 2.3 Define predictor matrix

```{r predictor-matrix}
predictor_matrix <- quickpred(df[, impute_vars])

# Prevent excluded variables from being used as predictors
predictor_matrix[, excluded_vars] <- 0
predictor_matrix[excluded_vars, ] <- 0

# Prevent variables from predicting themselves
diag(predictor_matrix) <- 0
```

### 2.4 Run mice

```{r run-mice}
set.seed(1234)

plan(sequential)
plan(multisession, workers = 4)

imp <- futuremice(
  data = df[, impute_vars],
  method = method,
  predictorMatrix = predictor_matrix,
  m = 20,
  maxit = 50
)

# Inspect results
completed_df <- complete(imp, 1)
head(completed_df)

# Save mids object
saveRDS(imp, "imp.rds")
```

## 3. Check imputations

### 3.1 Convergence

```{r convergence}
View(mice::convergence(imp))
```

### 3.2 Check missingness post imputation

```{r missingness}
# Define variables
pds_vars <- c(
  "ph_p_pds_001", 
  "ph_p_pds_002", 
  "ph_p_pds_003",
  "ph_p_pds__f_001", 
  "ph_p_pds__f_002"
  )

demo_vars <- c(
  "age", 
  "household_income",
  "ph_y_anthr__weight_mean",
  "ph_y_anthr__height_mean"
  )

cbcl_vars  <- c(
  "mh_p_cbcl__synd__attn_sum",
  "mh_p_cbcl__synd__ext_sum",
  "mh_p_cbcl__synd__int_sum"
  )

# Select relevant waves
completed_sub <- completed_df %>%
  dplyr::select(wave, all_of(c(pds_vars, demo_vars, cbcl_vars)))

# Function to compute missingness by variable, wave
missing_summary <- function(df) {
  df %>%
    group_by(wave) %>%
    summarise(across(
      everything(), 
      list(n_missing = ~ sum(is.na(.)),
           percent_missing = ~ round(mean(is.na(.)) * 100, 2)),
      .names = "{.col}_{.fn}"
    ), .groups = "drop")
}

missing_after <- missing_summary(completed_sub)
View(missing_after)
```

### 3.3 Trace plots

```{r trace-plots}
continuous_vars <- c(
  "ph_y_anthr__weight_mean",
  "ph_y_anthr__height_mean",
  
  "household_income",
  "ab_p_demo__roster_001",
  "le_l_adi__addr1_wsum",

  "mh_p_cbcl__synd__attn_sum",
  "mh_p_cbcl__synd__ext_sum",
  "mh_p_cbcl__synd__int_sum"
  )


# Which variables were actually imputed
imputed_vars <- names(imp$chainMean)

# Keep only continuous vars that were actually imputed
continuous_vars_imputed <- intersect(continuous_vars, imputed_vars)

# Trace plots
plot(imp, continuous_vars_imputed)
```

### 3.4 Stripplots

```{r stripplots}
vars_to_plot <- intersect(continuous_vars, names(imp$imp))

for (v in vars_to_plot) {
  message("Plotting: ", v)
  
  print(
    stripplot(
      imp,
      as.formula(paste(v, "~ .imp")),
      main = paste("Observed vs Imputed:", v),
      pch = 20,
      cex = 1.1
    )
  )
}
```

### 3.5 Means/ranges comparison between observed and imputed

```{r mean-ranges-comparison}
compare_mean_range <- function(imp, var) {

  if (!var %in% names(imp$data)) return(NULL)

  # observed data
  obs <- imp$data[[var]]

  # store observed stats
  obs_n    <- sum(!is.na(obs))
  obs_mean <- mean(obs, na.rm = TRUE)
  obs_min  <- min(obs, na.rm = TRUE)
  obs_max  <- max(obs, na.rm = TRUE)

  # completed stats across imputations
  comp_means <- numeric(imp$m)
  comp_mins  <- numeric(imp$m)
  comp_maxs  <- numeric(imp$m)

  for (i in 1:imp$m) {
    x <- complete(imp, i)[[var]]

    comp_means[i] <- mean(x, na.rm = TRUE)
    comp_mins[i]  <- min(x, na.rm = TRUE)
    comp_maxs[i]  <- max(x, na.rm = TRUE)
  }

  # return tidy row
  data.frame(
    variable = var,

    obs_n    = obs_n,
    obs_mean = round(obs_mean, 2),
    imp_mean = round(mean(comp_means), 2),
    obs_min  = round(obs_min, 2),
    imp_min  = round(min(comp_mins), 2),
    obs_max  = round(obs_max, 2),
    imp_max  = round(max(comp_maxs), 2)
  )
}

results_table <- purrr::map_dfr(continuous_vars, ~compare_mean_range(imp, .x))

results_table
```

## 4. Average imputed dfs

```{r average-imputed-dfs}
# Extract all completed datasets from the mids object
all_imputed <- lapply(1:imp$m, function(i) complete(imp, i))

# Process each imputed dataset
imp_dfs <- purrr::imap(all_imputed, ~ {
  df <- .x
  df$.imp <- .y

  # Drop ALL PDS variables
  df <- df %>% dplyr::select(-dplyr::starts_with("ph_p_pds_"))

  df
})

# Combine all imputations into one dataframe
combined <- dplyr::bind_rows(imp_dfs)

# Identify ID variables
id_vars <- c("participant_id", "wave")

# Identify numeric variables for averaging
numeric_vars <- combined %>%
  dplyr::select(-all_of(id_vars), -.imp) %>%
  dplyr::select(where(is.numeric)) %>%
  names()

# Average across imputations
averaged <- combined %>%
  group_by(dplyr::across(all_of(id_vars))) %>%
  summarise(
  across(all_of(numeric_vars), ~mean(.x, na.rm = TRUE)),
    .groups = "drop"
  )

# Use first completed dataset as template
original <- all_imputed[[1]] %>%
  dplyr::select(-dplyr::starts_with("ph_p_pds_"))  

# Merge averaged values
imp_df <- dplyr::left_join(original, averaged, by = id_vars, 
                           suffix = c("", "_avg"))

# Replace missing values with averaged imputations
for (v in numeric_vars) {
  imp_df[[v]] <- ifelse(
    is.na(imp_df[[v]]),
    imp_df[[paste0(v, "_avg")]],
    imp_df[[v]]
  )
  imp_df[[paste0(v, "_avg")]] <- NULL
}

head(imp_df)

dim(imp_df)
```

## 5. Check missingingness

```{r missingness-averaged-df}
# Define variables
pds_vars <- c(
  "menarche_status_p",
  "breast_development_p"
)

demo_vars <- c(
  "age", 
  "household_income",
  "ph_y_anthr__weight_mean",
  "ph_y_anthr__height_mean")

cbcl_vars  <- c(
  "mh_p_cbcl__synd__attn_sum",
  "mh_p_cbcl__synd__ext_sum",
  "mh_p_cbcl__synd__int_sum"
)

# Select relevant waves
completed_sub <- imp_df %>%
  dplyr::select(wave, all_of(c(pds_vars, demo_vars, cbcl_vars)))

# Function to compute missingness by variable, wave
missing_summary <- function(df) {
  df %>%
    group_by(wave) %>%
    summarise(across(
      everything(), 
      list(n_missing = ~ sum(is.na(.)),
           percent_missing = ~ round(mean(is.na(.)) * 100, 2)),
      .names = "{.col}_{.fn}"
    ), .groups = "drop")
}

missing_after <- missing_summary(completed_sub)
missing_after
```

## 6. Save averaged imputed df 

```{r export}
write_csv(imp_df, "imputed_dataset.csv")
```

