---
title: "ABCD ADHD menarche post imputation processing (step 5)"
author: "Emily Friedel"
date: "`r Sys.Date()`"
editor: source
---

##Setup

### Load libraries

```{r load-libraries}
library(tidyverse)
library(ggplot2)
library(mice)
library(brms)
library(purrr)
library(corrplot)
```

### Import imputed df

```{r import}
imp_df <- read.csv("imputed_dataset.csv")
```

## Initial checks

```{r structure-imp}
str(imp_df)
head(imp_df)
```

## 1. Post-imputation data processing

### 1.1 Apply exclusions

```{r exclusions}
n_original <- length(unique(imp_df$participant_id))

# Participant-level exclusions
excluded_ids <- imp_df %>%
  group_by(participant_id) %>%
  summarise(
    tic = any(tic_flag == 1, na.rm = TRUE),
    tbi = any(tbi_flag == 1, na.rm = TRUE),
    schizophrenia = any(schizophrenia_flag == 1, na.rm = TRUE),
    bipolar = any(bipolar_flag == 1, na.rm = TRUE),
    nihtb = any(nihtb_low_flag == 1, na.rm = TRUE),
    .groups = "drop"
  )

# Initialise remaining IDs (all participants at start)
remaining_ids <- excluded_ids$participant_id

# Initialise dataframe to track sequential exclusions
exclusion_counts <- data.frame(
  flag = character(),
  n_excluded = integer(),
  stringsAsFactors = FALSE
)

# Function to apply exclusion and count
exclude_flag <- function(flag_name, current_ids) {
  to_exclude <- excluded_ids %>%
    filter(participant_id %in% current_ids) %>%
    filter(.data[[flag_name]] == TRUE) %>%
    pull(participant_id)
  
  n_excl <- length(to_exclude)
  
  # Update exclusion counts
  exclusion_counts <<- rbind(exclusion_counts,
                             data.frame(flag = flag_name,
                                        n_excluded = n_excl))
  
  # Return IDs that remain after exclusion
  setdiff(current_ids, to_exclude)
}

# Exclude participant with missing ethnrace first
missing_ethnrace <- imp_df %>%
  filter(participant_id %in% remaining_ids) %>%
  filter(is.na(ab_g_stc__cohort_ethnrace__leg)) %>%
  distinct(participant_id) %>%
  pull(participant_id)

exclusion_counts <- rbind(exclusion_counts,
                          data.frame(flag = "missing_ethnrace",
                                     n_excluded = length(missing_ethnrace)))

remaining_ids <- setdiff(remaining_ids, missing_ethnrace)

# Apply exclusions sequentially
remaining_ids <- exclude_flag("tic", remaining_ids)
remaining_ids <- exclude_flag("tbi", remaining_ids)
remaining_ids <- exclude_flag("schizophrenia", remaining_ids)
remaining_ids <- exclude_flag("bipolar", remaining_ids)
remaining_ids <- exclude_flag("nihtb", remaining_ids)

# Final observed dataframe
imp_df <- imp_df %>%
  filter(participant_id %in% remaining_ids)

n_obs <- length(unique(imp_df$participant_id))

# Total excluded
total_excluded <- n_original - n_obs

# Print results
print(exclusion_counts)
cat("Total participants excluded:", total_excluded, "\n")
cat("Total participants remaining:", length(unique(imp_df$participant_id)), "\n")
```

### 1.2 Check sample size

```{r sample-size}
summary(unique(imp_df$participant_id))
```

### 1.3 Check number of ADHD left

```{r check-adhd}
adhd_life_remaining <- imp_df %>%
  group_by(participant_id) %>%
  filter(adhd_status_life == 1)

summary(unique(adhd_life_remaining$participant_id))

adhd_pres_remaining <- imp_df %>%
  group_by(participant_id) %>%
  filter(adhd_status_pres == 1)

summary(unique(adhd_pres_remaining$participant_id))
```

### 1.4 Convert character variables to factors

```{r recoding-factors}
imp_df <- imp_df %>%
  mutate(
    menarche_status_p = factor(
      menarche_status_p, 
      levels = c("N", "Y"), 
      labels = c("N", "Y")
    ),
    adhd_status_life = factor(
      adhd_status_life, 
      levels = c(0, 1), 
      labels = c("N", "Y")
    ),
    adhd_status_pres = factor(
      adhd_status_pres, 
      levels = c(0, 1), 
      labels = c("N", "Y")
    ),
    ethnrace = factor(
      ab_g_stc__cohort_ethnrace__leg,
      levels = c(2, 1, 3, 4, 13),
      labels = c(
        "white_nh",
        "hispanic",
        "black_nh",
        "other",
        "other"
      )
    )
  )

imp_df$breast_development_p <- as.numeric(imp_df$breast_development_p)
imp_df$ph_y_anthr__weight_mean <- as.numeric(imp_df$ph_y_anthr__weight_mean)
imp_df$ph_y_anthr__height_mean <- as.numeric(imp_df$ph_y_anthr__height_mean)
```

### 1.5 Calculate BMI

```{r bmi}
imp_df <- imp_df %>%
    mutate(bmi = (ph_y_anthr__weight_mean / (ph_y_anthr__height_mean^2)) * 703)

head(imp_df$bmi)
```

### 1.6 Create INR variable

```{r inr}
US_poverty_thresholds <- read.csv("US_poverty_thresholds.csv")

US_poverty_thresholds <- US_poverty_thresholds %>%
  mutate(wave_people = paste0(wave, "_", people))

income_mapping <- c(
  2500,   # <$5,000
  8500,   # $5,000–$11,999
  14000,  # $12,000–$15,999
  20500,  # $16,000–$24,999
  30000,  # $25,000–$34,999
  42500,  # $35,000–$49,999
  62500,  # $50,000–$74,999
  87500,  # $75,000–$99,999
  150000, # $100,000–$199,999
  300000  # $200,000+
)

imp_df <- imp_df %>%
    mutate(
      # Map ordinal income to midpoint
      mean_income = income_mapping[household_income],

      # Create merge key
      wave_people = paste0(wave, "_", ab_p_demo__roster_001)
    ) %>%

    # Join poverty thresholds
    left_join(US_poverty_thresholds, by = c("wave_people", "wave")) %>%

    # Calculate Income-to-Needs Ratio
    mutate(
      inr = mean_income / poverty_threshold
    )

names(imp_df)
head(imp_df$inr)
```

### 1.7 Create observation ID variable

```{r obs-id}
imp_df$obs_id <- seq_len(nrow(imp_df))
```

### 1.8  Rename analytic variables

```{r rename-analytic-variables}
imp_df <- imp_df %>%
    rename(
      family_id = ab_g_stc__design_id__fam,
      site_id = ab_g_dyn__design_site,
      adhd_traits = mh_p_cbcl__synd__attn_sum,
      internalising = mh_p_cbcl__synd__int_sum,
      externalising = mh_p_cbcl__synd__ext_sum
    )

names(imp_df)
```

### 1.9 Check variable types

```{r check-var-types}
check_vars <- c(
  "age",
  "bmi",
  "adhd_traits",
  "internalising",
  "externalising",
  "menarche_status_p",
  "breast_development_p",
  "inr",
  "ethnrace"
)

# Check class of these variables
sapply(imp_df[check_vars], class)

unique(imp_df$ethnrace)
```

## 2. Pre-analysis data checks

### 2.1 Histograms of continous variables

```{r histograms}
# Quick histograms
continuous_vars <- c(
    "age",
    "adhd_traits",
    "internalising",
    "externalising",
    "inr",
    "breast_development_p",
    "bmi"
)

sapply(imp_df[continuous_vars], function(x) sum(!is.na(x)))

continuous_vars[!continuous_vars %in% names(imp_df)]
```

### 2.2 Multicollinearity checks

```{r multicollinearity}
# Select continuous variables
cont_data <- imp_df %>% dplyr::select(all_of(continuous_vars))

# Compute correlation matrix (pairwise complete observations to handle any remaining NAs)
cor_matrix <- cor(cont_data, use = "pairwise.complete.obs")

# Print the correlation matrix
print(cor_matrix)

# Visualize with corrplot
corrplot::corrplot(cor_matrix, method = "color", type = "upper",
                   tl.col = "black", tl.srt = 45, addCoef.col = "black",
                   number.cex = 0.8)
```

### 2.3 Check cells for categorical interaction term

```{r categorical-cells-interaction}
with(
  imp_df,
  table(wave, adhd_status_life, menarche_status_p, useNA = "ifany")
)

with(
  imp_df,
  table(wave, adhd_status_pres, menarche_status_p, useNA = "ifany")
)
```

### 2.4 Scaling

```{r scaling}
imp_df <- imp_df %>%
  mutate(
    across(
      all_of(continuous_vars),
      ~ (.x - mean(.x, na.rm = TRUE)) / (2 * sd(.x, na.rm = TRUE)),
      .names = "{.col}_sc"
    )
  )

# Spot checks
mean(imp_df$inr_sc, na.rm = TRUE)
sd(imp_df$inr_sc, na.rm = TRUE)
mean(imp_df$age_sc, na.rm = TRUE)
sd(imp_df$age_sc, na.rm = TRUE)
mean(imp_df$breast_development_p_sc, na.rm = TRUE)
sd(imp_df$breast_development_p_sc, na.rm = TRUE)
summary(is.na(imp_df$breast_development_p_sc))
mean(imp_df$bmi_sc, na.rm = TRUE)
sd(imp_df$bmi_sc, na.rm = TRUE)
```

## 3. Save analytic df

```{r export-analytic-df}
write_csv(imp_df, "analytic_df.csv")
```

## 4. Wide analytic df

```{r wide-analytic-df}
# Pivot to wide format for longitudinal analyses
imp_df_wide <- imp_df %>%
  pivot_wider(
    id_cols = participant_id,
    names_from = wave,
    names_glue = "{.value}_{wave}",
    values_from = -c(participant_id, wave)
  )
```

### 4.1 Rename wide analytic variables

```{r rename-wide-vars}
imp_df_wide <- imp_df_wide %>%
  mutate(
    family_id = family_id_wave_0,
    site_id = site_id_wave_0,
    ethnrace = ethnrace_wave_0,
    adhd_status_life = adhd_status_life_wave_0,
    adhd_status_pres = adhd_status_pres_wave_0
  )
```

### 4.2 Wide variables scaling

```{r wide-scaling}
continuous_vars <- c(
    "age_wave_2",
    "adhd_traits_wave_0",
    "adhd_traits_wave_2",
    "internalising_wave_0",
    "externalising_wave_0",
    "inr_wave_0",
    "breast_development_p_wave_2",
    "bmi_wave_0"
)

imp_df_wide <- imp_df_wide %>%
  mutate(
    across(
      all_of(continuous_vars),
      ~ (.x - mean(.x, na.rm = TRUE)) / (2 * sd(.x, na.rm = TRUE)),
      .names = "{.col}_sc"
    )
  )

# Spot checks
mean(imp_df_wide$inr_wave_0_sc, na.rm = TRUE)
sd(imp_df_wide$inr_wave_0_sc, na.rm = TRUE)
mean(imp_df_wide$age_wave_2_sc, na.rm = TRUE)
sd(imp_df_wide$age_wave_2_sc, na.rm = TRUE)
sum(is.na(imp_df_wide$breast_development_p_wave_2_sc))
```

## 5. Save wide analytic df

```{r export-wide-analytic-df}
write_csv(imp_df_wide, "analytic_df_wide.csv")
```


