---
title: "ABCD ADHD menarche processing all other variables (step 3)"
author: "Emily Friedel"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r load libraries}
library(tidyverse)
library(dplyr)
```

## Read saved ADHD menarche dataframe 

```{r read/view menarche cleaned df}
adhd_menarche_clean_menstat <- read.csv("1_adhd_menarche_clean_menstat.csv")

head(adhd_menarche_clean_menstat)
```

### Copy and rename dataset for cleaning

```{r copy and rename for clean}
adhd_menarche_clean <- adhd_menarche_clean_menstat
```

## Pubertal Development Scale cleaning

### Change 777 and 999 to NA for parent and youth

```{r PDS 777 and 999 to NA}
### 777 and 999 to NA for parent and youth PDS 
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(across(c(pds_1_p, pds_2_p, pds_3_p, 
                  pds_f4_p, pds_f5b_p,
                  pds_ht2_y, pds_skin2_y, pds_bdyhair_y, 
                  pds_f4_2_y, pds_f6_y), 
                ~ na_if(., 999))) %>%
  mutate(across(c(pds_1_p, pds_2_p, pds_3_p, 
                  pds_f4_p, pds_f5b_p,
                  pds_ht2_y, pds_skin2_y, pds_bdyhair_y, 
                  pds_f4_2_y, pds_f6_y), 
                ~ na_if(., 777)))
```

### Breast development (parent report) summary statistics

```{r breast development summary}
# Summary statistics by wave
summary_stats_breast_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(pds_f4_p, na.rm = TRUE),
    median_score = median(pds_f4_p, na.rm = TRUE),
    sd_score = sd(pds_f4_p, na.rm = TRUE),
    min_score = min(pds_f4_p, na.rm = TRUE),
    max_score = max(pds_f4_p, na.rm = TRUE),
    n_valid = sum(!is.na(pds_f4_p))
  )

print(summary_stats_breast_by_wave)
```

### Check missing breast development (parent report) data

```{r NA missing breast development}
# Count missing breast development by wave
missing_bd_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    missing_pds_f4_p = sum(is.na(pds_f4_p)),
    total = n()
  ) %>%
  mutate(
    missing_pds_f4_p_percent = (missing_pds_f4_p / total) * 100
  )

# Print the results
print(missing_bd_by_wave)
```

### Check number of participants who have breast development that goes backwards (i.e., score is smaller than previous wave)

```{r bd backwards check}
# Create vector with number of participants with breast development that goes backwards
n_decreasing_bd <- adhd_menarche_clean %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  filter(!all(is.na(pds_f4_p))) %>%
  mutate(pds_diff = pds_f4_p - lag(pds_f4_p)) %>%
  summarise(has_decrease = any(pds_diff < 0, na.rm = TRUE)) %>%
  summarise(total_decreasing = sum(has_decrease))

# Print the total number
print(n_decreasing_bd)
```

## ADHD diagnosis processing

### Check NA and 888 for KSADS ADHD diagnosis at baseline and wave 1

```{r NA 888 adhd diagnosis}
# List of KSADS variables to check
ksads_vars <- c("ksads_14_853_p", "ksads_14_854_p", "ksads_14_856_p", "ksads_14_855_p")

# Count NA and 888 per variable, grouped by wave
adhd_menarche_clean %>%
  select(wave, all_of(ksads_vars)) %>%
  pivot_longer(
    cols = all_of(ksads_vars),
    names_to = "ksads_variable",
    values_to = "value"
  ) %>%
  group_by(wave, ksads_variable) %>%
  summarise(
    n_total = n(),
    n_NA = sum(is.na(value)),
    n_888 = sum(value == 888, na.rm = TRUE),
    .groups = "drop"
  )
```

### Count number of participants with KSADS ADHD variables missing across for waves 0, 1, 2

```{r count ADHD KSADS some missing}
# Number of participants with all KSDAS missing for any available waves
adhd_menarche_clean %>%
  filter(wave %in% c("wave_0", "wave_1", "wave_2")) %>%
  group_by(wave, src_subject_id) %>%
  summarise(
    all_ksads_missing = all(
      is.na(ksads_14_856_p) &
      is.na(ksads_14_855_p) &
      is.na(ksads_14_853_p) &
      is.na(ksads_14_854_p)
    ),
    .groups = "drop"
  ) %>%
  filter(all_ksads_missing) %>%
  count(wave, name = "n_participants_all_ksads_missing")
```

### Count the number that have all ksads missing for all of those waves

```{r all KSADS ADHD missing}
# Flag per wave whether all KSADS vars are missing
ksads_missing_flags <- adhd_menarche_clean %>%
  filter(wave %in% c("wave_0", "wave_1", "wave_2")) %>%
  group_by(src_subject_id, wave) %>%
  summarise(
    all_ksads_missing = all(
      is.na(ksads_14_856_p) &
      is.na(ksads_14_855_p) &
      is.na(ksads_14_853_p) &
      is.na(ksads_14_854_p)
    ),
    .groups = "drop"
  )

# Check for participants where all three waves are missing
fully_missing_participants <- ksads_missing_flags %>%
  group_by(src_subject_id) %>%
  summarise(missing_all_3_waves = all(all_ksads_missing), .groups = "drop") %>%
  filter(missing_all_3_waves)

# Count how many such participants
n_fully_missing_ksads <- nrow(fully_missing_participants)

# View result
print(n_fully_missing_ksads)
```

### Create ADHD diagnosis variable with consistent value across waves (0 = non-ADHD, 1 = ADHD) and convert 888 = NA

```{r adhd diagnosis variable}
# Identify participants with ADHD diagnosis (any item = 1 at waves 0–2)
adhd_1_ids <- adhd_menarche_clean %>%
  filter(wave %in% c("wave_0", "wave_1", "wave_2")) %>%
  filter(
    ksads_14_856_p == 1 |
    ksads_14_855_p == 1 |
    ksads_14_853_p == 1 |
    ksads_14_854_p == 1
  ) %>%
  pull(src_subject_id) %>%
  unique()

# Identify participants with all items = 0 at any wave (0–2), excluding ADHD positive IDs
adhd_0_ids <- adhd_menarche_clean%>%
  filter(wave %in% c("wave_0", "wave_1", "wave_2")) %>%
  filter(!src_subject_id %in% adhd_1_ids) %>%
  group_by(src_subject_id, wave) %>%
  summarise(
    all_zero = all(
      ksads_14_856_p == 0 &
      ksads_14_855_p == 0 &
      ksads_14_853_p == 0 &
      ksads_14_854_p == 0
    ),
    .groups = "drop"
  ) %>%
  filter(all_zero) %>%
  pull(src_subject_id) %>%
  unique()

# Assign ADHD diagnosis across all waves
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(adhd_diagnosis = case_when(
    src_subject_id %in% adhd_1_ids ~ 1,
    src_subject_id %in% adhd_0_ids ~ 0,
    TRUE ~ NA_real_
  ))
```

### Check new ADHD diagnosis column

```{r check adhd diganosis}
head(adhd_menarche_clean$adhd_diagnosis)
```

### Check missing for adhd diagnosis

```{r NA adhd diagnosis}
sum(is.na(adhd_menarche_clean$adhd_diagnosis))

total_adhd_diagnosis <- adhd_menarche_clean %>%
  filter(wave == "wave_0", adhd_diagnosis == 1) %>%
  count()

print(total_adhd_diagnosis)
```

## Check all waves still present

```{r check all waves}
unique(adhd_menarche_clean$wave)
```

## CBCL processing

### ADHD traits (CBCL attention) summary stats

```{r adhd traits summary}
# Summary statistics by wave
summary_stats_traits_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(cbcl_scr_syn_attention_r, na.rm = TRUE),
    median_score = median(cbcl_scr_syn_attention_r, na.rm = TRUE),
    sd_score = sd(cbcl_scr_syn_attention_r, na.rm = TRUE),
    min_score = min(cbcl_scr_syn_attention_r, na.rm = TRUE),
    max_score = max(cbcl_scr_syn_attention_r, na.rm = TRUE),
    n_valid = sum(!is.na(cbcl_scr_syn_attention_r))
  )

print(summary_stats_traits_by_wave)
```

### Check missing values for ADHD traits

```{r NA adhd traits}
# Count missing attention data by wave
missing_traits_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_traits = sum(is.na(cbcl_scr_syn_attention_r)),
            total = n(),
            missing_percent = (missing_traits / total) * 100)

# Print the results
print(missing_traits_by_wave)
```

### CBCL internalising summary

```{r internal summary}
# Summary statistics by wave
summary_stats_internalising_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(cbcl_scr_syn_internal_r, na.rm = TRUE),
    median_score = median(cbcl_scr_syn_internal_r, na.rm = TRUE),
    sd_score = sd(cbcl_scr_syn_internal_r, na.rm = TRUE),
    min_score = min(cbcl_scr_syn_internal_r, na.rm = TRUE),
    max_score = max(cbcl_scr_syn_internal_r, na.rm = TRUE),
    n_valid = sum(!is.na(cbcl_scr_syn_internal_r))
  )

print(summary_stats_internalising_by_wave)
```

### Check missing values for CBCL internalising

```{r NA cbcl internal}
# Count missing internalising data by wave
missing_internal_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_internal = sum(is.na(cbcl_scr_syn_internal_r)),
            total = n(),
            missing_percent = (missing_internal / total) * 100)

# Print the results
print(missing_internal_by_wave)
```

### CBCL externalising summary

```{r external summary}
# Summary statistics by wave
summary_stats_externalising_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(cbcl_scr_syn_external_r, na.rm = TRUE),
    median_score = median(cbcl_scr_syn_external_r, na.rm = TRUE),
    sd_score = sd(cbcl_scr_syn_external_r, na.rm = TRUE),
    min_score = min(cbcl_scr_syn_external_r, na.rm = TRUE),
    max_score = max(cbcl_scr_syn_external_r, na.rm = TRUE),
    n_valid = sum(!is.na(cbcl_scr_syn_external_r))
  )

print(summary_stats_externalising_by_wave)
```

### Check missing values for CBCL externalising

```{r NA cbcl external}
# Count missing externalising data by wave
missing_external_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_external = sum(is.na(cbcl_scr_syn_external_r)),
            total = n(),
            missing_percent = (missing_external / total) * 100)

# Print the results
print(missing_external_by_wave)
```

## Age processing

### Summary of age

```{r age summary}
# Summary statistics by wave
summary_stats_age_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(age_years, na.rm = TRUE),
    median_score = median(age_years, na.rm = TRUE),
    sd_score = sd(age_years, na.rm = TRUE),
    min_score = min(age_years, na.rm = TRUE),
    max_score = max(age_years, na.rm = TRUE),
    n_valid = sum(!is.na(age_years))
  )

print(summary_stats_age_by_wave)
```

### Check missing data for age

```{r NA age_years}
# Count missing age data by wave
missing_age_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_age = sum(is.na(age_years)),
            total = n(),
            missing_percent = (missing_age / total) * 100)

# Print the results
print(missing_age_by_wave)
```

## Processing for INR

Note that ABCD has income across all waves in 2 different variables so these need to be combined before summarising. The same applies for roster. 

### Income processing

### Check missing data for income by wave

```{r NA income by wave}
# Count missing income data by wave from v2
missing_income_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_income = sum(is.na(demo_comb_income_v2)),
            total = n(),
            missing_percent = (missing_income / total) * 100)

# Print the results
print(missing_income_by_wave)

# Count missing income data by wave from v2_l
missing_income_l_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_income_l = sum(is.na(demo_comb_income_v2_l)),
            total = n(),
            missing_percent = (missing_income_l / total) * 100)

# Print the results
print(missing_income_l_by_wave)
```

### Combine income variables so that income_v2 is used to fill in baseline and income_v2_l is used for all other waves and change 999, 777 to NA

```{r create new income variable}
adhd_menarche_clean <- adhd_menarche_clean %>% 
  mutate(income = ifelse(!is.na(demo_comb_income_v2), 
                                   demo_comb_income_v2, 
                                   demo_comb_income_v2_l))

# Count occurrences of 777 and 999 in the income column
count_777 <- sum(adhd_menarche_clean$income == 777, na.rm = TRUE)
count_999 <- sum(adhd_menarche_clean$income == 999, na.rm = TRUE)

# Display the counts
count_777
count_999
```

### Convert 999 and 777 to NA in income variable

```{r income 999/777 to NA}
adhd_menarche_clean <- adhd_menarche_clean %>% 
  mutate(income = ifelse(income >= 1 & income <= 10, 
                                  income, 
                                  NA))

head(adhd_menarche_clean$income)
```

### Summary of income

```{summary income}
# Summary statistics by wave
summary_stats_income_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(income, na.rm = TRUE),
    median_score = median(income, na.rm = TRUE),
    sd_score = sd(income, na.rm = TRUE),
    min_score = min(income, na.rm = TRUE),
    max_score = max(income, na.rm = TRUE)
  )

print(summary_stats_income_by_wave)
```

### Final check missing income by wave

```{r final check NA income by wave}
# Count missing income data by wave
missing_income_2_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_income_2 = sum(is.na(income)),
            total = n(),
            missing_percent = (missing_income_2 / total) * 100)

# Print the results
print(missing_income_2_by_wave)
```

### Roster processing

### Check missing household roster by wave 

```{r check NA roster by wave}
# Count missing roster data by wave
missing_roster_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_roster = sum(is.na(demo_roster_v2)),
            total = n(),
            missing_percent = (missing_roster / total) * 100)

# Print the results
print(missing_roster_by_wave)

# Count missing roster l data by wave
missing_roster_l_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_roster_l = sum(is.na(demo_roster_v2_l)),
            total = n(),
            missing_percent = (missing_roster_l / total) * 100)

# Print the results
print(missing_roster_l_by_wave)
```

### Combine roster variables so that roster_v2 is used to fill in baseline and roster_v2_l is used for all other waves

```{r create roster variable}
adhd_menarche_clean <- adhd_menarche_clean %>% 
  mutate(roster = ifelse(!is.na(demo_roster_v2), 
                                   demo_roster_v2, 
                                   demo_roster_v2_l))
      
head(adhd_menarche_clean$roster)
```

### Summary roster

```{r summary roster}
# Summary statistics by wave
summary_stats_roster_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(roster, na.rm = TRUE),
    median_score = median(roster, na.rm = TRUE),
    sd_score = sd(roster, na.rm = TRUE),
    min_score = min(roster, na.rm = TRUE),
    max_score = max(roster, na.rm = TRUE),
    n_valid = sum(!is.na(roster))
  )

print(summary_stats_roster_by_wave)
```

### Final check missing roster by wave

```{r final check missing roster by wave}
# Count missing income data by wave
missing_roster_2_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_roster_2 = sum(is.na(roster)),
            total = n(),
            missing_percent = (missing_roster_2 / total) * 100)

# Print the results
print(missing_roster_2_by_wave)
```

## Create INR variable

### Step 1 - convert coded income to mean for each income bracket

```{r create inr variable step 1}
# Define the income mapping
income_mapping <- c(
  2500,   # Midpoint of <$5,000
  8500,   # Midpoint of $5,000–$11,999
  14000,  # Midpoint of $12,000–$15,999
  20500,  # Midpoint of $16,000–$24,999
  30000,  # Midpoint of $25,000–$34,999
  42500,  # Midpoint of $35,000–$49,999
  62500,  # Midpoint of $50,000–$74,999
  87500,  # Midpoint of $75,000–$99,999
  150000, # Midpoint of $100,000–$199,999
  300000  # Capped value for $200,000+
)

# Mutate the dataframe to include the mean income
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(mean_income = income_mapping[income])

# View the updated dataframe
head(adhd_menarche_clean$mean_income)
```

### Households over 15 people

```{r households over 15}
# Count households with more than 20 people
households_over_15 <- adhd_menarche_clean %>%
  filter(demo_roster_v2_l > 15) %>%
  count()
# Display the count
print(households_over_15)
```

### Create wave_people variable in dataframe, exclude 2 x households over 15

```{r create wave_people variable}
# Create a new column for wave_people
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(
    wave_people = case_when(
      roster <= 15 ~ paste0(wave, "_", roster),
      TRUE ~ NA_character_ # Assign NA for households with more than 15 people
    )
  )

# View the updated dataframe
head(adhd_menarche_clean$wave_people)
```

### Import US_poverty_thresholds dataframe

```{r import poverty df}
US_poverty_thresholds <- read.csv("US_poverty_thresholds.csv")

# Create a new column combining wave and people
US_poverty_thresholds <- US_poverty_thresholds %>%
  mutate(wave_people = paste0(wave, "_", people))

# Print a preview of the updated dataset
head(US_poverty_thresholds)
```

### Join poverty threshold from US_poverty_thresholds based on wave_people

```{r merge poverty based on wave_people}
adhd_menarche_clean <- adhd_menarche_clean %>%
  left_join(US_poverty_thresholds, by = c("wave_people", "wave"))

# View the first few rows of the updated dataframe
head(adhd_menarche_clean)
```

## Create income-to-needs variable

```{r create inr}
# Calculate Income-to-Needs Ratio (INR)
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(
    inr = mean_income / poverty_threshold
  )

# View the first few rows of the updated dataframe
head(adhd_menarche_clean$inr)
```

### Summary inr

```{r summary inr}
# Summary statistics by wave
summary_stats_inr_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(inr, na.rm = TRUE),
    median_score = median(inr, na.rm = TRUE),
    sd_score = sd(inr, na.rm = TRUE),
    min_score = min(inr, na.rm = TRUE),
    max_score = max(inr, na.rm = TRUE),
    n_valid = sum(!is.na(inr))
  )

print(summary_stats_inr_by_wave)
```

### Check missing inr data

```{r NA inr}
# Count missing height data by wave
missing_inr_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_inr = sum(is.na(inr)),
            total = n(),
            missing_percent = (missing_inr / total) * 100)

# Print the results
print(missing_inr_by_wave)
```

### ADI processing

### Summary ADI

```{r summary adi}
# Summary statistics by wave
summary_stats_adi_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(reshist_addr1_adi_wsum, na.rm = TRUE),
    median_score = median(reshist_addr1_adi_wsum, na.rm = TRUE),
    sd_score = sd(reshist_addr1_adi_wsum, na.rm = TRUE),
    min_score = min(reshist_addr1_adi_wsum, na.rm = TRUE),
    max_score = max(reshist_addr1_adi_wsum, na.rm = TRUE)
  )

print(summary_stats_adi_by_wave)
```

### Check missing ADI data by wave 

```{r NA adi}
# Count missing led_l_adi_wsum data by wave
missing_adi_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_adi = sum(is.na(reshist_addr1_adi_wsum)),
            total = n(),
            missing_percent = (missing_adi / total) * 100)

# Print the results
print(missing_adi_by_wave)
```

### Create ADI variable consistent across all waves

```{r create adi variable}
# Propagate reshist_addr1_adi_wsum from wave 0 across all waves for each participant
adhd_menarche_clean <- adhd_menarche_clean %>%
  group_by(src_subject_id) %>%
  mutate(adi = first(reshist_addr1_adi_wsum)) %>%
  ungroup()

head(adhd_menarche_clean$adi)
```

### Check adi missing by wave again

```{r final check NA adi}
# Count missing led_l_adi_wsum data by wave
missing_adi_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_adi = sum(is.na(adi)),
            total = n(),
            missing_percent = (missing_adi / total) * 100)

print(missing_adi_by_wave)
```

## Create stimulant use variable 

### Step 1: search for stimulants in medication columns and create stimulant_meds vector 

```{r search for stimulants}
# list of stimulant medications
stimulant_meds <- c(
  "adderall",
  "adzenys",
  "amphetamine",
  "concerta",
  "daytrana",
  "desoxyn",
  "dexedrine",
  "dexmethylphenidate",
  "dextroamphetamine",
  "dyanavel",
  "evekeo",
  "focalin",
  "lisdexamfetamine",
  "metadate",
  "methamphetamine",
  "methylin",
  "methylphenidate",
  "mydayis",
  "procentra",
  "quillichew",
  "quillivant",
  "ritalin",
  "vyvanse",
  "zenzedi"
)
```

### Check all unique medications identified in dataset

```{r check identified stimulants}
# Ensure stimulant_meds are all lowercase
stimulant_meds <- tolower(stimulant_meds)

# Identify unique stimulant medications in the dataset (case-insensitive)
detected_stimulants <- unique(
  unlist(
    apply(
      adhd_menarche_clean[, paste0("med", 1:15, "_rxnorm_p")],
      1,
      function(row) {
        row_lower <- tolower(row)  # Lowercase row for case-insensitive matching
        meds_found <- sapply(stimulant_meds, function(med) {
          grep(med, row_lower, fixed = TRUE, value = TRUE)
        })
        unlist(meds_found)
      }
    )
  )
)

# Remove any empty strings from detected list
detected_stimulants <- detected_stimulants[detected_stimulants != ""]

# Print results
print(detected_stimulants)
```

### Create new stimulant_use variable, 1 = yes and 0 = no to stimulant use

```{r create stimulant_use}
# create a new column indicating stimulant medication use
adhd_menarche_clean$stimulant_use <- apply(
  adhd_menarche_clean[, paste0("med", 1:15, "_rxnorm_p")], 
  1, 
  function(row) ifelse(
    any(sapply(stimulant_meds, function(med) any(grepl(med, tolower(row), fixed = TRUE)))), 
    1, 
    0
  )
)

# check the first few rows to confirm the new column
head(adhd_menarche_clean[, c(paste0("med", 1:15, "_rxnorm_p"), "stimulant_use")])
```

### Count numbers of participants using stimulants at each wave

```{r stimulant count by wave}
# Count stimulant_use (1 = yes, 0 = no) by wave
stimulant_use_counts <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    stimulant_use_yes = sum(stimulant_use == 1, na.rm = TRUE),
    stimulant_use_no = sum(stimulant_use == 0, na.rm = TRUE),
    total = n()
  )

# View the counts
stimulant_use_counts
```

### Check why waves 3 and 4 have 0 stimulant use

```{r check waves 3 4 stimulants}
# Check missing meds data for waves 3 & 4
adhd_menarche_clean %>%
  filter(wave %in% c("wave_3", "wave_4")) %>%
  group_by(wave) %>%
  summarise(across(paste0("med", 1:15, "_rxnorm_p"), 
                   list(
                     missing_n = ~ sum(is.na(.)),
                     missing_pct = ~ round(mean(is.na(.)) * 100, 2)
                   )
  ))

```

Seems like medication data were not recorded at waves 3 and 4

## Cleaning for BMI

### Check missing data for calculated height and weight variables

```{r NA height and weight}
# Count missing height data by wave
missing_height_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_height = sum(is.na(anthroheightcalc)),
            total = n(),
            missing_percent = (missing_height / total) * 100)

# Print the results
print(missing_height_by_wave)


missing_weight_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_weight = sum(is.na(anthroweightcalc)),
            total = n(),
            missing_percent = (missing_weight / total) * 100)

# Print the results
print(missing_weight_by_wave)
```

### Check missing height for non-calculated columns

```{r check NA height non calc}
# Count missing height data by wave
missing_height1_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_height1 = sum(is.na(anthro_1_height_in)),
            total = n(),
            missing_percent = (missing_height1 / total) * 100)

# Print the results
print(missing_height1_by_wave)

missing_height2_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_height2 = sum(is.na(anthro2heightin)),
            total = n(),
            missing_percent = (missing_height2 / total) * 100)

# Print the results
print(missing_height2_by_wave)

missing_height3_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_height3 = sum(is.na(anthro3heightin)),
            total = n(),
            missing_percent = (missing_height3 / total) * 100)

# Print the results
print(missing_height3_by_wave)
```

### Check missing weight non-calculated variables

```{r check NA weight non calc}
# Count missing height data by wave
missing_weight1_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_weight1 = sum(is.na(anthroweight1lb)),
            total = n(),
            missing_percent = (missing_weight1 / total) * 100)

# Print the results
print(missing_weight1_by_wave)

missing_weight2_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_weight2 = sum(is.na(anthroweight2lb)),
            total = n(),
            missing_percent = (missing_weight2 / total) * 100)

# Print the results
print(missing_weight2_by_wave)

missing_weight3_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_weight3 = sum(is.na(anthroweight3lb)),
            total = n(),
            missing_percent = (missing_weight3 / total) * 100)

# Print the results
print(missing_weight3_by_wave)
```

## Create BMI variable

```{r create BMI variable}
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(bmi = (anthroweightcalc / (anthroheightcalc^2)) * 703)

head(adhd_menarche_clean$bmi)
```

### Summary BMI by wave 

```{r summary bmi}
# Summary BMI statistics by wave
summary_stats_bmi_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(bmi, na.rm = TRUE),
    median_score = median(bmi, na.rm = TRUE),
    sd_score = sd(bmi, na.rm = TRUE),
    min_score = min(bmi, na.rm = TRUE),
    max_score = max(bmi, na.rm = TRUE),
    n_valid = sum(!is.na(bmi)),
    n_missing = sum(is.na(bmi)),
    total = n(),
    missing_percent = (sum(is.na(bmi)) / n()) * 100
  )

print(summary_stats_bmi_by_wave)
```

### Check highly unlikely BMI values at baseline

```{r check bmi baseline <10}
# Filter for wave_0 and BMI < 10, then count
n_bmi_under_10_wave0 <- adhd_menarche_clean %>%
  filter(wave == "wave_0", bmi < 10) %>%
  summarise(count = n())

# Print the result
print(n_bmi_under_10_wave0)
```

### Convert bmi < 10 to NA

```{r convert bmi <10 to NA}
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(bmi = if_else(wave == "wave_0" & bmi < 10, NA_real_, bmi))
```

### Create baseline bmi variable

```{r create baseline bmi variable}
adhd_menarche_clean <- adhd_menarche_clean %>%
  group_by(src_subject_id) %>%
  mutate(bmi_baseline = first(na.omit(bmi[wave == "wave_0"]))) %>%
  ungroup()

head(adhd_menarche_clean$bmi_baseline)
```

## Race/ethnicity processing

### Create ethnicity variable and fill in for all waves for each participant

```{r create ethnicity variable}
adhd_menarche_clean <- adhd_menarche_clean %>%
  mutate(ethnicity = race_ethnicity) %>% 
  group_by(src_subject_id) %>%
  mutate(ethnicity = ifelse(all(is.na(ethnicity)), 
                                 NA_real_, max(ethnicity, na.rm = TRUE))) %>%
  ungroup()

head(adhd_menarche_clean$ethnicity)
```

### Race/ethnicity missing

```{r NA ethnicity}
# Count missing menarche data by wave
missing_ethnicity_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_ethnicity = sum(is.na(ethnicity)),
            total = n(),
            missing_percent = (missing_ethnicity / total) * 100)

# Print the results
print(missing_ethnicity_by_wave)
```

## Family and side id processing

### Create family and site id variables and propagate across all waves

```{r family/site id variables}
adhd_menarche_clean <- adhd_menarche_clean %>%
  group_by(src_subject_id) %>%
  mutate(family_id = first(na.omit(rel_family_id[wave == "wave_0"]))) %>%
  ungroup()

adhd_menarche_clean <- adhd_menarche_clean %>%
  group_by(src_subject_id) %>%
  mutate(site_id = first(na.omit(site_id_l[wave == "wave_0"]))) %>%
  ungroup()
```

## Check structure of final cleaned dataset

```{r check final df}
str(adhd_menarche_clean)
```

### Export copy of fully cleaned df

```{r export clean df}
write.csv(adhd_menarche_clean, file = "2_adhd_menarche_clean.csv", row.names = FALSE)
```
