---
title: "ABCD ADHD menarche processing menarche data (step 2)"
author: "Emily Friedel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r load libraries}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Read saved raw ADHD menarche dataframe and check

```{r read/view raw df}
adhd_menarche_raw <- read.csv("0_adhd_menarche_raw.csv")

head(adhd_menarche_raw)
```

### Copy and rename dataset for cleaning menarche status

```{r copy and rename for clean}
adhd_menarche_clean_menstat <- adhd_menarche_raw
```

### Check for duplicates

```{r check duplicates}
sum(duplicated(adhd_menarche_clean_menstat))
```

### Second check for duplicates

```{r second duplicates check}
# Identify participants with any duplicated eventname
duplicates <- adhd_menarche_clean_menstat %>%
  group_by(src_subject_id) %>%
  filter(duplicated(eventname) | duplicated(eventname, fromLast = TRUE))

# View the duplicated rows
print(duplicates)
```

### Keep only AFAB participants

```{r drop non females}
adhd_menarche_clean_menstat <- adhd_menarche_clean_menstat %>%
  group_by(src_subject_id) %>%
  mutate(
    consistent_sex = first(demo_sex_v2)
  ) %>%
  ungroup() %>%
  filter(consistent_sex == 2)
```

### select waves and recode using mutate

```{r select recode waves}
adhd_menarche_clean_menstat <- adhd_menarche_clean_menstat %>%
  filter(eventname %in% c("baseline_year_1_arm_1", 
                          "1_year_follow_up_y_arm_1", 
                          "2_year_follow_up_y_arm_1", 
                          "3_year_follow_up_y_arm_1", 
                          "4_year_follow_up_y_arm_1")) %>%
  mutate(wave = recode_factor(eventname,
                "baseline_year_1_arm_1" = "wave_0",
                "1_year_follow_up_y_arm_1" = "wave_1",
                "2_year_follow_up_y_arm_1" = "wave_2",
                "3_year_follow_up_y_arm_1" = "wave_3",
                "4_year_follow_up_y_arm_1" = "wave_4"),
         wave = factor(wave, levels = c("wave_0", "wave_1", "wave_2", "wave_3", "wave_4")))
```

### Create age in years variable

```{r create age in years}
adhd_menarche_clean_menstat <- adhd_menarche_clean_menstat %>%
  mutate(age_years = interview_age/12)
```

## BEGIN PARENT MENARCHE STATUS CLEANING

### Create new menarche status variables (youth and parent)

Note that menarche status needs to remain as a character for cleaning code to work properly

```{r recode and rename y menarche status}
# Create menarche status (youth report) variable column 1 = N, 4 = Y, other values = NA
adhd_menarche_clean_menstat <- adhd_menarche_clean_menstat %>%
  mutate(menarche_status_y = recode(as.character(pds_f5_y), 
                                    `1` = "N", `4` = "Y",
                                    .default = NA_character_))
# Create menarche status (parent report) variable column 1 = N, 4 = Y, other values = NA
adhd_menarche_clean_menstat <- adhd_menarche_clean_menstat %>%
  mutate(menarche_status_p = recode(as.character(pds_f5b_p), 
                                    `1` = "N", `4` = "Y",
                                    .default = NA_character_))
```

### Check base dataframe is ready for for cleaning

```{r check stucture clean df}
str(adhd_menarche_clean_menstat)
```

### Check missing values for menarche status (parent) by wave

```{r check NA menarche status}
# Count missing menarche status data by wave
missing_menarche_by_wave <- adhd_menarche_clean_menstat %>%
  group_by(wave) %>%
  summarise(missing_menarche = sum(is.na(menarche_status_p)),
            total = n(),
            missing_percent = (missing_menarche / total) * 100)

# Print the results
print(missing_menarche_by_wave)
```

### Visualise menarche status (parent) by wave

```{r bar graph status by wave}
# Create a bar graph for menarche status by wave
ggplot(adhd_menarche_clean_menstat, aes(x = wave, fill = menarche_status_p)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Menarche Status by Wave",
    x = "Wave",
    y = "Count",
    fill = "Menarche Status"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("Y" = "lightblue", "N" = "salmon"))
```

## IDENTIFY ALL PARTICIPANTS WITH MENARCHE STATUS DATA THAT CAN BE CORRECTED

This includes those with impossible changes in menarche status (going from post-menarche back to pre-menarche) and NAs that can be reasonably inferred.

```{r impossible change Y N}
# Identify participants with "Y" at one wave followed by "N" at any later wave
inconsistent_menarche_status_p <- adhd_menarche_clean_menstat %>%
  arrange(src_subject_id, wave) %>%
  filter(!is.na(menarche_status_p)) %>%
  group_by(src_subject_id) %>%
  summarise(
    has_Y_before_N = any(menarche_status_p == "Y" & lead(menarche_status_p) == "N"),
    .groups = "drop"
  ) %>%
  filter(has_Y_before_N) %>%
  select(src_subject_id) %>%
  distinct()

# Print the result
print(inconsistent_menarche_status_p)

# Create vector with subject IDs
inconsistent_menarche_participants <- inconsistent_menarche_status_p$src_subject_id
```

### Create df to visually inspect menarche status data for participants with Y N inconsistency and identify patterns that can be corrected, export as csv to visually inspect

```{r visual inspection menarche status inconsistencies df}
# Create dataframe of of participants with inconsistent menarche status including subject ID, wave, age, parent & youth menarche status reports, and parent and youth age at menarche report  
inconsistent_menarche_status_p_df <- adhd_menarche_clean_menstat %>%
  filter(src_subject_id %in% inconsistent_menarche_status_p$src_subject_id) %>%
  select(src_subject_id, wave, age_years, menarche_status_p, menarche_status_y, 
         pds_f6_y, pds_f6_p) 

# Export as csv to inspect
write.csv(inconsistent_menarche_status_p_df, file = "inconsistent_menarche_status_p_df.csv", row.names = FALSE)
```

### Check for patterns

Three patterns identified for inconsistent menarche status group:

1)  Participants with N at baseline and YN elsewhere across waves. In all these cases, the N in YN appears to be incorrect.
2)  Participants with Y at baseline followed by N at wave 1. Comparison of youth reports and parent reports of status and age at menarche suggests there are two patterns to correct: where the Y at baseline is correct and where it is incorrect.
3)  Participants with NA at baseline and YN elsewhere, either wave 1 or wave 2 needs to be corrected so Y is N

### Identify those with NA at baseline and N at wave 1

```{r baseline NA wave 1 N}
# Identify participants where baseline = NA and wave 1 = N
baseline_na_wave1_n <- adhd_menarche_clean_menstat %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  reframe(
    baseline_status = menarche_status_p[wave == "wave_0"], 
    wave1_status = menarche_status_p[wave == "wave_1"],
    .groups = "drop"
  ) %>%
  filter(is.na(baseline_status) & wave1_status == "N")

# View results
print(baseline_na_wave1_n)

# Create vector with participant ids
baseline_na_participants <- baseline_na_wave1_n$src_subject_id
```

### Identify N NA N and Y NA Y (sandwich NA status)

```{r sandwich NA status}
# Collapse the sequences for each participant into a single string
pattern_data <- adhd_menarche_clean_menstat %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  summarise(pattern = paste(menarche_status_p, collapse = "_"), .groups = 'drop')

# Identify participants who have the exact "N_NA_N" or "Y_NA_Y" patterns
sandwich_NA <- pattern_data %>%
  filter(grepl("(?<=^|_)(N_NA_N|Y_NA_Y)(?=$|_)", pattern, perl = TRUE)) %>%
  select(src_subject_id, pattern)

# Print the participants with the identified pattern
print(sandwich_NA)

# Create vector with only participant ids
sandwich_NA_participants <- sandwich_NA$src_subject_id
```

### Check unique patterns of sandwich NAs identified

```{r unique sandwich NA patterns}
# Get unique sandwich patterns
unique_sandwich_patterns <- sandwich_NA %>%
  distinct(pattern)

# View the unique sandwich patterns
print(unique_sandwich_patterns)
```

### Put all participants who can have menarche status corrected into a vector

```{r list participants to correct}
# Combine all subject IDs into one vector
participants_to_correct <- c(sandwich_NA_participants, 
                                inconsistent_menarche_participants, 
                                baseline_na_participants)

# Print the list of all unique subject IDs
str(participants_to_correct)

# Check for duplicates
sum(duplicated(participants_to_correct))
```

### Remove duplicates from list

```{r removed duplicated participants to correct}
participants_to_correct <- unique(participants_to_correct)

str(participants_to_correct)
```

### Extract rows for participants with menarche status data to correct from main df

```{r extract participants to correct}
# Extract data for participants needing corrections
participants_to_correct_data <- adhd_menarche_clean_menstat %>%
  filter(src_subject_id %in% participants_to_correct)

# View extracted data
str(participants_to_correct_data)

# Check correct no. of participants
length(unique(participants_to_correct_data$src_subject_id))
```

### Create correction type variable so that each participant can be assigned this to avoid correcting more than once

```{r create correction type}
# Create correction_type variable
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(correction_type = NA_character_)
```

## CORRECTION TYPES

### Correction 1: Identify participants with N at baseline, YN elsewhere

```{r correction 1}
# Identify participants with "N" at baseline
participants_with_N_baseline <- participants_to_correct_data %>%
  filter(wave == "wave_0" & menarche_status_p == "N") %>%
  pull(src_subject_id)

# Check for a "Y" followed by an "N" in later waves
correction_1_ids <- participants_to_correct_data %>%
  filter(src_subject_id %in% participants_with_N_baseline) %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  summarise(
    has_Y_before_N = any(menarche_status_p == "Y" & lead(menarche_status_p) == "N"),
    .groups = "drop"
  ) %>%
  filter(has_Y_before_N) %>%
  pull(src_subject_id)

# Assign "correction_1" only to those participants
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(correction_type = ifelse(src_subject_id %in% correction_1_ids, "correction_1", correction_type))

# Count unique participants who were assigned correction_1
num_correction_1 <- participants_to_correct_data %>%
  filter(correction_type == "correction_1") %>%
  summarise(unique_subjects = n_distinct(src_subject_id)) %>%
  pull(unique_subjects)

# Print the count
print(num_correction_1)
```

### Correction 2: Y at baseline, N at other waves

```{r correction 2}
# Identify participants with "Y" at baseline (wave_0)
participants_with_Y_baseline <- participants_to_correct_data %>%
  filter(wave == "wave_0" & menarche_status_p == "Y") %>%
  pull(src_subject_id)

# Check if they have "N" in any later wave
correction_2_ids <- participants_to_correct_data %>%
  filter(src_subject_id %in% participants_with_Y_baseline) %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  summarise(has_N_later = any(wave != "wave_0" & menarche_status_p == "N"), .groups = "drop") %>%
  filter(has_N_later) %>%
  pull(src_subject_id)

# Assign "correction_2" to these participants
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(correction_type = ifelse(is.na(correction_type) & src_subject_id %in% correction_2_ids, "correction_2", correction_type))

# Count unique participants who were assigned correction_2
num_correction_2 <- participants_to_correct_data %>%
  filter(correction_type == "correction_2") %>%
  summarise(unique_subjects = n_distinct(src_subject_id)) %>%
  pull(unique_subjects)

# Print the count
print(num_correction_2)
```

### Visual check of correction 2 participants

```{r view correction 2 participants}
correction_2_view <- participants_to_correct_data %>%
  filter(src_subject_id %in% correction_2_ids) %>%
  select(src_subject_id, wave, menarche_status_p, menarche_status_y,
         pds_f6_p, pds_f6_y, age_years, correction_type)

print(correction_2_view)
```

### Correction 3: NA at baseline and N at wave 1

```{r correction 3}
# Identify participants with NA at baseline (wave_0)
participants_with_NA_baseline <- participants_to_correct_data %>%
  filter(wave == "wave_0" & is.na(menarche_status_p)) %>%
  pull(src_subject_id)

# Check if they have "N" at wave 1
correction_3_ids <- participants_to_correct_data %>%
  filter(src_subject_id %in% participants_with_NA_baseline) %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  summarise(has_N_at_wave_1 = any(wave == "wave_1" & menarche_status_p == "N"), .groups = "drop") %>%
  filter(has_N_at_wave_1) %>%
  pull(src_subject_id)

# Assign "correction_3" to these participants
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(correction_type = ifelse(is.na(correction_type) & src_subject_id %in% correction_3_ids, "correction_3", correction_type))

# Count unique participants who were assigned correction_4
num_correction_3 <- participants_to_correct_data %>%
  filter(correction_type == "correction_3") %>%
  summarise(unique_subjects = n_distinct(src_subject_id)) %>%
  pull(unique_subjects)

# Print the count
print(num_correction_3)
```

### Correction 4: N NA N pattern

```{r correction 4}
# Collapse the menarche_status_p for each participant across waves
pattern_data <- participants_to_correct_data %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  summarise(pattern = paste(menarche_status_p, collapse = "_"), .groups = 'drop')

# Identify participants with the pattern "N_NA_N"
correction_4_ids <- pattern_data %>%
  filter(grepl("(?<=^|_)N_NA_N(?=$|_)", pattern, perl = TRUE)) %>%
  pull(src_subject_id)

# Assign "correction_4" only to these participants (if no correction already exists)
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(correction_type = ifelse(is.na(correction_type) & src_subject_id %in% correction_4_ids, "correction_4", correction_type))

# Count unique participants who were assigned correction_4
num_correction_4 <- participants_to_correct_data %>%
  filter(correction_type == "correction_4") %>%
  summarise(unique_subjects = n_distinct(src_subject_id)) %>%
  pull(unique_subjects)

# Print the count
print(num_correction_4)
```

### Correction 5: Y NA Y

```{r correction 5}
# Collapse the menarche_status_p for each participant across waves
pattern_data <- participants_to_correct_data %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  summarise(pattern = paste(menarche_status_p, collapse = "_"), .groups = 'drop')

# Identify participants with the pattern "Y_NA_Y"
correction_5_ids <- pattern_data %>%
  filter(grepl("(?<=^|_)Y_NA_Y(?=$|_)", pattern, perl = TRUE)) %>%
  pull(src_subject_id)

# Assign "correction_5" only to these participants (if no correction already exists)
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(correction_type = ifelse(is.na(correction_type) & src_subject_id %in% correction_5_ids, "correction_5", correction_type))

# Count unique participants who were assigned correction_5
num_correction_5 <- participants_to_correct_data %>%
  filter(correction_type == "correction_5") %>%
  summarise(unique_subjects = n_distinct(src_subject_id)) %>%
  pull(unique_subjects)

# Print the count
print(num_correction_5)
```

### Check number of participants with each correction type

```{r check correction type totals}
participants_to_correct_data %>%
  group_by(correction_type) %>%
  summarise(unique_subjects = n_distinct(src_subject_id))
```

## APPLY CORRECTIONS

### Apply Correction 1

```{r apply correction 1}
# Function to correct "Y N" to "Y Y" for participants in correction_1
correct_yn_pattern <- function(wave_status, participant_id) {
  if (participant_id %in% correction_1_ids) {  
    for (i in seq_len(length(wave_status) - 1)) {
      if (!is.na(wave_status[i]) && !is.na(wave_status[i + 1]) &&
          wave_status[i] == "Y" && wave_status[i + 1] == "N") {
        wave_status[i + 1] <- "Y"  # Change the "N" to "Y"
      }
    }
  }
  return(wave_status)
}

# Apply correction only to participants assigned correction_1
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(menarche_status_p = as.character(menarche_status_p)) %>%
  group_by(src_subject_id) %>%
  mutate(
    menarche_status_p = if (first(correction_type) == "correction_1") {
      correct_yn_pattern(menarche_status_p, first(src_subject_id))
    } else {
      menarche_status_p
    }
  ) %>%
  ungroup()
```

### Check correction 1 worked

```{r check correction 1}
# Check for a "Y" followed by an "N" in later waves, only for correction_1 participants
correction_1_check <- participants_to_correct_data %>%
  filter(correction_type == "correction_1") %>%
  arrange(src_subject_id, wave) %>% 
  group_by(src_subject_id) %>%
  summarise(
    has_Y_before_N = any(menarche_status_p == "Y" & lead(menarche_status_p) == "N"),
    .groups = "drop"
  ) %>%
  filter(has_Y_before_N) %>%
  pull(src_subject_id)

# Count how many participants still have the "Y N" pattern after correction
print(length(correction_1_check))
```

### Apply Correction 2 individual corrections

```{r apply correction 2}
# Apply individual corrections to menarche_status_p based on participant ID and wave
participants_to_correct_data <- participants_to_correct_data %>%
  mutate(menarche_status_p = case_when(
    src_subject_id %in% c("NDAR_INV0HL5D3DF", "NDAR_INVL48MK0DM") & wave == "wave_1" ~ "Y",
    src_subject_id %in% c("NDAR_INV1CYTVBWT", "NDAR_INVZ1WNHT63") & wave == "wave_2" ~ "Y",
    src_subject_id == "NDAR_INVB8G9KFMC" & wave %in% c("wave_0", "wave_1") ~ "N",
    src_subject_id %in% c("NDAR_INVBU31791C", "NDAR_INVDETP7ENE", "NDAR_INVGE3NT7ZK") & wave == "wave_0" ~ "N",
    src_subject_id %in% c("NDAR_INVECM7TBEL", "NDAR_INVLG5EUWL6", "NDAR_INVWMM5T2PL") & wave == "wave_3" ~ "Y",
    TRUE ~ menarche_status_p
  ))

# Print the modified rows to check the changes
print(participants_to_correct_data %>%
  filter(src_subject_id %in% c(
    "NDAR_INV0HL5D3DF", "NDAR_INVL48MK0DM",
    "NDAR_INV1CYTVBWT", "NDAR_INVZ1WNHT63",
    "NDAR_INVB8G9KFMC",
    "NDAR_INVBU31791C", "NDAR_INVDETP7ENE", "NDAR_INVGE3NT7ZK",
    "NDAR_INVECM7TBEL", "NDAR_INVLG5EUWL6", "NDAR_INVWMM5T2PL"
  )) %>%
  select(src_subject_id, wave, menarche_status_p))
```

### Check Correction 2 worked

```{r check correction 2}
# Identify participants with "Y" at baseline (wave_0)
participants_with_Y_baseline_check <- participants_to_correct_data %>%
  filter(wave == "wave_0" & menarche_status_p == "Y") %>%
  pull(src_subject_id)

# Check if they have "N" in any later wave
correction_2_check <- participants_to_correct_data %>%
  filter(src_subject_id %in% participants_with_Y_baseline_check) %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  summarise(has_N_later = any(wave != "wave_0" & menarche_status_p == "N"), .groups = "drop") %>%
  filter(has_N_later) %>%
  pull(src_subject_id)

# Count unique participants with correction_2
print(length(correction_2_check))
```

### Apply Correction 3

```{r apply correction 3}
# Correct baseline when baseline is NA and wave_1 is N, only for correction_3 participants
participants_to_correct_data <- participants_to_correct_data %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  mutate(
    menarche_status_p = ifelse(
      correction_type == "correction_3" & 
      wave == "wave_0" & is.na(menarche_status_p) &
      any(wave == "wave_1" & menarche_status_p == "N"),
      "N",
      menarche_status_p
    )
  ) %>%
  ungroup()
```

### Check Correction 3 worked

```{r check correction 3}
# Identify participants who still have NA at baseline when wave_1 is N
correction_3_check <- participants_to_correct_data %>%
  filter(correction_type == "correction_4") %>% 
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  summarise(
    still_na_at_baseline = any(wave == "wave_0" & is.na(menarche_status_p) &  
                               any(wave == "wave_1" & menarche_status_p == "N")),
    .groups = "drop"
  ) %>%
  filter(still_na_at_baseline) %>%
  pull(src_subject_id)

# Count the number of participants who still need correction
print(length(correction_3_check))
```

### Apply Correction 4

```{r apply correction 4}
# Apply correction for "N_NA_N" pattern across waves for participants needing correction 4
participants_to_correct_data <- participants_to_correct_data %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  mutate(
    menarche_status_p = case_when(
      correction_type == "correction_4" &
      is.na(menarche_status_p) &
      lag(menarche_status_p, default = "X") == "N" &
      lead(menarche_status_p, default = "X") == "N" ~ "N",
      TRUE ~ menarche_status_p
    )
  ) %>%
  ungroup()
```

### Check Correction 4 worked

```{r check correction 4 worked}
# Collapse the sequences for each participant into a single string
correction_4_check <- participants_to_correct_data %>%
  filter(correction_type == "correction_4") %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  summarise(pattern = paste(menarche_status_p, collapse = "_"), .groups = 'drop')

# Identify participants who have the exact "N_NA_N"
correction_4_check <- correction_4_check %>%
  filter(grepl("(?<=^|_)N_NA_N(?=$|_)", pattern, perl = TRUE)) %>%
  pull(src_subject_id)

# Print the number of participants who still have "N_NA_N"
print(length(correction_4_check))
```

### Apply Correction 5

```{r apply correction 5}
# Apply correction for "Y_NA_Y" pattern across waves for participants needing correction 5
participants_to_correct_data <- participants_to_correct_data %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  mutate(
    menarche_status_p = case_when(
      correction_type == "correction_5" &
      is.na(menarche_status_p) &
      lag(menarche_status_p, default = "X") == "Y" &
      lead(menarche_status_p, default = "X") == "Y" ~ "Y",
      TRUE ~ menarche_status_p
    )
  ) %>%
  ungroup()
```

### Check Correction 5 worked

```{r check correction 5 worked}
# Collapse the sequences for each participant into a single string
correction_5_check <- participants_to_correct_data %>%
  filter(correction_type == "correction_5") %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  summarise(pattern = paste(menarche_status_p, collapse = "_"), .groups = 'drop')

# Identify participants who have the exact subsequence "N_NA_N"
correction_5_check <- correction_5_check %>%
  filter(grepl("(?<=^|_)Y_NA_Y(?=$|_)", pattern, perl = TRUE)) %>%
  pull(src_subject_id)

# Print the number of participants who still have "N_NA_N"
print(length(correction_5_check))
```

## CHECK NUMBERS FOR ALL CORRECTIONS ORIGINALLY IDENTIFIED

### Check inconsistencies in menarche status (parent)

```{r check menarche status y inconsistencies 1}
# Check participants with inconsistent menarche status
inconsistent_menarche_status_p_check <- participants_to_correct_data %>%
  arrange(src_subject_id, wave) %>%
  filter(!is.na(menarche_status_p)) %>%
  group_by(src_subject_id) %>%
  mutate(
    prev_status = lag(menarche_status_p),
    inconsistent = ifelse(
      menarche_status_p == "N" & prev_status == "Y",
      TRUE,
      FALSE
    )
  ) %>%
  ungroup() %>% 
  filter(inconsistent == TRUE) %>%
  select(src_subject_id) %>%
  distinct()

# Display the result
print(inconsistent_menarche_status_p_check)
```

### Check those with NA at baseline and N at wave 1

```{r baseline NA wave 1 N check 1}
# Check participants where baseline = NA and wave 1 = N
baseline_na_wave1_n_check <- participants_to_correct_data %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  reframe(
    baseline_status = menarche_status_p[wave == "wave_0"], 
    wave1_status = menarche_status_p[wave == "wave_1"]
  ) %>%
  filter(is.na(baseline_status) & wave1_status == "N")

# View results
print(baseline_na_wave1_n_check)
```

### Check N NA N and Y NA Y (sandwich NA status)

```{r check sandwich NA status}
# Collapse the sequences for each participant into a single string
pattern_data_check <- participants_to_correct_data %>%
  arrange(src_subject_id, wave) %>% 
  group_by(src_subject_id) %>%
  summarise(pattern = paste(menarche_status_p, collapse = "_"), .groups = 'drop')

# Identify participants who have the exact "N_NA_N" or "Y_NA_Y" pattern
sandwich_NA_check <- pattern_data_check %>%
  filter(grepl("(?<=^|_)(N_NA_N|Y_NA_Y)(?=$|_)", pattern, perl = TRUE)) %>%
  select(src_subject_id, pattern)

# Print the participants with the identified pattern
print(sandwich_NA_check)
```

## RETURN CORRECTED PARTICIPANTS TO MAIN DF

```{r return corrected participant data to main df}
adhd_menarche_clean_menstat <- adhd_menarche_clean_menstat %>%
  left_join(
    participants_to_correct_data %>%
      select(src_subject_id, wave, menarche_status_p),
    by = c("src_subject_id", "wave"),
    suffix = c(".original", ".corrected")
  ) %>%
  mutate(
    menarche_status_p = coalesce(menarche_status_p.corrected, as.character(menarche_status_p.original)),
  ) %>%
  select(-menarche_status_p.original, -menarche_status_p.corrected)

# Check the updated dataframe
head(adhd_menarche_clean_menstat)
```

### Check menarche status variable

```{r check menarche status}
head(adhd_menarche_clean_menstat$menarche_status_p)
```

### Check numbers of corrections identified in main df align with what was in the corrected df after cleaning

### Check inconsistencies in menarche status (youth) in main df

```{r check menarche status y inconsistencies}
# Check participants with inconsistent menarche status
inconsistent_menarche_status_p_check_2 <- adhd_menarche_clean_menstat %>%
  arrange(src_subject_id, wave) %>%
  filter(!is.na(menarche_status_p)) %>%
  group_by(src_subject_id) %>%
  mutate(
    prev_status = lag(menarche_status_p),
    inconsistent = ifelse(
      menarche_status_p == "N" & prev_status == "Y",
      TRUE,
      FALSE
    )
  ) %>%
  ungroup() %>% 
  filter(inconsistent == TRUE) %>%
  select(src_subject_id) %>%
  distinct()

# Display the result
print(inconsistent_menarche_status_p_check_2)
```

### Check those with NA at baseline and N at wave 1

```{r baseline NA wave 1 N check}
# Check participants where baseline = NA and wave 1 = N
baseline_na_wave1_n_check_2 <- adhd_menarche_clean_menstat %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  reframe(
    baseline_status = menarche_status_p[wave == "wave_0"], 
    wave1_status = menarche_status_p[wave == "wave_1"],
    .groups = "drop"
  ) %>%
  filter(is.na(baseline_status) & wave1_status == "N")

# View results
print(baseline_na_wave1_n_check)
```

### Check N NA N and Y NA Y (sandwich NA status)

```{r check main df sandwich NA status}
# Collapse the sequences for each participant into a single string
pattern_data_check_2 <- adhd_menarche_clean_menstat %>%
  arrange(src_subject_id, wave) %>%
  group_by(src_subject_id) %>%
  summarise(pattern = paste(menarche_status_p, collapse = "_"), .groups = 'drop')

# Identify participants who have the exact subsequence "N_NA_N" or "Y_NA_Y"
sandwich_NA_check_2 <- pattern_data_check_2 %>%
  filter(grepl("(?<=^|_)(N_NA_N|Y_NA_Y)(?=$|_)", pattern, perl = TRUE)) %>%
  select(src_subject_id, pattern)

# Print the participants with the identified pattern
print(sandwich_NA_check_2)
```

### Check that there are no duplicates after including corrected data

```{r check corrected duplicates}
sum(duplicated(adhd_menarche_clean_menstat))
```

### Final check for duplicates in clean df

```{r second corrected duplicates check}
# Identify participants with duplicated eventname
duplicates_2 <- adhd_menarche_clean_menstat %>%
  group_by(src_subject_id) %>%
  filter(duplicated(eventname) | duplicated(eventname, fromLast = TRUE))

# View the duplicated rows
print(duplicates_2)
```

### Check reduction in missing menarche status

```{r final check NA menarche status}
# Count missing menarche data by wave
missing_menarche_by_wave_2 <- adhd_menarche_clean_menstat %>%
  group_by(wave) %>%
  summarise(missing_menarche = sum(is.na(menarche_status_p)),
            total = n(),
            missing_percent = (missing_menarche / total) * 100)

# Print the results
print(missing_menarche_by_wave_2)
```

### Create recoded numerical version of cleaned parent menarche status for imputation

```{r numerical menarche status p}
adhd_menarche_clean_menstat <- adhd_menarche_clean_menstat %>%
  mutate(menarche_status_pds = dplyr::recode((menarche_status_p), 
                                      "N" = 1, "Y" = 4))
```

### Final check of cleaned df

```{r final check clean df}
head(adhd_menarche_clean_menstat)
```

## Export copy of csv with clean menarche status (parent) data

```{r export df with cleaned menarche status}
write.csv(adhd_menarche_clean_menstat, file = "1_adhd_menarche_clean_menstat.csv", row.names = FALSE)
```
