---
title: "ABCD ADHD menarche raw dataset preparation (step 1)"
author: "Emily Friedel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r load libraries}
library(tidyverse)
library(dplyr)
```

## Make raw dataframe with all variables relevent to analyses

### Read all relevant ABCD files

```{r read abcd files}
abcd_p_demo <- read_csv("ABCD files/abcd_p_demo.csv", show_col_types = FALSE)
abcd_p_screen <- read_csv("ABCD files/abcd_p_screen.csv", show_col_types = FALSE)
abcd_y_lt <- read_csv("ABCD files/abcd_y_lt.csv", show_col_types = FALSE)
led_l_adi <- read_csv("ABCD files/led_l_adi.csv", show_col_types = FALSE)
mh_p_cbcl <- read_csv("ABCD files/mh_p_cbcl.csv", show_col_types = FALSE)
mh_p_ksads_ss <- read_csv("ABCD files/mh_p_ksads_ss.csv", show_col_types = FALSE)
ph_p_meds <- read_csv("ABCD files/ph_p_meds.csv", show_col_types = FALSE)
ph_p_pds <- read_csv("ABCD files/ph_p_pds.csv", show_col_types = FALSE)
ph_y_pds <- read_csv("ABCD files/ph_y_pds.csv", show_col_types = FALSE)
ph_y_anthro <- read_csv("ABCD files/ph_y_anthro.csv", show_col_types = FALSE)
```
### Check for parsing problems

```{r check parsing problems}
# Check for parsing problems
datasets <- list(
  abcd_p_demo = abcd_p_demo,
  abcd_p_screen = abcd_p_screen,
  abcd_y_lt = abcd_y_lt,
  led_l_adi = led_l_adi,
  mh_p_cbcl = mh_p_cbcl,
  mh_p_ksads_ss = mh_p_ksads_ss,
  ph_p_meds = ph_p_meds,
  ph_p_pds = ph_p_pds,
  ph_y_pds = ph_y_pds,
  ph_y_anthro = ph_y_anthro
)

# Run problems() on each and print only if issues exist
for (name in names(datasets)) {
  dat <- datasets[[name]]
  prob <- problems(dat)
  
  if (nrow(prob) > 0) {
    cat("\nParsing issues in", name, ":\n")
    
    # Add column names to the problems dataframe
    prob$col_name <- names(dat)[prob$col]
    
    # Print full problem details including column names
    print(as.data.frame(prob))
    
    # Print unique columns affected
    cat("\nColumns with parsing problems in", name, ":\n")
    print(unique(prob$col_name))
    
  } else {
    cat("\nNo parsing issues in", name, "\n")
  }
}
```

Note: No columns with parsing issues are relevant to our analyses

### Select columns for all relevant variables from various ABCD files

```{r select columns from abcd files}
demo_select <- abcd_p_demo %>%
  dplyr::select(src_subject_id, eventname, 
         demo_sex_v2, demo_comb_income_v2_l, demo_comb_income_v2, 
         demo_roster_v2_l, demo_roster_v2, race_ethnicity)

pds_p_select <- ph_p_pds %>%
  dplyr::select(src_subject_id, eventname,
         pds_1_p, pds_2_p, pds_3_p, pds_f4_p, pds_f5b_p, pds_f6_p)

pds_y_select <- ph_y_pds %>%
  dplyr::select(src_subject_id, eventname,
         pds_ht2_y, pds_skin2_y, pds_bdyhair_y, pds_f4_2_y, 
         pds_f5_y, pds_f6_y)

cbcl_select <- mh_p_cbcl %>%
  dplyr::select(src_subject_id, eventname, 
         cbcl_scr_syn_attention_r, cbcl_scr_syn_internal_r, 
         cbcl_scr_syn_external_r,
       )

ksads_select <- mh_p_ksads_ss %>%
  dplyr::select(src_subject_id, eventname, 
         ksads_14_853_p,ksads_14_854_p,
         ksads_14_856_p, ksads_14_855_p 
      )

anthro_select <- ph_y_anthro %>%
  dplyr::select(src_subject_id, eventname, 
         anthro_1_height_in, anthro2heightin, anthro3heightin, anthroheightcalc, 
         anthroweight1lb, anthroweight2lb, anthroweight3lb, anthroweightcalc)

lt_select <- abcd_y_lt %>%
  dplyr::select(src_subject_id, eventname, 
         rel_family_id, site_id_l, interview_age)

led_l_adi <- led_l_adi %>%
  dplyr::select(src_subject_id, eventname, reshist_addr1_adi_wsum)

meds_select <- ph_p_meds %>%
  dplyr::select(src_subject_id, eventname, 
         med1_rxnorm_p, med2_rxnorm_p, med3_rxnorm_p, med4_rxnorm_p, med1_rxnorm_p,
         med2_rxnorm_p, med3_rxnorm_p, med4_rxnorm_p, med5_rxnorm_p, med6_rxnorm_p,
         med7_rxnorm_p, med8_rxnorm_p, med9_rxnorm_p, med10_rxnorm_p, med11_rxnorm_p, 
         med12_rxnorm_p, med13_rxnorm_p, med14_rxnorm_p, med15_rxnorm_p)
```

### Join columns to make a data frame with all relevant variables

```{r join selected columns}
adhd_menarche_raw <- demo_select %>%
  full_join(pds_p_select, by = c("src_subject_id", "eventname")) %>%
  full_join(pds_y_select, by = c("src_subject_id", "eventname")) %>%
  full_join(cbcl_select, by = c("src_subject_id", "eventname")) %>%
  full_join(ksads_select, by = c("src_subject_id", "eventname")) %>%
  full_join(anthro_select, by = c("src_subject_id", "eventname")) %>%
  full_join(lt_select, by = c("src_subject_id", "eventname")) %>%
  full_join(led_l_adi, by = c("src_subject_id", "eventname")) %>%
  full_join(meds_select, by = c("src_subject_id", "eventname"))
```

### Check raw dataframe

```{r head raw df}
head(adhd_menarche_raw)
```

## Save raw dataframe

```{r export raw df}
write.csv(adhd_menarche_raw, "0_adhd_menarche_raw.csv", row.names = FALSE)
```
