---
title: "ABCD ADHD menarche imputed dataset"
format: html
editor: visual
---

## Load libraries

```{r Load Libraries}
# load libraries
library(tidyverse)
library(knitr)
library(tidyr)
library(readxl)
library(readr)
library(mice)
library(future)
library(future.apply)
```

## Import cleaned dataframe

```{r read cleaned df csv}
# Read csv
df <- read_csv("2_adhd_menarche_clean.csv", show_col_types = FALSE)
```

### Check structure of cleaned dataframe

```{r check structure clean df}
str(df)
```

### Select relevant variables

```{r select variables}
df <- df[c("src_subject_id", 
           "wave",
           "age_years",
           "menarche_status_p",
           "pds_1_p",
           "pds_2_p",
           "pds_3_p",
           "pds_f4_p",
           "pds_f6_p",
           "pds_ht2_y", 
           "pds_skin2_y", 
           "pds_bdyhair_y", 
           "pds_f4_2_y",
           "pds_f5_y",
           "menarche_status_pds",
           "cbcl_scr_syn_attention_r",
           "cbcl_scr_syn_internal_r",  
           "cbcl_scr_syn_external_r",
           "adhd_diagnosis", 
           "ethnicity",
           "adi",
           "inr",
           "income",
           "roster",
           "family_id",
           "site_id",
           "bmi_baseline",
           "stimulant_use")]
```

### Visual check of working dataframe

```{r check working df}
head(df)
```

## Imputation

### Perform imputations

```{r imputations}
df$pds_f4_p <- factor(df$pds_f4_p, levels = 1:4, ordered = TRUE)

# Identify variables with missing data, excluding some variables
excluded_vars <- c("wave", "src_subject_id", "family_id", "site_id", 
                   "menarche_status_p", "adhd_diagnosis", "stimulant_use")

vars_to_impute <- names(df)[
  sapply(df, function(x) any(is.na(x))) & !(names(df) %in% excluded_vars)
]

# Create method vector
methods <- make.method(df)
methods[] <- ""  # initialise all as excluded
methods[vars_to_impute] <- "pmm"  # default to PMM
methods["pds_f4_p"] <- "polr" # use ordinal logistic regression for breast development

# Define predictor matrix
predictor_matrix <- make.predictorMatrix(df)

# Prevent ID and wave variables from being used as predictors
predictor_matrix[, excluded_vars] <- 0

# Prevent variables from predicting themselves
diag(predictor_matrix) <- 0

# Set seed
set.seed(1234)

# Run imputation using futuremice (parallelised)
imp <- futuremice(
  df, 
  method = methods, 
  predictorMatrix = predictor_matrix, 
  m = 20, 
  n.core = 4, 
  n.imp.core = 1
)
```

### Average imputed datasets

```{r average imputed datasets}
# Extract all 20 completed datasets
all_imputed <- lapply(1:imp$m, function(i) complete(imp, i))

# Bind all datasets and track imputation number
for (i in seq_along(all_imputed)) {
  all_imputed[[i]]$.imp <- i
  # Convert ordered factors to numeric for averaging
  all_imputed[[i]]$pds_f4_p <- as.numeric(as.character(all_imputed[[i]]$pds_f4_p))
}

combined <- bind_rows(all_imputed)

# Identify variables to average
id_vars <- c("src_subject_id", "wave")
imputed_vars <- names(imp$method)[imp$method != ""]

# Average imputed values across imputations
averaged <- combined %>%
  group_by(across(all_of(id_vars))) %>%
  summarise(across(all_of(imputed_vars), ~mean(.x, na.rm = TRUE)), .groups = "drop")

# Get original dataset
original <- complete(imp, 0)

# Merge and replace missing values
imp_df <- left_join(original, averaged, by = id_vars, suffix = c("", "_avg"))

for (v in imputed_vars) {
  imp_df[[v]] <- ifelse(is.na(imp_df[[v]]), imp_df[[paste0(v, "_avg")]], imp_df[[v]])
  imp_df[[paste0(v, "_avg")]] <- NULL
}
```

### Check imputed dataframe structure

```{r check imputed df structure}
str(imp_df)
```

### Remove wave 4 data (only half collected)

```{r remove wave 4}
imp_df <- imp_df %>%
  filter(wave != "wave_4")
```

### Grand mean centre continuous predictors

```{r grand mean centre}
# Variables to grand mean center
vars_to_center <- c("age_years", "inr", "cbcl_scr_syn_attention_r",
                    "cbcl_scr_syn_internal_r", "cbcl_scr_syn_external_r",  
                    "pds_f4_p", "bmi_baseline")

# Compute grand means
grand_means <- imp_df %>%
  summarise(across(all_of(vars_to_center), \(x) mean(x, na.rm = TRUE)))

# Center the variables
imp_df <- imp_df %>%
  mutate(across(all_of(vars_to_center), 
                ~ . - grand_means[[cur_column()]], 
                .names = "{.col}_c"))

# View first few rows to check
head(imp_df)
```

### Check INR missingness

```{r NA inr}
# Count missing led_l_inr_wsum data by wave
missing_inr_imp_by_wave <- imp_df %>%
  group_by(wave) %>%
  summarise(missing_inr_c = sum(is.na(inr_c)),
            total = n(),
            missing_percent = (missing_inr_c / total) * 100)

# Print the results
print(missing_inr_imp_by_wave)
```

### Save imputed dataset

```{r export df}
write.csv(imp_df, "study_2_imputed_dataset.csv", row.names = FALSE)
```
