---
title: "ABCD ADHD menarche cross-sectional sensitivity analyses"
author: "Emily Friedel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Setup

### Load libraries

```{r libraries}
library(tidyverse)
library(ggplot2)
library(mice)
library(brms)
library(posterior)
library(bayestestR)
library(emmeans)
```

### Import saved imputed dfs

```{r import-data}
analytic_df <- read_csv("analytic_df.csv")
obs_df <- read_csv("obs_df.csv")
```

## Initial checks

```{r initial-checks}
length(analytic_df)
colnames(analytic_df)
summary(unique(obs_df$participant_id))
length(obs_df)
colnames(obs_df)
```

## 1. Pre model variable checks

### 1.1 Convert to factors: subject ID, site ID, family ID, ethnicity, and menarche status

```{r factors}
analytic_df <- analytic_df %>%
  mutate(across(c(participant_id, family_id, site_id, ethnrace, menarche_status_p, adhd_status_life), as.factor))

# set white_nh as reference for enthrace
analytic_df$ethnrace <- relevel(analytic_df$ethnrace, ref = "white_nh")
levels(analytic_df$ethnrace)

obs_df <- obs_df %>%
  mutate(across(c(participant_id, family_id, site_id, ethnrace, menarche_status_p, adhd_status_life), as.factor))

# collapse "asian" into "other" for ethnrace
obs_df$ethnrace <- forcats::fct_collapse(
  obs_df$ethnrace,
  other = c("asian", "other")
)

# order levels alphabetically, then set white_nh as reference
obs_df$ethnrace <- forcats::fct_relevel(
  obs_df$ethnrace,
  "white_nh", "black_nh", "hispanic", "other"
)
levels(obs_df$ethnrace)
```

### 1.2 Convert to integers: all outcomes

```{r outcomes-as-integers}
analytic_df <- analytic_df %>%
  mutate(across(c(adhd_traits, internalising, externalising), as.integer))

obs_df <- obs_df %>%
  mutate(across(c(adhd_traits, internalising, externalising), as.integer))
```

### 1.2 Create observation ID variable for observed dataset

```{r obs id}
obs_df$obs_id <- seq_len(nrow(obs_df))
```

### 1.3 Create baseline BMI variable across waves

```{r baseline-bmi}
analytic_df <- analytic_df %>%
  group_by(participant_id) %>%
  mutate(bmi_wave_0_sc =
           min(if_else(wave == "wave_0", bmi_sc, NA), na.rm = TRUE)
         )
head(analytic_df$bmi_wave_0_sc)
```

### 1.4 Structure of dataframes

```{r df-structure}
str(analytic_df)
str(obs_df)
```

## 2. Set up hypothesis formulas and prior

```{r hypothesis-formulas}

formula_2_adhd_pres <- bf(internalising ~ menarche_status_p * adhd_status_pres
                + age_sc + ethnrace + inr_sc
                + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                + (1 | obs_id)
                )

formula_2b_adhd_pres <- bf(internalising ~ breast_development_p_sc * adhd_status_pres
                 + age_sc + ethnrace + inr_sc
                 + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                 + (1 | obs_id)
                 )

formula_3_adhd_pres <- bf(externalising ~ menarche_status_p * adhd_status_pres
                + age_sc + ethnrace + inr_sc
                + (1 | participant_id) + (1 | family_id) + (1 | site_id) 
                + (1| obs_id)
                )

formula_3b_adhd_pres <- bf(externalising ~ breast_development_p_sc * adhd_status_pres
                 + age_sc + ethnrace + inr_sc 
                 + (1 | participant_id) + (1 | family_id) + (1 | site_id) 
                 + (1| obs_id)
                 )

formula_1b_obs <- bf(adhd_traits ~ breast_development_p_sc
                  + age_sc + ethnrace + inr_sc
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )

formula_4_obs <- bf(internalising ~ menarche_status_p * adhd_traits_sc
                  + age_sc + ethnrace + inr_sc
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )

formula_5_obs <- bf(externalising ~ menarche_status_p * adhd_traits_sc
                  + age_sc + ethnrace + inr_sc
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )

formula_1b_bmi <- bf(adhd_traits ~ breast_development_p_sc
                  + age_sc + ethnrace + inr_sc + bmi_wave_0_sc
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )

formula_4_bmi <- bf(internalising ~ menarche_status_p * adhd_traits_sc
                  + age_sc + ethnrace + inr_sc + bmi_wave_0_sc
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )

formula_5_bmi <- bf(externalising ~ menarche_status_p * adhd_traits_sc
                  + age_sc + ethnrace + inr_sc + bmi_wave_0_sc
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )

formula_1b_stimulant <- bf(adhd_traits ~ breast_development_p_sc
                  + age_sc + ethnrace + inr_sc + stimulant_status
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )

formula_4_stimulant <- bf(internalising ~ menarche_status_p * adhd_traits_sc
                  + age_sc + ethnrace + inr_sc + stimulant_status
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )

formula_5_stimulant <- bf(externalising ~ menarche_status_p * adhd_traits_sc
                  + age_sc + ethnrace + inr_sc + stimulant_status
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )

formula_1b_prior1 <- bf(adhd_traits ~ breast_development_p_sc
                  + age_sc + ethnrace + inr_sc
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )

formula_4_prior1 <- bf(internalising ~ menarche_status_p * adhd_traits_sc
                  + age_sc + ethnrace + inr_sc 
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )

formula_5_prior1 <- bf(externalising ~ menarche_status_p * adhd_traits_sc
                  + age_sc + ethnrace + inr_sc 
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )


formula_1b_prior2 <- bf(adhd_traits ~ breast_development_p_sc
                  + age_sc + ethnrace + inr_sc
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )

formula_4_prior2 <- bf(internalising ~ menarche_status_p * adhd_traits_sc
                  + age_sc + ethnrace + inr_sc 
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )

formula_5_prior2 <- bf(externalising ~ menarche_status_p * adhd_traits_sc
                  + age_sc + ethnrace + inr_sc 
                  + (1 | participant_id) + (1 | family_id) + (1 | site_id)
                  + (1| obs_id)
                  )

# Set prior
prior <- prior(normal(0, 1),class = "b")
prior1 <- prior(normal(0, 0.25), class = "b")
prior2 <- prior(normal(0, 10), class = "b")
```

### 2.1 Check default prior

```{r default-prior}
get_prior(
  formula = formula_2_adhd_pres,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_2b_adhd_pres,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_3_adhd_pres,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_3b_adhd_pres,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1b_obs,
  data = obs_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_4_obs,
  data = obs_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5_obs,
  data = obs_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1b_bmi,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_4_bmi,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5_bmi,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1b_stimulant,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_4_stimulant,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5_stimulant,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1b_prior1,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_4_prior1,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5_prior1,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1b_prior2,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_4_prior2,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5_prior2,
  data = analytic_df,
  family = poisson(),
  unique = TRUE
)
```

## 3. Run models

### Model 2 ADHD Present

```{r model-2-adhd-pres}
fit_2_adhd_pres <- brm(
  formula = formula_2_adhd_pres,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_2_adhd_pres, "fit_2_adhd_pres.rds")
```

### Model 2 ADHD Present Output

```{r model-2-output-adhd-pres}
# Model R2
bayes_R2(fit_2_adhd_pres)

# Model estimates
print(fit_2_adhd_pres, digits = 3)
```

### Model 2 ADHD Present Diagnostics

```{r model-2-diagnostics-adhd-pres}
# Prior Summary
prior_summary(fit_2_adhd_pres)

# MCMC sampling diagnostics
plot(fit_2_adhd_pres)

# Posterior predictive checks
pp_check(fit_2_adhd_pres, type = "bars", ndraws = 100)
pp_check(fit_2_adhd_pres, type = "hist", ndraws = 10)
pp_check(fit_2_adhd_pres, type = "error_hist")
pp_check(fit_2_adhd_pres, type = "stat", stat = "mean")
pp_check(fit_2_adhd_pres, type = "stat", stat = "sd")
pp_check(fit_2_adhd_pres, type = "scatter_avg")
```

### Model 2B ADHD Present

```{r model-2b-adhd-pres}
fit_2b_adhd_pres <- brm(
  formula = formula_2b_adhd_pres,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_2b_adhd_pres, "fit_2b_adhd_pres.rds")
```

### Model 2B ADHD Present Output

```{r model-2b-output-adhd-pres}
# Model R2
bayes_R2(fit_2b_adhd_pres)

# Model estimates
print(fit_2b_adhd_pres, digits = 3)
```

### Model 2B ADHD Present Diagnostics

```{r model-2b-diagnostics-adhd-pres}
# Prior Summary
prior_summary(fit_2b_adhd_pres)

# Posterior Predictive Checks
plot(fit_2b_adhd_pres, ask = FALSE)

# PPC plots
pp_check(fit_2b_adhd_pres, type = "bars", ndraws = 100)
pp_check(fit_2b_adhd_pres, type = "hist", ndraws = 10)
pp_check(fit_2b_adhd_pres, type = "error_hist")
pp_check(fit_2b_adhd_pres, type = "stat", stat = "mean")
pp_check(fit_2b_adhd_pres, type = "stat", stat = "sd")
pp_check(fit_2b_adhd_pres, type = "scatter_avg")
```

### Model 3 ADHD Present

```{r model-3-adhd-pres}
fit_3_adhd_pres <- brm(
  formula = formula_3_adhd_pres,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_3_adhd_pres, "fit_3_adhd_pres.rds") 
```

### Model 3 ADHD Present Output

```{r model-3-output-adhd-pres}
# Model R2
bayes_R2(fit_3_adhd_pres)

# Model estimates
print(fit_3_adhd_pres, digits = 3)
```

### Model 3 ADHD Present Diagnostics

```{r model-3-diagnostics-adhd-pres}
# Prior Summary
prior_summary(fit_3_adhd_pres)

# Posterior Predictive Checks
plot(fit_3_adhd_pres, ask = FALSE)

# PPC plots
pp_check(fit_3_adhd_pres, type = "bars", ndraws = 100)
pp_check(fit_3_adhd_pres, type = "hist", ndraws = 10)
pp_check(fit_3_adhd_pres, type = "error_hist")
pp_check(fit_3_adhd_pres, type = "stat", stat = "mean")
pp_check(fit_3_adhd_pres, type = "stat", stat = "sd")
pp_check(fit_3_adhd_pres, type = "scatter_avg")
```

### Model 3B ADHD Present

```{r model-3b-adhd-pres}
fit_3b_adhd_pres <- brm(
  formula = formula_3b_adhd_pres,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_3b_adhd_pres, "fit_3b_adhd_pres.rds")
```

### Model 3B ADHD Present Output

```{r model-3b-output-adhd-pres}
# Model R2
bayes_R2(fit_3b_adhd_pres)

# Model estimates
print(fit_3b_adhd_pres, digits = 3)
```

### Model 3B ADHD Present Diagnostics

```{r model-3b-diagnostics-adhd-pres}
# Prior Summary
prior_summary(fit_3b_adhd_pres)

# Posterior Predictive Checks
plot(fit_3b_adhd_pres, ask = FALSE)

# PPC plots
pp_check(fit_3b_adhd_pres, type = "bars", ndraws = 100)
pp_check(fit_3b_adhd_pres, type = "hist", ndraws = 10)
pp_check(fit_3b_adhd_pres, type = "error_hist")
pp_check(fit_3b_adhd_pres, type = "stat", stat = "mean")
pp_check(fit_3b_adhd_pres, type = "stat", stat = "sd")
pp_check(fit_3b_adhd_pres, type = "scatter_avg")
```

### Model 1B Observed

```{r model-1b-obs}
fit_1b_obs <- brm(
  formula = formula_1b_obs,
  data = obs_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1b_obs, "fit_1b_obs.rds")
```

### Model 1B Observed Output

```{r model-1b-ob-output}
# Model R2
bayes_R2(fit_1b_obs)

# Model estimates
print(fit_1b_obs, digits = 3)
```

### Model 1B Observed Diagnostics

```{r model-1b-ob-diagnostics}
# Prior Summary
prior_summary(fit_1b_obs)

# Posterior Predictive Checks
plot(fit_1b_obs, ask = FALSE)

# PPC plots
pp_check(fit_1b_obs, type = "bars", ndraws = 100)
pp_check(fit_1b_obs, type = "hist", ndraws = 10)
pp_check(fit_1b_obs, type = "error_hist")
pp_check(fit_1b_obs, type = "stat", stat = "mean")
pp_check(fit_1b_obs, type = "stat", stat = "sd")
pp_check(fit_1b_obs, type = "scatter_avg")
```

### Model 4 Observed

```{r model-4-obs}
fit_4_obs <- brm(
  formula = formula_4_obs,
  data = obs_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_4_obs, "fit_4_obs.rds")
```

### Model 4 Observed Output

```{r model-4-obs-output}
# Model R2
bayes_R2(fit_4_obs)

# Model estimates
print(fit_4_obs, digits = 3)
```

### Model 4 Observed Diagnostics

```{r model-4-obs-diagnostics}
# Prior Summary
prior_summary(fit_4_obs)

# Posterior Predictive Checks
plot(fit_4_obs, ask = FALSE)

# PPC plots
pp_check(fit_4_obs, type = "bars", ndraws = 100)
pp_check(fit_4_obs, type = "hist", ndraws = 10)
pp_check(fit_4_obs, type = "error_hist")
pp_check(fit_4_obs, type = "stat", stat = "mean")
pp_check(fit_4_obs, type = "stat", stat = "sd")
pp_check(fit_4_obs, type = "scatter_avg")
```

### Model 5 Observed

```{r model-5-obs}
fit_5_obs <- brm(
  formula = formula_5_obs,
  data = obs_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_5_obs, "fit_5_obs.rds")
```

### Model 5 Observed Output

```{r model-5-obs-output}
# Model R2
bayes_R2(fit_5_obs)

# Model estimates
print(fit_5_obs, digits = 3)
```

### Model 5 Observed Diagnostics

```{r model-5-obs-diagnostics}
# Prior Summary
prior_summary(fit_5_obs)

# Posterior Predictive Checks
plot(fit_5_obs, ask = FALSE)

# PPC plots
pp_check(fit_5_obs, type = "bars", ndraws = 100)
pp_check(fit_5_obs, type = "hist", ndraws = 10)
pp_check(fit_5_obs, type = "error_hist")
pp_check(fit_5_obs, type = "stat", stat = "mean")
pp_check(fit_5_obs, type = "stat", stat = "sd")
pp_check(fit_5_obs, type = "scatter_avg")
```

### Model 1B BMI

```{r model-1b-bmi}
fit_1b_bmi <- brm(
  formula = formula_1b_bmi,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1b_bmi, "fit_1b_bmi.rds")
```

### Model 1B BMI Output

```{r model-1b-bmi-output}
# Model R2
bayes_R2(fit_1b_bmi)

# Model estimates
print(fit_1b_bmi, digits = 3)
```

### Model 1B BMI Diagnostics

```{r model-1b-bmi-diagnostics}
# Prior Summary
prior_summary(fit_1b_bmi)

# Posterior Predictive Checks
plot(fit_1b_bmi, ask = FALSE)

# PPC plots
pp_check(fit_1b_bmi, type = "bars", ndraws = 100)
pp_check(fit_1b_bmi, type = "hist", ndraws = 10)
pp_check(fit_1b_bmi, type = "error_hist")
pp_check(fit_1b_bmi, type = "stat", stat = "mean")
pp_check(fit_1b_bmi, type = "stat", stat = "sd")
pp_check(fit_1b_bmi, type = "scatter_avg")
```

### Model 4 BMI

```{r model-4-bmi}
fit_4_bmi <- brm(
  formula = formula_4_bmi,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_4_bmi, "fit_4_bmi.rds")
```

### Model 4 BMI Output

```{r model-4-bmi-output}
# Model R2
bayes_R2(fit_4_bmi)

# Model estimates
print(fit_4_bmi, digits = 3)
```

### Model 4 BMI Diagnostics

```{r model-4-bmi-diagnostics}
# Prior Summary
prior_summary(fit_4_bmi)

# Posterior Predictive Checks
plot(fit_4_bmi, ask = FALSE)

# PPC plots
pp_check(fit_4_bmi, type = "bars", ndraws = 100)
pp_check(fit_4_bmi, type = "hist", ndraws = 10)
pp_check(fit_4_bmi, type = "error_hist")
pp_check(fit_4_bmi, type = "stat", stat = "mean")
pp_check(fit_4_bmi, type = "stat", stat = "sd")
pp_check(fit_4_bmi, type = "scatter_avg")
```

### Model 5 BMI

```{r model-5-bmi}
fit_5_bmi <- brm(
  formula = formula_5_bmi,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_5_bmi, "fit_5_bmi.rds")
```

### Model 5 BMI Output

```{r model-5-bmi-output}
# Model R2
bayes_R2(fit_5_bmi)

# Model estimates
print(fit_5_bmi, digits = 3)
```

### Model 5 BMI Diagnostics

```{r model-5-bmi-diagnostics}
# Prior Summary
prior_summary(fit_5_bmi)

# Posterior Predictive Checks
plot(fit_5_bmi, ask = FALSE)

# PPC plots
pp_check(fit_5_bmi, type = "bars", ndraws = 100)
pp_check(fit_5_bmi, type = "hist", ndraws = 10)
pp_check(fit_5_bmi, type = "error_hist")
pp_check(fit_5_bmi, type = "stat", stat = "mean")
pp_check(fit_5_bmi, type = "stat", stat = "sd")
pp_check(fit_5_bmi, type = "scatter_avg")
```

### Model 1B Stimulant

```{r model-1b-stimulant}
fit_1b_stimulant <- brm(
  formula = formula_1b_stimulant,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1b_stimulant, "fit_1b_stimulant.rds")
```

### Model 1B Stimulant Output

```{r model-1b--stimulant-output}
# Model R2
bayes_R2(fit_1b_stimulant)

# Model estimates
print(fit_1b_stimulant, digits = 3)
```

### Model 1B Stimulant Diagnostics

```{r model-1b-stimulant-diagnostics}
# Prior Summary
prior_summary(fit_1b_stimulant)

# Posterior Predictive Checks
plot(fit_1b_stimulant, ask = FALSE)

# PPC plots
pp_check(fit_1b_stimulant, type = "bars", ndraws = 100)
pp_check(fit_1b_stimulant, type = "hist", ndraws = 10)
pp_check(fit_1b_stimulant, type = "error_hist")
pp_check(fit_1b_stimulant, type = "stat", stat = "mean")
pp_check(fit_1b_stimulant, type = "stat", stat = "sd")
pp_check(fit_1b_stimulant, type = "scatter_avg")
```

### Model 4 Stimulant

```{r model-4-stimulant}
fit_4_stimulant <- brm(
  formula = formula_4_stimulant,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_4_stimulant, "fit_4_stimulant.rds")
```

### Model 4 Stimulant Output

```{r model-4-stimulant-output}
# Model R2
bayes_R2(fit_4_stimulant)

# Model estimates
print(fit_4_stimulant, digits = 3)
```

### Model 4 Stimulant Diagnostics

```{r model-4-stimulant-diagnostics}
# Prior Summary
prior_summary(fit_4_stimulant)

# Posterior Predictive Checks
plot(fit_4_stimulant, ask = FALSE)

# PPC plots
pp_check(fit_4_stimulant, type = "bars", ndraws = 100)
pp_check(fit_4_stimulant, type = "hist", ndraws = 10)
pp_check(fit_4_stimulant, type = "error_hist")
pp_check(fit_4_stimulant, type = "stat", stat = "mean")
pp_check(fit_4_stimulant, type = "stat", stat = "sd")
pp_check(fit_4_stimulant, type = "scatter_avg")
```

### Model 5 Stimulant

```{r model-5-stimulant}
fit_5_stimulant <- brm(
  formula = formula_5_stimulant,
  data = analytic_df,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_5_stimulant, "fit_5_stimulant.rds")
```

### Model 5 Stimulant Output

```{r model-5-stimulant-output}
# Model R2
bayes_R2(fit_5_stimulant)

# Model estimates
print(fit_5_stimulant, digits = 3)
```

### Model 5 Stimulant Diagnostics

```{r model-5-stimulant-diagnostics}
# Prior Summary
prior_summary(fit_5_stimulant)

# Posterior Predictive Checks
plot(fit_5_stimulant, ask = FALSE)

# PPC plots
pp_check(fit_5_stimulant, type = "bars", ndraws = 100)
pp_check(fit_5_stimulant, type = "hist", ndraws = 10)
pp_check(fit_5_stimulant, type = "error_hist")
pp_check(fit_5_stimulant, type = "stat", stat = "mean")
pp_check(fit_5_stimulant, type = "stat", stat = "sd")
pp_check(fit_5_stimulant, type = "scatter_avg")
```

### Model 1B Prior 1

```{r model-1b-prior1}
fit_1b_prior1 <- brm(
  formula = formula_1b_prior1,
  data = analytic_df,
  family = poisson(),  
  prior = prior1,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1b_prior1, "fit_1b_prior1.rds")
```

### Model 1B Prior 1 Output

```{r model-1b-prior1-output}
# Model R2
bayes_R2(fit_1b_prior1)

# Model estimates
print(fit_1b_prior1, digits = 3)
```

### Model 1B Prior 1 Diagnostics

```{r model-1b-prior1-diagnostics}
# Prior Summary
prior_summary(fit_1b_prior1)

# Posterior Predictive Checks
plot(fit_1b_prior1, ask = FALSE)

# PPC plots
pp_check(fit_1b_prior1, type = "bars", ndraws = 100)
pp_check(fit_1b_prior1, type = "hist", ndraws = 10)
pp_check(fit_1b_prior1, type = "error_hist")
pp_check(fit_1b_prior1, type = "stat", stat = "mean")
pp_check(fit_1b_prior1, type = "stat", stat = "sd")
pp_check(fit_1b_prior1, type = "scatter_avg")
```

### Model 4 Prior 1

```{r model-4-prior1}
fit_4_prior1 <- brm(
  formula = formula_4_prior1,
  data = analytic_df,
  family = poisson(),  
  prior = prior1,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_4_prior1, "fit_4_prior1.rds")
```

### Model 4 Prior 1 Output

```{r model-4-prior1-output}
# Model R2
bayes_R2(fit_4_prior1)

# Model estimates
print(fit_4_prior1, digits = 3)
```

### Model 4 Prior 1 Diagnostics

```{r model-4-prior1-diagnostics}
# Prior Summary
prior_summary(fit_4_prior1)

# Posterior Predictive Checks
plot(fit_4_prior1, ask = FALSE)

# PPC plots
pp_check(fit_4_prior1, type = "bars", ndraws = 100)
pp_check(fit_4_prior1, type = "hist", ndraws = 10)
pp_check(fit_4_prior1, type = "error_hist")
pp_check(fit_4_prior1, type = "stat", stat = "mean")
pp_check(fit_4_prior1, type = "stat", stat = "sd")
pp_check(fit_4_prior1, type = "scatter_avg")
```

### Model 5 Prior 1

```{r model-5-prior1}
fit_5_prior1 <- brm(
  formula = formula_5_prior1,
  data = analytic_df,
  family = poisson(),  
  prior = prior1,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_5_prior1, "fit_5_prior1.rds")
```

### Model 5 Prior 1 Output

```{r model-5-prior1-output}
# Model R2
bayes_R2(fit_5_prior1)

# Model estimates
print(fit_5_prior1, digits = 3)
```

### Model 5 Prior 1 Diagnostics

```{r model-5-prior1-diagnostics}
# Prior Summary
prior_summary(fit_5_prior1)

# Posterior Predictive Checks
plot(fit_5_prior1, ask = FALSE)

# PPC plots
pp_check(fit_5_prior1, type = "bars", ndraws = 100)
pp_check(fit_5_prior1, type = "hist", ndraws = 10)
pp_check(fit_5_prior1, type = "error_hist")
pp_check(fit_5_prior1, type = "stat", stat = "mean")
pp_check(fit_5_prior1, type = "stat", stat = "sd")
pp_check(fit_5_prior1, type = "scatter_avg")
```

### Model 1B Prior 2

```{r model-1b-prior2}
fit_1b_prior2 <- brm(
  formula = formula_1b_prior2,
  data = analytic_df,
  family = poisson(),  
  prior = prior2,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1b_prior2, "fit_1b_prior2.rds")
```

### Model 1B Prior 2 Output

```{r model-1b-prior2-output}
# Model R2
bayes_R2(fit_1b_prior2)

# Model estimates
print(fit_1b_prior2, digits = 3)
```

### Model 1B Prior 2 Diagnostics

```{r model-1b-prior2-diagnostics}
# Prior Summary
prior_summary(fit_1b_prior2)

# Posterior Predictive Checks
plot(fit_1b_prior2, ask = FALSE)

# PPC plots
pp_check(fit_1b_prior2, type = "bars", ndraws = 100)
pp_check(fit_1b_prior2, type = "hist", ndraws = 10)
pp_check(fit_1b_prior2, type = "error_hist")
pp_check(fit_1b_prior2, type = "stat", stat = "mean")
pp_check(fit_1b_prior2, type = "stat", stat = "sd")
pp_check(fit_1b_prior2, type = "scatter_avg")
```

### Model 4 Prior 2

```{r model-4-prior2}
fit_4_prior2 <- brm(
  formula = formula_4_prior2,
  data = analytic_df,
  family = poisson(),  
  prior = prior2,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_4_prior2, "fit_4_prior2.rds")
```

### Model 4 Prior 2 Output

```{r model-4-prior2-output}
# Model R2
bayes_R2(fit_4_prior2)

# Model estimates
print(fit_4_prior2, digits = 3)
```

### Model 4 Prior 2 Diagnostics

```{r model-4-prior2-diagnostics}
# Prior Summary
prior_summary(fit_4_prior2)

# Posterior Predictive Checks
plot(fit_4_prior2, ask = FALSE)

# PPC plots
pp_check(fit_4_prior2, type = "bars", ndraws = 100)
pp_check(fit_4_prior2, type = "hist", ndraws = 10)
pp_check(fit_4_prior2, type = "error_hist")
pp_check(fit_4_prior2, type = "stat", stat = "mean")
pp_check(fit_4_prior2, type = "stat", stat = "sd")
pp_check(fit_4_prior2, type = "scatter_avg")
```

### Model 5 Prior 2

```{r model-5-prior2}
fit_5_prior2 <- brm(
  formula = formula_5_prior2,
  data = analytic_df,
  family = poisson(),  
  prior = prior2,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_5_prior2, "fit_5_prior2.rds")
```

### Model 5 Prior 2 Output

```{r model-5-prior2-output}
# Model R2
bayes_R2(fit_5_prior2)

# Model estimates
print(fit_5_prior2, digits = 3)
```

### Model 5 Prior 2 Diagnostics

```{r model-5-prior2-diagnostics}
# Prior Summary
prior_summary(fit_5_prior2)

# Posterior Predictive Checks
plot(fit_5_prior2, ask = FALSE)

# PPC plots
pp_check(fit_5_prior2, type = "bars", ndraws = 100)
pp_check(fit_5_prior2, type = "hist", ndraws = 10)
pp_check(fit_5_prior2, type = "error_hist")
pp_check(fit_5_prior2, type = "stat", stat = "mean")
pp_check(fit_5_prior2, type = "stat", stat = "sd")
pp_check(fit_5_prior2, type = "scatter_avg")
```
