---
title: "ABCD ADHD menarche longitudinal supplementary/sensitivity analyses"
author: "Emily Friedel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Setup

### Load libraries

```{r libraries}
library(tidyverse)
library(ggplot2)
library(mice)
library(brms)
library(posterior)
library(bayestestR)
library(emmeans)
```

### Import saved imputed dfs

```{r import-data}
analytic_df_wide <- read_csv("analytic_df_wide.csv")
obs_df_wide <- read_csv("obs_df_wide.csv")
```

## Initial checks

```{r initial-checks}
summary(unique(analytic_df_wide$participant_id))
length(analytic_df_wide)
colnames(analytic_df_wide)

summary(unique(obs_df_wide$participant_id))
length(obs_df_wide)
colnames(obs_df_wide)
```

## 1. Pre model processing

### Convert to factors: site ID, family ID, ethnicity, and menarche status factors

```{r factors}
analytic_df_wide <- analytic_df_wide %>%
  mutate(
    family_id = as.factor(family_id),
    site_id = as.factor(site_id),
    ethnrace = as.factor(ethnrace),
    adhd_status_life = as.factor(adhd_status_life),
    menarche_status_p_wave_2 = as.factor(menarche_status_p_wave_2),
    stimulant_status_wave_3 = as.factor(stimulant_status_wave_3)
  )

# set white_nh as reference for enthrace
analytic_df_wide$ethnrace <- relevel(analytic_df_wide$ethnrace, ref = "white_nh")
levels(analytic_df_wide$ethnrace)

obs_df_wide <- obs_df_wide %>%
  mutate(
    family_id = as.factor(family_id),
    site_id = as.factor(site_id),
    ethnrace = as.factor(ethnrace),
    adhd_status_life = as.factor(adhd_status_life),
    menarche_status_p_wave_2 = as.factor(menarche_status_p_wave_2),
    stimulant_status_wave_3 = as.factor(stimulant_status_wave_3)
  )
# collapse "asian" into "other" for ethnrace
obs_df_wide$ethnrace <- forcats::fct_collapse(
  obs_df_wide$ethnrace,
  other = c("asian", "other")
)

# order levels alphabetically, then set white_nh as reference
obs_df_wide$ethnrace <- forcats::fct_relevel(
  obs_df_wide$ethnrace,
  "white_nh", "black_nh", "hispanic", "other"
)
levels(obs_df_wide$ethnrace)
```

### Convert to integers: all outcomes

```{r outcomes as integers}
analytic_df_wide <- analytic_df_wide %>%
  mutate(
    adhd_traits_wave_3 = as.integer(adhd_traits_wave_3),
    internalising_wave_3 = as.integer(internalising_wave_3),
    externalising_wave_3 = as.integer(externalising_wave_3)
  )

obs_df_wide <- obs_df_wide %>%
  mutate(
    adhd_traits_wave_3 = as.integer(adhd_traits_wave_3),
    internalising_wave_3 = as.integer(internalising_wave_3),
    externalising_wave_3 = as.integer(externalising_wave_3)
  )
```

### Check head of wide dataframe

```{r check head wide df}
head(analytic_df_wide)
head(obs_df_wide)
```

### Check wide imputed df structure

```{r check structure wide df}
str(analytic_df_wide)
str(obs_df_wide)
```

## 2. Set up hypothesis formulas and prior

```{r hypothesis-formulas}

formula_2_long_adhd_pres <- bf(internalising_wave_3 ~ 
                                 menarche_status_p_wave_2 * adhd_status_pres
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc 
                  + internalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_2b_long_adhd_pres <- bf(internalising_wave_3 ~ breast_development_p_wave_2 * adhd_status_pres
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc 
                  + internalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_3_long_adhd_pres <- bf(externalising_wave_3 ~ menarche_status_p_wave_2 * adhd_status_pres + 
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc 
                  + externalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_3b_long_adhd_pres <- bf(externalising_wave_3 ~ breast_development_p_wave_2 * adhd_status_pres
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc 
                  + externalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )


formula_1_long_obs <- bf(adhd_traits_wave_3 ~ menarche_status_p_wave_2
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + adhd_traits_wave_0_sc 
                  + (1 | family_id) + (1 | site_id)
                  )

formula_1b_long_obs <- bf(adhd_traits_wave_3 ~ breast_development_p_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + adhd_traits_wave_0_sc 
                  + (1 | family_id) + (1 | site_id)
                  )

formula_5b_long_obs <- bf(externalising_wave_3 ~ 
                            breast_development_p_wave_2_sc * adhd_traits_wave_2_sc
                          + age_wave_2_sc + ethnrace + inr_wave_0_sc 
                  + externalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_1_long_bmi <- bf(adhd_traits_wave_3 ~ menarche_status_p_wave_2
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + adhd_traits_wave_0_sc + bmi_wave_0_sc 
                  + (1 | family_id) + (1 | site_id)
                  )

formula_1b_long_bmi <- bf(adhd_traits_wave_3 ~ breast_development_p_wave_2_sc
                 + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + adhd_traits_wave_0_sc + bmi_wave_0_sc 
                  + (1 | family_id) + (1 | site_id)
                  )

formula_5b_long_bmi <- bf(externalising_wave_3 ~ 
                  breast_development_p_wave_2_sc * adhd_traits_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc 
                  + externalising_wave_0_sc + bmi_wave_0_sc 
                  + (1 | family_id) + (1 | site_id)
                  )

formula_1_long_stimulant <- bf(adhd_traits_wave_3 ~ menarche_status_p_wave_2
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + adhd_traits_wave_0_sc + stimulant_status_wave_3
                  + (1 | family_id) + (1 | site_id)
                  )

formula_1b_long_stimulant <- bf(adhd_traits_wave_3 ~ breast_development_p_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + adhd_traits_wave_0_sc + stimulant_status_wave_3
                  + (1 | family_id) + (1 | site_id)
                  )

formula_5b_long_stimulant <- bf(externalising_wave_3 ~
                  breast_development_p_wave_2_sc * adhd_traits_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc 
                  + externalising_wave_0_sc + stimulant_status_wave_3
                  + (1 | family_id) + (1 | site_id)
                  )

formula_1_long_prior1 <- bf(adhd_traits_wave_3 ~ menarche_status_p_wave_2
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + adhd_traits_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_1b_long_prior1 <- bf(adhd_traits_wave_3 ~ breast_development_p_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + adhd_traits_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_5b_long_prior1 <- bf(externalising_wave_3 ~ 
                  breast_development_p_wave_2_sc * adhd_traits_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc 
                  + externalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_1_long_prior2 <- bf(adhd_traits_wave_3 ~ menarche_status_p_wave_2
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + adhd_traits_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_1b_long_prior2 <- bf(adhd_traits_wave_3 ~ breast_development_p_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc
                  + adhd_traits_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

formula_5b_long_prior2 <- bf(externalising_wave_3 ~ 
                  breast_development_p_wave_2_sc * adhd_traits_wave_2_sc
                  + age_wave_2_sc + ethnrace + inr_wave_0_sc 
                  + externalising_wave_0_sc
                  + (1 | family_id) + (1 | site_id)
                  )

# Set prior
prior <- prior(normal(0, 1), class = "b")
prior1 <- prior(normal(0, 0.25), class = "b")
prior2 <- prior(normal(0, 10), class = "b")
```

### 2.1 Check default prior

```{r default-prior}
get_prior(
  formula = formula_2_long_adhd_pres,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_2b_long_adhd_pres,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_3_long_adhd_pres,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_3b_long_adhd_pres,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1_long_obs,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1b_long_obs,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5b_long_obs,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1_long_bmi,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1b_long_bmi,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5b_long_bmi,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1_long_stimulant,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1b_long_stimulant,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5b_long_stimulant,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1_long_prior1,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1b_long_prior1,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5b_long_prior1,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1_long_prior2,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_1b_long_prior2,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)

get_prior(
  formula = formula_5b_long_prior2,
  data = analytic_df_wide,
  family = poisson(),
  unique = TRUE
)
```

## 3. Run long models

### Model 2 Long ADHD Present

```{r model-2-long-adhd-pres}
fit_2_long_adhd_pres <- brm(
  formula = formula_2_long_adhd_pres,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_2_long_adhd_pres, "fit_2_long_adhd_pres.rds")
```

### Model 2 Long ADHD Present Output

```{r model-1-long-adhd-pres-output}
# Posterior Summary Table
bayes_R2(fit_2_long_adhd_pres)

# Model estimates
print(fit_2_long_adhd_pres, digits = 3)
```

### Model 2 Long ADHD Present Diagnostics

```{r model-2-long-adhd-pres-diagnostics}
# Prior Summary
prior_summary(fit_2_long_adhd_pres)

# Posterior Predictive Checks
plot(fit_2_long_adhd_pres, ask = FALSE)

# PPC plots
pp_check(fit_2_long_adhd_pres, type = "bars", ndraws = 100)
pp_check(fit_2_long_adhd_pres, type = "hist", ndraws = 10)
pp_check(fit_2_long_adhd_pres, type = "error_hist")
pp_check(fit_2_long_adhd_pres, type = "stat", stat = "mean")
pp_check(fit_2_long_adhd_pres, type = "stat", stat = "sd")
pp_check(fit_2_long_adhd_pres, type = "scatter_avg")
```

### Model 2B Long ADHD Present

```{r model-2b-long-adhd-pres}
fit_2b_long_adhd_pres <- brm(
  formula = formula_2b_long_adhd_pres,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_2b_long_adhd_pres, "fit_2b_long_adhd_pres.rds")
```

### Model 2B Long ADHD Present Output

```{r model-2b-long-adhd-pres-output}
# Posterior Summary Table
bayes_R2(fit_2b_long_adhd_pres)

# Model estimates
print(fit_2b_long_adhd_pres, digits = 3)
```

### Model 2B Long ADHD Present Diagnostics

```{r model-2b-long-adhd-pres-diagnostics}
# Prior Summary
prior_summary(fit_2b_long_adhd_pres)

# Posterior Predictive Checks
plot(fit_2b_long_adhd_pres, ask = FALSE)

# PPC plots
pp_check(fit_2b_long_adhd_pres, type = "bars", ndraws = 100)
pp_check(fit_2b_long_adhd_pres, type = "hist", ndraws = 10)
pp_check(fit_2b_long_adhd_pres, type = "error_hist")
pp_check(fit_2b_long_adhd_pres, type = "stat", stat = "mean")
pp_check(fit_2b_long_adhd_pres, type = "stat", stat = "sd")
pp_check(fit_2b_long_adhd_pres, type = "scatter_avg")
```

### Model 3 Long ADHD Present

```{r model-3-long}
fit_3_long_adhd_pres <- brm(
  formula = formula_3_long_adhd_pres,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  control = list(adapt_delta = 0.90),
  cores = 2,
  seed = 1234
)

saveRDS(fit_3_long_adhd_pres, "fit_3_long_adhd_pres.rds")
```


### Model 3 Long ADHD Present Output

```{r model-3-long-output}
# Posterior Summary Table
bayes_R2(fit_3_long_adhd_pres)

# Model estimates
print(fit_3_long_adhd_pres, digits = 3)
```

### Model 3 Long ADHD Present Diagnostics

```{r model-3-long-adhd-pres-diagnostics}
# Prior Summary
prior_summary(fit_3_long_adhd_pres)

# Posterior Predictive Checks
plot(fit_3_long_adhd_pres, ask = FALSE)

# PPC plots
pp_check(fit_3_long_adhd_pres, type = "bars", ndraws = 100)
pp_check(fit_3_long_adhd_pres, type = "hist", ndraws = 10)
pp_check(fit_3_long_adhd_pres, type = "error_hist")
pp_check(fit_3_long_adhd_pres, type = "stat", stat = "mean")
pp_check(fit_3_long_adhd_pres, type = "stat", stat = "sd")
pp_check(fit_3_long_adhd_pres, type = "scatter_avg")
```

### Model 3B Long ADHD Present 

```{r model-3b-long-adhd-pres}
fit_3b_long_adhd_pres <- brm(
  formula = formula_3b_long_adhd_pres,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_3b_long_adhd_pres, "fit_3b_long_adhd_pres.rds")
```

### Model 3B Long ADHD Present Output

```{r model-3b-long--adhd-pres-output}
# Posterior Summary Table
bayes_R2(fit_3b_long_adhd_pres)

# Model estimates
print(fit_3b_long_adhd_pres, digits = 3)
```

### Model 3B Long ADHD Present Diagnostics

```{r model-3b-diagnostics}
# Prior Summary
prior_summary(fit_3b_long_adhd_pres)

# Posterior Predictive Checks
plot(fit_3b_long_adhd_pres, ask = FALSE)

# PPC plots
pp_check(fit_3b_long_adhd_pres, type = "bars", ndraws = 100)
pp_check(fit_3b_long_adhd_pres, type = "hist", ndraws = 10)
pp_check(fit_3b_long_adhd_pres, type = "error_hist")
pp_check(fit_3b_long_adhd_pres, type = "stat", stat = "mean")
pp_check(fit_3b_long_adhd_pres, type = "stat", stat = "sd")
pp_check(fit_3b_long_adhd_pres, type = "scatter_avg")
```

### Model 1 Long Observed

```{r model-1-long-obs}
fit_1_long_obs <- brm(
  formula = formula_1_long_obs,
  data = obs_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  control = list(adapt_delta = 0.95),
  cores = 2,
  seed = 1234
)

saveRDS(fit_1_long_obs, "fit_1_long_obs.rds")
```

### Model 1 Long Observed Output

```{r model-1-long-obs-output}
# Posterior Summary Table
bayes_R2(fit_1_long_obs)

# Model estimates
print(fit_1_long_obs, digits = 3)
```

### Model 1 Long Observed Diagnostics

```{r model-1-long-obs-diagnostics}
# Prior Summary
prior_summary(fit_1_long_obs)

# Posterior Predictive Checks
plot(fit_1_long_obs, ask = FALSE)

# PPC plots
pp_check(fit_1_long_obs, type = "bars", ndraws = 100)
pp_check(fit_1_long_obs, type = "hist", ndraws = 10)
pp_check(fit_1_long_obs, type = "error_hist")
pp_check(fit_1_long_obs, type = "stat", stat = "mean")
pp_check(fit_1_long_obs, type = "stat", stat = "sd")
pp_check(fit_1_long_obs, type = "scatter_avg")
```

### Model 1B Long Observed

```{r model-1b-long-obs}
fit_1b_long_obs <- brm(
  formula = formula_1b_long_obs,
  data = obs_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  control = list(adapt_delta = 0.95),
  cores = 2,
  seed = 1234
)

saveRDS(fit_1b_long_obs, "fit_1b_long_obs.rds")
```

### Model 1B Long Observed Output

```{r model-1b-long-obs-output}
# Posterior Summary Table
bayes_R2(fit_1b_long_obs)

# Model estimates
print(fit_1b_long_obs, digits = 3)
```

### Model 1B Long Observed Diagnostics

```{r model-1b-long-obs-diagnostics}
# Prior Summary
prior_summary(fit_1b_long_obs)

# Posterior Predictive Checks
plot(fit_1b_long_obs, ask = FALSE)

# PPC plots
pp_check(fit_1b_long_obs, ndraws = 50)
pp_check(fit_1b_long_obs, type = "error_hist")
pp_check(fit_1b_long_obs, type = "stat", stat = "mean")
pp_check(fit_1b_long_obs, type = "stat", stat = "sd")
pp_check(fit_1b_long_obs, type = "scatter_avg")
```

### Model 5B Long Observed

```{r model-3-long-obs}
fit_5b_long_obs <- brm(
  formula = formula_5b_long_obs,
  data = obs_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_5b_long_obs, "fit_5b_long_obs.rds")
```

### Model 5B Long Observed Output

```{r model-5b-long-obs-output}
# Posterior Summary Table
bayes_R2(fit_5b_long_obs)

# Model estimates
print(fit_5b_long_obs, digits = 3)
```

### Model 5B Long Observed Diagnostics

```{r model-5b-long-obs-diagnostics}
# Prior Summary
prior_summary(fit_5b_long_obs)

# Posterior Predictive Checks
plot(fit_5b_long_obs, ask = FALSE)

# PPC plots
pp_check(fit_5b_long_obs, type = "bars", ndraws = 100)
pp_check(fit_5b_long_obs, type = "hist", ndraws = 10)
pp_check(fit_5b_long_obs, type = "error_hist")
pp_check(fit_5b_long_obs, type = "stat", stat = "mean")
pp_check(fit_5b_long_obs, type = "stat", stat = "sd")
pp_check(fit_5b_long_obs, type = "scatter_avg")
```

### Model 1 Long BMI

```{r model-1-long-bmi}
fit_1_long_bmi <- brm(
  formula = formula_1_long_bmi,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1_long_bmi, "fit_1_long_bmi.rds")
```

### Model 1 Long BMI Output

```{r model-1-long-bmi-output}
# Posterior Summary Table
bayes_R2(fit_1_long_bmi)

# Model estimates
print(fit_1_long_bmi, digits = 3)
```

### Model 1 Long BMI Diagnostics

```{r model-1-long-bmi-diagnostics}
# Prior Summary
prior_summary(fit_1_long_bmi)

# Posterior Predictive Checks

plot(fit_1_long_bmi, ask = FALSE)

# PPC plots
pp_check(fit_1_long_bmi, type = "bars", ndraws = 100)
pp_check(fit_1_long_bmi, type = "hist", ndraws = 10)
pp_check(fit_1_long_bmi, type = "error_hist")
pp_check(fit_1_long_bmi, type = "stat", stat = "mean")
pp_check(fit_1_long_bmi, type = "stat", stat = "sd")
pp_check(fit_1_long_bmi, type = "scatter_avg")
```

### Model 1B Long BMI

```{r model-1b-long-bmi}
fit_1b_long_bmi <- brm(
  formula = formula_1b_long_bmi,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1b_long_bmi, "fit_1b_long_bmi.rds")
```

### Model 1B Long BMI Output

```{r model-1b-long-bmi-output}
# Posterior Summary Table
bayes_R2(fit_1b_long_bmi)

# Model estimates
print(fit_1b_long_bmi, digits = 3)
```

### Model 1B Long BMI Diagnostics

```{r model-1b-long-bmi-diagnostics}
# Prior Summary
prior_summary(fit_1b_long_bmi)

# Posterior Predictive Checks
plot(fit_1b_long_bmi, ask = FALSE)

# PPC plots
pp_check(fit_1b_long_bmi, ndraws = 50)
pp_check(fit_1b_long_bmi, type = "error_hist")
pp_check(fit_1b_long_bmi, type = "stat", stat = "mean")
pp_check(fit_1b_long_bmi, type = "stat", stat = "sd")
pp_check(fit_1b_long_bmi, type = "scatter_avg")
```

### Model 5B Long BMI

```{r model-5b-long-bmi}
fit_5b_long_bmi <- brm(
  formula = formula_5b_long_bmi,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  control = list(adapt_delta = 0.99),
  cores = 2,
  seed = 1234
)

saveRDS(fit_5b_long_bmi, "fit_5b_long_bmi.rds")
```

### Model 5B Long BMI Output

```{r model-5b-long-bmi-output}
# Posterior Summary Table
bayes_R2(fit_5b_long_bmi)

# Model estimates
print(fit_5b_long_bmi, digits = 3)
```

### Model 1B Long BMI Diagnostics

```{r model-5b-long-bmi-diagnostics}
# Prior Summary
prior_summary(fit_5b_long_bmi)

# Posterior Predictive Checks

plot(fit_5b_long_bmi, ask = FALSE)

# PPC plots
pp_check(fit_5b_long_bmi, ndraws = 50)
pp_check(fit_5b_long_bmi, type = "error_hist")
pp_check(fit_5b_long_bmi, type = "stat", stat = "mean")
pp_check(fit_5b_long_bmi, type = "stat", stat = "sd")
pp_check(fit_5b_long_bmi, type = "scatter_avg")
```

### Model 1 Long Stimulant

```{r model-1-long-stimulant}
fit_1_long_stimulant <- brm(
  formula = formula_1_long_stimulant,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  control = list(adapt_delta = 0.95),
  cores = 2,
  seed = 1234
)

saveRDS(fit_1_long_stimulant, "fit_1_long_stimulant.rds")
```

### Model 1 Long Stimulant Output

```{r model-1-long-stimulant-output}
# Model R2
bayes_R2(fit_1_long_stimulant)

# Model estimates
print(fit_1_long_stimulant, digits = 3)
```

### Model 1 Stimulant Diagnostics

```{r model-1-long-stimulant-diagnostics}
# Prior Summary
prior_summary(fit_1_long_stimulant)

# Posterior Predictive Checks
plot(fit_1_long_stimulant, ask = FALSE)

# PPC plots
pp_check(fit_1_long_stimulant, type = "bars", ndraws = 100)
pp_check(fit_1_long_stimulant, type = "hist", ndraws = 10)
pp_check(fit_1_long_stimulant, type = "error_hist")
pp_check(fit_1_long_stimulant, type = "stat", stat = "mean")
pp_check(fit_1_long_stimulant, type = "stat", stat = "sd")
pp_check(fit_1_long_stimulant, type = "scatter_avg")
```

### Model 1B Long Stimulant

```{r model-1b-long-stimulant}
fit_1b_long_stimulant <- brm(
  formula = formula_1b_long_stimulant,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  control = list(adapt_delta = 0.90),
  cores = 2,
  seed = 1234
)

saveRDS(fit_1b_long_stimulant, "fit_1b_long_stimulant.rds")
```

### Model 1B Long Stimulant Output

```{r model-1b-stimulant-output}
# Model R2
bayes_R2(fit_1b_long_stimulant)

# Model estimates
print(fit_1b_long_stimulant, digits = 3)
```

### Model 1B Long Stimulant Diagnostics

```{r model-1b-long-stimulant-diagnostics}
# Prior Summary
prior_summary(fit_1b_long_stimulant)

# Posterior Predictive Checks
plot(fit_1b_stimulant, ask = FALSE)

# PPC plots
pp_check(fit_1b_long_stimulant, type = "bars", ndraws = 100)
pp_check(fit_1b_long_stimulant, type = "hist", ndraws = 10)
pp_check(fit_1b_long_stimulant, type = "error_hist")
pp_check(fit_1b_long_stimulant, type = "stat", stat = "mean")
pp_check(fit_1b_long_stimulant, type = "stat", stat = "sd")
pp_check(fit_1b_long_stimulant, type = "scatter_avg")
```

### Model 5B Long Stimulant

```{r model-5b-long-stimulant}
fit_5b_long_stimulant <- brm(
  formula = formula_5b_long_stimulant,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  control = list(adapt_delta = 0.95),
  cores = 2,
  seed = 1234
)

saveRDS(fit_5b_long_stimulant, "fit_5b_long_stimulant.rds")
```

### Model 5B Long Stimulant Output

```{r model-5b-long-stimulant-output}
# Model R2
bayes_R2(fit_5b_long_stimulant)

# Model estimates
print(fit_5b_long_stimulant, digits = 3)
```

### Model 5B Long Stimulant Diagnostics

```{r model-5b-long-stimulant-diagnostics}
# Prior Summary
prior_summary(fit_5b_stimulant)

# Posterior Predictive Checks
plot(fit_5b_stimulant, ask = FALSE)

# PPC plots
pp_check(fit_5b_stimulant, type = "bars", ndraws = 100)
pp_check(fit_5b_stimulant, type = "hist", ndraws = 10)
pp_check(fit_5b_stimulant, type = "error_hist")
pp_check(fit_5b_stimulant, type = "stat", stat = "mean")
pp_check(fit_5b_stimulant, type = "stat", stat = "sd")
pp_check(fit_1b_stimulant, type = "scatter_avg")
```

### Model 1 Long Prior 1

```{r model-1-long-prior1}
fit_1_long_prior1 <- brm(
  formula = formula_1_long_prior1,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior1,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1_long_prior1, "fit_1_long_prior1.rds")
```

### Model 1 Prior 1 Output

```{r model-1-long-prior1-output}
# Posterior Summary Table
bayes_R2(fit_1_long_prior1)

# Model estimates
print(fit_1_long_prior1, digits = 3)
```

### Model 1 Long Prior 1 Diagnostics

```{r model-1-long-prior1-diagnostics}
# Prior Summary
prior_summary(fit_1_long_prior1)

# Posterior Predictive Checks
plot(fit_1_long_prior1, ask = FALSE)

# PPC plots
pp_check(fit_1_long_prior1, type = "bars", ndraws = 100)
pp_check(fit_1_long_prior1, type = "hist", ndraws = 10)
pp_check(fit_1_long_prior1, type = "error_hist")
pp_check(fit_1_long_prior1, type = "stat", stat = "mean")
pp_check(fit_1_long_prior1, type = "stat", stat = "sd")
pp_check(fit_1_long_prior1, type = "scatter_avg")
```

### Model 1B Long Prior 1

```{r model-1b-long-prior1}
fit_1b_long_prior1 <- brm(
  formula = formula_1b_long_prior1,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior1,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1b_long_prior1, "fit_1b_long_prior1.rds")
```

### Model 1B Long Prior 1 Output

```{r model-1b-long-output}
# Posterior Summary Table
bayes_R2(fit_1b_long_prior1)

# Model estimates
print(fit_1b_long_prior1, digits = 3)
```

### Model 1B Long Prior 1 Diagnostics

```{r model-1b-long-prior1-diagnostics}
# Prior Summary
prior_summary(fit_1b_long_prior1)

# Posterior Predictive Checks
plot(fit_1b_long_prior1, ask = FALSE)

# PPC plots
pp_check(fit_1b_long_prior1, ndraws = 50)
pp_check(fit_1b_long_prior1, type = "error_hist")
pp_check(fit_1b_long_prior1, type = "stat", stat = "mean")
pp_check(fit_1b_long_prior1, type = "stat", stat = "sd")
pp_check(fit_1b_long_prior1, type = "scatter_avg")
```

### Model 5B Long Prior 1

```{r model-1b-long-prior1}
fit_5b_long_prior1 <- brm(
  formula = formula_5b_long_prior1,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior1,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  control = list(adapt_delta = 0.99),
  cores = 2,
  seed = 1234
)

saveRDS(fit_5b_long_prior1, "fit_5b_long_prior1.rds")
```

### Model 5B Long Prior 1 Output

```{r model-5b-long-prior1-output}
# Posterior Summary Table
bayes_R2(fit_5b_long_prior1)

# Model estimates
print(fit_5b_long_prior1, digits = 3)
```

### Model 5B Long Diagnostics

```{r model-5b-long-prior1-diagnostics}
# Prior Summary
prior_summary(fit_5b_long_prior1)

# Posterior Predictive Checks
plot(fit_5b_long_prior1, ask = FALSE)

# PPC plots
pp_check(fit_5b_long_prior1, ndraws = 50)
pp_check(fit_5b_long_prior1, type = "error_hist")
pp_check(fit_5b_long_prior1, type = "stat", stat = "mean")
pp_check(fit_5b_long_prior1, type = "stat", stat = "sd")
pp_check(fit_5b_long_prior1, type = "scatter_avg")
```

### Model 1 Long Prior 2

```{r model-1-long-prior2}
fit_1_long_prior2 <- brm(
  formula = formula_1_long_prior2,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior2,
  chains = 2,            
  iter = 4000,
  warmup = 2000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1_long_prior2, "fit_1_long_prior2.rds")
```

### Model 1 Long Prior 2 Output

```{r model-1-long-prior2-output}
# Posterior Summary Table
bayes_R2(fit_1_long_prior2)

# Model estimates
print(fit_1_long_prior2, digits = 3)
```

### Model 1 Long Prior 2 Diagnostics

```{r model-1-long-prior1-diagnostics}
# Prior Summary
prior_summary(fit_1_long_prior2)

# Posterior Predictive Checks
plot(fit_1_long_prior2, ask = FALSE)

# PPC plots
pp_check(fit_1_long_prior2, type = "bars", ndraws = 100)
pp_check(fit_1_long_prior2, type = "hist", ndraws = 10)
pp_check(fit_1_long_prior2, type = "error_hist")
pp_check(fit_1_long_prior2, type = "stat", stat = "mean")
pp_check(fit_1_long_prior2, type = "stat", stat = "sd")
pp_check(fit_1_long_prior2, type = "scatter_avg")
```

### Model 1B Long Prior 2

```{r model-1b-long-prior2}
fit_1b_long_prior2 <- brm(
  formula = formula_1b_long_prior2,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior2,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  cores = 2,
  seed = 1234
)

saveRDS(fit_1b_long_prior2, "fit_1b_long_prior2.rds")
```

### Model 1B Long Prior 2 Output

```{r model-1b-long-prior2-output}
# Posterior Summary Table
bayes_R2(fit_1b_long_prior2)

# Model estimates
print(fit_1b_long_prior2, digits = 3)
```

### Model 1B Long Prior 2 Diagnostics

```{r model-1b-long-prior2-diagnostics}
# Prior Summary
prior_summary(fit_1b_long_prior2)

# Posterior Predictive Checks
plot(fit_1b_long_prior2, ask = FALSE)

# PPC plots
pp_check(fit_1b_long_prior2, ndraws = 50)
pp_check(fit_1b_long_prior2, type = "error_hist")
pp_check(fit_1b_long_prior2, type = "stat", stat = "mean")
pp_check(fit_1b_long_prior2, type = "stat", stat = "sd")
pp_check(fit_1b_long_prior2, type = "scatter_avg")
```

### Model 5B Long Prior 2

```{r model-5b-long-prior2}
fit_5b_long_prior2 <- brm(
  formula = formula_5b_long_prior2,
  data = analytic_df_wide,
  family = poisson(),  
  prior = prior2,
  chains = 2,            
  iter = 8000,
  warmup = 4000,
  control = list(adapt_delta = 0.95),
  cores = 2,
  seed = 1234
)

saveRDS(fit_5b_long_prior2, "fit_5b_long_prior2.rds")
```

### Model 5B Long Prior 2 Output

```{r model-5b-long-prior2-output}
# Posterior Summary Table
bayes_R2(fit_5b_long_prior2)

# Model estimates
print(fit_5b_long_prior2, digits = 3)
```

### Model 5B Long Prior 2 Diagnostics

```{r model-5b-long-prior2-diagnostics}
# Prior Summary
prior_summary(fit_5b_long_prior2)

# Posterior Predictive Checks
plot(fit_5b_long_prior2, ask = FALSE)

# PPC plots
pp_check(fit_5b_long_prior2, ndraws = 50)
pp_check(fit_5b_long_prior2, type = "error_hist")
pp_check(fit_5b_long_prior2, type = "stat", stat = "mean")
pp_check(fit_5b_long_prior2, type = "stat", stat = "sd")
pp_check(fit_5b_long_prior2, type = "scatter_avg")
```
