---
title: "ABCD menarche and ADHD data checks and processing (step 2)"
author: "Emily Friedel"
date: "`r Sys.Date()`"
output: html_document
---

## Setup

### Load libraries

```{r load libraries, include = FALSE}
# load libraries
library(dplyr)
library(tidyr)
library(readr)
library(purrr)
library(stringr)
library(broom)
library(AER)
library(effsize)
```

### Import data

```{r import}
# Read csv
obs_df <- read_csv("obs_df.csv", show_col_types = FALSE)

head(obs_df)
str(obs_df)
```

## 1. Descriptives

### 1.1 Sample size

```{r sample-size}
# Full sample N
n_full <- obs_df %>%
  group_by(wave) %>%
  summarise(N_Full = n(), .groups = "drop")

# ADHD lifetime Y / N
n_life <- obs_df %>%
  filter(!is.na(adhd_status_life)) %>%
  group_by(wave, adhd_status_life) %>%
  summarise(N = n(), .groups = "drop") %>%
  pivot_wider(
    names_from = adhd_status_life,
    values_from = N,
    names_prefix = "N_ADHD_life_"
  )

# ADHD present Y only
n_pres <- obs_df %>%
  filter(adhd_status_pres == "Y") %>%
  group_by(wave) %>%
  summarise(N_ADHD_pres_Y = n(), .groups = "drop")

# Combine all
group_ns <- n_full %>%
  full_join(n_life, by = "wave") %>%
  full_join(n_pres, by = "wave") %>%
  arrange(wave)

group_ns
```

### 1.2 Create variable groups

```{r var-groups}
continuous_vars <- c(
    "age",
    "adhd_traits",
    "internalising",
    "externalising",
    "inr",
    "breast_development_p",
    "bmi"
)

categorical_vars <- c(
  "menarche_status_p",
  "ethnrace",
  "depression",
  "gad",
  "cd",
  "odd"
)
```

### 1.3 Continuous variable descriptives

```{r continuous-descriptives}
# Summary function (continuous)
summarise_cont <- function(df, var) {
  x <- df[[var]]
  
  n_total <- length(x)
  n_missing <- sum(is.na(x))
  x_nonmiss <- x[!is.na(x)]
  
  missing_pct <- round(n_missing / n_total * 100, 1)
  
  if (length(x_nonmiss) == 0) {
    tibble(
      Range   = NA_character_,
      Mean_SD = NA_character_,
      Missing = paste0(missing_pct, "%")
    )
  } else {
    tibble(
      Range   = paste0(round(min(x_nonmiss), 1), "–", round(max(x_nonmiss), 1)),
      Mean_SD = paste0(round(mean(x_nonmiss), 1), " (", round(sd(x_nonmiss), 1), ")"),
      Missing = paste0(missing_pct, "%")
    )
  }
}

# Create table
continuous_table <- lapply(continuous_vars, function(var) {
  
  # Full sample
  full_df <- obs_df %>%
    mutate(Group = "Full") %>%
    group_by(wave, Group) %>%
    group_modify(~ summarise_cont(.x, var)) %>%
    ungroup()
  
  # ADHD groups
  group_df <- obs_df %>%
    filter(!is.na(adhd_status_life)) %>%
    mutate(Group = case_when(
      adhd_status_life == "Y" ~ "ADHD",
      adhd_status_life == "N" ~ "Non_ADHD"
    )) %>%
    group_by(wave, Group) %>%
    group_modify(~ summarise_cont(.x, var)) %>%
    ungroup()
  
  # bind long format
  long_table <- bind_rows(full_df, group_df) %>%
    mutate(Variable = var)
  
  long_table
}) %>%
  bind_rows()

# Convert to wide format
continuous_table_wide <- continuous_table %>%
  pivot_wider(
    names_from = Group,
    values_from = c(Range, Mean_SD, Missing),
    names_glue = "{Group}_{.value}"
  )

# Reorder columns
continuous_table_wide <- continuous_table_wide %>%
  select(
    Variable, wave,
    starts_with("Full"),
    starts_with("Non_ADHD"),
    starts_with("ADHD")
  )

# Export
readr::write_excel_csv(continuous_table_wide, "continuous_table.csv")

continuous_table_wide
```

### 1.4 Continuous group differences

```{r test-continuous-differences}
group_filters <- list(
  
  # Non-ADHD females
  `Non-ADHD_F` = function(df) {
    df %>%
      dplyr::filter(
        adhd_status_life == "N"
      )
  },
  
  # ADHD females
  ADHD_F = function(df) {
    df %>%
      dplyr::filter(
        adhd_status_life == "Y"
      )
  }
  
)

mann_whitney_wave <- function(df, var, group1_filter, group2_filter,
                              group1_name, group2_name) {
  
  df %>%
    group_by(wave) %>%
    group_modify(~ {
      
      x <- group1_filter(.x)[[var]]
      y <- group2_filter(.x)[[var]]
      
      x <- x[!is.na(x)]
      y <- y[!is.na(y)]
      
      if (length(x) < 2 || length(y) < 2) {
        return(tibble(
          U_stat = NA,
          mean_diff = NA,
          effect_r = NA,
          cliffs_delta = NA,
          p_value = NA
        ))
      }
      
      # Mann-Whitney / Wilcoxon rank-sum test
      wilc_res <- wilcox.test(x, y, exact = FALSE)
      
      # Rank-biserial correlation
      n1 <- length(x)
      n2 <- length(y)
      W <- wilc_res$statistic
      mu_W <- n1 * n2 / 2
      sigma_W <- sqrt(n1 * n2 * (n1 + n2 + 1) / 12)
      z <- (W - mu_W) / sigma_W
      r <- z / sqrt(n1 + n2)
      
      # Cliff's delta (note: group order matters for sign)
      cd <- cliff.delta(y, x)  # ADHD vs Non-ADHD
      
      tibble(
        U_stat = unname(W),
        mean_diff = round(mean(x) - mean(y), 2),
        effect_r = round(r, 2),
        cliffs_delta = round(cd$estimate, 2),
        p_value = wilc_res$p.value
      )
      
    }) %>%
    ungroup() %>%
    mutate(
      Variable = var,
      Comparison = "Females: ADHD vs Non-ADHD"
    ) %>%
    dplyr::select(Variable, wave, Comparison, p_value, U_stat, mean_diff, cliffs_delta)
}

# Run across all continuous variables
mw_tests_wave <- do.call(rbind, lapply(continuous_vars, function(var) {
  mann_whitney_wave(
    df = obs_df,
    var = var,
    group1_filter = group_filters$`Non-ADHD_F`,
    group2_filter = group_filters$ADHD_F,
    group1_name = "Non-ADHD_F",
    group2_name = "ADHD_F"
  )
}))

# Format p-values
format_p <- function(p) {
  ifelse(is.na(p), NA,
         ifelse(p < .001, "< .001",
                sprintf("%.3f", p)))
}

mw_tests_wave_clean <- mw_tests_wave %>%
  mutate(
    p_value = format_p(p_value)
  )

View(mw_tests_wave_clean)
```

### 1.5 Continuous ADHD life vs present

```{r adhd-life-pres}
# Summary function
summarise_cont_adhd <- function(df, var) {
  x <- df[[var]]
  
  n_total <- length(x)
  n_missing <- sum(is.na(x))
  x_nonmiss <- x[!is.na(x)]
  
  missing_pct <- round(n_missing / n_total * 100, 1)
  
  if (length(x_nonmiss) == 0) {
    tibble(
      Range   = NA_character_,
      Mean_SD = NA_character_,
      Missing = paste0(missing_pct, "%")
    )
  } else {
    tibble(
      Range   = paste0(round(min(x_nonmiss), 1), "-", round(max(x_nonmiss), 1)),
      Mean_SD = paste0(round(mean(x_nonmiss), 1), " (", round(sd(x_nonmiss), 1), ")"),
      Missing = paste0(missing_pct, "%")
    )
  }
}

# Table
continuous_table_adhd <- lapply(continuous_vars, function(var) {
  
  # ADHD lifetime = Y
  life_df <- obs_df %>%
    filter(adhd_status_life == "Y") %>%
    mutate(Group = "ADHD_Life") %>%
    group_by(wave, Group) %>%
    group_modify(~ summarise_cont(.x, var)) %>%
    ungroup()
  
  # ADHD present = Y
  pres_df <- obs_df %>%
    filter(adhd_status_pres == "Y") %>%
    mutate(Group = "ADHD_Present") %>%
    group_by(wave, Group) %>%
    group_modify(~ summarise_cont(.x, var)) %>%
    ungroup()
  
  bind_rows(life_df, pres_df) %>%
    mutate(Variable = var)
  
}) %>%
  bind_rows()

# Wide format (side-by-side)
continuous_table_adhd_wide <- continuous_table_adhd %>%
  pivot_wider(
    names_from = Group,
    values_from = c(Range, Mean_SD, Missing),
    names_glue = "{Group}_{.value}"
  ) %>%
  select(
    Variable, wave,
    starts_with("ADHD_Life"),
    starts_with("ADHD_Present")
  )

# Export
readr::write_excel_csv(continuous_table_adhd_wide, "continuous_table_adhd_life_vs_present.csv")

continuous_table_wide
```

### 1.6 Categorical variables descriptives

```{r cat-vars-descriptives}
var_levels <- list( 
  ethnrace = c( "white_nh", "hispanic", "black_nh", "asian", "other" ), 
  menarche_status_p = c("N", "Y"), 
  depression = c("0", "1"), 
  gad = c("0", "1"), 
  cd = c("0", "1"), 
  odd = c("0", "1") 
  )

# Define categorical table by wave
categorical_table_wave <- lapply(categorical_vars, function(var) {
  
  if (!(var %in% names(obs_df))) return(NULL)
  
  # Make sure the column is a factor with the full set of levels
  obs_df[[var]] <- factor(obs_df[[var]], levels = var_levels[[var]])
  
  obs_df %>%
    group_by(wave) %>%
    group_modify(~ {
      df <- .x
      
      # Full sample counts
      full_df <- df %>%
        count(Level = .data[[var]]) %>%
        tidyr::complete(Level = var_levels[[var]], fill = list(n = 0)) %>%
        mutate(
          Perc_full = round(100 * n / sum(n), 1),
          Full = paste0(n, " (", Perc_full, "%)")
        ) %>%
        select(Level, Full)
      
      # ADHD split
      adhd_df <- df %>%
        filter(!is.na(adhd_status_life)) %>%
        count(adhd_status_life, Level = .data[[var]]) %>%
        group_by(adhd_status_life) %>%
        tidyr::complete(Level = var_levels[[var]], fill = list(n = 0)) %>%
        mutate(
          Perc = round(100 * n / sum(n), 1),
          Value = paste0(n, " (", Perc, "%)")
        ) %>%
        ungroup() %>%
        select(Level, adhd_status_life, Value) %>%
        tidyr::pivot_wider(names_from = adhd_status_life, values_from = Value, values_fill = "0 (0%)")
      
      # Join full and ADHD counts
      full_df %>%
        left_join(adhd_df, by = "Level") %>%
        mutate(
          Variable = var,
          Level = factor(Level, levels = var_levels[[var]])
        ) %>%
        arrange(Level) %>%
        select(Variable, Level, Full, N = `N`, Y = `Y`)
      
    }) %>%
    ungroup()
  
}) %>% bind_rows()

# Save to CSV
write_csv(categorical_table_wave, "categorical_table_by_wave.csv")
```

### 1.7 Categorical ADHD lifetime and ADHD present

```{r cat-adhd-life-pres}
var_levels <- list(
  ethnrace = c("white_nh", "hispanic", "black_nh", "asian", "other"),
  menarche_status_p = c("N", "Y"),
  depression = c("0", "1"),
  gad = c("0", "1"),
  cd = c("0", "1"),
  odd = c("0", "1")
)

# Categorical table by wave
categorical_adhd_table_wave <- lapply(categorical_vars, function(var) {
  
  if (!(var %in% names(obs_df))) return(NULL)
  
  obs_df[[var]] <- factor(obs_df[[var]], levels = var_levels[[var]])
  
  obs_df %>%
    group_by(wave) %>%
    group_modify(~ {
      
      df <- .x
      
      # Full sample
      full_df <- df %>%
        count(Level = .data[[var]]) %>%
        complete(Level = var_levels[[var]], fill = list(n = 0)) %>%
        mutate(
          Perc_full = round(100 * n / sum(n), 1),
          Full = paste0(n, " (", Perc_full, "%)")
        ) %>%
        select(Level, Full)
      
      # ADHD lifetime Y
      adhd_life_df <- df %>%
        filter(adhd_status_life == "Y") %>%
        count(Level = .data[[var]]) %>%
        complete(Level = var_levels[[var]], fill = list(n = 0)) %>%
        mutate(
          Perc = round(100 * n / sum(n), 1),
          ADHD_Life_Y = paste0(n, " (", Perc, "%)")
        ) %>%
        select(Level, ADHD_Life_Y)
      
      # ADHD present Y
      adhd_pres_df <- df %>%
        filter(adhd_status_pres == "Y") %>%
        count(Level = .data[[var]]) %>%
        complete(Level = var_levels[[var]], fill = list(n = 0)) %>%
        mutate(
          Perc = round(100 * n / sum(n), 1),
          ADHD_Pres_Y = paste0(n, " (", Perc, "%)")
        ) %>%
        select(Level, ADHD_Pres_Y)
      
      # Combine
      full_df %>%
        left_join(adhd_life_df, by = "Level") %>%
        left_join(adhd_pres_df, by = "Level") %>%
        mutate(
          Variable = var,
          Level = factor(Level, levels = var_levels[[var]])
        ) %>%
        arrange(Level) %>%
        select(Variable, Level, Full, ADHD_Life_Y, ADHD_Pres_Y)
      
    }) %>%
    ungroup()
  
}) %>% bind_rows()

readr::write_excel_csv(categorical_adhd_table_wave, "categorical_table_adhd_life_vs_present.csv")
```

### 1.8 Test group diffs for categorical

```{r cat-group-diffs}
run_chisq <- function(data, var, group_var = "adhd_status_life") {
  
  df <- data %>%
    select(all_of(c(var, group_var))) %>%
    filter(!is.na(.data[[var]]), !is.na(.data[[group_var]]))
  
  tab <- table(df[[var]], df[[group_var]])
  
  test <- chisq.test(tab)
  
  tidy(test) %>%
    mutate(
      variable = var,
      n = sum(tab)
    )
}

wave0_vars <- c("ethnrace", "depression", "gad", "cd", "odd")

wave0_results <- map_dfr(wave0_vars, function(v) {
  df_wave0 <- obs_df %>% filter(wave == "wave_0")
  run_chisq(df_wave0, v)
})
 
wave0_results <- wave0_results %>%
  select(variable, statistic, p.value, n)

wave0_results

menarche_results <- obs_df %>%
  filter(!is.na(wave)) %>%
  group_split(wave) %>%
  map_dfr(function(df_wave) {
    wave_name <- unique(df_wave$wave)
    
    run_chisq(df_wave, "menarche_status_p") %>%
      mutate(wave = wave_name)
  })

menarche_results %>%
  select(variable, wave, statistic, p.value, parameter, n) %>%
  mutate(p.value = formatC(p.value, format = "f", digits = 3))
```


## 2. Overdispersion tests

```{r overdispersion-tests}
## Poisson models
poisson_model_traits <- glm(adhd_traits ~ 1, family = poisson, data = obs_df)
poisson_model_int <- glm(internalising ~ 1, family = poisson, data = obs_df)
poisson_model_ext <- glm(externalising ~ 1, family = poisson, data = obs_df)

## Dispersion test
dispersion_test_traits <- dispersiontest(poisson_model_traits)
dispersion_test_int <- dispersiontest(poisson_model_int)
dispersion_test_ext <- dispersiontest(poisson_model_ext)

print(dispersion_test_traits)
print(dispersion_test_int)
print(dispersion_test_ext)
```

## 3. Missingness for menarche and breast development exclusion flowchart

```{r puberty-missing}
# Missing across all waves
count_all_missing_menarche <- obs_df %>%
  group_by(participant_id) %>%
  summarise(all_missing = all(is.na(menarche_status_p))) %>%
  summarise(n = sum(all_missing)) 

count_all_missing_breast_development <- obs_df %>%
  group_by(participant_id) %>%
  summarise(all_missing = all(is.na(breast_development_p))) %>%
  summarise(n = sum(all_missing)) 

# Missing at wave 2 
missing_menarche_wave_2 <- obs_df %>%
  filter(wave == "wave_2") %>%
  count(is.na(menarche_status_p))

missing_breast_development_wave_2 <- obs_df %>%
  filter(wave == "wave_2") %>%
  count(is.na(breast_development_p))

count_all_missing_menarche
count_all_missing_breast_development 
missing_menarche_wave_2 
missing_breast_development_wave_2 
```

### 4. Menarche change from wave 2 --> 3

```{r menarche-2-3}
# Keep only participants with Wave 2 and Wave 3 menarche data
w2_w3 <- obs_df |>
  filter(wave %in% c("wave_2", "wave_3")) |>
  select(participant_id, wave, menarche_status_p) |>
  tidyr::pivot_wider(
    names_from = wave,
    values_from = menarche_status_p
  ) |>
  filter(!is.na(wave_2), !is.na(wave_3))

# Keep participants who were pre-menarche at Wave 2
eligible <- w2_w3 |>
  filter(wave_2 == "N")

# Calculate percentage who converted to post-menarche by Wave 3
conversion_pct <- mean(eligible$wave_3 == "Y") * 100

conversion_pct
```

