---
title: "ABCD ADHD menarche analysis descriptives"
format: pdf
execute: 
  echo: false
---

## Load libraries

```{r load libraries, include = FALSE}
# load libraries
library(dplyr)
library(knitr)
library(tidyr)
library(readxl)
library(readr)
library(AER)
library(ggplot2)
library(psych)
library(purrr)
library(stringr)
library(irr)
```

### Read csv for cleaned dataframe

```{r read csv}
# Read csv
adhd_menarche_clean <- read_csv("2_adhd_menarche_clean.csv", show_col_types = FALSE)
```

### Check cleaned dataframe

```{r head clen df}
head(adhd_menarche_clean)
```

### Remove wave 4

```{r remove wave 4}
adhd_menarche_clean <- adhd_menarche_clean %>%
  filter(wave != "wave_4")
```

## Sample size

```{r sample size}
# Function to count sample size by wave
get_n <- function(data) {
  data %>%
    group_by(wave) %>%
    summarise(n = n())
}

# Full sample
full_n <- get_n(adhd_menarche_clean) %>%
  rename(`Full sample` = n)

# ADHD group
adhd_n <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 1) %>%
  get_n() %>%
  rename(ADHD = n)

# Non-ADHD group
non_adhd_n <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 0) %>%
  get_n() %>%
  rename(`Non-ADHD` = n)

# Combine into one table
sample_size_table <- full_n %>%
  left_join(adhd_n, by = "wave") %>%
  left_join(non_adhd_n, by = "wave")

# View the table
print(sample_size_table)
```

## ADHD Diagnosis

### ADHD diagnosis missing

```{r NA adhd diganosis}
# Count missing adhd data by wave
missing_diagnosis_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_diagnosis = sum(is.na(adhd_diagnosis)),
            total = n(),
            missing_percent = (missing_diagnosis / total) * 100)

# Print the results
print(missing_diagnosis_by_wave)
```

## Age

```{r age}
# Summary statistics by wave
summary_stats_age_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(age_years, na.rm = TRUE),
    median_score = median(age_years, na.rm = TRUE),
    sd_score = sd(age_years, na.rm = TRUE),
    min_score = min(age_years, na.rm = TRUE),
    max_score = max(age_years, na.rm = TRUE),
    n_valid = sum(!is.na(age_years)),
    n_missing = sum(is.na(age_years))
  )

print(summary_stats_age_by_wave)

# ADHD group
adhd_age <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 1) %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(age_years, na.rm = TRUE),
    median_score = median(age_years, na.rm = TRUE),
    sd_score = sd(age_years, na.rm = TRUE),
    min_score = min(age_years, na.rm = TRUE),
    max_score = max(age_years, na.rm = TRUE),
    n_valid = sum(!is.na(age_years)),
    n_missing = sum(is.na(age_years))
  )

print(adhd_age)

# Non-ADHD group
non_adhd_age <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 0) %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(age_years, na.rm = TRUE),
    median_score = median(age_years, na.rm = TRUE),
    sd_score = sd(age_years, na.rm = TRUE),
    min_score = min(age_years, na.rm = TRUE),
    max_score = max(age_years, na.rm = TRUE),
    n_valid = sum(!is.na(age_years)),
    n_missing = sum(is.na(age_years))
  )

print(non_adhd_age)
```

## Menarche

### Menarche missing

```{r NA menarche status p}
# Count missing menarche data by wave
missing_menarche_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_menarche = sum(is.na(menarche_status_p)),
            total = n(),
            missing_percent = (missing_menarche / total) * 100)

# Print the results
print(missing_menarche_by_wave)
```

### Menarche histogram

```{r menarche graph}
# Create a bar graph for menarche status by wave
ggplot(adhd_menarche_clean, aes(x = wave, fill = menarche_status_p)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Menarche Status by Wave",
    x = "Wave",
    y = "Count",
    fill = "Menarche Status"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("Y" = "lightblue", "N" = "salmon"))
```

### Menarche status by ADHD diagnosis for each wave

```{r menarche status by ADHD}
# Function to calculate percentage with menarche_status_p == "Y"
get_percent_menarche <- function(data) {
  data %>%
    group_by(wave) %>%
    summarise(
      percent = mean(menarche_status_p == "Y", na.rm = TRUE) * 100
    )
}

# Full sample
full <- get_percent_menarche(adhd_menarche_clean) %>%
  rename(`Full sample` = percent)

# ADHD only
adhd <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 1) %>%
  get_percent_menarche() %>%
  rename(ADHD = percent)

# Non-ADHD only
non_adhd <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 0) %>%
  get_percent_menarche() %>%
  rename(`Non-ADHD` = percent)

# Join all into one table
menarche_table <- full %>%
  left_join(adhd, by = "wave") %>%
  left_join(non_adhd, by = "wave")

# Optional: round percentages to 1 decimal place
menarche_table <- menarche_table %>%
  mutate(across(-wave, \(x) round(x, 1)))

# View table
print(menarche_table)
```

### Concordance between (raw) youth and parent menarche status

```{r youth parent menarche agreement}
# Filter out cases where either parent or youth report is missing
df_agreement <- adhd_menarche_clean %>%
  filter(!is.na(pds_f5b_p) & !is.na(pds_f5_y)) %>%
  transmute(
    parent = as.character(pds_f5b_p),
    youth  = as.character(pds_f5_y)
  )

# Run Cohen's Kappa on valid data only
kappa_result <- cohen.kappa(as.matrix(df_agreement))

# Print the Kappa result
print(kappa_result)

# Calculate and print percentage agreement
percent_agreement <- mean(df_agreement$parent == df_agreement$youth) * 100
print(percent_agreement)
```

## Breast development

### Breast development score missing

```{r NA breast development}
# Count missing breast developmenbt data by wave
missing_breast_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_breast = sum(is.na(pds_f4_p)),
            total = n(),
            missing_percent = (missing_breast / total) * 100)

# Print the results
print(missing_breast_by_wave)
```

### Breast development histogram

```{r breast histogram}
histogram_breast_data <- adhd_menarche_clean %>%
  filter(!is.na(pds_f4_p))

ggplot(histogram_breast_data, aes(x = pds_f4_p, fill = factor(wave))) +
  geom_histogram(binwidth = 1, color = "black", alpha = 0.7) +
  facet_wrap(~ wave, scales = "free_y") +
  labs(
    title = "Distribution of Breast Scores by Wave",
    x = "Breast Scores",
    y = "Count",
    fill = "Wave"
  ) +
  theme_minimal()
```

### Breast development boxplots

```{r breast boxplots}
boxplot_breast_data <- adhd_menarche_clean %>%
  filter(!is.na(pds_f4_2_y))

ggplot(boxplot_breast_data, aes(x = wave, y = pds_f4_2_y)) +
  geom_boxplot(fill = "skyblue", alpha = 0.7) +
  labs(
    title = "Distribution of Breast Scores by Wave",
    x = "Wave",
    y = "Breast Scores"
  ) +
  theme_minimal()
```

### Breast development summary statistics

```{r breast summary}
# Summary statistics by wave
summary_stats_breast_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(pds_f4_p, na.rm = TRUE),
    median_score = median(pds_f4_p, na.rm = TRUE),
    sd_score = sd(pds_f4_p, na.rm = TRUE),
    min_score = min(pds_f4_p, na.rm = TRUE),
    max_score = max(pds_f4_p, na.rm = TRUE),
    n_valid = sum(!is.na(pds_f4_p)),
    n_missing = sum(is.na(pds_f4_p))
  )

print(summary_stats_breast_by_wave)

# ADHD group
adhd_breast <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 1) %>%
    group_by(wave) %>%
  summarise(
    mean_score = mean(pds_f4_p, na.rm = TRUE),
    median_score = median(pds_f4_p, na.rm = TRUE),
    sd_score = sd(pds_f4_p, na.rm = TRUE),
    min_score = min(pds_f4_p, na.rm = TRUE),
    max_score = max(pds_f4_p, na.rm = TRUE),
    n_valid = sum(!is.na(pds_f4_p)),
    n_missing = sum(is.na(pds_f4_p))
  )

print(adhd_breast)

# Non-ADHD group
non_adhd_breast <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 0) %>%
    group_by(wave) %>%
  summarise(
    mean_score = mean(pds_f4_p, na.rm = TRUE),
    median_score = median(pds_f4_p, na.rm = TRUE),
    sd_score = sd(pds_f4_p, na.rm = TRUE),
    min_score = min(pds_f4_p, na.rm = TRUE),
    max_score = max(pds_f4_p, na.rm = TRUE),
    n_valid = sum(!is.na(pds_f4_p)),
    n_missing = sum(is.na(pds_f4_p))
  )

print(non_adhd_breast)
```

## ADHD traits

### ADHD traits missing

```{r NA cbcl attention}
# Count missing adhd traits data by wave
missing_traits_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_traits = sum(is.na(cbcl_scr_syn_attention_r)),
            total = n(),
            missing_percent = (missing_traits / total) * 100)

# Print the results
print(missing_traits_by_wave)
```

### ADHD traits histogram

```{r adhd traits histogram}
histogram_traits_data <- adhd_menarche_clean %>%
  filter(!is.na(cbcl_scr_syn_attention_r))

ggplot(histogram_traits_data, aes(x = cbcl_scr_syn_attention_r, fill = factor(wave))) +
  geom_histogram(binwidth = 1, color = "black", alpha = 0.7) +
  facet_wrap(~ wave, scales = "free_y") +
  labs(
    title = "Distribution of ADHD Traits by Wave",
    x = "CBCL Attention Syndrome Raw Scores",
    y = "Count",
    fill = "Wave"
  ) +
  theme_minimal()
```

### ADHD traits boxplots

```{r traits boxplots}
boxplot_traits_data <- adhd_menarche_clean %>%
  filter(!is.na(cbcl_scr_syn_attention_r))

ggplot(boxplot_traits_data, aes(x = wave, y = cbcl_scr_syn_attention_r)) +
  geom_boxplot(fill = "skyblue", alpha = 0.7) +
  labs(
    title = "Distribution of Attention Syndrome Scores by Wave",
    x = "Wave",
    y = "CBCL Attention Syndrome Raw Scores"
  ) +
  theme_minimal()
```

### ADHD traits summary statistics

```{r adhd dummary}
# Summary statistics by wave
summary_stats_traits_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(cbcl_scr_syn_attention_r, na.rm = TRUE),
    median_score = median(cbcl_scr_syn_attention_r, na.rm = TRUE),
    sd_score = sd(cbcl_scr_syn_attention_r, na.rm = TRUE),
    min_score = min(cbcl_scr_syn_attention_r, na.rm = TRUE),
    max_score = max(cbcl_scr_syn_attention_r, na.rm = TRUE),
    n_valid = sum(!is.na(cbcl_scr_syn_attention_r)),
    n_missing = sum(is.na(cbcl_scr_syn_attention_r))
  )

print(summary_stats_traits_by_wave)

# ADHD group
adhd_traits <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 1) %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(cbcl_scr_syn_attention_r, na.rm = TRUE),
    median_score = median(cbcl_scr_syn_attention_r, na.rm = TRUE),
    sd_score = sd(cbcl_scr_syn_attention_r, na.rm = TRUE),
    min_score = min(cbcl_scr_syn_attention_r, na.rm = TRUE),
    max_score = max(cbcl_scr_syn_attention_r, na.rm = TRUE),
    n_valid = sum(!is.na(cbcl_scr_syn_attention_r)),
    n_missing = sum(is.na(cbcl_scr_syn_attention_r))
  )

print(adhd_traits)

# Non-ADHD group
non_adhd_traits <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 0) %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(cbcl_scr_syn_attention_r, na.rm = TRUE),
    median_score = median(cbcl_scr_syn_attention_r, na.rm = TRUE),
    sd_score = sd(cbcl_scr_syn_attention_r, na.rm = TRUE),
    min_score = min(cbcl_scr_syn_attention_r, na.rm = TRUE),
    max_score = max(cbcl_scr_syn_attention_r, na.rm = TRUE),
    n_valid = sum(!is.na(cbcl_scr_syn_attention_r)),
    n_missing = sum(is.na(cbcl_scr_syn_attention_r))
  )

print(non_adhd_traits)
```

### ADHD traits overdispersion test

```{r overdispersion test traits}
poisson_model <- glm(cbcl_scr_syn_attention_r ~ 1, family = poisson, data = adhd_menarche_clean)

dispersion_test <- dispersiontest(poisson_model)

print(dispersion_test)
```

## Internalising

### Internalising missing

```{r NA cbcl internal}
# Count missing internalising data by wave
missing_internal_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_internal = sum(is.na(cbcl_scr_syn_internal_r)),
            total = n(),
            missing_percent = (missing_internal / total) * 100)

# Print the results
print(missing_internal_by_wave)
```

### Internalising histogram

```{r internal histogram}
histogram_internal_data <- adhd_menarche_clean %>%
  filter(!is.na(cbcl_scr_syn_internal_r))

ggplot(histogram_internal_data, aes(x = cbcl_scr_syn_internal_r, fill = factor(wave))) +
  geom_histogram(binwidth = 1, color = "black", alpha = 0.7) +
  facet_wrap(~ wave, scales = "free_y") +
  labs(
    title = "Distribution of Internalising Syndrome Scores by Wave",
    x = "CBCL Internalising Syndrome Raw Scores",
    y = "Count",
    fill = "Wave"
  ) +
  theme_minimal()
```

### Internalising boxplots

```{r internal boxplots}
boxplot_internal_data <- adhd_menarche_clean %>%
  filter(!is.na(cbcl_scr_syn_internal_r))

ggplot(boxplot_internal_data, aes(x = wave, y = cbcl_scr_syn_internal_r)) +
  geom_boxplot(fill = "skyblue", alpha = 0.7) +
  labs(
    title = "Distribution of Internalising Syndrome Scores by Wave",
    x = "Wave",
    y = "CBCL Internalising Syndrome Raw Scores"
  ) +
  theme_minimal()
```

### Internalising summary

```{r internal summary}
# Summary statistics by wave
summary_stats_internalising_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(cbcl_scr_syn_internal_r, na.rm = TRUE),
    median_score = median(cbcl_scr_syn_internal_r, na.rm = TRUE),
    sd_score = sd(cbcl_scr_syn_internal_r, na.rm = TRUE),
    min_score = min(cbcl_scr_syn_internal_r, na.rm = TRUE),
    max_score = max(cbcl_scr_syn_internal_r, na.rm = TRUE),
    n_valid = sum(!is.na(cbcl_scr_syn_internal_r)),
    n_missing = sum(is.na(cbcl_scr_syn_internal_r))
  )

print(summary_stats_internalising_by_wave)

# ADHD group
adhd_internalising <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 1) %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(cbcl_scr_syn_internal_r, na.rm = TRUE),
    median_score = median(cbcl_scr_syn_internal_r, na.rm = TRUE),
    sd_score = sd(cbcl_scr_syn_internal_r, na.rm = TRUE),
    min_score = min(cbcl_scr_syn_internal_r, na.rm = TRUE),
    max_score = max(cbcl_scr_syn_internal_r, na.rm = TRUE),
    n_valid = sum(!is.na(cbcl_scr_syn_internal_r)),
    n_missing = sum(is.na(cbcl_scr_syn_internal_r))
  )

print(adhd_internalising)

# Non-ADHD group
non_adhd_internalising <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 0) %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(cbcl_scr_syn_internal_r, na.rm = TRUE),
    median_score = median(cbcl_scr_syn_internal_r, na.rm = TRUE),
    sd_score = sd(cbcl_scr_syn_internal_r, na.rm = TRUE),
    min_score = min(cbcl_scr_syn_internal_r, na.rm = TRUE),
    max_score = max(cbcl_scr_syn_internal_r, na.rm = TRUE),
    n_valid = sum(!is.na(cbcl_scr_syn_internal_r)),
    n_missing = sum(is.na(cbcl_scr_syn_internal_r))
  )

print(non_adhd_internalising)
```

### Internalising overdispersion test

```{r overdispersion test internal}
poisson_model <- glm(cbcl_scr_syn_internal_r ~ 1, family = poisson, data = adhd_menarche_clean)

dispersion_test <- dispersiontest(poisson_model)

print(dispersion_test)
```

## Externalising

### Externalising missing

```{r NA external}
# Count missing externalising data by wave
missing_external_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_external = sum(is.na(cbcl_scr_syn_external_r)),
            total = n(),
            missing_percent = (missing_external / total) * 100)

# Print the results
print(missing_external_by_wave)
```

### Externalising histogram

```{r external graph}
histogram_external_data <- adhd_menarche_clean %>%
  filter(!is.na(cbcl_scr_syn_external_r))

ggplot(histogram_external_data, aes(x = cbcl_scr_syn_external_r, fill = factor(wave))) +
  geom_histogram(binwidth = 1, color = "black", alpha = 0.7) +
  facet_wrap(~ wave, scales = "free_y") +
  labs(
    title = "Distribution of Externalising Syndrome Scores by Wave",
    x = "CBCL Externalising Syndrome Raw Scores",
    y = "Count",
    fill = "Wave"
  ) +
  theme_minimal()
```

### Externalising boxplots

```{r external boxplots}
boxplot_external_data <- adhd_menarche_clean %>%
  filter(!is.na(cbcl_scr_syn_external_r))

ggplot(boxplot_external_data, aes(x = wave, y = cbcl_scr_syn_external_r)) +
  geom_boxplot(fill = "skyblue", alpha = 0.7) +
  labs(
    title = "Distribution of Externalising Syndrome Scores by Wave",
    x = "Wave",
    y = "CBCL Externalising Syndrome Raw Scores"
  ) +
  theme_minimal()
```

### Externalising summary statistics

```{r external summary}
# Summary statistics by wave
summary_stats_externalising_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(cbcl_scr_syn_external_r, na.rm = TRUE),
    median_score = median(cbcl_scr_syn_external_r, na.rm = TRUE),
    sd_score = sd(cbcl_scr_syn_external_r, na.rm = TRUE),
    min_score = min(cbcl_scr_syn_external_r, na.rm = TRUE),
    max_score = max(cbcl_scr_syn_external_r, na.rm = TRUE),
    n_valid = sum(!is.na(cbcl_scr_syn_external_r)),
    n_missing = sum(is.na(cbcl_scr_syn_external_r))
  )

print(summary_stats_externalising_by_wave)

# ADHD group
adhd_externalising <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 1) %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(cbcl_scr_syn_external_r, na.rm = TRUE),
    median_score = median(cbcl_scr_syn_external_r, na.rm = TRUE),
    sd_score = sd(cbcl_scr_syn_external_r, na.rm = TRUE),
    min_score = min(cbcl_scr_syn_external_r, na.rm = TRUE),
    max_score = max(cbcl_scr_syn_external_r, na.rm = TRUE),
    n_valid = sum(!is.na(cbcl_scr_syn_external_r)),
    n_missing = sum(is.na(cbcl_scr_syn_external_r))
  )

print(adhd_externalising)

# Non-ADHD group
non_adhd_externalising <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 0) %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(cbcl_scr_syn_external_r, na.rm = TRUE),
    median_score = median(cbcl_scr_syn_external_r, na.rm = TRUE),
    sd_score = sd(cbcl_scr_syn_external_r, na.rm = TRUE),
    min_score = min(cbcl_scr_syn_external_r, na.rm = TRUE),
    max_score = max(cbcl_scr_syn_external_r, na.rm = TRUE),
    n_valid = sum(!is.na(cbcl_scr_syn_external_r)),
    n_missing = sum(is.na(cbcl_scr_syn_external_r))
  )

print(non_adhd_externalising)
```

### Externalising overdispersion test

```{r overdispersion test external}
poisson_model <- glm(cbcl_scr_syn_external_r ~ 1, family = poisson, data = adhd_menarche_clean)

dispersion_test <- dispersiontest(poisson_model)

print(dispersion_test)
```

## Income-to-needs ratio (INR)

### INR missing

```{r NA inr}
# Count missing inr data by wave
missing_inr_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_inr = sum(is.na(inr)),
            total = n(),
            missing_percent = (missing_inr / total) * 100)

# Print the results
print(missing_inr_by_wave)
```

### INR histogram

```{r inr histogram}
histogram_inr_data <- adhd_menarche_clean %>%
  filter(!is.na(adi))

ggplot(histogram_inr_data, aes(x = inr, fill = factor(wave))) +
  geom_histogram(binwidth = 1, color = "black", alpha = 0.7) +
  facet_wrap(~ wave, scales = "free_y") +
  labs(
    title = "Distribution of INR by Wave",
    x = "INR",
    y = "Count",
    fill = "Wave"
  ) +
  theme_minimal()
```

### INR boxplots

```{r inr boxplots}
boxplot_external_data <- adhd_menarche_clean %>%
  filter(!is.na(inr))

ggplot(boxplot_external_data, aes(x = wave, y = inr)) +
  geom_boxplot(fill = "skyblue", alpha = 0.7) +
  labs(
    title = "Distribution of Income-to-Needs Ratio by Wave",
    x = "Wave",
    y = "Income-to-Needs Ratio"
  ) +
  theme_minimal()
```

### INR summary statistics

```{r inr summary}
# Summary statistics by wave
summary_stats_inr_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(inr, na.rm = TRUE),
    sd_score = sd(inr, na.rm = TRUE),
    min_score = min(inr, na.rm = TRUE),
    max_score = max(inr, na.rm = TRUE),
    n_valid = sum(!is.na(inr)),
    n_missing = sum(is.na(inr))
  )

print(summary_stats_inr_by_wave)

# ADHD group
adhd_inr <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 1) %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(inr, na.rm = TRUE),
    sd_score = sd(inr, na.rm = TRUE),
    min_score = min(inr, na.rm = TRUE),
    max_score = max(inr, na.rm = TRUE),
    n_valid = sum(!is.na(inr)),
    n_missing = sum(is.na(inr))
  )

print(adhd_inr)

# Non-ADHD group
non_adhd_inr <- adhd_menarche_clean %>%
  filter(adhd_diagnosis == 0) %>%
  group_by(wave) %>%
  summarise(
    mean_score = mean(inr, na.rm = TRUE),
    sd_score = sd(inr, na.rm = TRUE),
    min_score = min(inr, na.rm = TRUE),
    max_score = max(inr, na.rm = TRUE),
    n_valid = sum(!is.na(inr)),
    n_missing = sum(is.na(inr))
  )

print(non_adhd_inr)
```

## Race/Ethnicity

### Race/ethnicity missing

```{r NA ethnicity}
# Count missing ethnicity data by wave
missing_ethnicity_by_wave <- adhd_menarche_clean %>%
  group_by(wave) %>%
  summarise(missing_ethnicity = sum(is.na(ethnicity)),
            total = n(),
            missing_percent = (missing_ethnicity / total) * 100)

# Print the results
print(missing_ethnicity_by_wave)
```

### Race/ethnicity table

```{r ethnicity table}
# Function to calculate ethnicity percentages
get_ethnicity_percent <- function(data) {
  data %>%
    filter(wave == "wave_0") %>%
    group_by(ethnicity) %>%
    summarise(percent = n() / nrow(.) * 100)
}

# Full sample
full_ethnicity <- adhd_menarche_clean %>%
  filter(wave == "wave_0") %>%
  count(ethnicity) %>%
  mutate(`Full sample` = n / sum(n) * 100) %>%
  select(-n)

# ADHD group
adhd_ethnicity <- adhd_menarche_clean %>%
  filter(wave == "wave_0", adhd_diagnosis == 1) %>%
  count(ethnicity) %>%
  mutate(ADHD = n / sum(n) * 100) %>%
  select(-n)

# Non-ADHD group
non_adhd_ethnicity <- adhd_menarche_clean %>%
  filter(wave == "wave_0", adhd_diagnosis == 0) %>%
  count(ethnicity) %>%
  mutate(`Non-ADHD` = n / sum(n) * 100) %>%
  select(-n)

# Join all tables
ethnicity_table <- full_ethnicity %>%
  full_join(adhd_ethnicity, by = "ethnicity") %>%
  full_join(non_adhd_ethnicity, by = "ethnicity") %>%
  mutate(across(-ethnicity, \(x) round(x, 1))) %>%
  arrange(ethnicity)

# View the table
print(ethnicity_table)
```

# Group differences

## Age

```{r age t-tests}
wave_0_df <- adhd_menarche_clean %>%
  filter(wave == "wave_0") %>%
  select(adhd_diagnosis, age_years) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for age
age_years_wilcox_0 <- wilcox.test(age_years ~ adhd_diagnosis, 
                               data = wave_0_df,
                               exact = FALSE)
print("Age Wilcoxon Test baseline:")
print(age_years_wilcox_0)


wave_1_df <- adhd_menarche_clean %>%
  filter(wave == "wave_1") %>%
  select(adhd_diagnosis, age_years) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for age
age_years_wilcox_1 <- wilcox.test(age_years ~ adhd_diagnosis, 
                               data = wave_1_df,
                               exact = FALSE)
print("Age Wilcoxon Test wave 1:")
print(age_years_wilcox_1)

wave_2_df <- adhd_menarche_clean %>%
  filter(wave == "wave_2") %>%
  select(adhd_diagnosis, age_years) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for age
age_years_wilcox_2 <- wilcox.test(age_years ~ adhd_diagnosis, 
                               data = wave_2_df,
                               exact = FALSE)
print("Age Wilcoxon Test wave 2:")
print(age_years_wilcox_2)

wave_3_df <- adhd_menarche_clean %>%
  filter(wave == "wave_3") %>%
  select(adhd_diagnosis, age_years) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for age
age_years_wilcox_3 <- wilcox.test(age_years ~ adhd_diagnosis, 
                               data = wave_3_df,
                               exact = FALSE)
print("Age Wilcoxon Test wave 3:")
print(age_years_wilcox_3)
```

## Menarche status

```{r menarche status chi-squared}
waves <- paste0("wave_", 0:3)

for (w in waves) {
  cat("\n====================\n")
  cat("Wave:", w, "\n")

  # Filter for wave, valid ADHD diagnosis, and menarche status
  df <- adhd_menarche_clean %>%
    filter(wave == w,
           adhd_diagnosis %in% c(0, 1),
           menarche_status_p %in% c("Y", "N"))

  # Create contingency table
  tab <- table(df$adhd_diagnosis, df$menarche_status_p)
  colnames(tab) <- c("Pre-menarche", "Post-menarche")
  rownames(tab) <- c("No ADHD", "ADHD")
  print(tab)

  # Use Fisher's exact test if any cell has <5, otherwise chi-squared
  if (any(tab < 5)) {
    test_result <- fisher.test(tab)
    cat("Fisher's Exact Test p-value:", round(test_result$p.value, 4), "\n")
  } else {
    test_result <- chisq.test(tab)
    cat("Chi-squared Test p-value:", round(test_result$p.value, 4), "\n")
  }
}

```

## Breast development

```{r breast development t-tests}
wave_0_df <- adhd_menarche_clean %>%
  filter(wave == "wave_0") %>%
  select(adhd_diagnosis, pds_f4_p) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for breast development
breast_wilcox_0 <- wilcox.test(pds_f4_p ~ adhd_diagnosis, 
                                data = wave_0_df,
                                exact = FALSE)
print("Breast Wilcoxon Test baseline:")
print(breast_wilcox_0)


wave_1_df <- adhd_menarche_clean %>%
  filter(wave == "wave_1") %>%
  select(adhd_diagnosis, pds_f4_p) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for breast development
breast_wilcox_1 <- wilcox.test(pds_f4_p ~ adhd_diagnosis, 
                                data = wave_1_df,
                                exact = FALSE)
print("Breast development Wilcoxon Test wave 1:")
print(breast_wilcox_1)

wave_2_df <- adhd_menarche_clean %>%
  filter(wave == "wave_2") %>%
  select(adhd_diagnosis, pds_f4_p) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for breast development
breast_wilcox_2 <- wilcox.test(pds_f4_p ~ adhd_diagnosis, 
                                data = wave_2_df,
                                exact = FALSE)
print("Breast development Wilcoxon Test wave 1:")
print(breast_wilcox_2)

wave_3_df <- adhd_menarche_clean %>%
  filter(wave == "wave_3") %>%
  select(adhd_diagnosis, pds_f4_p) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for breast development
breast_wilcox_3 <- wilcox.test(pds_f4_p ~ adhd_diagnosis, 
                                data = wave_3_df,
                                exact = FALSE)
print("Breast development Wilcoxon Test wave 3:")
print(breast_wilcox_3)

```

## CBCL

### CBCL scores baseline

```{r cbcl baseline t-test}
wave_0_df <- adhd_menarche_clean %>%
  filter(wave == "wave_0") %>%
  select(adhd_diagnosis, cbcl_scr_syn_attention_r, 
         cbcl_scr_syn_internal_r, cbcl_scr_syn_external_r) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for cbcl_scr_syn_attention_r
attention_wilcox <- wilcox.test(cbcl_scr_syn_attention_r ~ adhd_diagnosis, 
                                data = wave_0_df,
                                exact = FALSE)
print("Attention Wilcoxon Test:")
print(attention_wilcox)

# Perform Wilcoxon test for cbcl_scr_syn_internal_r
internal_wilcox <- wilcox.test(cbcl_scr_syn_internal_r ~ adhd_diagnosis, 
                                data = wave_0_df,
                                exact = FALSE)
print("Internal Wilcoxon Test:")
print(internal_wilcox)

# Perform Wilcoxon test for cbcl_scr_syn_external_r
external_wilcox <- wilcox.test(cbcl_scr_syn_external_r ~ adhd_diagnosis, 
                                data = wave_0_df,
                                exact = FALSE)
print("External Wilcoxon Test:")
print(external_wilcox)
```

### CBCL wave 1

```{r cbcl wave 1 t-tests}
wave_1_df <- adhd_menarche_clean %>%
  filter(wave == "wave_1") %>%
  select(adhd_diagnosis, cbcl_scr_syn_attention_r, 
         cbcl_scr_syn_internal_r, cbcl_scr_syn_external_r) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for cbcl_scr_syn_attention_r
attention_wilcox <- wilcox.test(cbcl_scr_syn_attention_r ~ adhd_diagnosis, 
                                data = wave_1_df,
                                exact = FALSE)
print("Attention Wilcoxon Test:")
print(attention_wilcox)

# Perform Wilcoxon test for cbcl_scr_syn_internal_r
internal_wilcox <- wilcox.test(cbcl_scr_syn_internal_r ~ adhd_diagnosis, 
                                data = wave_1_df,
                                exact = FALSE)
print("Internal Wilcoxon Test:")
print(internal_wilcox)

# Perform Wilcoxon test for cbcl_scr_syn_external_r
external_wilcox <- wilcox.test(cbcl_scr_syn_external_r ~ adhd_diagnosis, 
                                data = wave_1_df,
                                exact = FALSE)
print("External Wilcoxon Test:")
print(external_wilcox)
```

### CBCL wave 2

```{r cbcl wave 2 t-tests}
wave_2_df <- adhd_menarche_clean %>%
  filter(wave == "wave_2") %>%
  select(adhd_diagnosis, cbcl_scr_syn_attention_r, 
         cbcl_scr_syn_internal_r, cbcl_scr_syn_external_r) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for cbcl_scr_syn_attention_r
attention_wilcox <- wilcox.test(cbcl_scr_syn_attention_r ~ adhd_diagnosis, 
                                data = wave_2_df,
                                exact = FALSE)
print("Attention Wilcoxon Test:")
print(attention_wilcox)

# Perform Wilcoxon test for cbcl_scr_syn_internal_r
internal_wilcox <- wilcox.test(cbcl_scr_syn_internal_r ~ adhd_diagnosis, 
                                data = wave_2_df,
                                exact = FALSE)
print("Internal Wilcoxon Test:")
print(internal_wilcox)

# Perform Wilcoxon test for cbcl_scr_syn_external_r
external_wilcox <- wilcox.test(cbcl_scr_syn_external_r ~ adhd_diagnosis, 
                                data = wave_2_df,
                                exact = FALSE)
print("External Wilcoxon Test:")
print(external_wilcox)
```

### CBCL wave 3

```{r cbcl wave 3 t-tests}
wave_3_df <- adhd_menarche_clean %>%
  filter(wave == "wave_3") %>%
  select(adhd_diagnosis, cbcl_scr_syn_attention_r, 
         cbcl_scr_syn_internal_r, cbcl_scr_syn_external_r) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for cbcl_scr_syn_attention_r
attention_wilcox <- wilcox.test(cbcl_scr_syn_attention_r ~ adhd_diagnosis, 
                                data = wave_3_df,
                                exact = FALSE)
print("Attention Wilcoxon Test:")
print(attention_wilcox)

# Perform Wilcoxon test for cbcl_scr_syn_internal_r
internal_wilcox <- wilcox.test(cbcl_scr_syn_internal_r ~ adhd_diagnosis, 
                                data = wave_3_df,
                                exact = FALSE)
print("Internal Wilcoxon Test:")
print(internal_wilcox)

# Perform Wilcoxon test for cbcl_scr_syn_external_r
external_wilcox <- wilcox.test(cbcl_scr_syn_external_r ~ adhd_diagnosis, 
                                data = wave_3_df,
                                exact = FALSE)
print("External Wilcoxon Test:")
print(external_wilcox)
```

## INR

```{r inr t-tests}
wave_0_df <- adhd_menarche_clean %>%
  filter(wave == "wave_0") %>%
  select(adhd_diagnosis, inr) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for inr
inr_wilcox_0 <- wilcox.test(inr ~ adhd_diagnosis, 
                               data = wave_0_df,
                               exact = FALSE)
print("INR Wilcoxon Test baseline:")
print(inr_wilcox_0)


wave_1_df <- adhd_menarche_clean %>%
  filter(wave == "wave_1") %>%
  select(adhd_diagnosis, inr) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for inr
inr_wilcox_1 <- wilcox.test(inr ~ adhd_diagnosis, 
                               data = wave_1_df,
                               exact = FALSE)
print("INR Wilcoxon Test wave 1:")
print(inr_wilcox_1)

wave_2_df <- adhd_menarche_clean %>%
  filter(wave == "wave_2") %>%
  select(adhd_diagnosis, inr) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for inr
inr_wilcox_2 <- wilcox.test(inr ~ adhd_diagnosis, 
                               data = wave_2_df,
                               exact = FALSE)
print("INR Wilcoxon Test wave 2")
print(inr_wilcox_2)

wave_3_df <- adhd_menarche_clean %>%
  filter(wave == "wave_3") %>%
  select(adhd_diagnosis, inr) %>%
  mutate(adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1),
                                 labels = c("No ADHD", "ADHD")))

# Perform Wilcoxon test for inr
inr_wilcox_3 <- wilcox.test(inr ~ adhd_diagnosis, 
                               data = wave_3_df,
                               exact = FALSE)
print("INR Wilcoxon Test wave 3")
print(inr_wilcox_3)
```

## Ethnicity

```{r ethnicity chi-squared}
# Define ethnicity label mapping
ethnicity_labels <- c(
  "1" = "Non-Hispanic White",
  "2" = "Non-Hispanic Black",
  "3" = "Hispanic",
  "4" = "Asian",
  "5" = "Other"
)

# Filter to wave_0 and valid ADHD diagnosis and ethnicity codes
wave_0_ethnicity_df <- adhd_menarche_clean %>%
  filter(wave == "wave_0",
         adhd_diagnosis %in% c(0, 1),
         ethnicity %in% 1:5) %>%
  mutate(
    ethnicity = factor(ethnicity_labels[as.character(ethnicity)], levels = ethnicity_labels),
    adhd_diagnosis = factor(adhd_diagnosis, levels = c(0, 1), labels = c("No ADHD", "ADHD"))
  )

# Display proportions
ethnicity_props <- prop.table(table(wave_0_ethnicity_df$adhd_diagnosis, wave_0_ethnicity_df$ethnicity), margin = 1) * 100
print(round(ethnicity_props, 1))

# Chi-squared or Fisher's test depending on counts
ethnicity_counts <- table(wave_0_ethnicity_df$adhd_diagnosis, wave_0_ethnicity_df$ethnicity)

if (any(ethnicity_counts < 5)) {
  test_result <- fisher.test(ethnicity_counts)
  cat("\nFisher's Exact Test p-value:", round(test_result$p.value, 4), "\n")
} else {
  test_result <- chisq.test(ethnicity_counts)
  cat("\nChi-squared Test p-value:", round(test_result$p.value, 4), "\n")
}

```
